---
title: "Optimisation in R"
author: "Joanna"
date: "2025-07-01"
output: html_document
---

## Read the data

```{r}
# Read Data
roots_df <- read_csv("yiyang_roots.csv")  # adjust path
# Select relevant columns (adjust based on output of colnames)
roots_data <- roots_df %>%
  select(D, L, RD, `C (%)`, `N(%)`, `C/N`, `ECM biomass Summer`, `ECM biomass whole year`, `Longevity Roots`, `Longevity Root Tips`, `Root biomass <2`, `Root biomass 2-5`) %>%
  rename(
    C_pct = `C (%)`,
    N_pct = `N(%)`,
    CN_ratio = `C/N`,
    ECM_summer = `ECM biomass Summer`,
    ECM_whole_year = `ECM biomass whole year`,
    Longevity_Roots = `Longevity Roots`,
    Longevity_Root_Tips = `Longevity Root Tips`,
    Root_biomass_2 = `Root biomass <2`,
    Root_biomass_2_5 = `Root biomass 2-5`
  )

soil_df  <- read_csv("yiyang_soil.csv")   # adjust path

# Convert to long format for ggplot
roots_long <- roots_df %>%
  mutate(D = roots_data$D,
         L = roots_data$L,
         RD = roots_data$RD)

roots_ostonen_2011 <- data.frame(
  D = c(0.279,  0.318,  0.381,  0.314,  0.336,  0.396,  0.297,  0.339,  0.358,  0.341,  0.319,  0.377),
  L = c(2.3,  2.2,  2.4,  1.8,  2.3,  1.5,  1.8,  1.4,  1.5,  1.5,  1.5,  1.10),
  RD = c(353, 350, 273, 308, 294, 239, 340, 263, 272, 292, 236, 244)
)

stele_ostonen_2003 <- data.frame(
  D = c(422.5, 370.0, 342.2, 315.7, 361.2)/1000,
  PRS = c(14.9, 13.4, 14.9, 15.2, 15.8)*0.01
)

Sys.setenv("VROOM_CONNECTION_SIZE" = 500000)
FRED <- read_csv("~/Downloads/FRED3_Entire_Database_2021.csv", n_max = 8)
FRED_small <- read_csv("~/Downloads/FRED3_Entire_Database_2021.csv", skip = 9)
names(FRED_small) <- names(FRED)
row_names = c(names(grep("cortex", FRED[1,], value = TRUE)[1]), names(grep("Root diameter", FRED[1,], value = TRUE)[3]), names(grep("stele", FRED[1,], value = TRUE)[c(1, 4, 8, 12)]))

c(grep("cortex", FRED[1,], value = TRUE)[1], grep("Root diameter", FRED[1,], value = TRUE)[3], grep("stele", FRED[1,], value = TRUE)[c(1, 4, 8, 12)])
summary(pine_stele)

pine_stele <- FRED_small %>%
  filter(F01286 == "Pinus",
         F01287 == "sylvestris",
         F00060 == "absorptive") %>%
  select(row_names)

spruce_stele <- FRED_small %>%
  filter(F01286 == "Picea",
         F01287 == "abies",
         F00060 == "absorptive") %>%
  select(row_names)

```

## Create the initial conditions

```{r setup, include=FALSE}
# Load required packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(nloptr)  # for constrained optimization
library(patchwork)  # for combining ggplot objects
library(tibble)
library(purrr)
library(rphydro)
library(callr)
library(parallel)
library(future.apply)
library(viridis)

# Environmental constants
TAverage <- 5.035873
Naverage <- 0.03
p0 <- 2.982752
y <- 365
A_const <- 0.05

# Constants
e_u_myco <- 0.5      # gN gC-1 d-1 mycorrhiza
e_u_root <- 0.17     # gN gC-1 d-1 roots
u_cB <- 10
u_cB_root <- 10
pi <- base::pi
beta_myco = 50
beta_root = 1000
mycorrhized = 0.9  # Heinonsalo, joskus
resp = 0
invetsment_tree = 0
invetsment_mcyo = 1

# CT VT MT
kg_fungi <- c(92.20, 116.49, 174.22)  # kg fungi per ha
number_of_trees <- c(407, 422, 684)   # number of trees per ha
fungi_per_tree <- kg_fungi / number_of_trees  # kg fungi per tree

fine_root_mass_per_ha <- c(189.01, 173.04, 90.56)  # g fine roots per ha
fine_root_mass_per_tree_g <- fine_root_mass_per_ha / number_of_trees  # g fine roots per tree
# More fungal compared to fine root biomass in more fertile sites!
tip_mass_mg <- c(0.12191424104274, 0.159271194001111, 0.175429654602955)  # mg
tip_mass_g <- tip_mass_mg / 1000  # convert mg to g

roots_per_tree <- fine_root_mass_per_tree_g / tip_mass_g  # number of roots per tree
fungi_per_root_kg <- fungi_per_tree / roots_per_tree  # kg fungi per root
fungi_per_root_g <- fungi_per_root_kg * 1000  # g fungi per root

# Fit a linear model
model <- lm(fungi_per_root_g ~ roots_per_tree)
# View the model summary
summary(model)
# Extract the gradient (slope)
gradient <- coef(model)[["roots_per_tree"]]
c_biomass <- coef(model)[["(Intercept)"]]
```

## Key relationships

```{r cars}
# Fit g using nonlinear least squares
fit_pine <- nls(
  D ~ g / L^(1/5),
  data = roots_df,
  start = list(g = 0.25)
)

g_pine <- coef(fit_pine)["g"]

fit_spruce <- nls(
  D ~ g / L^(1/5),
  data = roots_ostonen_2011,
  start = list(g = 0.25)
)

g_spruce <- coef(fit_spruce)["g"]

d_pine <- function(l) {
  l_safe <- pmax(l, 1e-4)
  g_pine / l_safe^(1/5)
}

d_spruce <- function(l) {
  l_safe <- pmax(l, 1e-4)
  g_spruce / l_safe^(1/5)
}

l_seq <- seq(0.5,
             max(c(roots_df$L, roots_ostonen_2011$L), na.rm = TRUE) + 0.5,
             length.out = 500)

curve_df <- data.frame(
  L = l_seq,
  Pine = d_pine(l_seq),
  Spruce = d_spruce(l_seq)
)

# Prepare point data
pine_points <- roots_df %>%
  select(L, D) %>%
  mutate(Species = "Pine")

spruce_points <- roots_ostonen_2011 %>%
  select(L, D) %>%
  mutate(Species = "Spruce")

# Pine
pine_points <- pine_points %>%
  mutate(Fit = d_pine(L),
         PercError = (D - Fit) / Fit * 100)

# Spruce
spruce_points <- spruce_points %>%
  mutate(Fit = d_spruce(L),
         PercError = (D - Fit) / Fit * 100)

# Combine for convenience
roots_all_res <- bind_rows(pine_points, spruce_points)

# Quick summary
diameter_percentage_area <- roots_all_res %>%
  group_by(Species) %>%
  summarise(
    MeanPercError = mean(PercError)
  )

# --- Create legend labels with MPE ---
pine_label <- paste0("Pine (MPE = ", round(diameter_percentage_area$MeanPercError[diameter_percentage_area$Species=="Pine"], 3), "%)")
spruce_label <- paste0("Spruce (MPE = ", round(diameter_percentage_area$MeanPercError[diameter_percentage_area$Species=="Spruce"], 3), "%)")

# Update species labels in points
pine_points$Species <- pine_label
spruce_points$Species <- spruce_label

roots_all <- bind_rows(pine_points, spruce_points)

# Update species labels in curve data
curve_df_long <- curve_df %>%
  pivot_longer(cols = c(Pine, Spruce),
               names_to = "Species",
               values_to = "RootDiameter")

curve_df_long$Species <- ifelse(
  curve_df_long$Species == "Pine",
  pine_label,
  spruce_label
)

# --- Plot with MPE in legend ---
p_diameter_length <- ggplot() +
  geom_line(data = curve_df_long,
            aes(x = L, y = RootDiameter, color = Species),
            linewidth = 1.4) +

  geom_point(data = roots_all,
             aes(x = L, y = D, color = Species),
             size = 5, alpha = 0.8) +

  labs(
    x = "Tip Length (mm)",
    y = "Root Diameter\n(mm)",
    color = "Species"
  ) +

  scale_color_manual(values = c("blue", "grey40")) +

  guides(color = guide_legend(override.aes = list(linetype = 0, size = 5))) +

  coord_cartesian(xlim = c(0.5, NA)) +

  theme_classic(base_size = 36) +
  theme(
    axis.text = element_text(size = 22),
    axis.title = element_text(size = 32),
    legend.position = "top",
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.text = element_text(size = 22),
    legend.key.size = unit(0.5, "cm")
  )
p_diameter_length
```

```{r}
# Step 1: Compute PRD for each species over its own L range
l_seq_pine <- seq(min(roots_df$L, na.rm=TRUE)-0.5,
                  max(roots_df$L, na.rm=TRUE)+0.5, length.out = 500)
l_seq_spruce <- seq(min(roots_ostonen_2011$L, na.rm=TRUE)-0.5,
                     max(roots_ostonen_2011$L, na.rm=TRUE)+0.5, length.out = 500)

# New coefficients
a_new <- 0.25   # baseline stele fraction (example)
b_new <- 0.02   # scaling for small roots (example)

PRD_pine <- function(l) (a_new + b_new / d_pine(l))^2
PRD_spruce <- function(l) (a_new + b_new / d_spruce(l))^2

nls_fit <- nls(PRS ~ (a + b / D)^2,
               data = stele_ostonen_2003,
               start = list(a = 0.2, b = 0.014))

a_new <- coef(nls_fit)["a"]
b_new <- coef(nls_fit)["b"]

PRS_curve_spruce <- PRD_spruce(l_seq_spruce)
PRS_curve_pine <- PRD_pine(l_seq_spruce)

ggplot() +
  geom_line(aes(x = d_spruce(l_seq_spruce),
                y = PRS_curve_spruce),
            color = "grey40", linewidth = 1.2) +
  geom_point(data = stele_ostonen_2003,
             aes(x = D, y = PRS),
             color = "grey40", size = 2, alpha = 0.7) +
  labs(x = "Root Diameter (mm)", y = "Stele Area Fraction (PRS)") +
  theme_classic()

```

```{r}
PRD_curve_spruce <- PRD_spruce(l_seq_spruce)
PRD_curve_pine <- PRD_pine(l_seq_pine)

# Pine PRD at measured points
roots_df <- roots_df %>%
  mutate(PRD_value = PRD_pine(L))

# Spruce PRD at measured points
roots_ostonen_2011 <- roots_ostonen_2011 %>%
  mutate(PRD_value = PRD_spruce(L))

# Step 2: Fit species-specific rho (slope + intercept)
# Linear model per species
# Step 2: Fit species-specific rho (NO intercept)
fit_pine_rho <- lm(RD ~ PRD_value, data = roots_df)
fit_spruce_rho <- lm(RD ~ PRD_value, data = roots_ostonen_2011)

# Now you have slope + intercept
s_pine <- coef(fit_pine_rho)["PRD_value"]
b_pine <- coef(fit_pine_rho)["(Intercept)"]

s_spruce <- coef(fit_spruce_rho)["PRD_value"]
b_spruce <- coef(fit_spruce_rho)["(Intercept)"]

# Compute curves
rho_curve_pine <- s_pine * PRD_curve_pine + b_pine
rho_curve_spruce <- s_spruce * PRD_curve_spruce + b_spruce

rho_pine <- function(PRD) s_pine * PRD + b_pine
rho_spruce <- function(PRD) s_spruce * PRD + b_spruce

# Step 4: Diameters
D_curve_pine <- d_pine(l_seq_pine)
D_curve_spruce <- d_spruce(l_seq_spruce)

# Pine MPE
roots_df <- roots_df %>%
  mutate(pred_RD = predict(fit_pine_rho, newdata = roots_df),
         abs_perc_err = abs(RD - pred_RD) / RD * 100)

pine_mpe <- mean(roots_df$abs_perc_err, na.rm = TRUE)

# Spruce MPE
roots_ostonen_2011 <- roots_ostonen_2011 %>%
  mutate(pred_RD = predict(fit_spruce_rho, newdata = roots_ostonen_2011),
         abs_perc_err = abs(RD - pred_RD) / RD * 100)

spruce_mpe <- mean(roots_ostonen_2011$abs_perc_err, na.rm = TRUE)

pine_label <- paste0("Pine (MPE = ", round(pine_mpe, 3), "%)")
spruce_label <- paste0("Spruce (MPE = ", round(spruce_mpe, 3), "%)")

# Step 5: Create plotting dataframe with updated labels
curve_full <- data.frame(
  D = c(D_curve_pine, D_curve_spruce),
  Rho = c(rho_curve_pine, rho_curve_spruce),
  species = rep(c(pine_label, spruce_label),
                times = c(length(D_curve_pine), length(D_curve_spruce)))
)

# --- Plot with MPE in legend ---
p_density_diameter <- ggplot() +
  geom_line(data = curve_full,
            aes(x = D, y = Rho, color = species),
            linewidth = 1.4) +

  geom_point(data = roots_df %>% mutate(species = pine_label),
             aes(x = D, y = RD, color = species),
             size = 5, alpha = 0.8) +

  geom_point(data = roots_ostonen_2011 %>% mutate(species = spruce_label),
             aes(x = D, y = RD, color = species),
             size = 5, alpha = 0.8) +

  labs(
    x = "Root Diameter (mm)",
    y = "Root Tissue\nDensity (kg/m³)",
    color = "Species"
  ) +

  scale_color_manual(values = c("blue", "grey40")) +

  guides(color = guide_legend(override.aes = list(linetype = 0, size = 5))) +

  theme_classic(base_size = 36) +
  theme(
    axis.text = element_text(size = 22),
    axis.text.x = element_text(size = 22, angle = 45, hjust = 1),
    axis.title = element_text(size = 32),
    legend.position = "top",
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.text = element_text(size = 22),
    legend.key.size = unit(0.5, "cm")
  )
p_density_diameter
```

```{r}
# l VERSION
C_NEW_pine <- function(n, l) {
  rho_pine(PRD_pine(l)) * (0.001 * d_pine(l))^2 * 0.001 * l * pi * n * 1/4
}

C_NEW_spruce <- function(n, l) {
  rho_spruce(PRD_spruce(l)) * (0.001 * d_spruce(l))^2 * 0.001 * l * pi * n * 1/4
}

# old version
C_OLD <- function(n, l, diameter, density) {
  density * (0.001 * diameter)^2 * 0.001 * l * pi * n * 1/4
}
```

```{r}
# --- Species-specific biomass calculations ---
new_biomass_pine <- C_NEW_pine(1, roots_df$L)              
new_biomass_spruce <- C_NEW_spruce(1, roots_ostonen_2011$L)

old_biomass_pine <- C_OLD(1, roots_df$L, roots_df$D, roots_df$RD)
old_biomass_spruce <- C_OLD(1, roots_ostonen_2011$L, roots_ostonen_2011$D, roots_ostonen_2011$RD)

# Combine into a single data frame
df_species <- data.frame(
  old_biomass = c(old_biomass_pine, old_biomass_spruce),
  new_biomass = c(new_biomass_pine, new_biomass_spruce),
  Species = rep(c("Pine", "Spruce"),
                times = c(length(old_biomass_pine), length(old_biomass_spruce)))
) %>%
  mutate(Species = as.character(Species))  # ensure character for joining

# Compute linear fit stats per species
calc_stats <- function(df) {
  fit <- lm(new_biomass ~ old_biomass, data = df)
  data.frame(
    R2 = summary(fit)$r.squared,
    slope = coef(fit)[["old_biomass"]],
    MPE = mean((df$new_biomass - df$old_biomass) / df$old_biomass * 100, na.rm = TRUE)
  )
}

calc_stats <- function(df) {
  fit <- lm(new_biomass ~ old_biomass, data = df)
  summ <- summary(fit)
  coefs <- summ$coefficients
  
  data.frame(
    R2 = summ$r.squared,
    
    intercept = coefs["(Intercept)", "Estimate"],
    intercept_p = coefs["(Intercept)", "Pr(>|t|)"],
    
    slope = coefs["old_biomass", "Estimate"],
    slope_p = coefs["old_biomass", "Pr(>|t|)"],
    
    MPE = mean((df$new_biomass - df$old_biomass) /
               df$old_biomass * 100, na.rm = TRUE)
  )
}

stats_species <- df_species %>%
  group_by(Species) %>%
  group_modify(~ calc_stats(.x)) %>%
  ungroup()

# Add R², slope, and MPE to Species labels
df_species <- df_species %>%
  left_join(stats_species, by = "Species") %>%
  mutate(
    Species_label = paste0(
      Species,
      " (slope = ", round(slope, 3),
      ", MPE = ", round(MPE, 3), "%)"
    )
  )

# Create named vector for colors
color_values <- setNames(
  c("blue", "grey40"),
  df_species %>% distinct(Species_label) %>% pull(Species_label)
)

# --- Plot ---
p_biomass_comparison <- ggplot(df_species, aes(x = old_biomass, y = new_biomass, color = Species_label)) +

  # Points
  geom_point(size = 5, alpha = 0.8) +

  # 1:1 reference line
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +

  # Species-specific linear fits
  geom_smooth(method = "lm", se = FALSE, aes(color = Species_label), linewidth = 1.2) +

  # Colors
  scale_color_manual(values = color_values) +

  # Labels
  labs(
    x = "Cylindrical Mass\nEstimation (kg)",
    y = "Root Length Based\nEstimation (kg)",
    color = "Species"
  ) +

  theme_classic(base_size = 36) +
  theme(
    axis.text = element_text(size = 22),
    axis.title = element_text(size = 32),
    legend.position = "top",
    legend.direction = "vertical",
    legend.background = element_rect(fill = alpha("white", 0.5)),
    legend.title = element_text(size = 26),
    legend.text = element_text(size = 24)
  )
p_biomass_comparison
```

```{r save-combined-plots}
combined_plot <- (p_biomass_comparison + (p_density_diameter / p_diameter_length) +
  plot_layout(widths = c(1.6, 1)) +
  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')) &
  theme(plot.tag = element_text(size = 28, face = "bold"))
ggsave("pine_spruce_combined.png", plot = combined_plot, width = 17, height = 12, dpi = 300)
```

# Age related

```{r}
a_age <- function(l) 3 - 3 * exp(-3 * d(l))
uptake_age <- function(l) 1 / (1 + exp(a_age(l) - 2.7))
```

## Uptake

```{r}
# TODO: think of a new fungal biomass
B_m <- function(n, gradient = -9.368976e-05, c = 0.3909945) {
  # Ding data
  out = gradient * n + c
  out[out < 0] <- 0
  
  # Old version - mistaken conversion
  out = 1000 * 4.81000339 * 10^(-8) * n
  
  # Exudate based?
  # out = (carbon - turnover_repiration) * B

  return(out)
}

# U_c_myco(N_av, n)
U_c_myco <- function(N_av, n) {
  numerator <- B_m(n) / 2 * B_m(n) / 2 * u_cB # * N_av * e_u_myco
  denominator <- u_cB + beta_myco * B_m(n) / 2 # + N_av * e_u_myco
  return(numerator / denominator)
}

# U_c(N_av, n, l)
U_c <- function(N_av, n, l) {
  surface_area <- d(l) * l * pi * n
  numerator <- d(l) * l * pi * n * (1 / 8 * 10^(-6)) * rho(PRD(l)) * d(l) * u_cB_root * surface_area # * N_av * e_u_root
  denominator <- u_cB_root + beta_root * surface_area # + N_av * e_u_root 
  return(numerator / denominator)
}

# U_myco(N_av, n)
U_myco <- function(N_av, n) {
  val <- U_c_myco(N_av, n)
  return(val * N_av / (val + N_av))
}

# U(N_av, n, l)
U <- function(N_av, n, l) {
  val <- U_c(N_av, n, l)
  return(val * N_av / (val + N_av))
}

# gateway for the nitrogen(l)
nitrogen_gate <- function(n, l, K = 10) {
  area_total <- n * pi * d(l) * l
  return(area_total^2 / (area_total^2 + K^2))
}

# Final uptake function U_a(N_av, n, l)
U_a <- function(N_av, n, l) {
  uptake_age_val <- uptake_age(l)
  
  return(uptake_age_val * ((1 - mycorrhized) * U(N_av, n, l) + mycorrhized * nitrogen_gate(n, l) * invetsment_mcyo * U_myco(N_av, n)))
}
```

```{r}
# Define n and l ranges
n_vals <- seq(0, 50, length.out = 100)
l_vals <- seq(0, 5, length.out = 100)

# Vectorize the function for grid application
gate_vec <- Vectorize(nitrogen_gate)

# Create grid of n and l
grid <- expand.grid(n = n_vals, l = l_vals)
grid$gate_val <- gate_vec(grid$n, grid$l)

# Plot
ggplot(grid, aes(x = n, y = l, fill = gate_val)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", name = "Gate Value") +
  labs(
    title = "Nitrogen Gate Value across Root Tip Number and Length",
    x = "Number of Root Tips (n)",
    y = "Root Length (l)"
  ) +
  theme_minimal()

```

## Heatmap of the root and mycorrhizal uptake

```{r}
# Define parameter ranges
n_vals <- seq(0, 50, length.out = 100)
l_vals <- seq(0, 5, length.out = 100)

N_av_vals <- c(0.05, 0.1, 0.3, 0.6)

# Create grid
grid <- expand.grid(n = n_vals, l = l_vals, N_av = N_av_vals)

# Evaluate both functions (you already have these defined)
grid <- grid %>%
  mutate(
    U_myco_val = mapply(U_myco, N_av, n),
    U_root_val = mapply(U, N_av, n, l)
  )

plot_data_long <- grid %>%
  pivot_longer(cols = c(U_myco_val, U_root_val), 
               names_to = "source", 
               values_to = "uptake") %>%
  mutate(source = recode(source,
                         U_myco_val = "Mycorrhizal Uptake",
                         U_root_val = "Root Uptake"))

ggplot(plot_data_long, aes(x = n, y = l, fill = uptake)) +
  geom_tile() +
  scale_fill_viridis_c(option = "turbo", na.value = "grey90") +
  facet_grid(source ~ N_av, labeller = label_bquote(cols = N[av] == .(N_av))) +
  labs(
    title = "Nitrogen Uptake from Mycorrhiza vs Roots across N_av",
    x = "Number of root tips (n)",
    y = "Root length (l)",
    fill = "Uptake"
  ) +
  theme_minimal()

```

## Heatmap of the Uptake

```{r}
# Create grid of n and l values
# Define sequences
n_vals <- seq(0, 50, length.out = 100)
l_vals <- seq(0, 5, length.out = 100)

N_av_vals <- c(0.05, 0.1, 0.3, 0.6)

U_a_vec <- Vectorize(U_a)

# Build data for all N_av in one data frame:
all_grids <- lapply(N_av_vals, function(N_av) {
  grid <- expand.grid(n = n_vals, l = l_vals)
  grid$U_a_val <- U_a_vec(N_av, grid$n, grid$l)
  grid$N_av <- N_av
  return(grid)
})

grid_df_all <- do.call(rbind, all_grids)

# Free up unused memory
rm(all_grids)
gc()

# Plot with facets:
ggplot(grid_df_all, aes(x = n, y = l, fill = U_a_val)) +
  geom_tile() +
  facet_wrap(~ N_av, labeller = label_bquote(N[av] == .(N_av))) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  labs(
    title = expression("Uptake rate "~U[a]~"across n, l for various"~N[av]),
    x = "Number of root tips (n)",
    y = "Root length (l)",
    fill = expression(U[a])
  ) +
  theme_minimal()

```
## Heatmap of the Carbon Cost

```{r}
# Define sequences
n_vals <- seq(0, 50, length.out = 100)
l_vals <- seq(0.1, 5, length.out = 100)  # Avoid l = 0

# Create grid of n and l values
grid_df_C <- expand.grid(n = n_vals, l = l_vals)

# Vectorize your function C if not done already
C_vec <- Vectorize(C)

# Compute carbon cost C(n, l) efficiently
grid_df_C$C_val <- C_vec(grid_df_C$n, grid_df_C$l)

# Optional: remove intermediate variables and clean memory if needed
gc()

# Plot heatmap
ggplot(grid_df_C, aes(x = n, y = l, fill = C_val)) +
  geom_tile() +
  scale_fill_viridis_c(option = "magma") +
  labs(
    title = expression("Carbon Cost " ~ C(n, l)),
    x = "Number of root tips (n)",
    y = "Root tip length (l)",
    fill = expression(C(n, l))
  ) +
  theme_minimal()

```

## Photosynthesis from phydro

```{r}
photosynthesis <- function(U, j) {
  kphio = 0.087;        # quantum yield efficiency
  ppfd = 300;          # umol/m2/s
  ppfd_max = 1200;
  vpd  = 1000;         # Pa
  co2  = 400;          # ppm
  elv  = 0;            # m.a.s.l.
  fapar = 0.7;         # fraction
  rdark = 0.015;
  tc = 25;
  vwind = 3;
  netrad = ppfd/2
  psi_soil = -0.05
  pa = calc_patm(elv)
  
  par_cost = list(alpha=0.1, gamma=1, nitrogen_store_conversion = 1);
  par_plant = list(conductivity=3e-17, psi50=-2, b=2);
  options = list(gs_method = "GS_IGF", 
               et_method = "ET_DIFFUSION",
               ftemp_vj_method = "FV_kumarathunge19",
               ftemp_rd_method = "FR_heskel16",
               ftemp_br_method = "FB_atkin15",
               scale_alpha = F)

  ### New parameters TODO: should work with 800!
  a_jmax = 800
  
  out = rphydro_nitrogen(tc, tc, ppfd, netrad, vpd, co2, pa, j*U, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, par_cost, options)
  return(out$a)
}
```

## Timeout

```{r}
# --- Define photosynthesis function ---
photosynthesis <- function(U, j) {
  kphio = 0.087        # quantum yield efficiency
  ppfd = 300           # umol/m2/s
  ppfd_max = 1200
  vpd  = 1000          # Pa
  co2  = 400           # ppm
  elv  = 0             # m.a.s.l.
  fapar = 0.7          # fraction
  rdark = 0.015
  tc = 25
  vwind = 3
  netrad = ppfd / 2
  psi_soil = -0.05
  pa = calc_patm(elv)

  par_cost = list(alpha = 0.1, gamma = 1, nitrogen_store_conversion = 1)
  par_plant = list(conductivity = 3e-17, psi50 = -2, b = 2)
  options = list(
    gs_method = "GS_IGF",
    et_method = "ET_DIFFUSION",
    ftemp_vj_method = "FV_kumarathunge19",
    ftemp_rd_method = "FR_heskel16",
    ftemp_br_method = "FB_atkin15",
    scale_alpha = FALSE
  )

  a_jmax = 800

  out <- rphydro_nitrogen(tc, tc, ppfd, netrad, vpd, co2, pa,
                          j * U, fapar, kphio, psi_soil, rdark, vwind,
                          a_jmax, par_plant, par_cost, options)

  return(out$a)
}

```

## Looking at the j parameter

```{r}
# --- Define fixed root parameters ---
n <- 20
l <- 1
j_vals <- seq(1, 1000000, length.out = 100)

# --- Define N_av values to vary ---
N_av_vals <- c(0.05, 0.1, 0.3, 0.6)

# --- Set up parallel backend ---
num_cores <- detectCores() - 1
cl <- makeCluster(num_cores)

# --- Load libraries and export functions/vars to cluster ---
invisible(clusterEvalQ(cl, library(rphydro)))
clusterExport(cl, varlist = c("photosynthesis", "U_a", "n", "l"), envir = .GlobalEnv)

# --- Collect data for each N_av ---
photosynthesis_list <- lapply(N_av_vals, function(N_av) {
  # Compute U for fixed n, l at this N_av
  Uval <- U_a(N_av, n, l)
  clusterExport(cl, varlist = "Uval", envir = environment())
  
  # Run photosynthesis(Uval, j) across j_vals
  P_vals <- parSapply(cl, j_vals, function(j) {
    tryCatch(photosynthesis(Uval, j), error = function(e) NA)
  })
  
  # Return data frame for plotting
  data.frame(
    j = j_vals,
    P = P_vals,
    N_av = N_av
  )
})

# --- Shut down cluster ---
stopCluster(cl)

# --- Combine results ---
df_photosynthesis <- bind_rows(photosynthesis_list)

# --- Plot: Photosynthesis vs j, faceted by N_av ---
ggplot(df_photosynthesis, aes(x = j, y = P)) +
  geom_line(color = "darkblue", linewidth = 1.2) +
  facet_wrap(~ N_av, labeller = label_bquote(N[av] == .(N_av))) +
  labs(
    title = expression("Photosynthesis vs j for different" ~ N[av]),
    x = "j (efficiency / light input)",
    y = "Photosynthesis rate"
  ) +
  theme_minimal()
```

## Looking at the l and n

```{r}
# --- Fixed parameters ---
j <- 100000
n_vals <- seq(5, 50, length.out = 100)
l_vals <- seq(0.5, 5, length.out = 100)

# --- N_av values to loop over ---
N_av_vals <- c(0.05, 0.1, 0.3, 0.6)

# --- Create empty list to store results ---
results_list <- list()

# --- Loop over N_av values ---
for (N_av in N_av_vals) {
  # Step 1: Create grid of n and l
  grid_df <- expand.grid(n = n_vals, l = l_vals)
  
  # Step 2: Compute Uval = U_a(N_av, n, l)
  grid_df$Uval <- mapply(function(n, l) {
    tryCatch(U_a(N_av, n, l), error = function(e) NA)
  }, grid_df$n, grid_df$l)
  
  # Step 3: Parallel setup
  num_cores <- detectCores() - 1
  cl <- makeCluster(num_cores)
  
  invisible(clusterEvalQ(cl, library(rphydro)))
  clusterExport(cl, varlist = c("photosynthesis", "j"), envir = environment())
  clusterExport(cl, varlist = "Uvals", envir = list2env(list(Uvals = grid_df$Uval)))
  
  # Step 4: Parallel photosynthesis computation
  grid_df$P_val <- parSapply(cl, grid_df$Uval, function(Uval) {
    tryCatch(photosynthesis(Uval, j), error = function(e) NA)
  })
  
  stopCluster(cl)
  
  # Step 5: Add N_av identifier
  grid_df$N_av <- N_av
  
  # Step 6: Store
  results_list[[as.character(N_av)]] <- grid_df
}

# --- Combine all into one dataframe ---
combined_df <- bind_rows(results_list)

# --- Plot all four as facets ---
ggplot(combined_df, aes(x = n, y = l, fill = P_val)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  facet_wrap(~ N_av, labeller = label_bquote(N[av] == .(N_av))) +
  labs(
    title = expression("Photosynthesis " ~ P(n, l) ~ "across N"[av]),
    x = "Number of root tips (n)",
    y = "Root tip length (l)",
    fill = "Photosynthesis"
  ) +
  theme_minimal()
```

## Objective being defined

```{r}
# Sequence of N_av values
nav_seq <- seq(0.01, 1, by = 0.01)

# Define functions first
objective <- function(x, N_av, j, q) {
  n <- x[1]; l <- x[2]

  # Reject invalid values early
  if (is.na(n) || is.na(l) || n <= 0 || l <= 0) {
    message("Invalid n or l:", n, l)
    return(1e6)
  }

  # Compute uptake value
  Uval <- tryCatch(U_a(N_av, n, l), error = function(e) NA)
  if (is.na(Uval) || !is.finite(Uval)) {
    message("Invalid Uval:", Uval, " for n=", n, " l=", l)
    return(1e6)
  }

  # Compute photosynthesis, wrapped with tryCatch
  Pval <- tryCatch(photosynthesis(Uval, j), error = function(e) NA)
  if (is.na(Pval) || !is.finite(Pval)) {
    message("Invalid Pval:", Pval, " for Uval=", Uval)
    return(1e6)
  }

  # Compute carbon cost
  Cval <- tryCatch(C(n, l), error = function(e) NA)
  if (is.na(Cval) || !is.finite(Cval)) {
    message("Invalid Cval:", Cval, " for n=", n, " l=", l)
    return(1e6)
  }

  val <- Pval - invetsment_tree * nitrogen_gate(n, l) * Pval - q * (1 + resp) * Cval
  if (!is.finite(val)) {
    message("Non-finite objective val:", val)
    return(1e6)
  }

  return(val)
}
```

## Objective function vs q

```{r}
# --- Define a range of q values ---
q_vals <- seq(0.1, 5000, length.out = 200)

# --- Compute objective for each q ---
obj_vals <- sapply(q_vals, function(q) {
  val <- tryCatch(
    objective(c(n, l), N_av, j, q),
    error = function(e) NA
  )
  if (is.na(val) || !is.finite(val)) return(1e6)  # Penalize invalid
  return(val)
})

# --- Create a dataframe for plotting ---
df <- data.frame(q = q_vals, objective = obj_vals)

# --- Plot the results ---
ggplot(df, aes(x = q, y = objective)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  labs(
    title = paste("Objective function vs q"),
    subtitle = paste("n =", n, "l =", l, "N_av =", N_av, "j =", j),
    x = "q (carbon cost multiplier)",
    y = "Objective value"
  ) +
  theme_minimal()

N_av_values <- c(0.1, 0.3, 0.5, 0.7)

# --- Grid of n and l ---
n_vals <- seq(5, 50, length.out = 100)
l_vals <- seq(0.1, 5, length.out = 100)
grid_base <- expand.grid(n = n_vals, l = l_vals)

q = 50000

# --- Compute objective for each N_av ---
all_grids <- lapply(N_av_values, function(N_av) {
  grid <- grid_base
  grid$N_av <- N_av
  grid$objective <- mapply(function(n, l) {
    val <- tryCatch(objective(c(n, l), N_av, j, q), error = function(e) NA)
    if (is.na(val) || !is.finite(val)) return(NA)
    return(val)
  }, grid$n, grid$l)
  return(grid)
})

# --- Combine into one dataframe ---
plot_data <- bind_rows(all_grids)

plot_data <- plot_data[plot_data$objective > 0,]

summary(plot_data$objective)

# --- Plot all in a 2x2 grid ---
# Step 1: Find max points
max_points <- plot_data %>%
  group_by(N_av) %>%
  filter(objective == max(objective, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(label = paste0("n = ", round(n, 2), "\n", "l = ", round(l, 4)))

# Step 2: Plot
ggplot(plot_data, aes(x = n, y = l, fill = objective)) +
  geom_tile() +
  scale_fill_viridis_c(option = "turbo", na.value = "grey90") +
  facet_wrap(~ N_av, labeller = label_bquote(N[av] == .(N_av))) +
  geom_point(data = max_points, aes(x = n, y = l), 
             shape = 21, size = 2, stroke = 1, fill = "white", color = "black") +
  geom_text(data = max_points, aes(label = label), 
            vjust = -0.75, hjust = 1.1, size = 3, lineheight = 0.9, col = "black") +
  labs(
    title = "Objective(n, l) across different N_av values",
    x = "Number of root tips (n)",
    y = "Root length (l)",
    fill = "Objective"
  ) +
  theme_minimal()

```

## Constraint

```{r}
A_const <- 3000  # example value

# === COPY YOUR CONSTRAINT FUNCTION ===
constraint_fun <- function(x, A_const) {
  n <- x[1]; l <- x[2]
  if (l <= 0 || n <= 0) return(-1e6)  # violation
  
  dval <- d(l)
  if (dval <= 0 || !is.finite(dval)) return(-1e6)
  
  PRD_val <- (0.2 + 0.014 / dval)^2
  rho_val <- s * PRD_val + b
  
  lhs <- rho_val * dval^2 * l * pi * 1e-6 * n
  rhs <- 4 * A_const
  val <- rhs - lhs
  
  if (!is.finite(val)) return(-1e6)
  return(val)
}

# === CREATE A GRID OF (n, l) VALUES ===
n_vals <- seq(5, 50, length.out = 100)
l_vals <- seq(0.1, 4, length.out = 100)
grid <- expand.grid(n = n_vals, l = l_vals)

# === EVALUATE CONSTRAINT FUNCTION ===
grid$constraint_val <- apply(grid, 1, function(row) {
  constraint_fun(c(row['n'], row['l']), A_const)
})

# === ADD FEASIBILITY COLUMN ===
grid$feasible <- grid$constraint_val >= 0

# === PLOT ===
ggplot(grid, aes(x = n, y = l, fill = feasible)) +
  geom_raster() +
  scale_fill_manual(values = c("red", "green"), labels = c("Infeasible", "Feasible")) +
  labs(
    title = "Feasible Region for Constraint",
    x = "n (number of units)",
    y = "l (length)",
    fill = "Constraint"
  ) +
  theme_minimal()

```

## Objective to maximise

```{r}
A_const = 0.05

constraint_fun <- function(x, A_const) {
  n <- x[1]; l <- x[2]
  if (l <= 0 || n <= 0) return(-1e6)  # violation
  
  PRD_val <- (0.2 + 0.014 / d(l))^2
  rho_val <- s * PRD_val + b
  
  lhs <- rho_val * d(l)^2 * l * pi * 1e-6 * n
  rhs <- 4 * A_const
  val <- rhs - lhs
  
  if (!is.finite(val)) return(-1e6)
  return(val)
}

optimize_for_nav <- function(Nav) {
  tryCatch({
    res <- nloptr(
      x0 = c(15, 1),
      eval_f = function(x) {
        val <- -objective(x, Nav, j, q)
        if (is.na(val) || !is.finite(val)) {
          message(sprintf("Bad objective for Nav=%.3f, x=%s: %s", Nav, paste(round(x,3), collapse = ","), val))
          return(1e6)  # large penalty for optimizer to avoid
        }
        return(val)
      },
      lb = c(5, 0.1),
      ub = c(50, 4),
      # eval_g_ineq = function(x) constraint_fun(x, A_const),
      opts = list(
        algorithm = "NLOPT_LN_COBYLA",
        xtol_rel = 1e-6,
        maxeval = 1000
      )
    )
    
    if (any(is.na(res$solution)) || !is.finite(res$objective)) {
      message(sprintf("Optimization returned NA/Inf for Nav=%.3f", Nav))
    }
    
    tibble(
      Nav = Nav,
      opt_n = res$solution[1],
      opt_l = res$solution[2],
      obj_val = -res$objective,
      status = res$status
    )
  }, error = function(e) {
    warning(sprintf("Optimization failed for Nav=%.3f : %s", Nav, e$message))
    tibble(Nav = Nav, opt_n = NA, opt_l = NA, obj_val = NA, status = -1)
  })
}

# Number of cores (use one less than max)
num_cores <- max(1, detectCores() - 1)
cl <- makeCluster(num_cores)

# Load required packages quietly on workers
invisible(clusterEvalQ(cl, {
  suppressPackageStartupMessages({
    library(nloptr)
    library(tibble)
    library(rphydro)
  })
}))

# Export all needed variables and functions to workers
clusterExport(cl, varlist = c(
  "j", "q", "s", "b", "A_const",
  "d", "U_a", "C", "photosynthesis",
  "objective", "constraint_fun", "optimize_for_nav"
), envir = .GlobalEnv)

# Run optimization in parallel
results_list <- parLapply(cl, nav_seq, optimize_for_nav)

# Stop cluster
stopCluster(cl)

# Combine results
results_df <- do.call(rbind, results_list)
```

## Plot the results

```{r}
# Remove failed optimization rows (optional)
opt_results <- results_df |> filter(!is.na(opt_n), status > 0)

# Plot 1: Optimal n vs Nav
p1 <- ggplot(opt_results, aes(x = Nav, y = opt_n)) +
  geom_line(color = "steelblue", linewidth = 1) +
  labs(
    title = expression("Optimal number of root tips "~italic(n)~" vs "~N[av]),
    x = expression(N[av]),
    y = expression("Optimal "~italic(n))
  ) +
  theme_minimal()

# Plot 2: Optimal l vs Nav
p2 <- ggplot(opt_results, aes(x = Nav, y = opt_l)) +
  geom_line(color = "firebrick", linewidth = 1) +
  labs(
    title = expression("Optimal root length "~italic(l)~" vs "~N[av]),
    x = expression(N[av]),
    y = expression("Optimal "~italic(l))
  ) +
  theme_minimal()

# Combine plots vertically
p1 / p2


```

## Parameter testing

```{r}
# Constants
N_av_fixed <- 0.2
j_fixed <- 250
q_fixed <- 250
A_fixed <- 100

# Parameter grids (adjust ranges as needed)
j_vals <- seq(100, 1000, length.out = 20)
q_vals <- seq(50, 5000000, length.out = 20)
A_vals <- seq(50, 500, length.out = 20)

# Define optimization function with debugging messages
optimize_for_nav2 <- function(N_av, j, q, A_const) {
  tryCatch({
    res <- nloptr(
      x0 = c(15, 1),
      eval_f = function(x) -objective(x, N_av, j, q),
      lb = c(0.1, 0.1),
      ub = c(50, 5),
      eval_g_ineq = function(x) constraint_fun(x, A_const),
      opts = list(
        algorithm = "NLOPT_LN_COBYLA",
        xtol_rel = 1e-6,
        maxeval = 1000
      )
    )
    message(sprintf("Success: j=%.1f, q=%.1f, A=%.1f, n=%.2f, l=%.2f",
                    j, q, A_const, res$solution[1], res$solution[2]))
    tibble(
      param = NA_real_,
      variable = NA_character_,
      Nav = N_av,
      opt_n = res$solution[1],
      opt_l = res$solution[2],
      obj_val = -res$objective,
      status = res$status
    )
  }, error = function(e) {
    message(sprintf("FAIL: j=%.1f, q=%.1f, A=%.1f | %s", j, q, A_const, e$message))
    tibble(param = NA_real_, variable = NA_character_, Nav = N_av, opt_n = NA, opt_l = NA, obj_val = NA, status = -1)
  })
}

# Helper to run over parameter vector and label results
run_param_grid <- function(param_vec, param_name) {
  cl <- makeCluster(max(1, detectCores() - 1))
  
  # Ensure cluster stops no matter what
  on.exit(stopCluster(cl))
  
  clusterEvalQ(cl, {
    suppressPackageStartupMessages({
      library(nloptr)
      library(tibble)
      library(dplyr)
      library(rphydro)
    })
  })
  
  clusterExport(cl, varlist = c(
    "objective", "constraint_fun", "optimize_for_nav2",
    "U_a", "photosynthesis", "C", "d", "s", "b",
    "j_fixed", "q_fixed", "A_fixed",
    "N_av_fixed"
  ), envir = .GlobalEnv)
  
  results <- parLapply(cl, param_vec, function(pval) {
    if(param_name == "j") {
      res <- optimize_for_nav2(N_av_fixed, pval, q_fixed, A_fixed)
    } else if(param_name == "q") {
      res <- optimize_for_nav2(N_av_fixed, j_fixed, pval, A_fixed)
    } else if(param_name == "A_const") {
      res <- optimize_for_nav2(N_av_fixed, j_fixed, q_fixed, pval)
    }
    res$param <- pval
    res$variable <- param_name
    return(res)
  })
  
  bind_rows(results)
}

# Run parameter sweeps in parallel safely
results_j <- run_param_grid(j_vals, "j")
results_q <- run_param_grid(q_vals, "q")
results_A <- run_param_grid(A_vals, "A_const")

all_results <- bind_rows(results_j, results_q, results_A)

# Remove failed optimizations
all_results_clean <- all_results %>% filter(!is.na(opt_n), status > 0)

# Plotting function (requires patchwork for p1 / p2 layout)
plot_param <- function(df, param_name) {
  p1 <- ggplot(df, aes(x = param, y = opt_n)) +
    geom_line(color = "steelblue", size = 1) +
    labs(title = paste("Optimal n vs", param_name), x = param_name, y = "Optimal n") +
    theme_minimal()
  
  p2 <- ggplot(df, aes(x = param, y = opt_l)) +
    geom_line(color = "firebrick", size = 1) +
    labs(title = paste("Optimal l vs", param_name), x = param_name, y = "Optimal l") +
    theme_minimal()
  
  return(p1 / p2)
}

# Display plots
print(plot_param(filter(all_results_clean, variable == "j"), "j"))
print(plot_param(filter(all_results_clean, variable == "q"), "q"))
print(plot_param(filter(all_results_clean, variable == "A_const"), "A_const"))
```


```{r}
optimize_for_nav2_minimal <- function() {
  nloptr(
    x0 = c(15, 1),
    eval_f = function(x) sum(x^2),
    lb = c(0.1, 0.1),
    ub = c(50, 5),
    opts = list(
      algorithm = "NLOPT_LN_COBYLA",
      xtol_rel = 1e-6,
      maxeval = 100
    )
  )
}

cl <- makeCluster(2)
clusterEvalQ(cl, library(nloptr))
clusterExport(cl, varlist = c(
    "objective", "constraint_fun", "optimize_for_nav2_minimal",
    "U_a", "photosynthesis", "C", "d", "s", "b",
    "j_fixed", "q_fixed", "A_fixed",
    "N_av_fixed"
  ), envir = .GlobalEnv)
res <- parLapply(cl, 1:2, function(i) optimize_for_nav2_minimal())
stopCluster(cl)
print(res)


```
