---
title: "Nitrogen Uptake Model — Whole-System Diffusion Budget"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(gridExtra)
library(patchwork)
library(viridis)
dir.create("vignettes/plots", showWarnings = FALSE, recursive = TRUE)
```

## Overview

This document implements a nitrogen uptake model based on a whole-system
diffusion budget. The total nitrogen available to the root system per unit
time is limited by diffusive flux into the root zone from bulk soil. A
mean soil concentration $\bar{N}$ adjusts self-consistently until diffusive
supply equals total biochemical demand.

A fraction $m$ of root surface is colonised by ECM fungi. The colonised
surface is covered by the Hartig net and does not contribute to direct
uptake. Active absorbing surface area is therefore:

$$\mathrm{SA}_\text{active} = (1-m)\,\mathrm{SA}_{fr} + \mathrm{SA}_m$$

The steady-state balance is:

$$\frac{D \cdot (N_s - \bar{N}) \cdot A_\text{zone}}{r_\text{zone}}
= \mathrm{SA}_\text{active} \cdot u_{A\max}
  \cdot \frac{\bar{N}}{\bar{N} + k_{15}}$$

Cross-multiplying, expanding, and collecting terms gives a quadratic in
$\bar{N}$. Defining $\alpha = u_{A\max} \cdot \mathrm{SA}_\text{active}
\cdot r_\text{zone} / (D \cdot A_\text{zone})$:

$$\bar{N}^2 + (k_{15} + \alpha - N_s)\,\bar{N} - k_{15}\,N_s = 0$$

with positive root:

$$\bar{N} = \frac{1}{2}\left[(N_s - k_{15} - \alpha)
  + \sqrt{(N_s - k_{15} - \alpha)^2 + 4\,k_{15}\,N_s}\right]$$

Uptake by each organ uses $\bar{N}$ and its own active surface:

$$U_{fr} = (1-m)\,\mathrm{SA}_{fr} \cdot u_{A\max}
           \cdot \frac{\bar{N}}{\bar{N} + k_{15}}$$

$$U_m = \mathrm{SA}_m \cdot u_{A\max}
        \cdot \frac{\bar{N}}{\bar{N} + k_{15}}$$

---

## Parameters

```{r parameters}
###
### SOIL PARAMETERS
###

u_Amax    <- 0.088   # Area-based max uptake (kg N m^-2 year^-1)
D_default <- 0.01    # Effective soil diffusion coefficient (m^2 year^-1)
k_15      <- 0.05    # Biochemical half-saturation constant (kg N m^-3)
depth     <- 0.1     # Rooting depth (m)

###
### ROOT PARAMETERS
###

myco_diameter <- 1e-5   # Hyphal diameter (m)
myco_density  <- 330    # Hyphal tissue density (kg m^-3)
l_hyp         <- 1e-4   # Hyphal segment length (m)
T_m           <- 2      # Mycorrhizal turnover rate (year^-1)
mycorrhized   <- 0.9    # Fraction of roots colonised by ECM

n <- 3e5
l <- 1.5

C_r <- 2

###
### EXTRA PARAMETERS
###

k_9  <- 0.01
k_8  <- 2.7
k_13 <- 2.5
k_14 <- 0.44

traits <- list(
  k_0 = 0.5, k_1 = 0.35,       k_2 = 0.2,           k_3 = 0.014,
  k_4 = 11632.91276, k_5 = -334.3626755, k_6 = 3,    k_7 = 3,
  k_8 = k_8,  k_9 = k_9
)

crown_area_fixed <- pi * C_r^2
B_cr_fixed       <- 0
rho_cr_fixed     <- 500
```

---

## Equations

```{r equations}
############################################################
## ROOT ARCHITECTURE FUNCTIONS
############################################################

root_diameter <- function(root_length, traits) {
  traits$k_1 / (root_length ^ traits$k_0)
}

root_cortex_proportion <- function(root_length, traits) {
  d <- root_diameter(root_length, traits)
  (traits$k_2 + traits$k_3 / d)^2
}

root_density <- function(root_length, traits) {
  PRD <- root_cortex_proportion(root_length, traits)
  traits$k_4 * PRD + traits$k_5
}

root_age <- function(root_length, traits) {
  d <- root_diameter(root_length, traits)
  traits$k_6 * (1 - exp(-traits$k_7 * d))
}

root_biomass_per_crown_area <- function(root_length, root_no, traits) {
  d_m   <- root_diameter(root_length, traits) / 1000
  rho   <- root_density(root_length, traits)
  l_tip <- root_length / 1000
  rho * (d_m / 2)^2 * l_tip * pi * root_no
}

root_mass <- function(root_length, root_no, crown_area, traits) {
  2 * root_biomass_per_crown_area(root_length, root_no, traits) * crown_area
}

############################################################
## SURFACE AREAS
############################################################

root_surface_area <- function(root_length, root_no, traits) {
  d_m   <- root_diameter(root_length, traits) / 1000
  l_tip <- root_length / 1000
  pi * d_m * l_tip * root_no
}

myco_surface_area <- function(B_m, myco_diameter, myco_density) {
  L_total <- B_m / (myco_density * pi * (myco_diameter / 2)^2)
  pi * myco_diameter * L_total
}

############################################################
## ROOT ZONE GEOMETRY
##
## Semi-ellipsoidal root zone of radius R = k_13 * C_r
## and depth q.
##   A_zone = 2*pi*R^2 + pi*R*q  (lateral + base)
##   r_zone = R
############################################################

zone_geometry <- function(C_r, q, k_13) {
  R      <- k_13 * C_r
  A_zone <- 2 * pi * R^2 + pi * R * q
  r_zone <- R
  list(A_zone = A_zone, r_zone = r_zone)
}

############################################################
## ACTIVE SURFACE AREA
##
## SA_active = (1 - m) * SA_fr + SA_m
##
## Colonised root surface (m * SA_fr) is excluded from
## direct uptake — nitrogen there is transferred via the
## fungal pool N_m rather than absorbed directly.
############################################################

active_surface_area <- function(SA_fr, SA_m, m) {
  (1 - m) * SA_fr + SA_m
}

############################################################
## MEAN SOIL CONCENTRATION N_bar
##
## From steady-state balance between diffusive supply
## and biochemical demand:
##
##   D*(N_s - N_bar)*A_zone/r_zone
##       = SA_active * u_Amax * N_bar/(N_bar + k_15)
##
## Cross-multiplying and collecting gives quadratic:
##   N_bar^2 + (k_15 + alpha - N_s)*N_bar - k_15*N_s = 0
##
## where alpha = u_Amax * SA_active * r_zone / (D * A_zone)
##
## Positive root:
##   b     = N_s - k_15 - alpha
##   N_bar = 0.5 * (b + sqrt(b^2 + 4*k_15*N_s))
############################################################

N_bar <- function(N_s, D, u_Amax, k_15,
                  SA_active, A_zone, r_zone) {
  if (SA_active <= 0) return(N_s)
  alpha <- u_Amax * SA_active * r_zone / (D * A_zone)
  b     <- N_s - k_15 - alpha
  disc  <- sqrt(b^2 + 4 * k_15 * N_s)
  if (b >= 0) 0.5 * (b + disc) else 2 * k_15 * N_s / (disc - b)
}

############################################################
## ORGAN UPTAKE
##
## U_fr = (1-m) * SA_fr * u_Amax * N_bar/(N_bar + k_15)
## U_m  =         SA_m  * u_Amax * N_bar/(N_bar + k_15)
##
## Both use N_bar. The (1-m) on U_fr is already embedded
## in SA_active used to compute N_bar, but also applied
## here to allocate the shared flux correctly between
## the two organ types.
############################################################

uptake_root <- function(SA_fr, m, u_Amax, k_15, Nbar) {
  (1 - m) * SA_fr * u_Amax * (Nbar / (Nbar + k_15))
}

uptake_myco <- function(SA_m, u_Amax, k_15, Nbar) {
  SA_m * u_Amax * (Nbar / (Nbar + k_15))
}

############################################################
## UPTAKE MODIFIERS
############################################################

# Arrhenius temperature response for u_max (root N transporter kinetics).
# Ea = 50 kJ/mol => Q10 ~ 2 in boreal range. Normalised to 1 at 25 degC.
# Hard-zero below 0 degC (frozen soil).
Ea_uptake <- 50000  # J/mol
R_gas     <- 8.3145 # J/mol/K
T_ref     <- 298.15 # K (25 degC)

ftemp_arrhenius_uptake <- function(tc) {
  if (tc <= 0) return(0)
  tk <- tc + 273.15
  exp(Ea_uptake * (tk - T_ref) / (T_ref * R_gas * tk))
}

uptake_age_modifier <- function(root_length, traits) {
  age <- root_age(root_length, traits)
  1 / (1 + exp(age - traits$k_8))
}

nitrogen_gate <- function(root_no, root_length, traits) {
  d  <- root_diameter(root_length, traits)
  SA <- root_no * pi * d * root_length
  SA / (SA + traits$k_9)
}

############################################################
## WHOLE-PLANT UPTAKE
##
## U_a = a_u * [U_fr + m * g(n,l) * T_m * max(0, N_m)]
##
## Note: (1-m) is already inside U_fr via uptake_root().
## The ECM transfer term handles nitrogen from colonised
## roots via the fungal pool N_m.
############################################################

plant_uptake <- function(N_s, N_m,
                         crown_area, root_length, root_no,
                         B_m, B_cr, rho_cr,
                         traits,
                         mycorrhized, T_m,
                         myco_diameter, myco_density, l_hyp,
                         C_r, q = 0.5, k_13 = 2, k_14 = 0.5,
                         u_Amax, D, k_15,
                         tc = 25) {

  B_fr  <- root_mass(root_length, root_no, crown_area, traits)
  SA_fr <- root_surface_area(root_length, root_no, traits) * crown_area
  SA_m  <- myco_surface_area(B_m, myco_diameter, myco_density)

  geom    <- zone_geometry(C_r, q, k_13)
  SA_act  <- active_surface_area(SA_fr, SA_m, mycorrhized)

  f_temp <- ftemp_arrhenius_uptake(tc)

  Nbar     <- N_bar(N_s, D, u_Amax * f_temp, k_15,
                    SA_act, geom$A_zone, geom$r_zone)
  F_supply <- D * N_s * geom$A_zone / geom$r_zone

  U_fr <- uptake_root(SA_fr, mycorrhized, u_Amax * f_temp, k_15, Nbar)
  U_m  <- uptake_myco(SA_m, u_Amax * f_temp, k_15, Nbar)

  a_u <- uptake_age_modifier(root_length, traits)
  g   <- nitrogen_gate(root_no, root_length, traits)

  ECM_transfer <- mycorrhized * g * T_m * max(0, N_m)
  U_a <- a_u * (U_fr + ECM_transfer)

  return(list(
    N_total      = U_a,
    U_fr         = U_fr,
    U_m          = U_m,
    SA_fr        = SA_fr,
    SA_m         = SA_m,
    SA_active    = SA_act,
    ECM_transfer = ECM_transfer,
    N_bar        = Nbar,
    F_supply     = F_supply,
    A_zone       = geom$A_zone,
    r_zone       = geom$r_zone,
    a_u          = a_u,
    g            = g,
    B_fr         = B_fr
  ))
}
```

---

## Diagnostic: $\bar{N}$ and Uptake Behaviour

```{r diagnostics, fig.height=8}
N_s_vals    <- seq(0.001, 2,  length.out = 200)
n_vals      <- 10^seq(3, 7,   length.out = 5)
D_vals_diag <- c(1e-5, 1e-4, 1e-3, 1e-2)

geom_fixed <- zone_geometry(C_r, depth, k_13)
D_fmt <- formatC(D_vals_diag, format = "e", digits = 2)
cb_D  <- setNames(c("#E69F00", "#56B4E9", "#009E73", "#CC79A7"), D_fmt)

# N_bar vs N_s by D
df_D <- expand_grid(N_s = N_s_vals, D = D_vals_diag) %>%
  rowwise() %>%
  mutate(
    SA_fr   = root_surface_area(l, n, traits) * crown_area_fixed,
    SA_m    = myco_surface_area(0, myco_diameter, myco_density),
    SA_act  = active_surface_area(SA_fr, SA_m, mycorrhized),
    Nbar    = N_bar(N_s, D, u_Amax, k_15,
                    SA_act, geom_fixed$A_zone, geom_fixed$r_zone),
    D_label = factor(formatC(D, format = "e", digits = 2), levels = D_fmt)
  ) %>%
  ungroup()

# N_bar and uptake vs tip density
df_diag <- expand_grid(N_s = N_s_vals, n_tip = n_vals) %>%
  rowwise() %>%
  mutate(
    B_fr   = root_mass(l, n_tip, crown_area_fixed, traits),
    SA_fr  = root_surface_area(l, n_tip, traits) * crown_area_fixed,
    SA_act = active_surface_area(SA_fr, 0, mycorrhized),
    Nbar   = N_bar(N_s, D_default, u_Amax, k_15,
                   SA_act, geom_fixed$A_zone, geom_fixed$r_zone),
    U_fr   = uptake_root(SA_fr, mycorrhized, u_Amax, k_15, Nbar),
    U_per_B = U_fr / B_fr,
    F_sup  = D_default * N_s * geom_fixed$A_zone / geom_fixed$r_zone
  ) %>%
  ungroup() %>%
  mutate(n_label = factor(signif(n_tip, 2)))

nbar_bounds <- data.frame(
  xintercept = c(0.0012, 0.003),
  label      = c("Korhonen (mineral)", "With amino acids")
)

p_Nbar_D <- ggplot(df_D, aes(N_s, Nbar, colour = D_label)) +
  geom_line(linewidth = 1.1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  scale_linetype_manual(name = "N_s reference",
                        values = c("Korhonen (mineral)" = "dashed",
                                   "With amino acids"   = "dotdash")) +
  scale_colour_manual(values = cb_D,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  labs(title = expression(bar(N)~"vs"~N[s]~"by D"),
       subtitle = bquote(n == .(signif(n, 2)) ~ "tips m"^{-2} * ", " ~ l == .(l) ~ "mm"),
       x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  theme_bw(base_size = 14)

p_Nbar_n <- ggplot(df_diag, aes(N_s, Nbar, colour = n_label)) +
  geom_line(linewidth = 1.1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  scale_linetype_manual(name = "N_s reference",
                        values = c("Korhonen (mineral)" = "dashed",
                                   "With amino acids"   = "dotdash")) +
  scale_colour_viridis_d(name = "n (tips/m\u00b2)") +
  labs(title = expression(bar(N)~"vs"~N[s]~"by tip density"),
       x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  theme_bw(base_size = 14)

p_U_Ns <- ggplot(df_diag, aes(N_s, U_fr, colour = n_label)) +
  geom_line(linewidth = 1.1) +
  geom_line(aes(y = F_sup), colour = "black",
            linetype = "dashed", linewidth = 0.8) +
  scale_colour_viridis_d(name = "n (tips/m\u00b2)") +
  scale_y_log10(labels = scales::scientific) +
  labs(title = "Total root uptake vs N_s",
       x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) +
  theme_bw(base_size = 14)

# Continuous B_fr sweep: max uptake (at highest N_s) as a function of biomass
n_cont   <- 10^seq(3, 7, length.out = 300)
N_s_max  <- max(N_s_vals)

df_diag_cont <- tibble(n_tip = n_cont) %>%
  rowwise() %>%
  mutate(
    B_fr    = root_mass(l, n_tip, crown_area_fixed, traits),
    SA_fr   = root_surface_area(l, n_tip, traits) * crown_area_fixed,
    SA_act  = active_surface_area(SA_fr, 0, mycorrhized),
    Nbar    = N_bar(N_s_max, D_default, u_Amax, k_15,
                    SA_act, geom_fixed$A_zone, geom_fixed$r_zone),
    U_fr    = uptake_root(SA_fr, mycorrhized, u_Amax, k_15, Nbar),
    U_per_B = U_fr / B_fr
  ) %>%
  ungroup()

p_U_B <- ggplot(df_diag_cont, aes(B_fr, U_per_B)) +
  geom_line(linewidth = 1.1, colour = "steelblue") +
  scale_x_log10(labels = scales::scientific) +
  scale_y_log10(labels = scales::scientific) +
  labs(title = bquote("Max"~U[fr]/B[fr]~"vs"~B[fr]~"("*N[s]==.(N_s_max)*")"),
       x = expression(B[fr]~"(kg)"),
       y = expression(max~U[fr]~B[fr]^{-1}~"(kg N kg"^{-1}~"yr"^{-1}*")")) +
  theme_bw(base_size = 14)

fig_diagnostics <- (p_Nbar_D | p_Nbar_n) / (p_U_Ns | p_U_B) +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") &
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_diagnostics
ggsave(here::here("manuscript/oup-authoring-template/Figures/01_diagnostics.png"), plot = fig_diagnostics, width = 12, height = 8)
```

---

## Root and Mycorrhizal Uptake

```{r uptake_plots, fig.height=10}
# =====================================================================
# PARAMETERS
# =====================================================================
D_vals        <- c(1e-5, 1e-4, 1e-3, 1e-2)
D_levels      <- formatC(D_vals, format = "e", digits = 2)
u_max_fixed   <- 0.88 / 10
N_s_vals      <- seq(0.001, 1.0, length.out = 200)
N_s_highlight <- c(0.2, 0.6)
B_m_highlight <- c(0.5, 1.0)
n_a_vals      <- c(800, 1e7)          # tip densities for a/c panels
B_m_sweep     <- 10^seq(-3, log10(2), length.out = 60)

cb_four <- setNames(c("#E69F00", "#56B4E9", "#009E73", "#CC79A7"), D_levels)

D_labels_plot <- setNames(
  list(expression(10^{-5}), expression(10^{-4}),
       expression(10^{-3}), expression(10^{-2})),
  D_levels
)

n_key_labels <- formatC(n_a_vals, format = "e", digits = 0)  # "8e+02", "1e+07"

df_basis <- tibble(n = 10^seq(1, 10, length.out = 60), l = 1.5) %>%
  rowwise() %>%
  mutate(
    B_fr  = root_mass(l, n, crown_area_fixed, traits),
    SA_fr = root_surface_area(l, n, traits) * crown_area_fixed
  ) %>%
  ungroup()

# SA_fr at the two specific n values used in a/c panels
SA_at_n <- sapply(n_a_vals, function(nv)
  root_surface_area(l, nv, traits) * crown_area_fixed)

# SA reference (low n) used as the "fixed root" in c/d panels
SA_highlight_ref <- SA_at_n[1]

theme_pub <- theme_bw(base_size = 18) +
  theme(panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text        = element_text(size = 16),
        legend.position   = "right",
        legend.text       = element_text(size = 15),
        legend.title      = element_text(size = 16),
        legend.key.size   = unit(0.55, "cm"))

get_Nbar <- function(N_s, D, SA_fr, B_m, C_r, q, k_13, m) {
  SA_m   <- myco_surface_area(B_m, myco_diameter, myco_density)
  SA_act <- active_surface_area(SA_fr, SA_m, m)
  geom   <- zone_geometry(C_r, q, k_13)
  N_bar(N_s, D, u_max_fixed, k_15, SA_act, geom$A_zone, geom$r_zone)
}

# =====================================================================
# LIMIT RIBBON DATA
# =====================================================================
geom_ref <- zone_geometry(C_r, depth, k_13)
D_max    <- max(D_vals)

# b panels: (b.i) roots only; (b.ii) roots + ECM at B_m_highlight[2]
N_s_ref_b  <- N_s_highlight[2]
SA_m_b_ref <- myco_surface_area(B_m_highlight[2], myco_diameter, myco_density)
df_lim_b <- df_basis %>%
  select(B_fr, SA_fr) %>%
  crossing(
    tibble(
      myco_case = c("no myco", "myco"),
      panel     = factor(c("(b.i) Without ECM", "(b.ii) With ECM"),
                         levels = c("(b.i) Without ECM", "(b.ii) With ECM"))
    )
  ) %>%
  mutate(
    SA_m_use = ifelse(myco_case == "myco", SA_m_b_ref, 0),
    m_use    = ifelse(myco_case == "myco", mycorrhized, 0),
    root_eff = (1 - m_use) * SA_fr,
    SA_total = root_eff + SA_m_use,
    U_bio    = SA_fr * u_max_fixed * N_s_ref_b / (N_s_ref_b + k_15),
    F_diff   = D_max * N_s_ref_b * geom_ref$A_zone / geom_ref$r_zone * root_eff / SA_total
  )

# d panels: (d.i) ECM only; (d.ii) ECM + roots at SA_highlight_ref
N_s_ref_d <- N_s_highlight[2]
df_lim_d <- crossing(
  tibble(B_m = B_m_sweep),
  tibble(
    root_case = c("no root", "root"),
    panel     = factor(c("(d.i) Without roots", "(d.ii) With roots"),
                       levels = c("(d.i) Without roots", "(d.ii) With roots"))
  )
) %>%
  rowwise() %>%
  mutate(
    SA_m_val  = myco_surface_area(B_m, myco_diameter, myco_density),
    SA_fr_use = ifelse(root_case == "root", SA_highlight_ref, 0),
    m_use     = ifelse(root_case == "root", mycorrhized, 0),
    root_eff  = (1 - m_use) * SA_fr_use,
    SA_total  = root_eff + SA_m_val,
    U_bio     = SA_m_val * u_max_fixed * N_s_ref_d / (N_s_ref_d + k_15),
    F_diff    = D_max * N_s_ref_d * geom_ref$A_zone / geom_ref$r_zone * SA_m_val / SA_total
  ) %>%
  ungroup()

# =====================================================================
# DATA: U_fr and U_m vs N_s  (a and c panels)
# =====================================================================

# ---- (a) U_fr vs N_s: two n values, faceted by ECM presence ----
df_root_Ns <- crossing(
    tibble(n_val = n_a_vals, SA_fr = SA_at_n),
    myco = c("no myco", "myco"),
    N_s  = N_s_vals,
    D    = D_vals) %>%
  rowwise() %>%
  mutate(
    B_m_use = ifelse(myco == "myco", B_m_highlight[2], 0),
    m_use   = ifelse(myco == "myco", mycorrhized, 0),
    Nbar    = get_Nbar(N_s, D, SA_fr, B_m_use, C_r, depth, k_13, m_use),
    U_fr    = uptake_root(SA_fr, m_use, u_max_fixed, k_15, Nbar),
    D_label = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    n_label = factor(formatC(n_val, format = "e", digits = 0),
                     levels = n_key_labels)
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(myco == "no myco",
                               "(a.i) Without ECM", "(a.ii) With ECM"),
                        levels = c("(a.i) Without ECM", "(a.ii) With ECM")))

# ---- (c) U_m vs N_s: two B_m values, faceted by root presence ----
df_myco_Ns <- crossing(
    tibble(B_m = B_m_highlight),
    root = c("no root", "root"),
    N_s  = N_s_vals,
    D    = D_vals) %>%
  rowwise() %>%
  mutate(
    SA_m      = myco_surface_area(B_m, myco_diameter, myco_density),
    SA_fr_use = ifelse(root == "root", SA_highlight_ref, 0),
    m_use     = ifelse(root == "root", mycorrhized, 0),
    Nbar      = get_Nbar(N_s, D, SA_fr_use, B_m, C_r, depth, k_13, m_use),
    U_m       = uptake_myco(SA_m, u_max_fixed, k_15, Nbar),
    D_label   = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    Bm_label  = factor(as.character(B_m), levels = as.character(B_m_highlight))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(root == "no root",
                               "(c.i) Without roots", "(c.ii) With roots"),
                        levels = c("(c.i) Without roots", "(c.ii) With roots")))

# =====================================================================
# DATA: U_fr and U_m vs biomass  (b and d panels)
# =====================================================================

# ---- (b) U_fr vs B_fr: N_s as linewidth, faceted by ECM ----
df_root_B <- df_basis %>%
  crossing(myco = c("no myco", "myco"), N_s = N_s_highlight, D = D_vals) %>%
  rowwise() %>%
  mutate(
    B_m_use = ifelse(myco == "myco", B_m_highlight[2], 0),
    m_use   = ifelse(myco == "myco", mycorrhized, 0),
    Nbar    = get_Nbar(N_s, D, SA_fr, B_m_use, C_r, depth, k_13, m_use),
    U_fr    = uptake_root(SA_fr, m_use, u_max_fixed, k_15, Nbar),
    D_label = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    N_label = factor(as.character(N_s), levels = as.character(N_s_highlight))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(myco == "no myco",
                               "(b.i) Without ECM", "(b.ii) With ECM"),
                        levels = c("(b.i) Without ECM", "(b.ii) With ECM")))

# ---- (d) U_m vs B_m: N_s as linetype, faceted by root presence ----
df_myco_B <- crossing(
    tibble(B_m = B_m_sweep),
    root = c("no root", "root"),
    N_s  = N_s_highlight,
    D    = D_vals) %>%
  rowwise() %>%
  mutate(
    SA_fr_use = ifelse(root == "root", SA_highlight_ref, 0),
    m_use     = ifelse(root == "root", mycorrhized, 0),
    SA_m      = myco_surface_area(B_m, myco_diameter, myco_density),
    Nbar      = get_Nbar(N_s, D, SA_fr_use, B_m, C_r, depth, k_13, m_use),
    U_m       = uptake_myco(SA_m, u_max_fixed, k_15, Nbar),
    D_label   = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    N_label   = factor(as.character(N_s), levels = as.character(N_s_highlight))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(root == "no root",
                               "(d.i) Without roots", "(d.ii) With roots"),
                        levels = c("(d.i) Without roots", "(d.ii) With roots")))
```

---

```{r uptake_total, fig.height=10}
fmt_log   <- scales::trans_format("log10", scales::math_format(10^.x))
x_lim_B   <- c(1e-4, 2)           # shared x-limit for b and d biomass panels

Bm_key_labels <- as.character(B_m_highlight)

# Filtered limit frames clipped to the biomass x-range
df_lim_b_clip <- filter(df_lim_b, B_fr >= x_lim_B[1], B_fr <= x_lim_B[2])

# a panels: (a.i) roots only; (a.ii) roots + ECM at B_m_highlight[2]
SA_m_high <- myco_surface_area(B_m_highlight[2], myco_diameter, myco_density)
df_lim_a <- crossing(
  tibble(N_s = N_s_vals),
  tibble(
    myco_case = c("no myco", "myco"),
    panel     = factor(c("(a.i) Without ECM", "(a.ii) With ECM"),
                       levels = c("(a.i) Without ECM", "(a.ii) With ECM"))
  )
) %>%
  mutate(
    SA_m_use = ifelse(myco_case == "myco", SA_m_high, 0),
    m_use    = ifelse(myco_case == "myco", mycorrhized, 0),
    root_eff = (1 - m_use) * SA_at_n[2],
    SA_total = root_eff + SA_m_use,
    U_bio    = SA_at_n[2] * u_max_fixed * N_s / (N_s + k_15),
    F_diff   = D_max * N_s * geom_ref$A_zone / geom_ref$r_zone * root_eff / SA_total
  )

# c panels: (c.i) ECM only; (c.ii) ECM + roots at SA_highlight_ref
df_lim_c <- crossing(
  tibble(N_s = N_s_vals),
  tibble(
    root_case = c("no root", "root"),
    panel     = factor(c("(c.i) Without roots", "(c.ii) With roots"),
                       levels = c("(c.i) Without roots", "(c.ii) With roots"))
  )
) %>%
  mutate(
    SA_fr_use = ifelse(root_case == "root", SA_highlight_ref, 0),
    m_use     = ifelse(root_case == "root", mycorrhized, 0),
    root_eff  = (1 - m_use) * SA_fr_use,
    SA_total  = root_eff + SA_m_high,
    U_bio     = SA_m_high * u_max_fixed * N_s / (N_s + k_15),
    F_diff    = D_max * N_s * geom_ref$A_zone / geom_ref$r_zone * SA_m_high / SA_total
  )

# Datasets for limits figure: orange (D=10^-5) + purple (D=10^-2) only
D_edge_vals   <- c(D_vals[1], D_vals[4])
D_edge_labels <- D_levels[c(1, 4)]
cb_edge       <- cb_four[D_edge_labels]

df_root_Ns_hi <- filter(df_root_Ns, D %in% D_edge_vals)
df_myco_Ns_hi <- filter(df_myco_Ns, D %in% D_edge_vals)
df_root_B_hi  <- filter(df_root_B,  D %in% D_edge_vals)
df_myco_B_hi  <- filter(df_myco_B,  D %in% D_edge_vals)

# Larger text theme with diagonal x-axis labels (used for limits figure)
theme_pub_lg <- theme_pub + theme(
  text           = element_text(size = 22),
  axis.title     = element_text(size = 22),
  axis.text      = element_text(size = 20),
  axis.text.x    = element_text(size = 20, angle = 45, hjust = 1),
  strip.text     = element_text(size = 14),
  legend.text    = element_text(size = 18),
  legend.title   = element_text(size = 20),
  legend.key.size = unit(0.7, "cm")
)

# ==================================================================
# SHARED SCALE OBJECTS
# N_s uses LINETYPE in BOTH b and d panels so legends merge
# ==================================================================
sc_col_all  <- scale_colour_manual(values = cb_four, labels = D_labels_plot,
                                   name = expression(D~"(m"^2~"yr"^{-1}*")"))
sc_col_edge <- scale_colour_manual(
  values = setNames(c("white", "black"), D_edge_labels),
  labels = D_labels_plot[D_edge_labels],
  name   = expression(D~"(m"^2~"yr"^{-1}*")")
)
sc_lt_n     <- scale_linetype_manual(
                 values = setNames(c(3, 1), n_key_labels),
                 labels = setNames(n_key_labels, n_key_labels),
                 name   = expression(n~"(tips m"^{-2}*")"))
sc_lt_Bm    <- scale_linetype_manual(
                 values = setNames(c(3, 1), Bm_key_labels),
                 labels = setNames(paste(Bm_key_labels, "kg"), Bm_key_labels),
                 name   = expression(B[m]~"(kg)"))
sc_lt_Ns    <- scale_linetype_manual(          # same for b AND d panels
                 values = c("0.2" = 2, "0.6" = 1),
                 labels = c("0.2" = expression(N[s] == 0.2),
                            "0.6" = expression(N[s] == 0.6)),
                 name   = expression(N[s]~"(kg N m"^{-3}*")"))

# Ribbon style: fill upward from the bottom so uptake lines ride the limit ceiling
# Biochemical limit (light) fills higher; diffusion limit (darker) fills lower when
# diffusion constrains — uptake lines sit at the top edge of the binding constraint.
y_min_rib <- 1e-15

# High-ref ribbons: grey fills
make_ribbons <- function(df, x_col) {
  list(
    geom_ribbon(data = df,
                aes(x = .data[[x_col]], ymin = y_min_rib, ymax = U_bio,
                    fill = "Biochemical limit"),
                alpha = 0.55, inherit.aes = FALSE),
    geom_ribbon(data = df,
                aes(x = .data[[x_col]], ymin = y_min_rib, ymax = F_diff,
                    fill = "Diffusion limit"),
                alpha = 0.55, inherit.aes = FALSE)
  )
}

# Low-ref ribbons: purple fills (lighter alpha for bio, darker for diff)
make_ribbons_lo <- function(df, x_col) {
  list(
    geom_ribbon(data = df,
                aes(x = .data[[x_col]], ymin = y_min_rib, ymax = U_bio,
                    fill = "Biochemical limit (low ref.)"),
                alpha = 0.35, inherit.aes = FALSE),
    geom_ribbon(data = df,
                aes(x = .data[[x_col]], ymin = y_min_rib, ymax = F_diff,
                    fill = "Diffusion limit (low ref.)"),
                alpha = 0.55, inherit.aes = FALSE)
  )
}

sc_fill_combined <- scale_fill_manual(
  values = c(
    "Biochemical limit"            = "grey82",
    "Diffusion limit"              = "grey58",
    "Biochemical limit (low ref.)" = "#0072B2",
    "Diffusion limit (low ref.)"   = "#0072B2"
  ),
  name = expression("Limits (D = 10"^{-2}*")")
)

ribbon_layers_a <- make_ribbons(df_lim_a, "N_s")
ribbon_layers_b <- make_ribbons(df_lim_b_clip, "B_fr")
ribbon_layers_c <- make_ribbons(df_lim_c, "N_s")
ribbon_layers_d <- make_ribbons(df_lim_d, "B_m")

# ==================================================================
# FIGURE 1: All four D values, NO limit ribbons
# N_s uses linetype in both b and d panels → single merged legend entry
# ==================================================================
p_a1 <- ggplot(df_root_Ns,
               aes(x = N_s, y = U_fr, colour = D_label, linetype = n_label)) +
  geom_line(linewidth = 0.9) + facet_wrap(~ panel) +
  sc_col_all + sc_lt_n +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) + theme_pub_lg

p_b1 <- ggplot(df_root_B,
               aes(x = B_fr, y = U_fr, colour = D_label, linetype = N_label)) +
  geom_line(linewidth = 0.9) + facet_wrap(~ panel) +
  sc_col_all + sc_lt_Ns +
  scale_x_log10(labels = fmt_log, limits = x_lim_B) +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(B[fr]~"(kg)"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) + theme_pub_lg

p_c1 <- ggplot(df_myco_Ns,
               aes(x = N_s, y = U_m, colour = D_label, linetype = Bm_label)) +
  geom_line(linewidth = 0.9) + facet_wrap(~ panel) +
  sc_col_all + sc_lt_Bm +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[m]~"(kg N yr"^{-1}*")")) + theme_pub_lg

p_d1 <- ggplot(df_myco_B,
               aes(x = B_m, y = U_m, colour = D_label, linetype = N_label)) +
  geom_line(linewidth = 0.9) + facet_wrap(~ panel) +
  sc_col_all + sc_lt_Ns +
  scale_x_log10(labels = fmt_log) +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(B[m]~"(kg)"),
       y = expression(U[m]~"(kg N yr"^{-1}*")")) + theme_pub_lg

fig_all_lines <- (p_a1 | p_b1) / (p_c1 | p_d1) +
  plot_layout(guides = "collect") & theme(legend.position = "right")
fig_all_lines
ggsave("vignettes/plots/03_all_lines.png",
       plot = fig_all_lines, width = 16, height = 10)

# ==================================================================
# FIGURE 2: Orange + purple, WITH limit ribbons
#   Grey  = high reference (high n / high B_m / N_s = 0.6)
#   Purple = low reference (low n / low B_m / N_s = 0.2), panel-specific
# ==================================================================

# ==================================================================
# FIGURE 3: Lower-purple limits — D = 10^-2, lower reference:
#   LHS: low n (800) / low B_m (0.5 kg)
#   RHS: low N_s = 0.2
#   Biochemical limit (U_bio): root/ECM SA × u_max × N_s/(N_s+k15)
#     — same in .i and .ii (property of the absorber alone)
#   Diffusion limit (F_diff): total flux × absorber_share
#     — lower in .ii because partner organ competes for the same flux
# ==================================================================
SA_ref_lo_a <- SA_at_n[1]   # n = 800
SA_m_ref_lo <- myco_surface_area(B_m_highlight[1], myco_diameter, myco_density)  # B_m = 0.5
SA_m_hi     <- myco_surface_area(B_m_highlight[2], myco_diameter, myco_density)  # B_m = 1.0 (b.ii)
N_s_ref_lo  <- N_s_highlight[1]  # N_s = 0.2

# a panels: absorber = roots; (a.ii) adds ECM so root gets reduced share of flux
df_lim_a_lo <- crossing(
  tibble(N_s = N_s_vals),
  tibble(
    myco_case = c("no myco", "myco"),
    panel     = factor(c("(a.i) Without ECM", "(a.ii) With ECM"),
                       levels = c("(a.i) Without ECM", "(a.ii) With ECM"))
  )
) %>%
  mutate(
    SA_m_use  = ifelse(myco_case == "myco", SA_m_ref_lo, 0),
    m_use     = ifelse(myco_case == "myco", mycorrhized, 0),
    root_eff  = (1 - m_use) * SA_ref_lo_a,
    SA_total  = root_eff + SA_m_use,
    U_bio     = SA_ref_lo_a * u_max_fixed * N_s / (N_s + k_15),
    F_diff    = D_max * N_s * geom_ref$A_zone / geom_ref$r_zone * root_eff / SA_total
  )

# b panels: absorber = roots sweeping SA_fr; (b.ii) adds ECM at B_m_highlight[2]; low N_s
df_lim_b_lo_clip <- df_lim_b_clip %>%
  select(B_fr, SA_fr) %>%
  crossing(
    tibble(
      myco_case = c("no myco", "myco"),
      panel     = factor(c("(b.i) Without ECM", "(b.ii) With ECM"),
                         levels = c("(b.i) Without ECM", "(b.ii) With ECM"))
    )
  ) %>%
  mutate(
    SA_m_use = ifelse(myco_case == "myco", SA_m_hi, 0),
    m_use    = ifelse(myco_case == "myco", mycorrhized, 0),
    root_eff = (1 - m_use) * SA_fr,
    SA_total = root_eff + SA_m_use,
    U_bio    = SA_fr * u_max_fixed * N_s_ref_lo / (N_s_ref_lo + k_15),
    F_diff   = D_max * N_s_ref_lo * geom_ref$A_zone / geom_ref$r_zone * root_eff / SA_total
  )

# c panels: absorber = ECM at low B_m; (c.ii) adds roots so ECM gets reduced share of flux
df_lim_c_lo <- crossing(
  tibble(N_s = N_s_vals),
  tibble(
    root_case = c("no root", "root"),
    panel     = factor(c("(c.i) Without roots", "(c.ii) With roots"),
                       levels = c("(c.i) Without roots", "(c.ii) With roots"))
  )
) %>%
  mutate(
    SA_fr_use = ifelse(root_case == "root", SA_highlight_ref, 0),
    m_use     = ifelse(root_case == "root", mycorrhized, 0),
    root_eff  = (1 - m_use) * SA_fr_use,
    SA_total  = root_eff + SA_m_ref_lo,
    U_bio     = SA_m_ref_lo * u_max_fixed * N_s / (N_s + k_15),
    F_diff    = D_max * N_s * geom_ref$A_zone / geom_ref$r_zone * SA_m_ref_lo / SA_total
  )

# d panels: absorber = ECM sweeping B_m; (d.ii) adds roots at SA_highlight_ref; low N_s
df_lim_d_lo <- crossing(
  tibble(B_m = B_m_sweep),
  tibble(
    root_case = c("no root", "root"),
    panel     = factor(c("(d.i) Without roots", "(d.ii) With roots"),
                       levels = c("(d.i) Without roots", "(d.ii) With roots"))
  )
) %>%
  rowwise() %>%
  mutate(
    SA_m_val  = myco_surface_area(B_m, myco_diameter, myco_density),
    SA_fr_use = ifelse(root_case == "root", SA_highlight_ref, 0),
    m_use     = ifelse(root_case == "root", mycorrhized, 0),
    root_eff  = (1 - m_use) * SA_fr_use,
    SA_total  = root_eff + SA_m_val,
    U_bio     = SA_m_val * u_max_fixed * N_s_ref_lo / (N_s_ref_lo + k_15),
    F_diff    = D_max * N_s_ref_lo * geom_ref$A_zone / geom_ref$r_zone * SA_m_val / SA_total
  ) %>%
  ungroup()

ribbon_layers_a_lo <- make_ribbons_lo(df_lim_a_lo,     "N_s")
ribbon_layers_b_lo <- make_ribbons_lo(df_lim_b_lo_clip, "B_fr")
ribbon_layers_c_lo <- make_ribbons_lo(df_lim_c_lo,     "N_s")
ribbon_layers_d_lo <- make_ribbons_lo(df_lim_d_lo,     "B_m")

p_a2 <- ggplot(df_root_Ns_hi,
               aes(x = N_s, y = U_fr, colour = D_label, linetype = n_label)) +
  ribbon_layers_a + ribbon_layers_a_lo +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  sc_col_edge + sc_lt_n + sc_fill_combined +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) + theme_pub_lg

p_b2 <- ggplot(df_root_B_hi,
               aes(x = B_fr, y = U_fr, colour = D_label, linetype = N_label)) +
  ribbon_layers_b + ribbon_layers_b_lo +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  sc_col_edge + sc_lt_Ns + sc_fill_combined +
  scale_x_log10(labels = fmt_log, limits = x_lim_B) +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(B[fr]~"(kg)"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) + theme_pub_lg

p_c2 <- ggplot(df_myco_Ns_hi,
               aes(x = N_s, y = U_m, colour = D_label, linetype = Bm_label)) +
  ribbon_layers_c + ribbon_layers_c_lo +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  sc_col_edge + sc_lt_Bm + sc_fill_combined +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[m]~"(kg N yr"^{-1}*")")) + theme_pub_lg

p_d2 <- ggplot(df_myco_B_hi,
               aes(x = B_m, y = U_m, colour = D_label, linetype = N_label)) +
  ribbon_layers_d + ribbon_layers_d_lo +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  sc_col_edge + sc_lt_Ns + sc_fill_combined +
  scale_x_log10(labels = fmt_log) +
  scale_y_log10(labels = fmt_log) +
  labs(x = expression(B[m]~"(kg)"),
       y = expression(U[m]~"(kg N yr"^{-1}*")")) + theme_pub_lg

fig_limits <- (p_a2 | p_b2) / (p_c2 | p_d2) +
  plot_layout(guides = "collect") & theme(legend.position = "right")
fig_limits
ggsave(here::here("manuscript/oup-authoring-template/Figures/03_limits.png"),
       plot = fig_limits, width = 16, height = 10)
```

## Sensitivity Analysis

```{r sensitivity, fig.height=8}
N_s_fixed <- 1.0
N_m_fixed <- 0.1
B_m_fixed <- 0.5

run_uptake <- function(depth_val, k15_val, k13_val, k14_val) {
  plant_uptake(
    N_s = N_s_fixed, N_m = N_m_fixed,
    crown_area = crown_area_fixed,
    root_length = l, root_no = n,
    B_m = B_m_fixed, B_cr = B_cr_fixed, rho_cr = rho_cr_fixed,
    traits = traits, mycorrhized = mycorrhized, T_m = T_m,
    myco_diameter = myco_diameter, myco_density = myco_density, l_hyp = l_hyp,
    C_r = C_r, q = depth_val, k_13 = k13_val, k_14 = k14_val,
    u_Amax = u_Amax, D = D_default, k_15 = k15_val
  )$N_total
}

baseline_uptake <- run_uptake(depth, k_15, k_13, k_14)

n_pts <- 60
param_ranges <- list(
  depth = seq(0.05, 2.0,  length.out = n_pts),
  k_15  = seq(0.05, 3.0,  length.out = n_pts),
  k_13  = seq(0.5,  20.0, length.out = n_pts),
  k_14  = seq(0.05, 0.95, length.out = n_pts)
)

df_sa <- bind_rows(
  tibble(param = "depth", value = param_ranges$depth,
         uptake = sapply(param_ranges$depth, \(v) run_uptake(v,     k_15, k_13, k_14))),
  tibble(param = "k_15",  value = param_ranges$k_15,
         uptake = sapply(param_ranges$k_15,  \(v) run_uptake(depth, v,    k_13, k_14))),
  tibble(param = "k_13",  value = param_ranges$k_13,
         uptake = sapply(param_ranges$k_13,  \(v) run_uptake(depth, k_15, v,    k_14))),
  tibble(param = "k_14",  value = param_ranges$k_14,
         uptake = sapply(param_ranges$k_14,  \(v) run_uptake(depth, k_15, k_13, v   )))
) %>%
  mutate(param = factor(param, levels = c("depth", "k_15", "k_13", "k_14")))

eps <- 0.01
si_df <- tibble(
  param       = c("depth", "k_15", "k_13", "k_14"),
  default_val = c(depth,   k_15,   k_13,   k_14),
  SI = c(
    (run_uptake(depth*(1+eps), k_15,          k_13,          k_14        ) -
     run_uptake(depth*(1-eps), k_15,          k_13,          k_14        )) / (2*eps*baseline_uptake),
    (run_uptake(depth,          k_15*(1+eps),  k_13,          k_14        ) -
     run_uptake(depth,          k_15*(1-eps),  k_13,          k_14        )) / (2*eps*baseline_uptake),
    (run_uptake(depth,          k_15,          k_13*(1+eps),  k_14        ) -
     run_uptake(depth,          k_15,          k_13*(1-eps),  k_14        )) / (2*eps*baseline_uptake),
    (run_uptake(depth,          k_15,          k_13,          k_14*(1+eps)) -
     run_uptake(depth,          k_15,          k_13,          k_14*(1-eps))) / (2*eps*baseline_uptake)
  )
)

cb_sa <- c(depth = "#E69F00", k_15 = "#56B4E9", k_13 = "#009E73", k_14 = "#CC79A7")

df_spider <- df_sa %>%
  left_join(tibble(param = c("depth","k_15","k_13","k_14"),
                   default_val = c(depth, k_15, k_13, k_14)), by = "param") %>%
  mutate(rel_param  = value / default_val,
         rel_uptake = uptake / baseline_uptake)

p_spider <- ggplot(df_spider, aes(rel_param, rel_uptake, colour = param)) +
  geom_line(linewidth = 1.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_colour_manual(values = cb_sa) +
  labs(title = "Sensitivity spider plot",
       x = "Relative parameter value (\u00d7 baseline)",
       y = "Relative uptake (\u00d7 baseline)") + theme_bw()

p_tornado <- ggplot(si_df, aes(x = param, y = SI, fill = param)) +
  geom_col(width = 0.6, show.legend = FALSE) + coord_flip() +
  scale_fill_manual(values = cb_sa) +
  labs(title = "Normalised sensitivity index", x = NULL, y = "SI") + theme_bw()

baseline_vlines <- tibble(param = c("depth","k_15","k_13","k_14"),
                           xint  = c(depth,   k_15,  k_13,  k_14))

p_sweeps <- ggplot(df_sa, aes(value, uptake, colour = param)) +
  geom_line(linewidth = 1.1, show.legend = FALSE) +
  geom_vline(data = baseline_vlines, aes(xintercept = xint), linetype = "dashed") +
  facet_wrap(~param, scales = "free_x") +
  scale_colour_manual(values = cb_sa) +
  labs(title = "Parameter sweeps", x = "Parameter value",
       y = "N uptake (kg N yr\u207b\u00b9)") + theme_bw()

fig_sensitivity <- p_sweeps / (p_spider | p_tornado)
fig_sensitivity
ggsave("vignettes/plots/04_sensitivity.png", plot = fig_sensitivity, width = 12, height = 8)
```

---

## Elasticity Analysis

```{r elasticity, fig.height=6}
eps       <- 1e-4
N_vals    <- seq(0.01, 2, length.out = 80)
N_m_fixed <- 0.1
B_m_fixed <- 0.5

params     <- c("depth", "k_15", "k_13", "k_14", "u_Amax")
components <- c("U_root", "U_myco")

run_U <- function(N, depth, k_15, k_13, k_14, u_Amax) {
  out <- plant_uptake(
    N_s = N, N_m = N_m_fixed, crown_area = crown_area_fixed,
    root_length = l, root_no = n,
    B_m = B_m_fixed, B_cr = B_cr_fixed, rho_cr = rho_cr_fixed,
    traits = traits, mycorrhized = mycorrhized, T_m = T_m,
    myco_diameter = myco_diameter, myco_density = myco_density, l_hyp = l_hyp,
    C_r = C_r, q = depth, k_13 = k_13, k_14 = k_14,
    u_Amax = u_Amax, D = D_default, k_15 = k_15
  )
  c(U_total = out$N_total, U_root = out$U_fr, U_myco = out$U_m)
}

elasticity <- function(N, param, component, depth, k_15, k_13, k_14, u_Amax) {
  component <- as.character(component)
  base <- run_U(N, depth, k_15, k_13, k_14, u_Amax)[component]
  pert <- switch(param,
    depth  = run_U(N, depth*(1+eps),  k_15,          k_13,          k_14,          u_Amax        )[component],
    k_15   = run_U(N, depth,           k_15*(1+eps),  k_13,          k_14,          u_Amax        )[component],
    k_13   = run_U(N, depth,           k_15,           k_13*(1+eps),  k_14,          u_Amax        )[component],
    k_14   = run_U(N, depth,           k_15,           k_13,          k_14*(1+eps),  u_Amax        )[component],
    u_Amax = run_U(N, depth,           k_15,           k_13,          k_14,          u_Amax*(1+eps))[component]
  )
  (pert - base) / (eps * base + 1e-12)
}

grid <- expand.grid(N = N_vals, param = params,
                    component = components, KEEP.OUT.ATTRS = FALSE)

df <- grid %>%
  rowwise() %>%
  mutate(SI = elasticity(N, param, component,
                         depth = depth, k_15 = k_15,
                         k_13 = k_13, k_14 = k_14, u_Amax = u_Amax)) %>%
  ungroup() %>%
  mutate(SI_abs    = abs(SI),
         SI_scaled = SI / max(abs(SI), na.rm = TRUE))

importance <- df %>%
  group_by(param, component) %>%
  summarise(mean_abs_SI = mean(abs(SI), na.rm = TRUE), .groups = "drop")

cb_el <- c(depth = "#E69F00", k_15 = "#56B4E9", k_13 = "#009E73",
           k_14 = "#CC79A7", u_Amax = "#D55E00")

param_labels <- c(
  depth  = "Depth",
  k_15   = expression(k[15]~"(half-saturation)"),
  k_13   = expression(k[13]~"(soil scaling)"),
  k_14   = expression(k[14]~"(porosity)"),
  u_Amax = expression(u[max]~"(uptake capacity)")
)

comp_labels <- c(U_root = "Root uptake", U_myco = "Mycorrhizal uptake")

fig_elasticity <- ggplot(df, aes(x = N, y = SI_scaled, colour = param)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey60") +
  geom_line(linewidth = 1) +
  facet_wrap(~ component, labeller = labeller(component = comp_labels)) +
  scale_colour_manual(values = cb_el, labels = param_labels) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(limits = c(-1, 1), oob = scales::squish) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = "Standardised elasticity", colour = "Parameter") +
  theme_bw(base_size = 12) +
  theme(legend.position = "right", strip.background = element_blank(),
        panel.grid.minor = element_blank())
fig_elasticity
ggsave("vignettes/plots/05_elasticity.png", plot = fig_elasticity, width = 12, height = 6)

fig_importance <- ggplot(importance, aes(param, mean_abs_SI, fill = param)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ component) +
  scale_fill_manual(values = cb_el) +
  theme_bw() +
  labs(x = NULL, y = "Mean absolute elasticity")
fig_importance
ggsave("vignettes/plots/06_elasticity_importance.png", plot = fig_importance, width = 10, height = 5)
```
