---
title: "Nitrogen Updates"
author: "YSSP"
date: "2023-05-27"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)
```
## Parameters

Carbon nitrogen_store_conversion was 0 for fertilised, 0.2 for unfertilised (Fransson 2024)
Original jmax = 86.02379

So set a_jmax as 800 to get the equilibrium point around 80 as an original.

Nitrogen leaf content upper limit is limited by the nitrogen store in the leaf!

```{r}
kphio = 0.087;        # quantum yield efficiency
ppfd = 300;          # umol/m2/s
ppfd_max = 1200;
vpd  = 1000;         # Pa
co2  = 400;          # ppm
elv  = 0;            # m.a.s.l.
fapar = 0.7;         # fraction
rdark = 0.015;
tc = 25;
vwind = 3;
netrad = ppfd/2
psi_soil = -0.05
pa = calc_patm(elv)
nitrogen = 1.5

### New parameters TODO: should work with 800!
a_jmax = 80

# biomass original value is 0.1
par_cost_original = list(alpha=0.1, gamma=1, nitrogen_store_conversion = 1)
par_cost = list(alpha=0.01, gamma=1, nitrogen_store_conversion = 1);
par_plant = list(conductivity=3e-17, psi50=-2, b=2);
options = list(gs_method = "GS_IGF", 
               et_method = "ET_DIFFUSION",
               ftemp_vj_method = "FV_kumarathunge19",
               ftemp_rd_method = "FR_heskel16",
               ftemp_br_method = "FB_atkin15",
               scale_alpha = F)
```

## Acclimating (Weekly) Response

```{r}
nitrogen_store_conversion = 1.5
dat_high = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 1.2
dat_medium = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 1
dat_medium_low = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 0.75
dat_low = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

dat_original = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_numerical(tc, tc, .x, netrad, vpd, co2, pa, fapar, kphio, psi_soil, rdark, vwind, par_plant, par_cost_original, options))) %>% unnest_wider(dat)
```


```{r}
df_high <- dat_high %>% melt("ppfd")
df_medium <- dat_medium %>% melt("ppfd")
df_medium_low <- dat_medium_low %>% melt("ppfd")
df_low <- dat_low %>% melt("ppfd")
df_original.m <- dat_original %>% melt("ppfd")

ggplot() + 
  geom_line(data = df_high, aes(y = value, x = ppfd, color = "a = 1.5")) +
  geom_line(data = df_medium, aes(y = value, x = ppfd, color = "a = 1.2")) +
  geom_line(data = df_medium_low, aes(y = value, x = ppfd, color = "a = 1")) +
  geom_line(data = df_low, aes(y = value, x = ppfd, color = "a = 0.75")) +
  geom_line(data = df_original.m, aes(y = value, x = ppfd, color = "Numeric"), linetype="dotted") +
  scale_color_manual(values = c("a = 1.5" = "red",
                                "a = 1.2" = "orange",
                                "a = 1" = "green",
                                "a = 0.75" = "blue",
                                "Numeric" = "black")) +
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_blank()) +
  facet_wrap(~variable, scales = "free")
```
## Formatted weekly response

```{r}
cols = c("ppfd", "a", "n_leaf", "gs", "jmax")
df_high <- dat_high[,cols] %>% melt("ppfd")
df_medium <- dat_medium[,cols] %>% melt("ppfd")
df_medium_low <- dat_medium_low[,cols] %>% melt("ppfd")
df_low <- dat_low[,cols] %>% melt("ppfd")
#df_original.m <- dat_original[,cols] %>% melt("ppfd")
df_original.m <- dat_original[,cols[-c(3,6)]] %>% melt("ppfd")

new_titles <- c(
  "ppfd" = "Photosynthetic Photon Flux Density (μmol m⁻² s⁻¹)",
  "a" = "Assimilation rate\n(A, μmol m⁻² s⁻¹)",
  "n_leaf" = "Nitrogen Content\nin Needle (g g-2)",
  "gs" = "Stomatal Conductance\n(gs, mol m⁻² s⁻¹)",
  "jmax" = "Jmax Parameter\n(Jmax, μmol m⁻² s⁻¹)"
)

ggplot() + 
  geom_line(data = df_high, aes(y = value, x = ppfd, color = "Root cost = 1.5")) +
  geom_line(data = df_medium, aes(y = value, x = ppfd, color = "Root cost = 1.2")) +
  geom_line(data = df_medium_low, aes(y = value, x = ppfd, color = "Root cost = 1")) +
  geom_line(data = df_low, aes(y = value, x = ppfd, color = "Root cost = 0.75")) +
  geom_line(data = df_original.m, aes(y = value, x = ppfd, color = "Numeric"), linetype="dotted") +
  scale_color_manual(values = c("Root cost = 1.5" = "red",
                                "Root cost = 1.2" = "orange",
                                "Root cost = 1" = "green",
                                "Root cost = 0.75" = "blue",
                                "Numeric" = "black")) +
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_text(face = "bold"),
        strip.text = element_text(size = 10, face = "bold"),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0, face = "bold")) +  # Added this line for the figure label
  labs(color = "Model Versions and\nInfrastructure Value",
       x = "Photosynthetic Photon Flux Density, (μmol m⁻² s⁻¹)") +  # Added the figure label here
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = new_titles))
```
# Including root biomass

```{r}
nitrogen_store_conversion = 1.5
dat_alloca_high = tibble(ppfd = seq(9,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 1.2
dat_alloca_medium = tibble(ppfd = seq(9,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 1
dat_alloca_medium_low = tibble(ppfd = seq(9,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 0.75
dat_alloca_low = tibble(ppfd = seq(9,1200,length.out=50)) %>% mutate(dat = purrr::map(.x=ppfd, .f = ~rphydro_nitrogen(tc, tc, .x, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion, root_biomass = root_biomass), options))) %>% unnest_wider(dat)

cols = c("ppfd", "a", "n_leaf", "gs", "jmax")
df_alloca_high <- dat_alloca_high[,cols] %>% melt("ppfd")
df_alloca_medium <- dat_alloca_medium[,cols] %>% melt("ppfd")
df_alloca_medium_low <- dat_alloca_medium_low[,cols] %>% melt("ppfd")
df_alloca_low <- dat_alloca_low[,cols] %>% melt("ppfd")

new_titles <- c(
  "ppfd" = "Photosynthetic Photon Flux Density (μmol m⁻² s⁻¹)",
  "a" = "Assimilation rate\n(A, μmol m⁻² s⁻¹)",
  "n_leaf" = "Nitrogen Content\nin Needle (g g-2)",
  "gs" = "Stomatal Conductance\n(gs, mol m⁻² s⁻¹)",
  "jmax" = "Jmax Parameter\n(Jmax, μmol m⁻² s⁻¹)"
)

ggplot() + 
  geom_line(data = df_alloca_high, aes(y = value, x = ppfd, color = "nitrogen conversion = 1.5")) +
  geom_line(data = df_alloca_medium, aes(y = value, x = ppfd, color = "nitrogen conversion = 1.2")) +
  geom_line(data = df_alloca_medium_low, aes(y = value, x = ppfd, color = "nitrogen conversion = 1")) +
  geom_line(data = df_alloca_low, aes(y = value, x = ppfd, color = "nitrogen conversion = 0.75")) +
  geom_line(data = df_original.m, aes(y = value, x = ppfd, color = "Numeric"), linetype="dotted") +
  scale_color_manual(values = c("nitrogen conversion = 1.5" = "red",
                                "nitrogen conversion = 1.2" = "orange",
                                "nitrogen conversion = 1" = "green",
                                "nitrogen conversion = 0.75" = "blue",
                                "Numeric" = "black")) +
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_text(face = "bold"),
        strip.text = element_text(size = 10, face = "bold"),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0, face = "bold")) +  # Added this line for the figure label
  labs(color = "Model Versions and\nInfrastructure Value",
       x = "Photosynthetic Photon Flux Density, (μmol m⁻² s⁻¹)") +  # Added the figure label here
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = new_titles))
```

## Weekly response with respect to temperature and VPD

```{r}
# Temperature
nitrogen_store_conversion = 1.5
dat_very_high_temp = tibble(ta = seq(-5,20,length.out=50)) %>% mutate(dat = purrr::map(.x=ta, .f = ~rphydro_nitrogen(.x, tc, ppfd, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion, root_biomass = root_biomass), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 1
dat_low_temp = tibble(ta = seq(-5,20,length.out=50)) %>% mutate(dat = purrr::map(.x=ta, .f = ~rphydro_nitrogen(.x, tc, ppfd, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion, root_biomass = root_biomass), options))) %>% unnest_wider(dat)

# VPD
nitrogen_store_conversion = 1.5
dat_very_high_vpd = tibble(vpd = seq(100,1000,length.out=50)) %>% mutate(dat = purrr::map(.x=vpd, .f = ~rphydro_nitrogen(tc, tc, ppfd, netrad, .x, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion, root_biomass = root_biomass), options))) %>% unnest_wider(dat)

nitrogen_store_conversion = 1
dat_low_vpd = tibble(vpd = seq(100,1000,length.out=50)) %>% mutate(dat = purrr::map(.x=vpd, .f = ~rphydro_nitrogen(tc, tc, ppfd, netrad, .x, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat)

```

```{r}
cols = c("ta", "a", "n_leaf", "gs", "jmax")
df_very_high_temp <- dat_very_high_temp[,cols] %>% melt("ta")
df_low_temp <- dat_low_temp[,cols] %>% melt("ta")

new_titles <- c(
  "ta" = "Air Temperature, 'C",
  "a" = "Assimilation rate\n(A, μmol m⁻² s⁻¹)",
  "n_leaf" = "Nitrogen Content\nin Needle (g g-2)",
  "gs" = "Stomatal Conductance\n(gs, mol m⁻² s⁻¹)",
  "jmax" = "Jmax Parameter\n(Jmax, μmol m⁻² s⁻¹)",
  "zeta" = "zeta, ratio"
)

ggplot() + 
  geom_line(data = df_very_high_temp, aes(y = value, x = ta, color = "Root cost = 9.9")) +
  geom_line(data = df_low_temp, aes(y = value, x = ta, color = "Root cost = 4.95")) +
  scale_color_manual(values = c("Root cost = 9.9" = "pink",
                                "Root cost = 4.95" = "green")) +
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_text(face = "bold"),
        strip.text = element_text(size = 10, face = "bold"),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0, face = "bold")) +  # Added this line for the figure label
  labs(color = "Model Versions and\nInfrastructure Value",
       x = "Air Temperature, 'C") +  # Added the figure label here
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = new_titles))

```
```{r}
cols = c("vpd", "a", "n_leaf", "gs", "jmax")
df_very_high_vpd <- dat_very_high_vpd[,cols] %>% melt("vpd")
df_low_vpd <- dat_low_vpd[,cols] %>% melt("vpd")

new_titles <- c(
  "vpd" = "VPD, Pa",
  "a" = "Assimilation rate\n(A, μmol m⁻² s⁻¹)",
  "n_leaf" = "Nitrogen Content\nin Needle (g g-2)",
  "gs" = "Stomatal Conductance\n(gs, mol m⁻² s⁻¹)",
  "jmax" = "Jmax Parameter\n(Jmax, μmol m⁻² s⁻¹)"
)

ggplot() + 
  geom_line(data = df_very_high_vpd, aes(y = value, x = vpd, color = "Root cost = 1.5")) +
  geom_line(data = df_low_vpd, aes(y = value, x = vpd, color = "Root cost = 1")) +
  scale_color_manual(values = c("Root cost = 1.5" = "pink",
                                "Root cost = 1" = "green")) +
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_text(face = "bold"),
        strip.text = element_text(size = 10, face = "bold"),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0, face = "bold")) +  # Added this line for the figure label
  labs(color = "Model Versions and\nInfrastructure Value",
       x = "VPD, Pa") +  # Added the figure label here
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = new_titles))

```

## Instantaneous Response

```{r}
nitrogen_store_conversion = 1.5
acc_medium_low = dat_medium_low %>% filter(ppfd == 1200)
dat_inst_medium_low = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat_medium_low = purrr::map(.x=ppfd, .f = ~rphydro_instantaneous_nitrogen(acc_medium_low$vcmax25, acc_medium_low$jmax25, tc, tc, .x, netrad, vpd, co2, pa, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat_medium_low)

nitrogen_store_conversion = 1.2
acc_low = dat_low %>% filter(ppfd == 1200)
dat_inst_low = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat_low = purrr::map(.x=ppfd, .f = ~rphydro_instantaneous_nitrogen(acc_low$vcmax25, acc_low$jmax25, tc, tc, .x, netrad, vpd, co2, pa, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat_low)

nitrogen_store_conversion = 1
acc_medium = dat_medium %>% filter(ppfd == 1200)
dat_inst_medium = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat_medium = purrr::map(.x=ppfd, .f = ~rphydro_instantaneous_nitrogen(acc_medium$vcmax25, acc_medium$jmax25, tc, tc, .x, netrad, vpd, co2, pa, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat_medium)

nitrogen_store_conversion = 0.75
acc_high = dat_high %>% filter(ppfd == 1200)
dat_inst_high = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(dat_high = purrr::map(.x=ppfd, .f = ~rphydro_instantaneous_nitrogen(acc_high$vcmax25, acc_high$jmax25, tc, tc, .x, netrad, vpd, co2, pa, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = nitrogen_store_conversion), options))) %>% unnest_wider(dat_high)

acc_original = dat_original %>% filter(ppfd == 1200)
dat_inst_original = tibble(ppfd = seq(1,1200,length.out=50)) %>% mutate(acc_original = purrr::map(.x=ppfd, .f = ~rphydro_instantaneous_numerical(acc_original$vcmax25, acc_original$jmax25, tc, tc, .x, netrad, vpd, co2, pa, fapar, kphio, psi_soil, rdark, vwind, par_plant, list(alpha=0.1, gamma=1), options))) %>% unnest_wider(acc_original)

df_inst_low <- dat_inst_low %>% melt("ppfd")
df_inst_medium_low <- dat_inst_medium_low %>% melt("ppfd")
df_inst_medium <- dat_inst_medium %>% melt("ppfd")
df_inst_high <- dat_inst_high %>% melt("ppfd")
df_inst_very_high <- dat_inst_very_high %>% melt("ppfd")
df_original.m.inst <- dat_inst_original %>% melt("ppfd")

ggplot() + 
  geom_line(data = df_inst_very_high, aes(y = value, x = ppfd, color = "Root cost = 10")) +
  geom_line(data = df_inst_low, aes(y = value, x = ppfd, color = "Root cost = 0.75")) +
  geom_line(data = df_inst_medium_low, aes(y = value, x = ppfd, color = "Root cost = 1.0")) +
  geom_line(data = df_inst_medium, aes(y = value, x = ppfd, color = "Root cost = 1.5")) +
  geom_line(data = df_inst_high, aes(y = value, x = ppfd, color = "Root cost = 2.0")) +
  geom_line(data = df_original.m.inst, aes(y = value, x = ppfd, color = "Numeric"), linetype="dotted") +
  scale_color_manual(values = c("Root cost = 0.75" = "red", 
                                "Root cost = 1.0" = "orange", 
                                "Root cost = 1.5" = "green", 
                                "Root cost = 2.0" = "blue", 
                                "Root cost = 10" = "pink",
                                "Numeric" = "black")) +
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_blank()) +
  facet_wrap(~variable, scales = "free")
```

### Compare ac, aj, in inst responses - TODO: check formula

```{r}
par(mfrow = c(1, 2))
dat_inst %>% select(ac, aj, isVcmaxLimited) %>% mutate(isVcmaxLimited = isVcmaxLimited*5) %>%  matplot(x=dat_inst$ppfd, type="l", lty=1, col=c("green4", "yellow3", "black"), main = "Nitrogen Instant", xlab = "ppfd")
abline(v=0, col="pink")
dat_inst_original %>% select(ac, aj, isVcmaxLimited) %>% mutate(isVcmaxLimited = isVcmaxLimited*5) %>%  matplot(x=dat_inst_original$ppfd, type="l", lty=2, col=c("green4", "yellow3", "black"), main = "Original Instant", xlab = "ppfd")
abline(v=0, col="pink")
```

### J Behaviour NOT CORRECTED CHECK THE VALUES

```{r}
J_n = tibble(ppfd = seq(0,1200, length.out=50)) %>% 
  mutate(J = kphio*ppfd/sqrt(1+(4*kphio*ppfd/(acc$n_leaf * a_jmax))^2))
J_n_low = tibble(ppfd = seq(0,1200, length.out=50)) %>% 
  mutate(J = kphio*ppfd/sqrt(1+(4*kphio*ppfd/(acc_low$n_leaf * a_jmax))^2))
J_n_medium = tibble(ppfd = seq(0,1200, length.out=50)) %>% 
  mutate(J = kphio*ppfd/sqrt(1+(4*kphio*ppfd/(acc_medium$n_leaf * a_jmax))^2))
J_n_high = tibble(ppfd = seq(0,1200, length.out=50)) %>% 
  mutate(J = kphio*ppfd/sqrt(1+(4*kphio*ppfd/(acc_high$n_leaf * a_jmax))^2))

J_original = tibble(ppfd = seq(0,1200, length.out=50)) %>% 
  mutate(J = kphio*ppfd/sqrt(1+(4*kphio*ppfd/(acc_original$jmax))^2))

ggplot() +
  geom_line(data = J_n, aes(x = ppfd, y = J, color = "Nitrogen Version, nitrogen_store_conversion = 1.0")) +
  geom_line(data = J_n_low, aes(x = ppfd, y = J, color = "Nitrogen Version, nitrogen_store_conversion = 0.75")) +
  geom_line(data = J_n_medium, aes(x = ppfd, y = J, color = "Nitrogen Version, nitrogen_store_conversion = 1.5")) +
  geom_line(data = J_n_high, aes(x = ppfd, y = J, color = "Nitrogen Version, nitrogen_store_conversion = 2.0")) +
  geom_line(data = J_original, aes(x = ppfd, y = J, color = "Numeric"), linetype="dotted") +
  scale_color_manual(values = c("Nitrogen Version, nitrogen_store_conversion = 0.75" = "red", 
                                "Nitrogen Version, nitrogen_store_conversion = 1.0" = "orange", 
                                "Nitrogen Version, nitrogen_store_conversion = 1.5" = "green", 
                                "Nitrogen Version, nitrogen_store_conversion = 2.0" = "blue", 
                                "Numeric" = "black")) +
  labs(x = "PPFD (mol m^-2 s^-1)", y = "J (mol m^-2 s^-1)", 
       title = "Electron Transport Rate (J) vs. PPFD")
```

### Testing parameters

```{r}
dat_nitrogen_conversion = tibble(nitrogen_conversion = seq(0.01, 2, length.out=15)) %>% mutate(dat = purrr::map(.x = nitrogen_conversion, .f = ~rphydro_nitrogen(tc, tc, ppfd, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.1, gamma=1, nitrogen_store_conversion = .x), options))) %>% unnest_wider(dat)

df_nitrogen_conversion <- dat_nitrogen_conversion %>% melt("nitrogen_conversion")

ggplot(df_nitrogen_conversion, aes(y=value, x=nitrogen_conversion)) + 
  geom_line(col="aquamarine4") + 
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1)) +
  facet_wrap(~variable, scales = "free")
```

## Formatted weekly response

```{r}
cols = c("nitrogen_conversion", "a", "n_leaf", "gs", "jmax")
df_nitrogen_conversion <- dat_nitrogen_conversion[,cols] %>% melt("nitrogen_conversion")

new_titles <- c(
  "nitrogen_conversion" = "Infrastructure",
  "a" = "Assimilation rate\n(A, μmol m⁻² s⁻¹)",
  "n_leaf" = "Nitrogen Content\nin Needle (g g-1)",
  "gs" = "Stomatal Conductance\n(gs, mol m⁻² s⁻¹)",
  "jmax" = "Jmax Parameter\n(Jmax, μmol m⁻² s⁻¹)"
)

ggplot(df_nitrogen_conversion, aes(y=value, x=nitrogen_conversion)) + 
  geom_line(col="aquamarine4") + 
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_text(face = "bold"),
        strip.text = element_text(size = 10, face = "bold"),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0, face = "bold")) +  # Added this line for the figure label
  labs(x = expression(paste("Nitrogen store conversion"))) +
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = new_titles))
```


```{r}
dat_a_jmax = tibble(a_jmax = seq(0.01,800,length.out=50)) %>% mutate(dat = purrr::map(.x = a_jmax, .f = ~rphydro_nitrogen(tc, tc, ppfd, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, .x, par_plant, par_cost, options))) %>% unnest_wider(dat)

df_ajmax <- dat_a_jmax %>% melt("a_jmax")

ggplot(df_ajmax, aes(y=value, x=a_jmax)) + 
  geom_line(col="aquamarine4") + 
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1))+
  facet_wrap(~variable, scales = "free")
```

```{r}
dat_alpha = tibble(alpha = seq(0.001,0.5,length.out=50)) %>% mutate(dat = purrr::map(.x = alpha, .f = ~rphydro_nitrogen(tc, tc, ppfd, netrad, vpd, co2, pa, nitrogen, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=.x, gamma=1, nitrogen_store_conversion = 1), options))) %>% unnest_wider(dat)

df_alpha <- dat_alpha %>% melt("alpha")

ggplot(df_alpha, aes(y=value, x=alpha)) + 
  geom_line(col="aquamarine4") + 
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1))+
  facet_wrap(~variable, scales = "free")
```

## Nitrogen store effect!

```{r}

dat_nitrogen = tibble(nitrogen = seq(0.01,1,length.out=50)) %>% mutate(dat = purrr::map(.x = nitrogen, .f = ~rphydro_nitrogen(tc, tc, ppfd, netrad, vpd, co2, pa, .x, fapar, kphio, psi_soil, rdark, vwind, a_jmax, par_plant, list(alpha=0.02, gamma=1, nitrogen_store_conversion = 1), options))) %>% unnest_wider(dat)

cols = c("nitrogen", "a", "n_leaf", "gs", "jmax")
df_nitrogen <- dat_nitrogen[,cols] %>% melt("nitrogen")

new_titles <- c(
  "nitrogen_conversion" = "Infrastructure",
  "a" = "Assimilation rate\n(A, μmol m⁻² s⁻¹)",
  "n_leaf" = "Nitrogen Content\nin Needle (g g-2)",
  "gs" = "Stomatal Conductance\n(gs, mol m⁻² s⁻¹)",
  "jmax" = "Jmax Parameter\n(Jmax, μmol m⁻² s⁻¹)",
  "zeta" = "zeta, ratio"
)

ggplot(df_nitrogen, aes(y=value, x=nitrogen)) + 
  geom_line(col="aquamarine4") + 
  theme_classic() +
  theme(strip.background = element_rect(color = "white", size = 1),
        legend.title = element_text(face = "bold"),
        strip.text = element_text(size = 10, face = "bold"),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0, face = "bold")) +  # Added this line for the figure label
  labs(x = expression(paste("Leaf Nitrogen Store (g g-1)"))) +
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = new_titles))
```
# Finding parameters

```{r}
# Set up parameters
nitrogen_conversion_values <- seq(0.02, 20, length.out = 1000)  
alpha_values <- seq(0.01, 0.04, length.out = 1000)     
param_grid <- expand.grid(nitrogen_conversion = nitrogen_conversion_values, alpha = alpha_values)

# Set chunk size and calculate number of chunks
chunk_size <- 10000  # You can adjust this based on your memory
n_chunks <- ceiling(nrow(param_grid) / chunk_size)

# Initialize results list
results_list <- vector("list", n_chunks)

for(chunk in 1:n_chunks) {
  start_idx <- ((chunk-1) * chunk_size) + 1
  end_idx <- min(chunk * chunk_size, nrow(param_grid))
  
  cat(sprintf("Processing chunk %d of %d\n", chunk, n_chunks))
  
  chunk_results <- parallel::mclapply(start_idx:end_idx, function(i) {
    tryCatch({
      params <- list(
        kphio = 0.087, ppfd = 500, ppfd_max = 1200,
        vpd = 1000, co2 = 400, elv = 0, fapar = 0.7,
        rdark = 0.015, tc = 25, vwind = 3, netrad = 150,
        psi_soil = -0.05, nitrogen = 1, a_jmax = 800
      )
      
      par_plant <- list(conductivity=3e-17, psi50=-2, b=2)
      options <- list(
        gs_method = "GS_IGF",
        et_method = "ET_DIFFUSION",
        ftemp_vj_method = "FV_kumarathunge19",
        ftemp_rd_method = "FR_heskel16",
        ftemp_br_method = "FB_atkin15",
        scale_alpha = FALSE
      )
      
      pa <- calc_patm(params$elv)
      
      par_cost_original <- list(alpha=0.1, gamma=1, nitrogen_store_conversion = 1)
      original <- as.data.frame(rphydro_numerical(
        params$tc, params$tc, params$ppfd, params$netrad, params$vpd, 
        params$co2, pa, params$fapar, params$kphio, params$psi_soil, 
        params$rdark, params$vwind, par_plant, par_cost_original, options
      ))
      
      current_params <- param_grid[i, ]
      result <- as.data.frame(rphydro_nitrogen(
        params$tc, params$tc, params$ppfd, params$netrad, params$vpd, 
        params$co2, pa, params$nitrogen, params$fapar, params$kphio, 
        params$psi_soil, params$rdark, params$vwind, params$a_jmax, 
        par_plant, list(
          alpha=current_params$alpha, 
          gamma=1, 
          nitrogen_store_conversion = current_params$nitrogen_conversion,
          root_biomass = 1
        ), 
        options
      ))
      
      result$rmse <- sqrt(sum((result[,-c(6, 10)] - original)^2)/20)
      result$alpha <- current_params$alpha
      result$nitrogen_conversion <- current_params$nitrogen_conversion
      result$dist_zeta <- abs(0.2 - result$zeta)
      result$dist_a <- abs(original$a - result$zeta)
      
      return(result)
    }, error = function(e) {
      return(NULL)  # Return NULL instead of error message
    })
  }, mc.cores = parallel::detectCores()-2)
  
  # Filter out NULL results and combine valid results
  valid_results <- chunk_results[!sapply(chunk_results, is.null)]
  if(length(valid_results) > 0) {
    results_list[[chunk]] <- do.call(rbind, valid_results)
  }
  
  # Clean up memory
  gc()
}

# Remove any NULL elements from results_list
results_list <- results_list[!sapply(results_list, is.null)]

# Combine all results
fitting_parameters_no_error_all_both <- do.call(rbind, results_list)
```

# Plots
```{r}
#plot(fitting_parameters_no_error_all_both$nitrogen_conversion, 
#     fitting_parameters_no_error_all_both$alpha,
#     xlab = "Root cost", ylab = "alpha")

library(ggplot2)
library(gridExtra)

# Create the plots
p1 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = a, color = alpha)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "a") +
  theme_minimal()

p2 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = alpha, y = a, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "alpha", y = "a") +
  theme_minimal()

p3 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = e, color = alpha)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "e") +
  theme_minimal()

p4 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = alpha, y = e, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "alpha", y = "e") +
  theme_minimal()

p5 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = n_leaf, color = alpha)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "N leaf") +
  theme_minimal()

p6 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = alpha, y = n_leaf, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "alpha", y = "N leaf") +
  theme_minimal()

p7 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = zeta, color = alpha)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "zeta") +
  theme_minimal()

p8 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = alpha, y = zeta, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "alpha", y = "zeta") +
  theme_minimal()

# Arrange all plots in a 4x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)
```

## Biomass vs nitrogen_conversion values

```{r}
# Set up parameters
nitrogen_conversion_values <- seq(0.02, 20, length.out = 100)  
alpha_values <- seq(0.005, 0.04, length.out = 100)     
root_biomass <- seq(0.01, 20, length.out = 100)     
param_grid <- expand.grid(nitrogen_conversion = nitrogen_conversion_values, alpha = alpha_values, root_biomass = root_biomass)

# Set chunk size and calculate number of chunks
chunk_size <- 10000  # You can adjust this based on your memory
n_chunks <- ceiling(nrow(param_grid) / chunk_size)

# Initialize results list
results_list <- vector("list", n_chunks)

for(chunk in 1:n_chunks) {
  start_idx <- ((chunk-1) * chunk_size) + 1
  end_idx <- min(chunk * chunk_size, nrow(param_grid))
  
  cat(sprintf("Processing chunk %d of %d\n", chunk, n_chunks))
  
  chunk_results <- parallel::mclapply(start_idx:end_idx, function(i) {
    tryCatch({
      params <- list(
        kphio = 0.087, ppfd = 300, ppfd_max = 1200,
        vpd = 1000, co2 = 400, elv = 0, fapar = 0.7,
        rdark = 0.015, tc = 25, vwind = 3, netrad = 150,
        psi_soil = -0.05, nitrogen = 1, a_jmax = 800
      )
      
      par_plant <- list(conductivity=3e-17, psi50=-2, b=2)
      options <- list(
        gs_method = "GS_IGF",
        et_method = "ET_DIFFUSION",
        ftemp_vj_method = "FV_kumarathunge19",
        ftemp_rd_method = "FR_heskel16",
        ftemp_br_method = "FB_atkin15",
        scale_alpha = FALSE
      )
      
      pa <- calc_patm(params$elv)
      
      par_cost_original <- list(alpha=0.1, gamma=1, nitrogen_store_conversion = 1)
      original <- as.data.frame(rphydro_numerical(
        params$tc, params$tc, params$ppfd, params$netrad, params$vpd, 
        params$co2, pa, params$fapar, params$kphio, params$psi_soil, 
        params$rdark, params$vwind, par_plant, par_cost_original, options
      ))
      
      current_params <- param_grid[i, ]
      result <- as.data.frame(rphydro_nitrogen(
        params$tc, params$tc, params$ppfd, params$netrad, params$vpd, 
        params$co2, pa, params$nitrogen, params$fapar, params$kphio, 
        params$psi_soil, params$rdark, params$vwind, params$a_jmax, 
        par_plant, list(
          alpha=current_params$alpha, 
          gamma=1, 
          nitrogen_store_conversion = current_params$nitrogen_conversion,
          root_biomass = current_params$root_biomass
        ), 
        options
      ))
      
      result$rmse <- sqrt(sum((result[,-c(6, 10)] - original)^2)/20)
      result$alpha <- current_params$alpha
      result$nitrogen_conversion <- current_params$nitrogen_conversion
      result$root_biomass <- current_params$root_biomass
      result$dist_zeta <- abs(0.2 - result$zeta)
      result$dist_a <- abs(original$a - result$zeta)
      
      return(result)
    }, error = function(e) {
      return(NULL)  # Return NULL instead of error message
    })
  }, mc.cores = parallel::detectCores()-2)
  
  # Filter out NULL results and combine valid results
  valid_results <- chunk_results[!sapply(chunk_results, is.null)]
  if(length(valid_results) > 0) {
    results_list[[chunk]] <- do.call(rbind, valid_results)
  }
  
  # Clean up memory
  gc()
}

# Remove any NULL elements from results_list
results_list <- results_list[!sapply(results_list, is.null)]

# Combine all results
fitting_parameters_no_error_all_both <- do.call(rbind, results_list)
```

# Plots
```{r}
#plot(fitting_parameters_no_error_all_both$nitrogen_conversion, 
#     fitting_parameters_no_error_all_both$root_biomass,
#     xlab = "Root cost", ylab = "root_biomass")

# Create the plots
p1 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = a, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "a") +
  theme_minimal()

p2 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = root_biomass, y = a, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "a") +
  theme_minimal()

p3 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = e, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "e") +
  theme_minimal()

p4 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = root_biomass, y = e, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "e") +
  theme_minimal()

p5 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = n_leaf, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "N leaf") +
  theme_minimal()

p6 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = root_biomass, y = n_leaf, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "N leaf") +
  theme_minimal()

p7 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = nitrogen_conversion, y = zeta, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "zeta") +
  theme_minimal()

p8 <- ggplot(fitting_parameters_no_error_all_both, 
             aes(x = root_biomass, y = zeta, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "zeta") +
  theme_minimal()

# Arrange all plots in a 4x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)

closest_rows <- fitting_parameters_no_error_all_both
closest_rows <- closest_rows[which(closest_rows$dist_a == min(closest_rows$dist_a)),]
closest_rows <- closest_rows[which(closest_rows$dist_zeta == min(closest_rows$dist_zeta)),]

p1 <- ggplot(closest_rows, 
             aes(x = nitrogen_conversion, y = a, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "a") +
  theme_minimal()

p2 <- ggplot(closest_rows, 
             aes(x = root_biomass, y = a, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "a") +
  theme_minimal()

p3 <- ggplot(closest_rows, 
             aes(x = nitrogen_conversion, y = e, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "e") +
  theme_minimal()

p4 <- ggplot(closest_rows, 
             aes(x = root_biomass, y = e, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "e") +
  theme_minimal()

p5 <- ggplot(closest_rows, 
             aes(x = nitrogen_conversion, y = n_leaf, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "N leaf") +
  theme_minimal()

p6 <- ggplot(closest_rows, 
             aes(x = root_biomass, y = n_leaf, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "N leaf") +
  theme_minimal()

p7 <- ggplot(closest_rows, 
             aes(x = nitrogen_conversion, y = zeta, color = root_biomass)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "Root cost", y = "zeta") +
  theme_minimal()

p8 <- ggplot(closest_rows, 
             aes(x = root_biomass, y = zeta, color = nitrogen_conversion)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(x = "root_biomass", y = "zeta") +
  theme_minimal()

# Arrange all plots in a 4x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)
```

## Added the root biomass value

```{r}
library(ggplot2)
library(gridExtra)
library(viridis)

# Function to create a plot for a specific response variable
create_plot_set <- function(data, response_var, response_name) {
  # Plot 1: x = root_cost, y = response, color = alpha
  p1 <- ggplot(data, aes(x = nitrogen_conversion, y = .data[[response_var]], color = alpha)) +
    geom_point(alpha = 0.5, size = 0.5) +
    scale_color_viridis_c(name = "Alpha") +
    labs(x = "Root cost", y = response_name) +
    theme_minimal() +
    theme(legend.position = "right")

  # Plot 2: x = alpha, y = response, color = root_cost
  p2 <- ggplot(data, aes(x = alpha, y = .data[[response_var]], color = nitrogen_conversion)) +
    geom_point(alpha = 0.5, size = 0.5) +
    scale_color_viridis_c(name = "Root cost") +
    labs(x = "Alpha", y = response_name) +
    theme_minimal() +
    theme(legend.position = "right")

  # Plot 3: x = root_biomass, y = response, color = root_cost
  p3 <- ggplot(data, aes(x = root_biomass, y = .data[[response_var]], color = nitrogen_conversion)) +
    geom_point(alpha = 0.5, size = 0.5) +
    scale_color_viridis_c(name = "Root cost") +
    labs(x = "Root biomass", y = response_name) +
    theme_minimal() +
    theme(legend.position = "right")

  return(list(p1, p2, p3))
}

# Create plots for each response variable
a_plots <- create_plot_set(fitting_parameters_no_error_all_both, "a", "Assimilation (a)")
e_plots <- create_plot_set(fitting_parameters_no_error_all_both, "e", "Transpiration (e)")
n_leaf_plots <- create_plot_set(fitting_parameters_no_error_all_both, "n_leaf", "N leaf")
zeta_plots <- create_plot_set(fitting_parameters_no_error_all_both, "zeta", "Zeta")

# Arrange all plots in a grid
# Each row represents one response variable, with three plots showing different parameter combinations
grid.arrange(
  # First row - Assimilation plots
  arrangeGrob(a_plots[[1]], a_plots[[2]], a_plots[[3]], ncol = 3),
  # Second row - Transpiration plots
  arrangeGrob(e_plots[[1]], e_plots[[2]], e_plots[[3]], ncol = 3),
  # Third row - N leaf plots
  arrangeGrob(n_leaf_plots[[1]], n_leaf_plots[[2]], n_leaf_plots[[3]], ncol = 3),
  # Fourth row - Zeta plots
  arrangeGrob(zeta_plots[[1]], zeta_plots[[2]], zeta_plots[[3]], ncol = 3),
  ncol = 1
)

# Create additional summary statistics plots
# RMSE plot
p_rmse <- ggplot(fitting_parameters_no_error_all_both, 
                 aes(x = nitrogen_conversion, y = alpha, color = rmse)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c() +
  facet_wrap(~cut(root_biomass, breaks = 4)) +
  labs(title = "RMSE across parameter space",
       x = "Root cost", y = "Alpha") +
  theme_minimal()

# Distance metrics plots
p_dist_zeta <- ggplot(fitting_parameters_no_error_all_both, 
                      aes(x = nitrogen_conversion, y = alpha, color = dist_zeta)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c() +
  facet_wrap(~cut(root_biomass, breaks = 4)) +
  labs(title = "Distance from target zeta",
       x = "Root cost", y = "Alpha") +
  theme_minimal()

p_dist_a <- ggplot(fitting_parameters_no_error_all_both, 
                   aes(x = nitrogen_conversion, y = alpha, color = dist_a)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c() +
  facet_wrap(~cut(root_biomass, breaks = 4)) +
  labs(title = "Distance from target assimilation",
       x = "Root cost", y = "Alpha") +
  theme_minimal()

# Arrange summary plots
grid.arrange(p_rmse, p_dist_zeta, p_dist_a, ncol = 1)

```