---
title: "Nitrogen Uptake Model — Whole-System Diffusion Budget"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(gridExtra)
library(patchwork)
library(viridis)
```

## Overview

This document implements a nitrogen uptake model based on a whole-system
diffusion budget. The total nitrogen available to the root system per unit
time is limited by diffusive flux into the root zone from bulk soil. A
mean soil concentration $\bar{N}$ adjusts self-consistently until diffusive
supply equals total biochemical demand.

A fraction $m$ of root surface is colonised by ECM fungi. The colonised
surface is covered by the Hartig net and does not contribute to direct
uptake. Active absorbing surface area is therefore:

$$\mathrm{SA}_\text{active} = (1-m)\,\mathrm{SA}_{fr} + \mathrm{SA}_m$$

The steady-state balance is:

$$\frac{D \cdot (N_s - \bar{N}) \cdot A_\text{zone}}{r_\text{zone}}
= \mathrm{SA}_\text{active} \cdot u_{A\max}
  \cdot \frac{\bar{N}}{\bar{N} + k_{15}}$$

Cross-multiplying, expanding, and collecting terms gives a quadratic in
$\bar{N}$. Defining $\alpha = u_{A\max} \cdot \mathrm{SA}_\text{active}
\cdot r_\text{zone} / (D \cdot A_\text{zone})$:

$$\bar{N}^2 + (k_{15} + \alpha - N_s)\,\bar{N} - k_{15}\,N_s = 0$$

with positive root:

$$\bar{N} = \frac{1}{2}\left[(N_s - k_{15} - \alpha)
  + \sqrt{(N_s - k_{15} - \alpha)^2 + 4\,k_{15}\,N_s}\right]$$

Uptake by each organ uses $\bar{N}$ and its own active surface:

$$U_{fr} = (1-m)\,\mathrm{SA}_{fr} \cdot u_{A\max}
           \cdot \frac{\bar{N}}{\bar{N} + k_{15}}$$

$$U_m = \mathrm{SA}_m \cdot u_{A\max}
        \cdot \frac{\bar{N}}{\bar{N} + k_{15}}$$

---

## Parameters

```{r parameters}
###
### SOIL PARAMETERS
###

u_Amax    <- 0.088   # Area-based max uptake (kg N m^-2 year^-1)
D_default <- 0.01    # Effective soil diffusion coefficient (m^2 year^-1)
k_15      <- 0.05    # Biochemical half-saturation constant (kg N m^-3)
depth     <- 0.1     # Rooting depth (m)

###
### ROOT PARAMETERS
###

myco_diameter <- 1e-5   # Hyphal diameter (m)
myco_density  <- 330    # Hyphal tissue density (kg m^-3)
l_hyp         <- 1e-4   # Hyphal segment length (m)
T_m           <- 2      # Mycorrhizal turnover rate (year^-1)
mycorrhized   <- 0.9    # Fraction of roots colonised by ECM

n <- 3e5
l <- 1.5

C_r <- 2

###
### EXTRA PARAMETERS
###

k_9  <- 0.01
k_8  <- 2.7
k_13 <- 2.5
k_14 <- 0.44

traits <- list(
  k_0 = 0.5, k_1 = 0.35,       k_2 = 0.2,           k_3 = 0.014,
  k_4 = 11632.91276, k_5 = -334.3626755, k_6 = 3,    k_7 = 3,
  k_8 = k_8,  k_9 = k_9
)

crown_area_fixed <- pi * C_r^2
B_cr_fixed       <- 0
rho_cr_fixed     <- 500
```

---

## Equations

```{r equations}
############################################################
## ROOT ARCHITECTURE FUNCTIONS
############################################################

root_diameter <- function(root_length, traits) {
  traits$k_1 / (root_length ^ traits$k_0)
}

root_cortex_proportion <- function(root_length, traits) {
  d <- root_diameter(root_length, traits)
  (traits$k_2 + traits$k_3 / d)^2
}

root_density <- function(root_length, traits) {
  PRD <- root_cortex_proportion(root_length, traits)
  traits$k_4 * PRD + traits$k_5
}

root_age <- function(root_length, traits) {
  d <- root_diameter(root_length, traits)
  traits$k_6 * (1 - exp(-traits$k_7 * d))
}

root_biomass_per_crown_area <- function(root_length, root_no, traits) {
  d_m   <- root_diameter(root_length, traits) / 1000
  rho   <- root_density(root_length, traits)
  l_tip <- root_length / 1000
  rho * (d_m / 2)^2 * l_tip * pi * root_no
}

root_mass <- function(root_length, root_no, crown_area, traits) {
  2 * root_biomass_per_crown_area(root_length, root_no, traits) * crown_area
}

############################################################
## SURFACE AREAS
############################################################

root_surface_area <- function(root_length, root_no, traits) {
  d_m   <- root_diameter(root_length, traits) / 1000
  l_tip <- root_length / 1000
  pi * d_m * l_tip * root_no
}

myco_surface_area <- function(B_m, myco_diameter, myco_density) {
  L_total <- B_m / (myco_density * pi * (myco_diameter / 2)^2)
  pi * myco_diameter * L_total
}

############################################################
## ROOT ZONE GEOMETRY
##
## Semi-ellipsoidal root zone of radius R = k_13 * C_r
## and depth q.
##   A_zone = 2*pi*R^2 + pi*R*q  (lateral + base)
##   r_zone = R
############################################################

zone_geometry <- function(C_r, q, k_13) {
  R      <- k_13 * C_r
  A_zone <- 2 * pi * R^2 + pi * R * q
  r_zone <- R
  list(A_zone = A_zone, r_zone = r_zone)
}

############################################################
## ACTIVE SURFACE AREA
##
## SA_active = (1 - m) * SA_fr + SA_m
##
## Colonised root surface (m * SA_fr) is excluded from
## direct uptake — nitrogen there is transferred via the
## fungal pool N_m rather than absorbed directly.
############################################################

active_surface_area <- function(SA_fr, SA_m, m) {
  (1 - m) * SA_fr + SA_m
}

############################################################
## MEAN SOIL CONCENTRATION N_bar
##
## From steady-state balance between diffusive supply
## and biochemical demand:
##
##   D*(N_s - N_bar)*A_zone/r_zone
##       = SA_active * u_Amax * N_bar/(N_bar + k_15)
##
## Cross-multiplying and collecting gives quadratic:
##   N_bar^2 + (k_15 + alpha - N_s)*N_bar - k_15*N_s = 0
##
## where alpha = u_Amax * SA_active * r_zone / (D * A_zone)
##
## Positive root:
##   b     = N_s - k_15 - alpha
##   N_bar = 0.5 * (b + sqrt(b^2 + 4*k_15*N_s))
############################################################

N_bar <- function(N_s, D, u_Amax, k_15,
                  SA_active, A_zone, r_zone) {
  if (SA_active <= 0) return(N_s)
  alpha <- u_Amax * SA_active * r_zone / (D * A_zone)
  b     <- N_s - k_15 - alpha
  disc  <- sqrt(b^2 + 4 * k_15 * N_s)
  if (b >= 0) 0.5 * (b + disc) else 2 * k_15 * N_s / (disc - b)
}

############################################################
## ORGAN UPTAKE
##
## U_fr = (1-m) * SA_fr * u_Amax * N_bar/(N_bar + k_15)
## U_m  =         SA_m  * u_Amax * N_bar/(N_bar + k_15)
##
## Both use N_bar. The (1-m) on U_fr is already embedded
## in SA_active used to compute N_bar, but also applied
## here to allocate the shared flux correctly between
## the two organ types.
############################################################

uptake_root <- function(SA_fr, m, u_Amax, k_15, Nbar) {
  (1 - m) * SA_fr * u_Amax * (Nbar / (Nbar + k_15))
}

uptake_myco <- function(SA_m, u_Amax, k_15, Nbar) {
  SA_m * u_Amax * (Nbar / (Nbar + k_15))
}

############################################################
## UPTAKE MODIFIERS
############################################################

# Arrhenius temperature response for u_max (root N transporter kinetics).
# Ea = 50 kJ/mol => Q10 ~ 2 in boreal range. Normalised to 1 at 25 degC.
# Hard-zero below 0 degC (frozen soil).
Ea_uptake <- 50000  # J/mol
R_gas     <- 8.3145 # J/mol/K
T_ref     <- 298.15 # K (25 degC)

ftemp_arrhenius_uptake <- function(tc) {
  if (tc <= 0) return(0)
  tk <- tc + 273.15
  exp(Ea_uptake * (tk - T_ref) / (T_ref * R_gas * tk))
}

uptake_age_modifier <- function(root_length, traits) {
  age <- root_age(root_length, traits)
  1 / (1 + exp(age - traits$k_8))
}

nitrogen_gate <- function(root_no, root_length, traits) {
  d  <- root_diameter(root_length, traits)
  SA <- root_no * pi * d * root_length
  SA / (SA + traits$k_9)
}

############################################################
## WHOLE-PLANT UPTAKE
##
## U_a = a_u * [U_fr + m * g(n,l) * T_m * max(0, N_m)]
##
## Note: (1-m) is already inside U_fr via uptake_root().
## The ECM transfer term handles nitrogen from colonised
## roots via the fungal pool N_m.
############################################################

plant_uptake <- function(N_s, N_m,
                         crown_area, root_length, root_no,
                         B_m, B_cr, rho_cr,
                         traits,
                         mycorrhized, T_m,
                         myco_diameter, myco_density, l_hyp,
                         C_r, q = 0.5, k_13 = 2, k_14 = 0.5,
                         u_Amax, D, k_15,
                         tc = 25) {

  B_fr  <- root_mass(root_length, root_no, crown_area, traits)
  SA_fr <- root_surface_area(root_length, root_no, traits) * crown_area
  SA_m  <- myco_surface_area(B_m, myco_diameter, myco_density)

  geom    <- zone_geometry(C_r, q, k_13)
  SA_act  <- active_surface_area(SA_fr, SA_m, mycorrhized)

  f_temp <- ftemp_arrhenius_uptake(tc)

  Nbar     <- N_bar(N_s, D, u_Amax * f_temp, k_15,
                    SA_act, geom$A_zone, geom$r_zone)
  F_supply <- D * N_s * geom$A_zone / geom$r_zone

  U_fr <- uptake_root(SA_fr, mycorrhized, u_Amax * f_temp, k_15, Nbar)
  U_m  <- uptake_myco(SA_m, u_Amax * f_temp, k_15, Nbar)

  a_u <- uptake_age_modifier(root_length, traits)
  g   <- nitrogen_gate(root_no, root_length, traits)

  ECM_transfer <- mycorrhized * g * T_m * max(0, N_m)
  U_a <- a_u * (U_fr + ECM_transfer)

  return(list(
    N_total      = U_a,
    U_fr         = U_fr,
    U_m          = U_m,
    SA_fr        = SA_fr,
    SA_m         = SA_m,
    SA_active    = SA_act,
    ECM_transfer = ECM_transfer,
    N_bar        = Nbar,
    F_supply     = F_supply,
    A_zone       = geom$A_zone,
    r_zone       = geom$r_zone,
    a_u          = a_u,
    g            = g,
    B_fr         = B_fr
  ))
}
```

---

## Diagnostic: $\bar{N}$ and Uptake Behaviour

```{r diagnostics, fig.height=8}
N_s_vals    <- seq(0.001, 2,  length.out = 200)
n_vals      <- 10^seq(3, 7,   length.out = 5)
D_vals_diag <- c(1e-5, 1e-4, 1e-3, 1e-2)

geom_fixed <- zone_geometry(C_r, depth, k_13)
D_fmt <- formatC(D_vals_diag, format = "e", digits = 2)
cb_D  <- setNames(c("#E69F00", "#56B4E9", "#009E73", "#CC79A7"), D_fmt)

# N_bar vs N_s by D
df_D <- expand_grid(N_s = N_s_vals, D = D_vals_diag) %>%
  rowwise() %>%
  mutate(
    SA_fr   = root_surface_area(l, n, traits) * crown_area_fixed,
    SA_m    = myco_surface_area(0, myco_diameter, myco_density),
    SA_act  = active_surface_area(SA_fr, SA_m, mycorrhized),
    Nbar    = N_bar(N_s, D, u_Amax, k_15,
                    SA_act, geom_fixed$A_zone, geom_fixed$r_zone),
    D_label = factor(formatC(D, format = "e", digits = 2), levels = D_fmt)
  ) %>%
  ungroup()

# N_bar and uptake vs tip density
df_diag <- expand_grid(N_s = N_s_vals, n_tip = n_vals) %>%
  rowwise() %>%
  mutate(
    B_fr   = root_mass(l, n_tip, crown_area_fixed, traits),
    SA_fr  = root_surface_area(l, n_tip, traits) * crown_area_fixed,
    SA_act = active_surface_area(SA_fr, 0, mycorrhized),
    Nbar   = N_bar(N_s, D_default, u_Amax, k_15,
                   SA_act, geom_fixed$A_zone, geom_fixed$r_zone),
    U_fr   = uptake_root(SA_fr, mycorrhized, u_Amax, k_15, Nbar),
    U_per_B = U_fr / B_fr,
    F_sup  = D_default * N_s * geom_fixed$A_zone / geom_fixed$r_zone
  ) %>%
  ungroup() %>%
  mutate(n_label = factor(signif(n_tip, 2)))

nbar_bounds <- data.frame(
  xintercept = c(0.0012, 0.003),
  label      = c("Korhonen (mineral)", "With amino acids")
)

p_Nbar_D <- ggplot(df_D, aes(N_s, Nbar, colour = D_label)) +
  geom_line(linewidth = 1.1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  scale_linetype_manual(name = "N_s reference",
                        values = c("Korhonen (mineral)" = "dashed",
                                   "With amino acids"   = "dotdash")) +
  scale_colour_manual(values = cb_D,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  labs(title = expression(bar(N)~"vs"~N[s]~"by D"),
       subtitle = "Dashed = no depletion; high D approaches dashed",
       x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  theme_bw(base_size = 12)

p_Nbar_n <- ggplot(df_diag, aes(N_s, Nbar, colour = n_label)) +
  geom_line(linewidth = 1.1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  scale_linetype_manual(name = "N_s reference",
                        values = c("Korhonen (mineral)" = "dashed",
                                   "With amino acids"   = "dotdash")) +
  scale_colour_viridis_d(name = "n (tips/m\u00b2)") +
  labs(title = expression(bar(N)~"vs"~N[s]~"by tip density"),
       subtitle = "More roots -> more depletion -> lower N_bar",
       x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  theme_bw(base_size = 12)

p_U_Ns <- ggplot(df_diag, aes(N_s, U_fr, colour = n_label)) +
  geom_line(linewidth = 1.1) +
  geom_line(aes(y = F_sup), colour = "black",
            linetype = "dashed", linewidth = 0.8) +
  scale_colour_viridis_d(name = "n (tips/m\u00b2)") +
  scale_y_log10(labels = scales::scientific) +
  labs(title = "Total root uptake vs N_s",
       subtitle = "Dashed = diffusive supply ceiling",
       x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) +
  theme_bw(base_size = 12)

p_U_B <- ggplot(df_diag, aes(B_fr, U_per_B, colour = n_label)) +
  geom_line(linewidth = 1.1) +
  scale_colour_viridis_d(name = "n (tips/m\u00b2)") +
  scale_x_log10(labels = scales::scientific) +
  scale_y_continuous(labels = scales::scientific) +
  labs(title = "Uptake per biomass vs B_fr — should saturate",
       x = expression(B[fr]~"(kg)"),
       y = expression(U[fr]~B[fr]^{-1}~"(kg N kg"^{-1}~"yr"^{-1}*")")) +
  theme_bw(base_size = 12)

(p_Nbar_D | p_Nbar_n) / (p_U_Ns | p_U_B)
```

---

## Root and Mycorrhizal Uptake

```{r uptake_plots, fig.height=8}
D_vals        <- c(1e-5, 1e-4, 1e-3, 1e-2)
D_levels      <- formatC(D_vals, format = "e", digits = 2)
u_max_fixed   <- 0.88/10
N_s_vals      <- seq(0.01, 1.0, length.out = 200)
N_s_highlight <- c(0.2, 0.6)

cb_two <- setNames(c("#E69F00", "#56B4E9", "#009E73", "#CC79A7"), D_levels)

D_labels_plot <- setNames(
  list(expression(10^{-5}), expression(10^{-4}),
       expression(10^{-3}), expression(10^{-2})),
  D_levels
)

df_basis <- tibble(n = seq(1e1, 1e10, length.out = 20), l = 1.5) %>%
  rowwise() %>%
  mutate(
    B_fr  = root_mass(l, n, crown_area_fixed, traits),
    SA_fr = root_surface_area(l, n, traits) * crown_area_fixed
  ) %>%
  ungroup()

B_highlight  <- c(df_basis$B_fr[7],  max(df_basis$B_fr))
SA_highlight <- c(df_basis$SA_fr[7], max(df_basis$SA_fr))
n_highlight  <- c(df_basis$n[7],     max(df_basis$n))

theme_pub <- theme_bw(base_size = 14) +
  theme(panel.grid.minor = element_blank(), strip.background = element_blank(),
        legend.position = "right", legend.text = element_text(size = 11),
        legend.title = element_text(size = 12), legend.key.size = unit(0.4, "cm"))

highlight_df <- tibble(B_fr  = B_highlight,
                       SA_fr = SA_highlight,
                       n_val = n_highlight)

# helper — uses mycorrhized and active_surface_area
get_Nbar <- function(N_s, D, SA_fr, B_m, C_r, q, k_13, m) {
  SA_m   <- myco_surface_area(B_m, myco_diameter, myco_density)
  SA_act <- active_surface_area(SA_fr, SA_m, m)
  geom   <- zone_geometry(C_r, q, k_13)
  N_bar(N_s, D, u_max_fixed, k_15,
        SA_act, geom$A_zone, geom$r_zone)
}

# ---- (a) U_fr per biomass vs N_s ----
df_root_Ns <- crossing(highlight_df,
                       myco = c("no myco", "myco"),
                       N_s  = N_s_vals,
                       D    = D_vals) %>%
  rowwise() %>%
  mutate(
    B_m_use    = ifelse(myco == "myco", 1.0, 0),
    m_use      = ifelse(myco == "myco", mycorrhized, 0),
    Nbar       = get_Nbar(N_s, D, SA_fr, B_m_use, C_r, depth, k_13, m_use),
    U_fr       = uptake_root(SA_fr, m_use, u_max_fixed, k_15, Nbar),
    U_fr_per_B = U_fr / B_fr,
    D_label    = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    B_label    = factor(ifelse(B_fr == min(B_highlight), "low", "high"),
                        levels = c("low", "high"))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(myco == "no myco",
                               "(a.i) Without ECM", "(a.ii) With ECM"),
                        levels = c("(a.i) Without ECM", "(a.ii) With ECM")))

p_root_Ns <- ggplot(df_root_Ns,
       aes(x = N_s, y = U_fr_per_B, colour = D_label, linetype = B_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("low" = 3, "high" = 1), name = expression(B[fr])) +
  scale_y_log10(labels = scales::scientific) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[fr]~B[fr]^{-1}~"(kg N kg"^{-1}~"yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

# ---- (b) U_fr per biomass vs B_fr ----
df_root_B <- df_basis %>%
  crossing(myco = c("no myco", "myco"), N_s = N_s_highlight, D = D_vals) %>%
  rowwise() %>%
  mutate(
    B_m_use    = ifelse(myco == "myco", 1.0, 0),
    m_use      = ifelse(myco == "myco", mycorrhized, 0),
    Nbar       = get_Nbar(N_s, D, SA_fr, B_m_use, C_r, depth, k_13, m_use),
    U_fr       = uptake_root(SA_fr, m_use, u_max_fixed, k_15, Nbar),
    U_fr_per_B = U_fr / B_fr,
    D_label    = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    N_label    = factor(as.character(N_s), levels = as.character(N_s_highlight))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(myco == "no myco",
                               "(b.i) Without ECM", "(b.ii) With ECM"),
                        levels = c("(b.i) Without ECM", "(b.ii) With ECM")))

p_root_B <- ggplot(df_root_B,
       aes(x = B_fr, y = U_fr_per_B, colour = D_label, linetype = N_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("0.2" = 3, "0.6" = 1),
                        labels = c("0.2" = expression(N[s] == 0.2),
                                   "0.6" = expression(N[s] == 0.6)),
                        name = expression(N[s]~"(kg N m"^{-3}*")")) +
  scale_x_log10(labels = scales::scientific) +
  scale_y_continuous(labels = scales::scientific) +
  labs(x = expression(B[fr]~"(kg)"),
       y = expression(U[fr]~B[fr]^{-1}~"(kg N kg"^{-1}~"yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

# ---- (c) U_m per biomass vs N_s ----
# B_m highlights are independent ECM biomass values, not derived from df_basis
B_m_highlight <- c(0.01, 1.0)   # low and high ECM biomass (kg)

df_myco_Ns <- crossing(
    tibble(B_m   = B_m_highlight),
    root = c("no root", "root"),
    N_s  = N_s_vals,
    D    = D_vals) %>%
  rowwise() %>%
  mutate(
    SA_m      = myco_surface_area(B_m, myco_diameter, myco_density),
    SA_fr_use = ifelse(root == "root",
                       root_surface_area(l, n, traits) * crown_area_fixed, 0),
    m_use     = ifelse(root == "root", mycorrhized, 0),
    Nbar      = get_Nbar(N_s, D, SA_fr_use, B_m, C_r, depth, k_13, m_use),
    U_m       = uptake_myco(SA_m, u_max_fixed, k_15, Nbar),
    U_m_per_B = U_m / B_m,
    D_label   = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    B_label   = factor(ifelse(B_m == min(B_m_highlight), "low", "high"),
                       levels = c("low", "high"))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(root == "no root",
                               "(c.i) Without roots", "(c.ii) With roots"),
                        levels = c("(c.i) Without roots", "(c.ii) With roots")))

p_myco_Ns <- ggplot(df_myco_Ns,
       aes(x = N_s, y = U_m_per_B, colour = D_label, linetype = B_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("low" = 3, "high" = 1), name = expression(B[m])) +
  scale_y_log10(labels = scales::scientific) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[m]~B[m]^{-1}~"(kg N kg"^{-1}~"yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

# ---- (d) U_m per biomass vs B_m ----
# B_m swept independently over a wide range so D controls
# where saturation occurs — not derived from df_basis
f_root <- 0.5   # fraction of total biomass allocated to roots vs ECM

df_myco_B <- crossing(
    tibble(B_total = 10^seq(-4, 2, length.out = 60)),
    root = c("no root", "root"),
    N_s  = N_s_highlight,
    D    = D_vals) %>%
  rowwise() %>%
  mutate(
    B_m       = ifelse(root == "root", (1 - f_root) * B_total, B_total),
    B_fr_use  = ifelse(root == "root", f_root * B_total, 0),
    n_use     = ifelse(root == "root",
                  # invert root_mass to get n from B_fr
                  B_fr_use / (2 * root_biomass_per_crown_area(l, 1, traits)
                               * crown_area_fixed), 0),
    SA_fr_use = root_surface_area(l, n_use, traits) * crown_area_fixed,
    m_use     = ifelse(root == "root", mycorrhized, 0),
    SA_m      = myco_surface_area(B_m, myco_diameter, myco_density),
    Nbar      = get_Nbar(N_s, D, SA_fr_use, B_m, C_r, depth, k_13, m_use),
    U_m       = uptake_myco(SA_m, u_max_fixed, k_15, Nbar),
    U_m_per_B = U_m / B_m,
    D_label   = factor(formatC(D, format = "e", digits = 2), levels = D_levels),
    N_label   = factor(as.character(N_s), levels = as.character(N_s_highlight))
  ) %>%
  ungroup() %>%
  mutate(panel = factor(ifelse(root == "no root",
                               "(d.i) Without roots", "(d.ii) With roots"),
                        levels = c("(d.i) Without roots", "(d.ii) With roots")))

p_myco_B <- ggplot(df_myco_B,
       aes(x = B_total, y = U_m_per_B, colour = D_label, linetype = N_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("0.2" = 3, "0.6" = 1),
                        labels = c("0.2" = expression(N[s] == 0.2),
                                   "0.6" = expression(N[s] == 0.6)),
                        name = expression(N[s]~"(kg N m"^{-3}*")")) +
  scale_x_log10(labels = scales::scientific) +
  scale_y_continuous(labels = scales::scientific) +
  labs(x = expression(B[tot]~"(kg)"),
       y = expression(U[m]~B[m]^{-1}~"(kg N kg"^{-1}~"yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

# per biomass figure
fig_uptake <- (p_root_Ns | p_root_B) / (p_myco_Ns | p_myco_B) +
  plot_layout(guides = "collect") & theme(legend.position = "right")
fig_uptake
```

```{r uptake_total, fig.height=8}
p_root_Ns_tot <- ggplot(df_root_Ns,
       aes(x = N_s, y = U_fr, colour = D_label, linetype = B_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("low" = 3, "high" = 1), name = expression(B[fr])) +
  scale_y_log10(labels = scales::scientific) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

p_root_B_tot <- ggplot(df_root_B,
       aes(x = B_fr, y = U_fr, colour = D_label, linetype = N_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("0.2" = 3, "0.6" = 1),
                        labels = c("0.2" = expression(N[s] == 0.2),
                                   "0.6" = expression(N[s] == 0.6)),
                        name = expression(N[s]~"(kg N m"^{-3}*")")) +
  scale_x_log10(labels = scales::scientific) +
  scale_y_continuous(labels = scales::scientific) +
  labs(x = expression(B[fr]~"(kg)"),
       y = expression(U[fr]~"(kg N yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

p_myco_Ns_tot <- ggplot(df_myco_Ns,
       aes(x = N_s, y = U_m, colour = D_label, linetype = B_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("low" = 3, "high" = 1), name = expression(B[m])) +
  scale_y_log10(labels = scales::scientific) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = expression(U[m]~"(kg N yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

p_myco_B_tot <- ggplot(df_myco_B,
       aes(x = B_total, y = U_m, colour = D_label, linetype = N_label)) +
  geom_line(linewidth = 1.1) + facet_wrap(~ panel) +
  scale_colour_manual(values = cb_two, labels = D_labels_plot,
                      name = expression(D~"(m"^2~"yr"^{-1}*")")) +
  scale_linetype_manual(values = c("0.2" = 3, "0.6" = 1),
                        labels = c("0.2" = expression(N[s] == 0.2),
                                   "0.6" = expression(N[s] == 0.6)),
                        name = expression(N[s]~"(kg N m"^{-3}*")")) +
  scale_x_log10(labels = scales::scientific) +
  scale_y_continuous(labels = scales::scientific) +
  labs(x = expression(B[tot]~"(kg)"),
       y = expression(U[m]~"(kg N yr"^{-1}*")")) +
  theme_pub + theme(strip.text = element_text(size = 10))

fig_uptake_total <- (p_root_Ns_tot | p_root_B_tot) /
                    (p_myco_Ns_tot | p_myco_B_tot) +
  plot_layout(guides = "collect") & theme(legend.position = "right")
fig_uptake_total
```

---

## Sensitivity Analysis

```{r sensitivity, fig.height=8}
N_s_fixed <- 1.0
N_m_fixed <- 0.1
B_m_fixed <- 0.5

run_uptake <- function(depth_val, k15_val, k13_val, k14_val) {
  plant_uptake(
    N_s = N_s_fixed, N_m = N_m_fixed,
    crown_area = crown_area_fixed,
    root_length = l, root_no = n,
    B_m = B_m_fixed, B_cr = B_cr_fixed, rho_cr = rho_cr_fixed,
    traits = traits, mycorrhized = mycorrhized, T_m = T_m,
    myco_diameter = myco_diameter, myco_density = myco_density, l_hyp = l_hyp,
    C_r = C_r, q = depth_val, k_13 = k13_val, k_14 = k14_val,
    u_Amax = u_Amax, D = D_default, k_15 = k15_val
  )$N_total
}

baseline_uptake <- run_uptake(depth, k_15, k_13, k_14)

n_pts <- 60
param_ranges <- list(
  depth = seq(0.05, 2.0,  length.out = n_pts),
  k_15  = seq(0.05, 3.0,  length.out = n_pts),
  k_13  = seq(0.5,  20.0, length.out = n_pts),
  k_14  = seq(0.05, 0.95, length.out = n_pts)
)

df_sa <- bind_rows(
  tibble(param = "depth", value = param_ranges$depth,
         uptake = sapply(param_ranges$depth, \(v) run_uptake(v,     k_15, k_13, k_14))),
  tibble(param = "k_15",  value = param_ranges$k_15,
         uptake = sapply(param_ranges$k_15,  \(v) run_uptake(depth, v,    k_13, k_14))),
  tibble(param = "k_13",  value = param_ranges$k_13,
         uptake = sapply(param_ranges$k_13,  \(v) run_uptake(depth, k_15, v,    k_14))),
  tibble(param = "k_14",  value = param_ranges$k_14,
         uptake = sapply(param_ranges$k_14,  \(v) run_uptake(depth, k_15, k_13, v   )))
) %>%
  mutate(param = factor(param, levels = c("depth", "k_15", "k_13", "k_14")))

eps <- 0.01
si_df <- tibble(
  param       = c("depth", "k_15", "k_13", "k_14"),
  default_val = c(depth,   k_15,   k_13,   k_14),
  SI = c(
    (run_uptake(depth*(1+eps), k_15,          k_13,          k_14        ) -
     run_uptake(depth*(1-eps), k_15,          k_13,          k_14        )) / (2*eps*baseline_uptake),
    (run_uptake(depth,          k_15*(1+eps),  k_13,          k_14        ) -
     run_uptake(depth,          k_15*(1-eps),  k_13,          k_14        )) / (2*eps*baseline_uptake),
    (run_uptake(depth,          k_15,          k_13*(1+eps),  k_14        ) -
     run_uptake(depth,          k_15,          k_13*(1-eps),  k_14        )) / (2*eps*baseline_uptake),
    (run_uptake(depth,          k_15,          k_13,          k_14*(1+eps)) -
     run_uptake(depth,          k_15,          k_13,          k_14*(1-eps))) / (2*eps*baseline_uptake)
  )
)

cb_sa <- c(depth = "#E69F00", k_15 = "#56B4E9", k_13 = "#009E73", k_14 = "#CC79A7")

df_spider <- df_sa %>%
  left_join(tibble(param = c("depth","k_15","k_13","k_14"),
                   default_val = c(depth, k_15, k_13, k_14)), by = "param") %>%
  mutate(rel_param  = value / default_val,
         rel_uptake = uptake / baseline_uptake)

p_spider <- ggplot(df_spider, aes(rel_param, rel_uptake, colour = param)) +
  geom_line(linewidth = 1.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_colour_manual(values = cb_sa) +
  labs(title = "Sensitivity spider plot",
       x = "Relative parameter value (\u00d7 baseline)",
       y = "Relative uptake (\u00d7 baseline)") + theme_bw()

p_tornado <- ggplot(si_df, aes(x = param, y = SI, fill = param)) +
  geom_col(width = 0.6, show.legend = FALSE) + coord_flip() +
  scale_fill_manual(values = cb_sa) +
  labs(title = "Normalised sensitivity index", x = NULL, y = "SI") + theme_bw()

baseline_vlines <- tibble(param = c("depth","k_15","k_13","k_14"),
                           xint  = c(depth,   k_15,  k_13,  k_14))

p_sweeps <- ggplot(df_sa, aes(value, uptake, colour = param)) +
  geom_line(linewidth = 1.1, show.legend = FALSE) +
  geom_vline(data = baseline_vlines, aes(xintercept = xint), linetype = "dashed") +
  facet_wrap(~param, scales = "free_x") +
  scale_colour_manual(values = cb_sa) +
  labs(title = "Parameter sweeps", x = "Parameter value",
       y = "N uptake (kg N yr\u207b\u00b9)") + theme_bw()

p_sweeps / (p_spider | p_tornado)
```

---

## Elasticity Analysis

```{r elasticity, fig.height=6}
eps       <- 1e-4
N_vals    <- seq(0.01, 2, length.out = 80)
N_m_fixed <- 0.1
B_m_fixed <- 0.5

params     <- c("depth", "k_15", "k_13", "k_14", "u_Amax")
components <- c("U_root", "U_myco")

run_U <- function(N, depth, k_15, k_13, k_14, u_Amax) {
  out <- plant_uptake(
    N_s = N, N_m = N_m_fixed, crown_area = crown_area_fixed,
    root_length = l, root_no = n,
    B_m = B_m_fixed, B_cr = B_cr_fixed, rho_cr = rho_cr_fixed,
    traits = traits, mycorrhized = mycorrhized, T_m = T_m,
    myco_diameter = myco_diameter, myco_density = myco_density, l_hyp = l_hyp,
    C_r = C_r, q = depth, k_13 = k_13, k_14 = k_14,
    u_Amax = u_Amax, D = D_default, k_15 = k_15
  )
  c(U_total = out$N_total, U_root = out$U_fr, U_myco = out$U_m)
}

elasticity <- function(N, param, component, depth, k_15, k_13, k_14, u_Amax) {
  component <- as.character(component)
  base <- run_U(N, depth, k_15, k_13, k_14, u_Amax)[component]
  pert <- switch(param,
    depth  = run_U(N, depth*(1+eps),  k_15,          k_13,          k_14,          u_Amax        )[component],
    k_15   = run_U(N, depth,           k_15*(1+eps),  k_13,          k_14,          u_Amax        )[component],
    k_13   = run_U(N, depth,           k_15,           k_13*(1+eps),  k_14,          u_Amax        )[component],
    k_14   = run_U(N, depth,           k_15,           k_13,          k_14*(1+eps),  u_Amax        )[component],
    u_Amax = run_U(N, depth,           k_15,           k_13,          k_14,          u_Amax*(1+eps))[component]
  )
  (pert - base) / (eps * base + 1e-12)
}

grid <- expand.grid(N = N_vals, param = params,
                    component = components, KEEP.OUT.ATTRS = FALSE)

df <- grid %>%
  rowwise() %>%
  mutate(SI = elasticity(N, param, component,
                         depth = depth, k_15 = k_15,
                         k_13 = k_13, k_14 = k_14, u_Amax = u_Amax)) %>%
  ungroup() %>%
  mutate(SI_abs    = abs(SI),
         SI_scaled = SI / max(abs(SI), na.rm = TRUE))

importance <- df %>%
  group_by(param, component) %>%
  summarise(mean_abs_SI = mean(abs(SI), na.rm = TRUE), .groups = "drop")

cb_el <- c(depth = "#E69F00", k_15 = "#56B4E9", k_13 = "#009E73",
           k_14 = "#CC79A7", u_Amax = "#D55E00")

param_labels <- c(
  depth  = "Depth",
  k_15   = expression(k[15]~"(half-saturation)"),
  k_13   = expression(k[13]~"(soil scaling)"),
  k_14   = expression(k[14]~"(porosity)"),
  u_Amax = expression(u[max]~"(uptake capacity)")
)

comp_labels <- c(U_root = "Root uptake", U_myco = "Mycorrhizal uptake")

ggplot(df, aes(x = N, y = SI_scaled, colour = param)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey60") +
  geom_line(linewidth = 1) +
  facet_wrap(~ component, labeller = labeller(component = comp_labels)) +
  scale_colour_manual(values = cb_el, labels = param_labels) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(limits = c(-1, 1), oob = scales::squish) +
  labs(x = expression(N[s]~"(kg N m"^{-3}*")"),
       y = "Standardised elasticity", colour = "Parameter") +
  theme_bw(base_size = 12) +
  theme(legend.position = "right", strip.background = element_blank(),
        panel.grid.minor = element_blank())

ggplot(importance, aes(param, mean_abs_SI, fill = param)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ component) +
  scale_fill_manual(values = cb_el) +
  theme_bw() +
  labs(x = NULL, y = "Mean absolute elasticity")
```
