---
title: "Publication Plots"
author: "Joanna"
date: "2026-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(data.table)
library(rphydro)
library(Rprebasso)
library(patchwork)
library(PlantFATE)
library(akima)
library(ggside)
library(future)
library(future.apply)
library(here)
```

```{r theme, include=FALSE}
pub_theme <- theme_classic(base_size = 24) +
  theme(
    axis.title       = element_text(size = 24, face = "bold"),
    axis.text        = element_text(size = 22, colour = "black"),
    axis.line        = element_line(colour = "black", linewidth = 0.6),
    axis.ticks       = element_line(colour = "black", linewidth = 0.5),
    legend.text      = element_text(size = 22),
    legend.key.size  = unit(0.9, "lines"),
    panel.grid.major = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background  = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.margin           = margin(8, 12, 8, 8),
    strip.text            = element_text(size = 24, face = "bold"),
    plot.tag              = element_text(size = 20, face = "bold"),
    legend.title.position = "top",
    legend.title          = element_text(size = 24, face = "bold", hjust = 0.5)
  )
```

## Data reading

```{r}
hyde_all = "~/Documents/Austria/Hyytiala_all_data/"
data <- lapply(paste0(hyde_all, list.files(hyde_all))[1:9], read.csv)
names(data) <- list.files(hyde_all)[1:9]
useful_data <- merge(data[[2]], data[[5]], c('plotID', 'eventID', 'eventYear'), all=TRUE)

```

## Simulations

```{r pressure, echo=FALSE}
# Set to TRUE for fast testing (fewer grid points, shorter run), FALSE for full publication run
testing_mode <- TRUE

plan(multicore, workers = parallel::detectCores() - 1)

# Parameter ranges
n_pts_main  <- if (testing_mode) 15 else 21
n_pts_joint <- if (testing_mode) 6 else 20   # per-axis for 3D co-optimisation sweep

nitrogen_values <- seq(0.01, 5, length.out = n_pts_main)

# Joint co-optimisation grid (coarser to keep run count tractable)
joint_ecto_allo_values    <- seq(0.0,  0.3, length.out = n_pts_joint)
joint_root_no_values      <- 3*10^5 #3*10^seq(3, 6, length.out = n_pts_joint)
joint_root_length_values  <- seq(0.05, 6,   length.out = n_pts_joint)
param_file <- "/home/josimms/Documents/Austria/Plant-FATE/tests/params/p_test_boreal.ini"
weather_file <- "/home/josimms/Documents/Austria/Plant-FATE/data/ERAS_Monthly.csv"
start_year <- 1960
end_year   <- 2022
k_mob_N <- new(LifeHistoryOptimizer, param_file)$traits0$k_mob_N
k_mine_default  <- as.numeric(sub("^k_mine\\s*=\\s*([0-9.e+\\-]+).*", "\\1",
                                   grep("^k_mine\\s*=", readLines(param_file), value = TRUE)[1]))
f_static_default <- as.numeric(sub("^f_static\\s*=\\s*([0-9.e+\\-]+).*", "\\1",
                                    grep("^f_static\\s*=", readLines(param_file), value = TRUE)[1]))

n_pts_sens <- if (testing_mode) 3 else 5
f_static_values <- seq(0.05, 0.5,              length.out = n_pts_sens)
k_mine_values   <- 10^seq(log10(0.3), log10(9), length.out = n_pts_sens)

start_year <- 1960
end_year   <- 2022

# -------------------------------
# Core simulation function with safe error handling
# -------------------------------
run_sim_parallel <- function(rl, rn, n, eb) {
  tryCatch({
    lho <- new(LifeHistoryOptimizer, param_file)
    lho$set_i_metFile(weather_file)
    lho$set_a_metFile(weather_file)
    lho$set_co2File("")
    lho$set_soil_nitrogen(n)
    lho$traits0$investment_from_tree <- eb
    lho$init()
    lho$root_override(rn, rl)

    dt <- 1/12
    start_year <- 1960
    end_year   <- 2022
    years_seq  <- seq(start_year, end_year, dt)
    
    results <- lapply(years_seq, function(t) {
      tryCatch({
        lho$grow_for_dt(t, dt)
        state <- lho$get_state(t + dt)
        df_row <- as.data.frame(t(state))
        names(df_row) <- lho$get_header()
        df_row
      }, error = function(e) {
        setNames(as.data.frame(as.list(rep(NA, length(lho$get_header())))),
                 lho$get_header())
      })
    })
    
    df <- do.call(rbind, results)
    df$date <- seq(as.Date(paste0(start_year, "-01-01")),
                   as.Date(paste0(end_year, "-01-01")),
                   by = "month")[1:nrow(df)]
    df$root_length   <- rl
    df$root_no       <- rn
    df$soil_nitrogen <- n
    df$ecto_allo     <- eb
    df
  }, error = function(e) {
    message("Failed run_sim_parallel: rl=", rl, " rn=", rn, " n=", n, " eb=", eb)
    NULL
  })
}

# -------------------------------
# Helper: safe wrapper for run_sim_parallel
# -------------------------------
run_sim_safe <- function(rl, rn, n, eb) {
  result <- run_sim_parallel(rl, rn, n, eb)
  if (is.null(result) || nrow(result) == 0) return(NULL)
  result
}

pub_fig_dir <- here::here("manuscript/oup-authoring-template/Figures")

pub_cache <- "plots/pub_plots_all_joint.csv"

if (file.exists(pub_cache)) {
  all_joint_df       <- read_csv(pub_cache, show_col_types = FALSE)
  all_joint_df$date  <- as.Date(all_joint_df$date)
  cat("Loaded", nrow(all_joint_df), "rows from cache:", pub_cache, "\n")

} else {
  # -------------------------------
  # Joint co-optimisation: ecto_allo × root_no × root_length × nitrogen
  # Parallelised over ecto_allo (outer axis)
  # -------------------------------
  all_joint_list <- future_lapply(joint_ecto_allo_values, function(eb) {
    inner_rn <- lapply(joint_root_no_values, function(rn) {
      inner_rl <- lapply(joint_root_length_values, function(rl) {
        inner_n <- lapply(nitrogen_values, function(n) {
          run_sim_safe(rl, rn, n, eb)
        })
        do.call(rbind, inner_n[!sapply(inner_n, is.null)])
      })
      do.call(rbind, inner_rl[!sapply(inner_rl, is.null)])
    })
    do.call(rbind, inner_rn[!sapply(inner_rn, is.null)])
  })
  all_joint_df <- do.call(rbind, all_joint_list[!sapply(all_joint_list, is.null)])
  write_csv(all_joint_df, pub_cache)
  cat("Saved", nrow(all_joint_df), "rows to cache:", pub_cache, "\n")
}
```

## Co-optimised trait optima

```{r joint_optimum}
# June–August 2021 growing-season mean per parameter combination.
# Rate variables (N uptake, GPP) are zero in winter (f_temp = 0), so January
# snapshots understate uptake; seasonal means give a fair comparison.
joint_gs <- all_joint_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  summarise(
    height                     = mean(height,                     na.rm = TRUE),
    ectomycorrhiza_mass        = mean(ectomycorrhiza_mass,        na.rm = TRUE),
    ectomycorrhiza_N_biomass   = mean(ectomycorrhiza_N_biomass,   na.rm = TRUE),
    ectomycorrhiza_C_free      = mean(ectomycorrhiza_C_free,      na.rm = TRUE),
    ectomycorrhiza_N_free      = mean(ectomycorrhiza_N_free,      na.rm = TRUE),
    crown_area                 = mean(crown_area,                 na.rm = TRUE),
    lai                        = mean(lai,                        na.rm = TRUE),
    N_bar_roots                = mean(N_bar_roots,                na.rm = TRUE),
    N_static                 = mean(N_static,                 na.rm = TRUE),
    C_export_to_myco           = mean(C_export_to_myco,           na.rm = TRUE),
    mycorrhizal_export_to_tree = mean(mycorrhizal_export_to_tree, na.rm = TRUE),
    root_uptake                = mean(root_uptake,                na.rm = TRUE),
    root_uptake_actual         = mean(root_uptake_actual,         na.rm = TRUE),
    myco_uptake                = mean(myco_uptake,                na.rm = TRUE),
    assim_gross                = mean(assim_gross,                na.rm = TRUE),
    optimal_leaf_nitrogen      = mean(optimal_leaf_nitrogen,      na.rm = TRUE),
    leaf_nitrogen_concentration = mean(leaf_nitrogen_concentration, na.rm = TRUE),
    tree_nitrogen               = mean(tree_nitrogen,              na.rm = TRUE),
    root_mass                   = mean(root_mass,                  na.rm = TRUE),
    coarse_root_mass            = mean(coarse_root_mass,           na.rm = TRUE),
    total_mass                  = mean(total_mass,                 na.rm = TRUE),
    fineroot_lifespan           = mean(fineroot_lifespan,          na.rm = TRUE),
    total_prod                  = mean(total_prod,                 na.rm = TRUE),
    rr                          = mean(rr,                         na.rm = TRUE),
    tr                          = mean(tr,                         na.rm = TRUE),
    ecm_frac_gpp                = mean(C_export_to_myco / pmax(assim_gross, 1e-12), na.rm = TRUE),
    .groups = "drop"
  )

opt_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(
    max_m        = max(height, na.rm = TRUE),
    near_opt     = height >= 0.99 * max_m,
    n_total      = n(),
    n_near_opt   = sum(near_opt),
    stab_frac    = n_near_opt / n_total,
    biomass_gain = max_m / sort(height, decreasing = TRUE)[2] - 1
  ) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup()


# --- Stability bands: all combos within 1% of maximum height per soil N ---
# This shows how flat the fitness landscape is around each optimum
stab_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_m    = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_m) %>%
  filter(near_opt) %>%
  summarise(
    ecto_lo     = min(ecto_allo,     na.rm = TRUE), ecto_hi     = max(ecto_allo,     na.rm = TRUE),
    ecm_gpp_lo  = min(ecm_frac_gpp,  na.rm = TRUE), ecm_gpp_hi  = max(ecm_frac_gpp,  na.rm = TRUE),
    rn_lo       = min(root_no,       na.rm = TRUE), rn_hi       = max(root_no,       na.rm = TRUE),
    rl_lo       = min(root_length,   na.rm = TRUE), rl_hi       = max(root_length,   na.rm = TRUE),
    .groups = "drop"
  )

joint_height_range <- range(opt_joint$height, na.rm = TRUE)

shape_vals <- c("Optimal parameters" = 3, "Model output" = 1, "Calculated" = 18)
ltype_vals <- c("Trend line"   = "dashed")

# Produces a 2-row data frame for a straight trend-line segment
tl <- function(x1, y1, x2, y2) data.frame(x = c(x1, x2), y = c(y1, y2))

colour_vals <- c("Stability (≥99% of max biomass)" = "grey70")

shared_scales <- list(
  scale_shape_manual(values = shape_vals, name = NULL),
  scale_linetype_manual(values = ltype_vals, name = NULL),
  scale_colour_manual(values = colour_vals, name = NULL)
)

# Fallback: ensures soil_nitrogen x-range is never NA when opt_joint is empty
soil_n_expand <- expand_limits(x = range(nitrogen_values))

stab_label <- geom_text(aes(label = paste0(round(stab_frac * 100), "%")),
                        vjust = -0.8, size = 3, inherit.aes = TRUE)

# --- Top row: crosses + stability bars, with dashed trend line ---
p_joint_ecto <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ecto_allo)) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = ecto_lo, ymax = ecto_hi,
                 colour = "Stability (≥99% of max biomass)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  scale_y_continuous(labels = scales::label_percent()) +
  expand_limits(y = c(0, max(joint_ecto_allo_values))) + soil_n_expand +
  labs(tag = "A", x = "Soil nitrogen (g N m⁻³)", y = "Optimal ECM C allocation\n(% NPP)") +
  pub_theme

p_joint_rn <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_no)) +
  geom_line(data = tl(0, 3e6, 2, 1e5), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rn_lo, ymax = rn_hi,
                 colour = "Stability (≥99% of max biomass)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  scale_y_log10(labels = scales::label_scientific(),
                limits = range(joint_root_no_values)) +
  soil_n_expand +
  labs(tag = "B", x = "Soil nitrogen (g N m⁻³)", y = "Optimal root number\nper crown area (no m⁻²)") +
  pub_theme

p_joint_rl <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_length)) +
  geom_line(data = tl(0, 5, 2, 1.2), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rl_lo, ymax = rl_hi,
                 colour = "Stability (≥99% of max biomass)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "C", x = "Soil nitrogen (g N m⁻³)", y = "Optimal root length (m)") +
  pub_theme

r_coarse_root <- 0.04   # coarse root respiration rate [yr^-1], same as rs (sapwood) in p_test_boreal.ini
rr_param      <- 0.4    # fine root respiration rate [yr^-1], rr in p_test_boreal.ini

# --- IQR ribbon across near-optimal combos (≥99% of max height) for panels D–I ---
stab_outputs <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_m    = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_m) %>%
  filter(near_opt) %>%
  summarise(
    ecm_lo          = min(ectomycorrhiza_mass, na.rm = TRUE),
    ecm_hi          = max(ectomycorrhiza_mass, na.rm = TRUE),
    ecm_area_lo     = min(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    ecm_area_hi     = max(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    total_mass_lo   = min(total_mass, na.rm = TRUE),
    total_mass_hi   = max(total_mass, na.rm = TRUE),
    height_lo       = min(height, na.rm = TRUE),
    height_hi       = max(height, na.rm = TRUE),
    nroots_lo     = min(N_bar_roots, na.rm = TRUE),
    nroots_hi     = max(N_bar_roots, na.rm = TRUE),
    nmyco_lo      = min(N_static,  na.rm = TRUE),
    nmyco_hi      = max(N_static,  na.rm = TRUE),
    cn_lo         = min((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) / pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
    cn_hi         = max((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) / pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
    gpp_lo        = min(assim_gross / crown_area, na.rm = TRUE),
    gpp_hi        = max(assim_gross / crown_area, na.rm = TRUE),
    n_upper_lo    = min(leaf_nitrogen_concentration, na.rm = TRUE),
    n_upper_hi    = max(leaf_nitrogen_concentration, na.rm = TRUE),
    n_realised_lo = min(optimal_leaf_nitrogen, na.rm = TRUE),
    n_realised_hi = max(optimal_leaf_nitrogen, na.rm = TRUE),
    n_free_lo     = min(tree_nitrogen, na.rm = TRUE),
    n_free_hi     = max(tree_nitrogen, na.rm = TRUE),
    bga_lo        = min((C_export_to_myco + root_mass * 0.44 / fineroot_lifespan + rr * 0.44 +
                           (r_coarse_root / rr_param) * rr * 0.44 * coarse_root_mass / root_mass) / pmax(assim_gross, 1e-12), na.rm = TRUE),
    bga_hi        = max((C_export_to_myco + root_mass * 0.44 / fineroot_lifespan + rr * 0.44 +
                           (r_coarse_root / rr_param) * rr * 0.44 * coarse_root_mass / root_mass) / pmax(assim_gross, 1e-12), na.rm = TRUE),
    .groups = "drop"
  )

p_joint_total_bga <- ggplot(
    opt_joint,
    aes(x = soil_nitrogen,
        y = (C_export_to_myco + root_mass * 0.44 / fineroot_lifespan + rr * 0.44 +
               (r_coarse_root / rr_param) * rr * 0.44 * coarse_root_mass / root_mass) / pmax(assim_gross, 1e-12))
  ) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = bga_lo, ymax = bga_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Calculated"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_percent()) +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "D", x = "Soil nitrogen (g N m⁻³)",
       y = "Total belowground C\nallocation (% GPP)") +
  pub_theme

# Mycorrhizal total C:N (structural + labile pools) at the optimal combo
opt_joint_cn <- opt_joint %>%
  mutate(CN = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
              pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12))

# --- Middle row: emergent biomass and height ---
p_joint_ecm <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = ecm_lo, ymax = ecm_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific()) +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "E", x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal\nbiomass (kg DM)") +
  pub_theme

p_joint_ecm_area <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = ecm_area_lo, ymax = ecm_area_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific()) +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "F", x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal biomass\nper crown area (kg DM m⁻²)") +
  pub_theme

p_joint_height <- ggplot(opt_joint, aes(x = soil_nitrogen, y = height)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = height_lo, ymax = height_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "I", x = "Soil nitrogen (g N m⁻³)", y = "Height (m)") +
  pub_theme

# --- Bottom row: N depletion and mycorrhizal economics ---
# Dual y-axis: N_bar_roots (left, black) = mineral pool depleted by roots + myco;
# N_static (right, steelblue) = organic pool depleted by myco only (f_myco_pool * N_s).
nroots_range <- range(opt_joint$N_bar_roots, na.rm = TRUE)
nmyco_range  <- range(opt_joint$N_static,  na.rm = TRUE)
# Guard against zero-range (e.g. N starvation makes all uptake values identical)
myco_to_roots <- function(x) {
  dm <- diff(nmyco_range); dr <- diff(nroots_range)
  if (!is.finite(dm) || dm == 0 || !is.finite(dr)) return(x * NA_real_)
  nroots_range[1] + (x - nmyco_range[1]) / dm * dr
}
roots_to_myco <- function(x) {
  dr <- diff(nroots_range); dm <- diff(nmyco_range)
  if (!is.finite(dr) || dr == 0 || !is.finite(dm)) return(x * NA_real_)
  nmyco_range[1] + (x - nroots_range[1]) / dr * dm
}

p_joint_nbar <- ggplot(opt_joint, aes(x = soil_nitrogen, y = N_bar_roots)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = nroots_lo, ymax = nroots_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = myco_to_roots(nmyco_lo), ymax = myco_to_roots(nmyco_hi)),
              fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  geom_point(aes(y = myco_to_roots(N_static)),
             colour = "steelblue", shape = 16, size = 4) +
  shared_scales +
  {
    # Only add a secondary axis when both N_bar series have real variation;
    # a degenerate transformation produces NaN limits that crash patchwork's
    # stats::approx axis-alignment for the right-hand y-axis.
    dr <- diff(nroots_range); dm <- diff(nmyco_range)
    if (is.finite(dr) && dr > 0 && is.finite(dm) && dm > 0)
      scale_y_continuous(
        name     = expression(bar(N)[roots]~"(kg N m"^{-3}*")"),
        sec.axis = sec_axis(~ roots_to_myco(.), name = expression(bar(N)[myco]~"(kg N m"^{-3}*")"))
      )
    else
      scale_y_continuous(name = expression(bar(N)[roots]~"(kg N m"^{-3}*")"))
  } +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "G", x = "Soil nitrogen (g N m⁻³)") +
  pub_theme +
  theme(
    axis.title.y.right = element_text(colour = "steelblue"),
    axis.text.y.right  = element_text(colour = "steelblue"),
    axis.ticks.y.right = element_line(colour = "steelblue"),
    axis.line.y.right  = element_line(colour = "steelblue")
  )

# Mycorrhizal total C:N (structural biomass + labile free pools)
# C (kg C) = ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free
# N (kg N) = ectomycorrhiza_N_biomass + ectomycorrhiza_N_free
p_joint_myco_cn <- ggplot(opt_joint_cn, aes(x = soil_nitrogen, y = CN)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = cn_lo, ymax = cn_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific(digits = 2)) +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "H", x = "Soil nitrogen (g N m⁻³)",
       y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

p_joint_gpp <- ggplot(opt_joint, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = gpp_lo, ymax = gpp_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "J", x = "Soil nitrogen (g N m⁻³)",
       y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

p_joint_n_limit <- ggplot(opt_joint, aes(x = leaf_nitrogen_concentration, y = optimal_leaf_nitrogen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(size = 4, colour = "black", shape = 16) +
  expand_limits(y = 0, x = 0) +
  labs(tag = "K", x = "Actual leaf N\n(kg N kg⁻¹ leaf)",
       y = "Optimal leaf N\n(kg N kg⁻¹ leaf)") +
  pub_theme

# Free nitrogen pool at maturity: if this is near nitrogen_start0 (0.039 kg) even at low soil N,
# the tree is running off its initial endowment rather than reaching a true N steady state.
p_joint_nfree <- ggplot(opt_joint, aes(x = soil_nitrogen, y = tree_nitrogen)) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_free_lo, ymax = n_free_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_hline(yintercept = 0.039, linetype = "dashed", colour = "grey40", linewidth = 0.8) +
  #annotate("text", x = 0.1, y = 0.039, label = "nitrogen_start0", vjust = -0.4,
  #         hjust = 0, size = 3.5, colour = "grey40") +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) + soil_n_expand +
  labs(tag = "L", x = "Soil nitrogen (g N m⁻³)",
       y = "Free N pool at maturity\n(kg N tree⁻¹)") +
  pub_theme

make_row_title <- function(label) {
  ggplot(data.frame(x = 0.5, y = 0.5), aes(x, y)) +
    geom_blank() +
    annotate("text", x = 0.5, y = 0.5, label = label,
             size = 7, fontface = "bold", hjust = 0.5, vjust = 0.5) +
    xlim(0, 1) + ylim(0, 1) +
    theme_void()
}

# 4-column layout: A-D parameters, E-H mycorrhizal & N depletion, I-L growth & N status
# Stability bands (grey bars) on A-C show the range of params achieving ≥99% of max height
joint_optima <-
  guide_area() /
  make_row_title("Parameters to be Optimised") /
  (p_joint_ecto | p_joint_rn | p_joint_rl | p_joint_total_bga) /
  make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
  (p_joint_ecm | p_joint_ecm_area | p_joint_nbar | p_joint_myco_cn) /
  make_row_title("Model results: Tree Status") /
  (p_joint_height | p_joint_gpp | p_joint_n_limit | p_joint_nfree) +
  plot_layout(heights = c(0.15, 0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(theme = theme(plot.margin = margin(5, 30, 5, 5))) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 16),
        legend.key.size = unit(1.0, "lines"),
        legend.position    = "top",
        legend.direction   = "horizontal",
        legend.box         = "horizontal")
joint_optima
```

## Objective function comparison

```{r objective_comparison}
# GPP minus belowground carbon costs per unit leaf area. All flux terms scale with
# A_c * LAI because fine root biomass is proportional to total leaf area (B_fr ∝ n * A_c * LAI).
# Note: only root-related costs are subtracted (not whole-tree autotrophic respiration).
#   rr             = fine root respiration (kg biomass / time unit); * 0.44 = carbon
#   tr             = fine root turnover   (kg biomass / time unit); * 0.44 = carbon
#   C_export_to_myco = C paid to mycorrhiza (kg C / time unit)
# Coarse root respiration excluded: invariant to root_no and ecto_allo.
add_net_gpp_ca <- function(df) {
  df %>% mutate(
    net_gpp_ca = (assim_gross - C_export_to_myco - rr * 0.44 - tr * 0.44) / (crown_area * lai)
  )
}

# Last-year snapshot: 2021 growing-season mean (joint_gs already computed above)
joint_gs_lastyear <- joint_gs %>% add_net_gpp_ca()


# Full 12-panel joint-optima figure for any objective column.
# gs_data must contain all columns from joint_gs plus tr.
make_optima_figure <- function(gs_data, var_col, fig_title, obj_label = NULL) {
  var_sym <- sym(var_col)

  opt <- gs_data %>%
    group_by(soil_nitrogen) %>%
    mutate(
      max_v      = max(!!var_sym, na.rm = TRUE),
      near_opt   = !!var_sym >= 0.99 * max_v,
      n_total    = n(),
      n_near_opt = sum(near_opt),
      stab_frac  = n_near_opt / n_total
    ) %>%
    filter(near_opt) %>%
    slice_max(!!var_sym, n = 1, with_ties = FALSE) %>%
    ungroup()

  stab <- gs_data %>%
    group_by(soil_nitrogen) %>%
    mutate(max_v = max(!!var_sym, na.rm = TRUE), near_opt = !!var_sym >= 0.99 * max_v) %>%
    filter(near_opt) %>%
    summarise(
      ecto_lo = min(ecto_allo,   na.rm = TRUE), ecto_hi = max(ecto_allo,   na.rm = TRUE),
      rn_lo   = min(root_no,     na.rm = TRUE), rn_hi   = max(root_no,     na.rm = TRUE),
      rl_lo   = min(root_length, na.rm = TRUE), rl_hi   = max(root_length, na.rm = TRUE),
      .groups = "drop"
    )

  stab_out <- gs_data %>%
    group_by(soil_nitrogen) %>%
    mutate(max_v = max(!!var_sym, na.rm = TRUE), near_opt = !!var_sym >= 0.99 * max_v) %>%
    filter(near_opt) %>%
    summarise(
      ecm_lo        = min(ectomycorrhiza_mass, na.rm = TRUE),
      ecm_hi        = max(ectomycorrhiza_mass, na.rm = TRUE),
      ecm_area_lo   = min(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
      ecm_area_hi   = max(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
      height_lo     = min(height,     na.rm = TRUE),
      height_hi     = max(height,     na.rm = TRUE),
      nroots_lo     = min(N_bar_roots, na.rm = TRUE),
      nroots_hi     = max(N_bar_roots, na.rm = TRUE),
      nmyco_lo      = min(N_static,  na.rm = TRUE),
      nmyco_hi      = max(N_static,  na.rm = TRUE),
      cn_lo         = min((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                            pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
      cn_hi         = max((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                            pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
      gpp_lo        = min(assim_gross / crown_area, na.rm = TRUE),
      gpp_hi        = max(assim_gross / crown_area, na.rm = TRUE),
      n_upper_lo    = min(leaf_nitrogen_concentration, na.rm = TRUE),
      n_upper_hi    = max(leaf_nitrogen_concentration, na.rm = TRUE),
      n_realised_lo = min(optimal_leaf_nitrogen, na.rm = TRUE),
      n_realised_hi = max(optimal_leaf_nitrogen, na.rm = TRUE),
      n_free_lo     = min(tree_nitrogen, na.rm = TRUE),
      n_free_hi     = max(tree_nitrogen, na.rm = TRUE),
      bga_lo        = min((C_export_to_myco + rr * 0.44 + tr * 0.44) / assim_gross, na.rm = TRUE),
      bga_hi        = max((C_export_to_myco + rr * 0.44 + tr * 0.44) / assim_gross, na.rm = TRUE),
      cn_exch_lo    = min(C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12), na.rm = TRUE),
      cn_exch_hi    = max(C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12), na.rm = TRUE),
      cexport_lo      = min(C_export_to_myco,                                           na.rm = TRUE),
      cexport_hi      = max(C_export_to_myco,                                           na.rm = TRUE),
      cexport_frac_lo = min(C_export_to_myco / assim_gross,                             na.rm = TRUE),
      cexport_frac_hi = max(C_export_to_myco / assim_gross,                             na.rm = TRUE),
      myco_up_lo      = min(myco_uptake,                                                na.rm = TRUE),
      myco_up_hi      = max(myco_uptake,                                                na.rm = TRUE),
      root_up_lo      = min(root_uptake_actual,                                          na.rm = TRUE),
      root_up_hi      = max(root_uptake_actual,                                          na.rm = TRUE),
      n_return_lo     = min(mycorrhizal_export_to_tree,                                 na.rm = TRUE),
      n_return_hi     = max(mycorrhizal_export_to_tree,                                 na.rm = TRUE),
      total_n_lo      = min(myco_uptake + root_uptake_actual,                           na.rm = TRUE),
      total_n_hi      = max(myco_uptake + root_uptake_actual,                           na.rm = TRUE),
      transf_eff_lo   = min(mycorrhizal_export_to_tree / pmax(myco_uptake, 1e-12),      na.rm = TRUE),
      transf_eff_hi   = max(mycorrhizal_export_to_tree / pmax(myco_uptake, 1e-12),      na.rm = TRUE),
      n_deficit_lo    = min(optimal_leaf_nitrogen - leaf_nitrogen_concentration,         na.rm = TRUE),
      n_deficit_hi    = max(optimal_leaf_nitrogen - leaf_nitrogen_concentration,         na.rm = TRUE),
      rr_lo           = min(rr,                                                          na.rm = TRUE),
      rr_hi           = max(rr,                                                          na.rm = TRUE),
      tr_c_lo         = min(tr * 0.44,                                                   na.rm = TRUE),
      tr_c_hi         = max(tr * 0.44,                                                   na.rm = TRUE),
      rr_frac_lo      = min(rr * 0.44 / assim_gross,                                      na.rm = TRUE),
      rr_frac_hi      = max(rr * 0.44 / assim_gross,                                     na.rm = TRUE),
      tr_c_frac_lo    = min(tr * 0.44 / assim_gross,                                     na.rm = TRUE),
      tr_c_frac_hi    = max(tr * 0.44 / assim_gross,                                     na.rm = TRUE),
      all_bg_c_lo     = min((rr * 0.44 + tr * 0.44 + C_export_to_myco) / assim_gross,   na.rm = TRUE),
      all_bg_c_hi     = max((rr * 0.44 + tr * 0.44 + C_export_to_myco) / assim_gross,   na.rm = TRUE),
      ecm_n_free_lo   = min(ectomycorrhiza_N_free,                                        na.rm = TRUE),
      ecm_n_free_hi   = max(ectomycorrhiza_N_free,                                        na.rm = TRUE),
      root_mass_lo    = min(root_mass,                                                    na.rm = TRUE),
      root_mass_hi    = max(root_mass,                                                    na.rm = TRUE),
      frl_lo          = min(fineroot_lifespan,                                            na.rm = TRUE),
      frl_hi          = max(fineroot_lifespan,                                            na.rm = TRUE),
      .groups = "drop"
    )

  opt_cn <- opt %>%
    mutate(CN = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                 pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12))

  opt_fluxes <- opt %>%
    mutate(
      ecm_frac_gpp   = C_export_to_myco / assim_gross,
      total_n_uptake = myco_uptake + root_uptake_actual,
      transfer_eff   = mycorrhizal_export_to_tree / pmax(myco_uptake + ectomycorrhiza_N_free * k_mob_N, 1e-12),
      n_deficit      = optimal_leaf_nitrogen - leaf_nitrogen_concentration,
      tr_c           = tr * 0.44,
      rr_frac        = rr * 0.44 / assim_gross,
      tr_c_frac      = tr * 0.44 / assim_gross,
      all_bg_c_frac  = (rr * 0.44 + tr * 0.44 + C_export_to_myco) / assim_gross,
      # max_N_transfer cap (k_0=0.2, k_1=0.35, k_8=2.7 from p_test_boreal.ini)
      root_diam_mm   = 0.35 / root_length^0.2,
      SA_tip_m2      = pi * (root_diam_mm / 1000) * (root_length / 1000),
      root_SA_m2     = SA_tip_m2 * root_no * crown_area * lai,
      myco_SA_m2     = 4 * ectomycorrhiza_mass / (1200 * 2e-5),
      age_factor_cap = 1 / (1 + exp(fineroot_lifespan - 2.7)),
      max_N_transfer = 0.033 * age_factor_cap * root_SA_m2
    )

  clean_nan  <- function(df) mutate(df, across(where(is.numeric), ~replace(., !is.finite(.), NA_real_)))
  opt        <- clean_nan(opt)
  stab       <- clean_nan(stab)
  stab_out   <- clean_nan(stab_out)
  opt_cn     <- clean_nan(opt_cn)
  opt_fluxes <- clean_nan(opt_fluxes)

  # x-axis fallback: ensures soil_nitrogen range is always non-degenerate even when
  # opt/stab are empty (all heights NaN → no rows after filter(near_opt))
  sn_expand  <- expand_limits(x = range(gs_data$soil_nitrogen, na.rm = TRUE))

  fn_colour <- c("Stability band" = "grey70")
  fn_scales <- list(
    scale_shape_manual(values = shape_vals, name = NULL),
    scale_linetype_manual(values = ltype_vals, name = NULL),
    scale_colour_manual(values = fn_colour, name = NULL)
  )
  stab_lbl <- geom_text(aes(label = paste0(round(stab_frac * 100), "%")),
                        vjust = -0.8, size = 3, inherit.aes = TRUE)

  pA <- ggplot(opt, aes(x = soil_nitrogen, y = ecto_allo)) +
    geom_linerange(data = stab, aes(x = soil_nitrogen, ymin = ecto_lo, ymax = ecto_hi,
                   colour = "Stability band"), linewidth = 2, inherit.aes = FALSE) +
    geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 3, stroke = 1.2) +
    stab_lbl + fn_scales + scale_y_continuous(labels = scales::label_percent()) +
    expand_limits(y = c(0, 1)) +
    labs(tag = "A", x = "Soil nitrogen (kg N / m3)", y = "Optimal ECM C allocation
(% NPP)") +
    pub_theme

  pB <- ggplot(opt, aes(x = soil_nitrogen, y = root_no)) +
    geom_linerange(data = stab, aes(x = soil_nitrogen, ymin = rn_lo, ymax = rn_hi,
                   colour = "Stability band"), linewidth = 2, inherit.aes = FALSE) +
    geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 3, stroke = 1.2) +
    stab_lbl + fn_scales +
    scale_y_log10(labels = scales::label_scientific(),
                  limits = range(gs_data$root_no, na.rm = TRUE)) +
    labs(tag = "B", x = "Soil nitrogen (kg N / m3)", y = "Optimal root number
(no / m2 crown)") +
    pub_theme

  pC <- ggplot(opt, aes(x = soil_nitrogen, y = root_length)) +
    geom_linerange(data = stab, aes(x = soil_nitrogen, ymin = rl_lo, ymax = rl_hi,
                   colour = "Stability band"), linewidth = 2, inherit.aes = FALSE) +
    geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 3, stroke = 1.2) +
    stab_lbl + fn_scales +
    expand_limits(y = range(gs_data$root_length, na.rm = TRUE)) +
    labs(tag = "C", x = "Soil nitrogen (kg N / m3)", y = "Optimal root length (m)") +
    pub_theme

  pD <- ggplot(opt,
      aes(x = soil_nitrogen, y = (C_export_to_myco + rr * 0.44 + tr * 0.44) / pmax(assim_gross, 1e-12))) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = bga_lo, ymax = bga_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Calculated"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_percent()) + expand_limits(y = c(0, 1)) +
    labs(tag = "D", x = "Soil nitrogen (kg N / m3)", y = "N-related belowground C
allocation (% GPP)") +
    pub_theme

  pE <- ggplot(opt, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = ecm_lo, ymax = ecm_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_scientific()) +
    expand_limits(y = c(0, 1e-6)) +
    labs(tag = "E", x = "Soil nitrogen (kg N / m3)", y = "ECM biomass (kg DM)") + pub_theme

  pF <- ggplot(opt, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = ecm_area_lo, ymax = ecm_area_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_scientific()) +
    expand_limits(y = c(0, 1e-6)) +
    labs(tag = "F", x = "Soil nitrogen (kg N / m3)", y = "ECM biomass per crown\n(kg DM m⁻²)") + pub_theme

  nr_rng    <- range(opt$N_bar_roots, na.rm = TRUE)
  nm_rng    <- range(opt$N_static,  na.rm = TRUE)
  has_dual_G <- all(is.finite(c(nr_rng, nm_rng))) && diff(nr_rng) > 0 && diff(nm_rng) > 0
  if (has_dual_G) {
    m2r <- function(x) nr_rng[1] + (x - nm_rng[1]) / diff(nm_rng) * diff(nr_rng)
    r2m <- function(x) nm_rng[1] + (x - nr_rng[1]) / diff(nr_rng) * diff(nm_rng)
  } else {
    m2r <- function(x) rep(NA_real_, length(x))
    r2m <- function(x) rep(NA_real_, length(x))
  }

  pG <- ggplot(opt, aes(x = soil_nitrogen, y = N_bar_roots)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = nroots_lo, ymax = nroots_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_ribbon(data = stab_out,
                aes(x = soil_nitrogen, ymin = m2r(nmyco_lo), ymax = m2r(nmyco_hi)),
                fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    geom_point(aes(y = m2r(N_static)), colour = "steelblue", shape = 16, size = 3) +
    fn_scales +
    {
      if (has_dual_G)
        scale_y_continuous(
          name = "N_bar roots (kg N / m3)",
          sec.axis = sec_axis(~ r2m(.), name = "N_bar myco (kg N / m3)")
        )
      else
        scale_y_continuous(name = "N_bar roots (kg N / m3)")
    } +
    expand_limits(y = c(0, 1e-10)) +
    labs(tag = "G", x = "Soil nitrogen (kg N / m3)") +
    pub_theme +
    {
      if (has_dual_G)
        theme(axis.title.y.right = element_text(colour = "steelblue"),
              axis.text.y.right  = element_text(colour = "steelblue"),
              axis.ticks.y.right = element_line(colour = "steelblue"),
              axis.line.y.right  = element_line(colour = "steelblue"))
      else
        theme()
    }

  cn_hi <- max(opt_cn$CN, na.rm = TRUE)
  if (!is.finite(cn_hi)) cn_hi <- 1

  pH <- ggplot(opt_cn, aes(x = soil_nitrogen, y = CN)) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales +
    scale_y_continuous(labels = scales::label_scientific(digits = 2),
                       limits = c(0, cn_hi * 1.1)) +
    labs(tag = "H", x = "Soil nitrogen (kg N / m3)", y = "Mycorrhizal C:N
(kg C / kg N)") + pub_theme

  pI <- ggplot(opt, aes(x = soil_nitrogen, y = height)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = height_lo, ymax = height_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "I", x = "Soil nitrogen (kg N / m3)", y = "Height (m)") + pub_theme

  pJ <- ggplot(opt, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = gpp_lo, ymax = gpp_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "J", x = "Soil nitrogen (kg N / m3)", y = "GPP per crown area
(kg C / m2 / yr)") + pub_theme

  pK <- ggplot(opt_fluxes, aes(x = leaf_nitrogen_concentration, y = optimal_leaf_nitrogen)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10), x = c(0, 1e-10)) +
    labs(tag = "K", x = "Actual leaf N\n(kg N / kg leaf)", y = "Optimal leaf N\n(kg N / kg leaf)") +
    pub_theme

  pL <- ggplot(opt, aes(x = soil_nitrogen, y = C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12))) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = cn_exch_lo, ymax = cn_exch_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "L", x = "Soil nitrogen (kg N / m3)", y = "C:N exchange cost\n(kg C / kg N received)") + pub_theme

  # ── Row 4: Mycorrhizal C/N Fluxes ──────────────────────────────────────────

  pN <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = ecm_frac_gpp)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = cexport_frac_lo, ymax = cexport_frac_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_percent()) + expand_limits(y = c(0, 1)) +
    labs(tag = "N", x = "Soil nitrogen (kg N / m3)", y = "C to ECM\n(% GPP)") + pub_theme

  pO <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = root_uptake_actual)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = root_up_lo, ymax = root_up_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "O", x = "Soil nitrogen (kg N / m3)", y = "Root N uptake\n(kg N / yr)") + pub_theme

  pO_myco <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = myco_uptake)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = myco_up_lo, ymax = myco_up_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "O", x = "Soil nitrogen (kg N / m3)", y = "ECM N uptake\n(kg N / yr)") + pub_theme

  pP <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = mycorrhizal_export_to_tree)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = n_return_lo, ymax = n_return_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "P", x = "Soil nitrogen (kg N / m3)", y = "N returned to tree\nfrom ECM (kg N / yr)") + pub_theme

  # ── Row 5: N Uptake & Demand ───────────────────────────────────────────────

  pQ <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = total_n_uptake)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = total_n_lo, ymax = total_n_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "Q", x = "Soil nitrogen (kg N / m3)", y = "Total N uptake\n(root + ECM, kg N / yr)") + pub_theme

  pT <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = tree_nitrogen)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = n_free_lo, ymax = n_free_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "T", x = "Soil nitrogen (kg N / m3)", y = "Tree free N pool\n(kg N)") + pub_theme

  # ── Row 6: Root Biomass & Costs ────────────────────────────────────────────

  pU <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = root_mass)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = root_mass_lo, ymax = root_mass_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "U", x = "Soil nitrogen (kg N / m3)", y = "Fine root mass\n(kg)") + pub_theme

  pW <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = rr)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = rr_lo, ymax = rr_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "W", x = "Soil nitrogen (kg N / m3)", y = "Root respiration\n(kg C / yr)") + pub_theme

  pX <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = tr_c)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = tr_c_lo, ymax = tr_c_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "X", x = "Soil nitrogen (kg N / m3)", y = "Root turnover C cost\n(kg C / yr)") + pub_theme

  # ── Row 7: C fractions of GPP ─────────────────────────────────────────────

  pBB <- ggplot(opt_fluxes, aes(x = soil_nitrogen, y = ectomycorrhiza_N_free)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = ecm_n_free_lo, ymax = ecm_n_free_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = c(0, 1e-10)) +
    labs(tag = "BB", x = "Soil nitrogen (kg N / m3)", y = "ECM free N pool\n(kg N — bottleneck if high)") + pub_theme

  guide_area() /
    make_row_title("Parameters to be Optimised") /
    (pA | pB | pC | pD) /
    make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
    (pE | pF | pG | pH) /
    make_row_title("Model results: Tree Status") /
    (pI | pJ | pK | pL) /
    make_row_title("Mycorrhizal C/N Fluxes") /
    (pN | pO | pO_myco | pP) /
    make_row_title("N Uptake & Demand") /
    (pQ | pT) /
    make_row_title("Root Biomass & Costs") /
    (pU | pW | pX) /
    make_row_title("ECM N Pool") /
    (pBB) +
    plot_layout(heights = c(0.15, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1),
                guides = "collect") +
    plot_annotation(
      title    = fig_title,
      subtitle = obj_label,
      theme    = theme(plot.margin   = margin(5, 30, 5, 5),
                       plot.title    = element_text(size = 18, face = "bold"),
                       plot.subtitle = element_text(size = 13, colour = "grey30"))
    ) &
    theme(plot.tag        = element_text(size = 18, face = "bold"),
          legend.text     = element_text(size = 14),
          legend.key.size = unit(0.9, "lines"),
          legend.position    = "top",
          legend.direction   = "horizontal",
          legend.box         = "horizontal") &
    sn_expand
}


```

## Yearly GPP-optimal: data wrangling

```{r annual_gpp_wrangling}
annual_gs_all <- all_joint_df %>%
  mutate(
    year  = as.integer(format(as.Date(date), "%Y")),
    month = as.integer(format(as.Date(date), "%m"))
  ) %>%
  filter(month %in% 6:8) %>%
  group_by(soil_nitrogen, year, ecto_allo, root_no, root_length) %>%
  summarise(
    height                      = mean(height,                      na.rm = TRUE),
    ectomycorrhiza_mass         = mean(ectomycorrhiza_mass,         na.rm = TRUE),
    ectomycorrhiza_N_biomass    = mean(ectomycorrhiza_N_biomass,    na.rm = TRUE),
    ectomycorrhiza_C_free       = mean(ectomycorrhiza_C_free,       na.rm = TRUE),
    ectomycorrhiza_N_free       = mean(ectomycorrhiza_N_free,       na.rm = TRUE),
    crown_area                  = mean(crown_area,                  na.rm = TRUE),
    lai                         = mean(lai,                         na.rm = TRUE),
    N_bar_roots                 = mean(N_bar_roots,                 na.rm = TRUE),
    N_static                  = mean(N_static,                  na.rm = TRUE),
    C_export_to_myco            = mean(C_export_to_myco,            na.rm = TRUE),
    mycorrhizal_export_to_tree  = mean(mycorrhizal_export_to_tree,  na.rm = TRUE),
    root_uptake                 = mean(root_uptake,                 na.rm = TRUE),
    root_uptake_actual          = mean(root_uptake_actual,          na.rm = TRUE),
    myco_uptake                 = mean(myco_uptake,                 na.rm = TRUE),
    assim_gross                 = mean(assim_gross,                 na.rm = TRUE),
    optimal_leaf_nitrogen       = mean(optimal_leaf_nitrogen,       na.rm = TRUE),
    leaf_nitrogen_concentration = mean(leaf_nitrogen_concentration, na.rm = TRUE),
    tree_nitrogen               = mean(tree_nitrogen,               na.rm = TRUE),
    root_mass                   = mean(root_mass,                   na.rm = TRUE),
    coarse_root_mass            = mean(coarse_root_mass,            na.rm = TRUE),
    total_mass                  = mean(total_mass,                  na.rm = TRUE),
    fineroot_lifespan           = mean(fineroot_lifespan,           na.rm = TRUE),
    rr                          = mean(rr,                          na.rm = TRUE),
    tr                          = mean(tr,                          na.rm = TRUE),
    ppfd                        = mean(ppfd,                        na.rm = TRUE),
    belowground_infrastructure  = mean(belowground_infrastructure,  na.rm = TRUE),
    myco_uptake_static          = mean(myco_uptake_static,          na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    net_gpp_ca    = (assim_gross - C_export_to_myco - rr * 0.44 - tr * 0.44) / (crown_area * lai),
    ecm_frac_gpp  = C_export_to_myco / assim_gross,
    myco_N_access = k_mine_default * ectomycorrhiza_mass * N_static,
    accessible_N  = N_bar_roots + N_static,
    mycorrhized   = 0.9
  ) %>%
  arrange(soil_nitrogen, ecto_allo, root_no, root_length, year) %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  mutate(height_increment = height - lag(height)) %>%
  ungroup()
annual_gs_all_full <- annual_gs_all
annual_gs_all <- annual_gs_all %>%
  filter(year >= min(year, na.rm = TRUE) + 5)

# For each year × soil_nitrogen: the parameter combo with the highest GPP minus root costs
annual_opt_gpp <- annual_gs_all %>%
  group_by(soil_nitrogen, year) %>%
  slice_max(net_gpp_ca, n = 1, with_ties = FALSE) %>%
  ungroup()
```

## Joint optima: trajectories over time

```{r joint_optima_trajectories}
# Use actual unique N values from the data; indices match 16c (annual_gpp_traj).
soil_N_unique  <- sort(unique(annual_opt_gpp$soil_nitrogen))
n_low_c  <- soil_N_unique[2]
n_high_c <- soil_N_unique[5]

n_label_low_c  <- paste0("Low N (",  round(n_low_c,  2), " g m⁻³)")
n_label_high_c <- paste0("High N (", round(n_high_c, 2), " g m⁻³)")

# Derive trajectory data from annual_opt_gpp (GPP minus root costs optimisation, consistent with 16a)
# filtered to the two N endpoints so trajectories show how optima shift as the tree ages.
annual_opt <- annual_opt_gpp %>%
  filter(soil_nitrogen %in% c(n_low_c, n_high_c)) %>%
  mutate(
    CN         = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                   pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12),
    gpp_per_ca = assim_gross / crown_area,
    N_label    = factor(
      ifelse(soil_nitrogen == n_low_c, "Low N", "High N"),
      levels = c("Low N", "High N"),
      labels = c(n_label_low_c, n_label_high_c)
    )
  )

traj_colour2 <- scale_colour_manual(
  name   = "Soil N",
  values = c("#2166ac", "#d6604d")
)

# --- Top row: how optimal parameter values shift as the tree grows ---
p_traj_ecto <- ggplot(annual_opt,
    aes(x = year, y = ecto_allo, colour = N_label, group = N_label)) +
  geom_point(size = 2, shape = 3, stroke = 1.5) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(tag = "(a)", x = NULL, y = "Optimal ECM C allocation\n(% NPP)") +
  pub_theme

p_traj_rn <- ggplot(annual_opt,
    aes(x = year, y = root_no, colour = N_label, group = N_label)) +
  geom_point(size = 2, shape = 3, stroke = 1.5) +
  traj_colour2 +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(tag = "(b)", x = NULL, y = "Optimal root number\nper crown area (no m⁻²)") +
  pub_theme

p_traj_rl <- ggplot(annual_opt,
    aes(x = year, y = root_length, colour = N_label, group = N_label)) +
  geom_point(size = 2, shape = 3, stroke = 1.5) +
  traj_colour2 +
  labs(tag = "(c)", x = NULL, y = "Optimal root length (m)") +
  pub_theme

p_traj_myco <- ggplot(annual_opt,
    aes(x = year, y = mycorrhized, colour = N_label, group = N_label)) +
  geom_point(size = 2, shape = 3, stroke = 1.5) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1)) +
  labs(tag = "(d)", x = NULL, y = "ECM colonisation\n(% roots covered)") +
  pub_theme

# --- Middle row: emergent mycorrhizal outputs ---
p_traj_ecm <- ggplot(annual_opt,
    aes(x = year, y = ectomycorrhiza_mass, colour = N_label, group = N_label)) +
  geom_point(size = 1.5) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "(e)", x = NULL, y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_traj_ecm_area <- ggplot(annual_opt,
    aes(x = year, y = ectomycorrhiza_mass / crown_area,
        colour = N_label, group = N_label)) +
  geom_point(size = 1.5) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "(f)", x = NULL, y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_traj_nbar <- ggplot(annual_opt,
    aes(x = year, y = N_bar_roots, colour = N_label, group = N_label)) +
  geom_point(size = 1.5) +
  traj_colour2 +
  labs(tag = "(g)", x = NULL, y = expression(bar(N)[roots]~"(kg N m"^{-3}*")")) +
  pub_theme

p_traj_cn <- ggplot(annual_opt,
    aes(x = year, y = CN, colour = N_label, group = N_label)) +
  geom_point(size = 1.5) +
  traj_colour2 +
  labs(tag = "(h)", x = NULL, y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

# --- Bottom row: tree status ---
p_traj_height <- {
  best_ht <- annual_gs_all %>%
    group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
    summarise(mean_gpp_ca = mean(net_gpp_ca, na.rm = TRUE), .groups = "drop") %>%
    group_by(soil_nitrogen) %>%
    slice_max(mean_gpp_ca, n = 1, with_ties = FALSE) %>%
    select(soil_nitrogen, ecto_allo, root_no, root_length)

  annual_gs_all_full %>%
    semi_join(best_ht, by = c("soil_nitrogen", "ecto_allo", "root_no", "root_length")) %>%
    filter(soil_nitrogen %in% c(n_low_c, n_high_c), !is.na(height_increment)) %>%
    mutate(N_label = factor(
      ifelse(soil_nitrogen == n_low_c, "Low N", "High N"),
      levels = c("Low N", "High N"),
      labels = c(n_label_low_c, n_label_high_c)
    )) %>%
    ggplot(aes(x = year, y = height_increment, colour = N_label, group = N_label)) +
    geom_point(size = 1.5) +
    traj_colour2 +
    labs(tag = "(i)", x = NULL, y = "Height increment\n(m yr⁻¹)") +
    pub_theme
}

p_traj_gpp <- ggplot(annual_opt,
    aes(x = year, y = gpp_per_ca, colour = N_label, group = N_label)) +
  geom_point(size = 1.5) +
  traj_colour2 +
  labs(tag = "(j)", x = NULL, y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

p_traj_n_limit <- ggplot(annual_opt,
    aes(x = year, colour = N_label)) +
  geom_point(aes(y = leaf_nitrogen_concentration), shape = 1, size = 1.5) +
  geom_point(aes(y = optimal_leaf_nitrogen), shape = 16, size = 1.5) +
  traj_colour2 +
  labs(tag = "(k)", x = NULL, y = "Leaf N concentration\n(kg N kg⁻¹ leaf)") +
  pub_theme

p_traj_nfree <- ggplot(annual_opt,
    aes(x = year, y = tree_nitrogen, colour = N_label, group = N_label)) +
  geom_point(size = 1.5) +
  traj_colour2 +
  labs(tag = "(l)", x = NULL, y = "Free N pool\n(kg N tree⁻¹)") +
  pub_theme

# Phantom plot used only for its legend: colour = Soil N, shape = plot type
legend_df_11b <- data.frame(
  x         = 1:4,
  y         = 1:4,
  soil_n    = factor(rep(c(n_label_low_c, n_label_high_c), 2),
                     levels = c(n_label_low_c, n_label_high_c)),
  plot_type = factor(rep(c("Parameters", "Model results"), each = 2),
                     levels = c("Parameters", "Model results"))
)

p_legend_11b <- ggplot(legend_df_11b, aes(x, y, colour = soil_n, shape = plot_type)) +
  geom_point(size = 3, stroke = 1.5) +
  scale_colour_manual(name = "Soil N", values = c("#2166ac", "#d6604d")) +
  scale_shape_manual(name = NULL, values = c("Parameters" = 3, "Model results" = 16)) +
  coord_cartesian(xlim = c(100, 200), ylim = c(100, 200)) +
  theme_void() +
  theme(legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.text      = element_text(size = 22),
        legend.title     = element_text(size = 22, face = "bold"),
        legend.key.size  = unit(1.2, "lines"))

joint_optima_traj <- wrap_plots(
  (make_row_title("Parameters to be Optimised") /
   wrap_plots(p_traj_ecto, p_traj_rn, p_traj_rl, p_traj_myco, ncol = 4) /
   make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
   (p_traj_ecm | p_traj_ecm_area | p_traj_nbar | p_traj_cn) /
   make_row_title("Model results: Tree Status") /
   (p_traj_height | p_traj_gpp | p_traj_n_limit | p_traj_nfree) +
   plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1)) &
   theme(plot.tag        = element_text(size = 22, face = "bold"),
         legend.text     = element_text(size = 22),
         legend.key.size = unit(1.2, "lines"),
         legend.position = "none",
         axis.text.x     = element_text(angle = 45, hjust = 1))),
  p_legend_11b,
  nrow    = 2,
  heights = c(17, 1)
)
joint_optima_traj

ggsave(file.path(pub_fig_dir, "11b_joint_optima_trajectories.png"),
       plot = joint_optima_traj, width = 20, height = 18, dpi = 300)
```

## Yearly GPP-optimal: plot function

```{r make_yearly_optima_figure}
make_yearly_optima_figure <- function(annual_data, x_col, x_lab, fig_title,
                                       obj_label = "Objective: max (GPP - C_myco - R_root*0.44 - T_root*0.44) / crown_area per year") {
  x_sym <- sym(x_col)

  dat <- annual_data %>%
    mutate(
      CN             = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                         pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12),
      ecm_frac_gpp   = C_export_to_myco / assim_gross,
      total_n_uptake = myco_uptake + root_uptake_actual,
      transfer_eff   = mycorrhizal_export_to_tree / pmax(myco_uptake + ectomycorrhiza_N_free * k_mob_N, 1e-12),
      tr_c           = tr * 0.44,
      rr_frac        = rr * 0.44 / assim_gross,
      tr_c_frac      = tr * 0.44 / assim_gross,
      all_bg_c_frac  = (rr * 0.44 + tr * 0.44 + C_export_to_myco) / assim_gross,
      cn_exchange    = C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12),
      bga            = (C_export_to_myco + rr * 0.44 + tr * 0.44) / assim_gross,
      # mycorrhizal CUE: 1 - maintenance / gross_C  (r_myco=0.5, conversion=2.27 from p_test_boreal.ini)
      gross_C_myco   = C_export_to_myco * 2.27 * 0.44,
      maint_myco     = 0.5 * 0.44 * ectomycorrhiza_mass,
      CUE_myco       = pmax(0, pmin(1, 1 - maint_myco / pmax(gross_C_myco, 1e-12))),
      eff_rF         = 1 / pmax(CN, 1e-3),
      # max_N_transfer cap (k_0=0.2, k_1=0.35, k_8=2.7 from p_test_boreal.ini)
      root_diam_mm   = 0.35 / root_length^0.2,
      SA_tip_m2      = pi * (root_diam_mm / 1000) * (root_length / 1000),
      root_SA_m2     = SA_tip_m2 * root_no * crown_area * lai,
      myco_SA_m2     = 4 * ectomycorrhiza_mass / (1200 * 2e-5),
      age_factor_cap = 1 / (1 + exp(fineroot_lifespan - 2.7)),
      max_N_transfer = 0.033 * age_factor_cap * root_SA_m2
    )

  yr_line <- list(scale_colour_viridis_c(name = "Year", option = "viridis",
                                         guide = guide_colorbar(barwidth  = unit(15, "cm"),
                                                                barheight = unit(0.5, "cm"))))
  pt  <- geom_point(aes(colour = year), size = 1.8, shape = 16, alpha = 0.7, na.rm = TRUE)

  viol <- geom_violin(fill = "grey85", colour = "grey50", linewidth = 0.4,
                      scale = "width", na.rm = TRUE)
  jit  <- geom_jitter(aes(colour = year), height = 0.2, size = 1.5,
                      alpha = 0.6, na.rm = TRUE)

  # ── Row 1: Parameters ────────────────────────────────────────────────────────
  pA <- ggplot(dat, aes(x = !!x_sym, y = factor(ecto_allo))) +
    viol + jit + yr_line +
    scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
    labs(tag = "A", x = x_lab, y = "Optimal ECM C allocation\n(% NPP)") + pub_theme

  pB <- ggplot(dat, aes(x = !!x_sym, y = factor(root_no))) +
    viol + jit + yr_line +
    scale_y_discrete(labels = function(x) scales::scientific(as.numeric(x))) +
    labs(tag = "B", x = x_lab, y = "Optimal root number\n(no / m2 crown)") + pub_theme

  pC <- ggplot(dat, aes(x = !!x_sym, y = factor(root_length))) +
    viol + jit + yr_line +
    labs(tag = "C", x = x_lab, y = "Optimal root length (m)") + pub_theme

  pD <- ggplot(dat, aes(x = !!x_sym, y = bga)) +
    pt + yr_line + scale_y_continuous(labels = scales::label_percent()) + expand_limits(y = 0) +
    labs(tag = "D", x = x_lab, y = "N-related belowground C\nallocation (% GPP)") + pub_theme

  # ── Row 2: Mycorrhizal dynamics ───────────────────────────────────────────────
  pE <- ggplot(dat, aes(x = !!x_sym, y = ectomycorrhiza_mass)) +
    pt + yr_line + scale_y_continuous(labels = scales::label_scientific()) +
    labs(tag = "E", x = x_lab, y = "ECM biomass (kg)") + pub_theme

  pF <- ggplot(dat, aes(x = !!x_sym, y = ectomycorrhiza_mass / crown_area)) +
    pt + yr_line + scale_y_continuous(labels = scales::label_scientific()) +
    labs(tag = "F", x = x_lab, y = "ECM biomass per crown\n(kg / m2)") + pub_theme

  # When x is N_bar_roots, show myco_N_access; when x is myco_N_access, show N_bar_roots.
  # When x is anything else, keep the dual-axis N_bar / myco_N_access panel.
  if (x_col == "N_bar_roots") {
    pG <- ggplot(dat, aes(x = !!x_sym, y = myco_N_access)) +
      pt + yr_line +
      labs(tag = "G", x = x_lab,
           y = "N mycorrhiza potential\n(k_mine·ECM·N_s)") + pub_theme
  } else if (x_col == "myco_N_access") {
    pG <- ggplot(dat, aes(x = !!x_sym, y = N_bar_roots)) +
      pt + yr_line +
      labs(tag = "G", x = x_lab,
           y = expression(bar(N)[roots]~"(kg N m"^{-3}*")")) + pub_theme
  } else {
    nr_rng <- range(dat$N_bar_roots,   na.rm = TRUE)
    nm_rng <- range(dat$myco_N_access, na.rm = TRUE)
    m2r    <- function(x) nr_rng[1] + (x - nm_rng[1]) / diff(nm_rng) * diff(nr_rng)
    r2m    <- function(x) nm_rng[1] + (x - nr_rng[1]) / diff(nr_rng) * diff(nm_rng)
    pG <- ggplot(dat, aes(x = !!x_sym, y = N_bar_roots)) +
      pt + yr_line +
      geom_point(aes(y = m2r(myco_N_access)), colour = "steelblue",
                 shape = 16, size = 1.8, na.rm = TRUE) +
      scale_y_continuous(
        name     = expression(bar(N)[roots]~"(kg N m"^{-3}*")"),
        sec.axis = sec_axis(~ r2m(.), name = "N mycorrhiza potential\n(k_mine·ECM·N_s)")
      ) +
      labs(tag = "G", x = x_lab) +
      pub_theme +
      theme(axis.title.y.right = element_text(colour = "steelblue"),
            axis.text.y.right  = element_text(colour = "steelblue"),
            axis.ticks.y.right = element_line(colour = "steelblue"),
            axis.line.y.right  = element_line(colour = "steelblue"))
  }

  pH <- ggplot(dat, aes(x = !!x_sym, y = CN)) +
    pt + yr_line + scale_y_continuous(labels = scales::label_scientific(digits = 2)) +
    labs(tag = "H", x = x_lab, y = "Mycorrhizal C:N\n(kg C / kg N)") + pub_theme

  # ── Row 3: Tree status ────────────────────────────────────────────────────────
  pI <- ggplot(dat, aes(x = !!x_sym, y = height_increment)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "I", x = x_lab, y = "Height increment (m yr⁻¹)") + pub_theme

  pJ <- ggplot(dat, aes(x = !!x_sym, y = assim_gross / crown_area)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "J", x = x_lab, y = "GPP per crown area\n(kg C / m2 / yr)") + pub_theme

  pObj <- ggplot(dat, aes(x = !!x_sym, y = net_gpp_ca)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "J2", x = x_lab, y = "GPP − root costs\n(kg C m⁻² yr⁻¹)") + pub_theme

  pK <- ggplot(dat, aes(x = leaf_nitrogen_concentration, y = optimal_leaf_nitrogen)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    pt + scale_colour_viridis_c(name = "Year", option = "viridis",
                                guide = guide_colorbar(barwidth  = unit(15, "cm"),
                                                       barheight = unit(0.5, "cm"))) +
    expand_limits(y = 0, x = 0) +
    labs(tag = "K", x = "Actual leaf N\n(kg N / kg leaf)", y = "Optimal leaf N\n(kg N / kg leaf)") + pub_theme

  pL <- ggplot(dat, aes(x = !!x_sym, y = cn_exchange)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "L", x = x_lab, y = "C:N exchange cost\n(kg C / kg N received)") + pub_theme

  # ── Row 4: Mycorrhizal C/N fluxes ─────────────────────────────────────────────
  pN <- ggplot(dat, aes(x = !!x_sym, y = ecm_frac_gpp)) +
    pt + yr_line + scale_y_continuous(labels = scales::label_percent()) + expand_limits(y = 0) +
    labs(tag = "N", x = x_lab, y = "C to ECM\n(% GPP)") + pub_theme

  pO <- ggplot(dat, aes(x = !!x_sym, y = root_uptake / root_SA_m2)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "O", x = x_lab, y = "Root N uptake per SA\n(kg N m⁻² yr⁻¹)") + pub_theme

  pO_myco <- ggplot(dat, aes(x = !!x_sym, y = myco_uptake / myco_SA_m2)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "O", x = x_lab, y = "ECM N uptake per SA\n(kg N m⁻² yr⁻¹)") + pub_theme

  pP <- ggplot(dat, aes(x = !!x_sym, y = mycorrhizal_export_to_tree)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "P", x = x_lab, y = "N returned to tree\nfrom ECM (kg N / yr)") + pub_theme

  # ── Row 5: N uptake & demand ──────────────────────────────────────────────────
  pQ <- ggplot(dat, aes(x = !!x_sym, y = total_n_uptake)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "Q", x = x_lab, y = "Total N uptake\n(root + ECM, kg N / yr)") + pub_theme

  pT_p <- ggplot(dat, aes(x = !!x_sym, y = tree_nitrogen)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "T", x = x_lab, y = "Tree free N pool\n(kg N)") + pub_theme

  # ── Row 6: Root biomass & costs ───────────────────────────────────────────────
  pU <- ggplot(dat, aes(x = !!x_sym, y = root_mass)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "U", x = x_lab, y = "Fine root mass\n(kg)") + pub_theme

  pW <- ggplot(dat, aes(x = !!x_sym, y = rr)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "W", x = x_lab, y = "Root respiration\n(kg C / yr)") + pub_theme

  pX <- ggplot(dat, aes(x = !!x_sym, y = tr_c)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "X", x = x_lab, y = "Root turnover C cost\n(kg C / yr)") + pub_theme

  # ── Row 7: ECM free N pool ────────────────────────────────────────────────────
  pBB <- ggplot(dat, aes(x = !!x_sym, y = ectomycorrhiza_N_free)) +
    pt + yr_line + expand_limits(y = 0) +
    labs(tag = "BB", x = x_lab, y = "ECM free N pool\n(kg N)") + pub_theme

  # ── Row 8: N transfer efficiency ──────────────────────────────────────────────
  pPhi_CUE <- ggplot(dat, aes(x = CUE_myco, y = transfer_eff)) +
    pt + yr_line + expand_limits(y = 0, x = 0) +
    labs(tag = "ΦA", x = "Mycorrhizal C-use efficiency (CUE)",
         y = "N transfer efficiency\nφ / (U_IN + U_ON)") + pub_theme

  pPhi_rF <- ggplot(dat, aes(x = eff_rF, y = transfer_eff)) +
    pt + yr_line + expand_limits(y = 0, x = 0) +
    labs(tag = "ΦB", x = expression("Effective fungal N:C, "*italic(r)[F]~"(kg N kg C"^{-1}*")"),
         y = "N transfer efficiency\nφ / (U_IN + U_ON)") + pub_theme

  guide_area() /
    make_row_title("Parameters to be Optimised") /
    (pA | pB | pC | pD) /
    make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
    (pE | pF | pG | pH) /
    make_row_title("Model results: Tree Status") /
    (pI | pJ | pObj | pK | pL) /
    make_row_title("Mycorrhizal C/N Fluxes") /
    (pN | pO | pO_myco | pP) /
    make_row_title("N Uptake & Demand / ECM N Pool") /
    (pT_p | pBB) /
    make_row_title("Root Biomass & Costs") /
    (pU | pW | pX) /
    make_row_title("N Transfer Efficiency") /
    (pPhi_CUE | pPhi_rF) +
    plot_layout(heights = c(0.15, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1, 0.07, 1),
                guides = "collect") +
    plot_annotation(
      title    = fig_title,
      subtitle = paste0(obj_label, "\nLines: loess mean of optimal points (shading = SE)"),
      theme    = theme(plot.margin   = margin(5, 30, 5, 5),
                       plot.title    = element_text(size = 18, face = "bold"),
                       plot.subtitle = element_text(size = 13, colour = "grey30"))
    ) &
    theme(plot.tag           = element_text(size = 18, face = "bold"),
          legend.text        = element_text(size = 14),
          legend.key.size    = unit(0.9, "lines"),
          legend.position    = "top",
          legend.direction   = "horizontal",
          legend.box         = "horizontal")
}
```

## Yearly GPP-optimal: x = N_bar

```{r gpp_xvar_diagnostic}
# Temperature and PPFD are identical across all parameter combinations (same met file),
# so read June-August means directly from the weather CSV and join by year.
met_summer <- read_csv(weather_file, show_col_types = FALSE) %>%
  filter(Month %in% 6:8) %>%
  group_by(year = Year) %>%
  summarise(temp_c = mean(Temp, na.rm = TRUE), .groups = "drop")

# Diagnostic: GPP per crown area vs candidate x-variables to find cleanest relationship
gpp_diag <- annual_opt_gpp %>%
  left_join(met_summer, by = "year") %>%
  mutate(
    gpp_ca         = assim_gross / crown_area,
    root_SA_m2     = pi * (0.35 / root_length^0.2 / 1000) * (root_length / 1000) * root_no * crown_area * lai,
    myco_SA_m2     = 4 * ectomycorrhiza_mass / (1200 * 2e-5),
    root_up_per_SA = root_uptake  / pmax(root_SA_m2,  1e-12),
    myco_up_per_SA = myco_uptake  / pmax(myco_SA_m2,  1e-12),
    total_n_uptake = root_uptake_actual + myco_uptake,
    myco_N_access  = k_mine_default * ectomycorrhiza_mass * N_static,
    accessible_N   = N_bar_roots + N_static
  )

candidates <- list(
  list(col = "accessible_N",              lab = "Accessible N (kg N m⁻³)"),
  list(col = "N_bar_roots",               lab = "N mineral soil (kg N m⁻³)"),
  list(col = "N_static",                   lab = "N organic pool (kg N m⁻³)"),
  list(col = "leaf_nitrogen_concentration",lab = "Actual leaf N (kg N kg⁻¹)"),
  list(col = "optimal_leaf_nitrogen",     lab = "Optimal leaf N (kg N kg⁻¹)"),
  list(col = "total_n_uptake",            lab = "Total N uptake (kg N yr⁻¹)"),
  list(col = "temp_c",                    lab = "Temperature Jun–Aug (°C)"),
  list(col = "ppfd",                      lab = "PPFD Jun–Aug (μmol m⁻² s⁻¹)"),
  list(col = "belowground_infrastructure",lab = "I_b (belowground N factor)")
)

tag_letters <- paste0("(", letters[seq_along(candidates)], ")")
diag_plots <- lapply(seq_along(candidates), function(i) {
  v <- candidates[[i]]
  ggplot(gpp_diag, aes(x = .data[[v$col]], y = gpp_ca, colour = year)) +
    geom_point(size = 1.5, alpha = 0.7, na.rm = TRUE) +
    scale_colour_viridis_c(name = "Year", option = "viridis",
                           guide = guide_colorbar(barwidth  = unit(15, "cm"),
                                                  barheight = unit(0.5, "cm"))) +
    expand_limits(y = 0) +
    labs(tag = tag_letters[i], x = v$lab, y = "GPP / crown area\n(kg C m⁻² yr⁻¹)") +
    pub_theme
})

fig_gpp_xvar <- wrap_plots(diag_plots, ncol = 3, guides = "collect") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
fig_gpp_xvar
ggsave(file.path(pub_fig_dir, "16a_gpp_xvar_diagnostic.png"), plot = fig_gpp_xvar,
       width = 18, height = 18, dpi = 300)
```

```{r annual_gpp_nbar}
fig_annual_nbar <- make_yearly_optima_figure(
  annual_opt_gpp,
  x_col     = "N_bar_roots",
  x_lab     = expression(bar(N)[roots]~"(kg N m"^{-3}*")"),
  fig_title = "Yearly GPP-optimal — x = N̄ roots (effective N at absorbing surface)"
)
fig_annual_nbar
ggsave(file.path(pub_fig_dir, "16a_annual_gpp_nbar.png"), plot = fig_annual_nbar,
       width = 22, height = 42, dpi = 300)

```


## Yearly GPP-optimal: trajectories

```{r annual_gpp_traj}
# Use smaller N indices: index 2 (~0.26 g m⁻³) and index 5 (~1.0 g m⁻³)
n_low_c  <- soil_N_unique[2]
n_high_c <- soil_N_unique[5]

# For each soil_nitrogen, find the single (ecto_allo, root_no, root_length) combo
# that maximises mean net_gpp_ca across all years, then follow that combo's full
# trajectory — this gives a physically consistent path (no year-to-year jumps).
best_combo_lifetime <- annual_gs_all_full %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  summarise(mean_gpp_ca = mean(net_gpp_ca, na.rm = TRUE), .groups = "drop") %>%
  group_by(soil_nitrogen) %>%
  slice_max(mean_gpp_ca, n = 1, with_ties = FALSE) %>%
  select(soil_nitrogen, ecto_allo, root_no, root_length)

traj_gpp <- annual_gs_all_full %>%
  semi_join(best_combo_lifetime,
            by = c("soil_nitrogen", "ecto_allo", "root_no", "root_length")) %>%
  filter(soil_nitrogen %in% c(n_low_c, n_high_c)) %>%
  mutate(
    CN           = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                     pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12),
    gpp_per_ca   = assim_gross / crown_area,
    ecm_frac_gpp = C_export_to_myco / assim_gross,
    N_label      = factor(
      ifelse(soil_nitrogen == n_low_c, "Low N", "High N"),
      levels = c("Low N", "High N"),
      labels = c(paste0("Low N (",  round(n_low_c,  2), " g m⁻³)"),
                 paste0("High N (", round(n_high_c, 2), " g m⁻³)"))
    )
  )

traj_colour_gpp <- scale_colour_manual(
  name   = "Soil N",
  values = c("#2166ac", "#d6604d"),
  guide  = guide_legend(override.aes = list(linetype = "solid", linewidth = 0.8))
)

# Row 1: lifetime-optimal parameter values (flat lines — fixed for whole trajectory)
p_tg_ecto <- ggplot(traj_gpp, aes(x = year, y = ecto_allo, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp + scale_y_continuous(labels = scales::label_percent()) +
  labs(x = NULL, y = "ECM C allocation\n(% NPP)", tag = "(a)") + pub_theme

p_tg_rn <- ggplot(traj_gpp, aes(x = year, y = root_no, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp + scale_y_log10(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Root number\nper crown area (no m⁻²)", tag = "(b)") + pub_theme

p_tg_rl <- ggplot(traj_gpp, aes(x = year, y = root_length, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = "Root length (m)", tag = "(c)") + pub_theme

p_tg_myco <- ggplot(traj_gpp, aes(x = year, y = mycorrhized, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1)) +
  labs(x = NULL, y = "ECM colonisation\n(% roots covered)", tag = "(d)") + pub_theme

# Row 2: mycorrhizal outputs at the yearly optimum
p_tg_ecm <- ggplot(traj_gpp,
    aes(x = year, y = ectomycorrhiza_mass, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp + scale_y_continuous(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Ectomycorrhizal\nbiomass (kg)", tag = "(e)") + pub_theme

p_tg_ecm_area <- ggplot(traj_gpp,
    aes(x = year, y = ectomycorrhiza_mass / crown_area, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp + scale_y_continuous(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)", tag = "(f)") + pub_theme

p_tg_nbar <- ggplot(traj_gpp, aes(x = year, y = N_bar_roots, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = expression(bar(N)[roots]~"(kg N m"^{-3}*")"), tag = "(g)") + pub_theme

p_tg_cn <- ggplot(traj_gpp, aes(x = year, y = CN, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)", tag = "(h)") + pub_theme

# Row 3: tree status at the yearly optimum
p_tg_height <- ggplot(traj_gpp, aes(x = year, y = height, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = "Height (m)", tag = "(i)") + pub_theme

p_tg_gpp <- ggplot(traj_gpp, aes(x = year, y = gpp_per_ca, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = "GPP per crown area\n(kg C m⁻² yr⁻¹)", tag = "(j)") + pub_theme

p_tg_nlim <- ggplot(traj_gpp, aes(x = year, colour = N_label, group = N_label)) +
  geom_line(aes(y = leaf_nitrogen_concentration), linetype = "dashed", linewidth = 0.8) +
  geom_line(aes(y = optimal_leaf_nitrogen),       linetype = "solid",  linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = "Leaf N concentration\n(kg N kg⁻¹ leaf)", tag = "(k)") + pub_theme

p_tg_nfree <- ggplot(traj_gpp, aes(x = year, y = tree_nitrogen, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  labs(x = NULL, y = "Free N pool\n(kg N tree⁻¹)", tag = "(l)") + pub_theme

# Phantom for colour legend
p_leg_col_gpp <- ggplot(traj_gpp, aes(x = year, y = ecto_allo, colour = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour_gpp +
  coord_cartesian(xlim = c(3000, 4000), ylim = c(100, 200)) +
  theme_void() +
  theme(legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.text      = element_text(size = 22),
        legend.title     = element_text(size = 22, face = "bold"),
        legend.key.size  = unit(1.2, "lines"))

# Phantom for linetype legend (panel k: dashed = actual leaf N, solid = optimal leaf N)
p_leg_lty_gpp <- ggplot(
  data.frame(x = 1:2, y = 1:2,
             type = factor(c("Actual leaf N", "Optimal leaf N"),
                           levels = c("Actual leaf N", "Optimal leaf N"))),
  aes(x = x, y = y, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_linetype_manual(name = NULL,
                        values = c("Actual leaf N" = "dashed", "Optimal leaf N" = "solid")) +
  coord_cartesian(xlim = c(100, 200), ylim = c(100, 200)) +
  theme_void() +
  theme(legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.text      = element_text(size = 22),
        legend.key.size  = unit(1.2, "lines"))

inner_gpp_traj <-
  make_row_title("Lifetime-optimal parameters (max mean GPP minus root costs across all years)") /
  (p_tg_ecto | p_tg_rn | p_tg_rl | p_tg_myco) /
  make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
  (p_tg_ecm | p_tg_ecm_area | p_tg_nbar | p_tg_cn) /
  make_row_title("Model results: Tree Status") /
  (p_tg_height | p_tg_gpp | p_tg_nlim | p_tg_nfree) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1)) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 22),
        legend.key.size = unit(1.2, "lines"),
        legend.position = "none",
        axis.text.x     = element_text(angle = 45, hjust = 1))

joint_gpp_traj <- wrap_plots(
  inner_gpp_traj,
  (p_leg_col_gpp | p_leg_lty_gpp),
  nrow    = 2,
  heights = c(17, 1)
)
joint_gpp_traj

ggsave(file.path(pub_fig_dir, "16c_annual_gpp_traj.png"),
       plot = joint_gpp_traj, width = 20, height = 18, dpi = 300)
```

## Stability of the optimum

```{r stability_metrics}
# Compute per (soil_nitrogen, year) stability metrics from the full parameter grid.
stab_metrics <- annual_gs_all %>%
  group_by(soil_nitrogen, year) %>%
  summarise(
    n_total      = n(),
    max_obj      = max(net_gpp_ca,  na.rm = TRUE),
    min_obj      = min(net_gpp_ca,  na.rm = TRUE),
    mean_obj     = mean(net_gpp_ca, na.rm = TRUE),
    sd_obj       = sd(net_gpp_ca,   na.rm = TRUE),
    # Flatness: what fraction of combos sits within X% of the maximum?
    stab_frac_1  = sum(net_gpp_ca >= 0.99 * max(net_gpp_ca, na.rm = TRUE), na.rm = TRUE) / n(),
    stab_frac_5  = sum(net_gpp_ca >= 0.95 * max(net_gpp_ca, na.rm = TRUE), na.rm = TRUE) / n(),
    stab_frac_10 = sum(net_gpp_ca >= 0.90 * max(net_gpp_ca, na.rm = TRUE), na.rm = TRUE) / n(),
    # Parameter IQR among 5%-near-optimal combos — wide IQR = insensitive parameter
    iqr_ecto     = IQR(ecto_allo[net_gpp_ca >= 0.95 * max(net_gpp_ca, na.rm = TRUE)],   na.rm = TRUE),
    iqr_rn       = IQR(root_no[net_gpp_ca >= 0.95 * max(net_gpp_ca, na.rm = TRUE)],     na.rm = TRUE),
    iqr_rl           = IQR(root_length[net_gpp_ca >= 0.95 * max(net_gpp_ca, na.rm = TRUE)], na.rm = TRUE),
    iqr_mycorrhized  = 0,
    .groups = "drop"
  ) %>%
  mutate(
    # Curvature: how much does the objective vary across the full grid?
    cv_obj    = sd_obj / pmax(mean_obj, 1e-12),
    obj_range = (max_obj - min_obj) / pmax(max_obj, 1e-12)
  )

# Year-to-year optimal trajectories (already one row per year × soil N)
yy_opt <- annual_opt_gpp %>%
  filter(year >= start_year)

n_col <- scale_colour_viridis_c(name = "Soil N\n(kg m⁻³)", option = "viridis")
pt_n  <- geom_point(size = 1.8, shape = 16, alpha = 0.7, na.rm = TRUE)

# --- Row 1: Flatness ---
pStabA <- ggplot(stab_metrics, aes(x = year, y = stab_frac_5, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(a)", x = "Year",
       y = "Flatness (5% threshold)\nfraction of combos near-optimal") + pub_theme

pStabB <- ggplot(stab_metrics, aes(x = year, y = stab_frac_1, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(b)", x = "Year",
       y = "Flatness (1% threshold)\nfraction of combos near-optimal") + pub_theme

# --- Row 2: Curvature (sharpness of the optimum) ---
pStabC <- ggplot(stab_metrics, aes(x = year, y = cv_obj, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(c)", x = "Year",
       y = "Curvature: CV of GPP − root costs\nacross full parameter grid") + pub_theme

pStabD <- ggplot(stab_metrics, aes(x = year, y = obj_range, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(d)", x = "Year",
       y = "Objective range\n(max − min) / max") + pub_theme

# --- Row 3: Parameter spread among near-optimal (5%) combos ---
pStabE <- ggplot(stab_metrics, aes(x = year, y = iqr_ecto, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(e)", x = "Year",
       y = "IQR ECM C allocation\namong 5%-near-optimal") + pub_theme

pStabF <- ggplot(stab_metrics, aes(x = year, y = iqr_rn, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "(f)", x = "Year",
       y = "IQR root number\namong 5%-near-optimal") + pub_theme

pStabG <- ggplot(stab_metrics, aes(x = year, y = iqr_rl, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(g)", x = "Year",
       y = "IQR root length (m)\namong 5%-near-optimal") + pub_theme

pStabG2 <- ggplot(stab_metrics, aes(x = year, y = iqr_mycorrhized, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(h)", x = "Year",
       y = "IQR ECM colonisation (%)\namong 5%-near-optimal") + pub_theme

# --- Row 4: Year-to-year consistency of optimal parameters ---
pStabH <- ggplot(yy_opt, aes(x = year, y = ecto_allo, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(i)", x = "Year",
       y = "Optimal ECM C allocation\n(year-to-year consistency)") + pub_theme

pStabI <- ggplot(yy_opt, aes(x = year, y = root_no, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "(j)", x = "Year",
       y = "Optimal root number\n(year-to-year consistency)") + pub_theme

pStabJ <- ggplot(yy_opt, aes(x = year, y = root_length, colour = soil_nitrogen)) +
  pt_n + n_col + expand_limits(y = 0) +
  labs(tag = "(k)", x = "Year",
       y = "Optimal root length (m)\n(year-to-year consistency)") + pub_theme

pStabJ2 <- ggplot(yy_opt, aes(x = year, y = mycorrhized, colour = soil_nitrogen)) +
  pt_n + n_col +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1)) +
  labs(tag = "(l)", x = "Year",
       y = "ECM colonisation (%)\n(year-to-year consistency)") + pub_theme

fig_stability <-
  make_row_title("Flatness: fraction of parameter grid within X% of maximum") /
  (pStabA | pStabB) /
  make_row_title("Curvature: how sharply GPP minus root costs drops away from the optimum") /
  (pStabC | pStabD) /
  make_row_title("Parameter spread: IQR of each parameter among 5%-near-optimal combos") /
  (pStabE | pStabF | pStabG | pStabG2) /
  make_row_title("Year-to-year consistency: which parameter combo wins each year") /
  (pStabH | pStabI | pStabJ | pStabJ2) +
  plot_layout(
    heights = c(0.06, 1, 0.06, 1, 0.06, 1, 0.06, 1),
    guides  = "collect"
  ) &
  theme(
    plot.tag        = element_text(size = 18, face = "bold"),
    legend.text     = element_text(size = 14),
    legend.key.size = unit(1.2, "lines"),
    legend.position = "right"
  )
fig_stability

ggsave(file.path(pub_fig_dir, "16e_stability_metrics.png"), plot = fig_stability,
       width = 20, height = 28, dpi = 300)
```

## Stability metrics: N axis

```{r stability_metrics_N}
yr_col <- scale_colour_viridis_c(name = "Year", option = "viridis")

pStabNA <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = stab_frac_5, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(a)", x = "Soil N (g m⁻³)",
       y = "Flatness (5% threshold)\nfraction of combos near-optimal") + pub_theme

pStabNB <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = stab_frac_1, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(b)", x = "Soil N (g m⁻³)",
       y = "Flatness (1% threshold)\nfraction of combos near-optimal") + pub_theme

pStabNC <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = cv_obj, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(c)", x = "Soil N (g m⁻³)",
       y = "Curvature: CV of GPP − root costs\nacross full parameter grid") + pub_theme

pStabND <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = obj_range, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(d)", x = "Soil N (g m⁻³)",
       y = "Objective range\n(max − min) / max") + pub_theme

pStabNE <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = iqr_ecto, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(e)", x = "Soil N (g m⁻³)",
       y = "IQR ECM C allocation\namong 5%-near-optimal") + pub_theme

pStabNF <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = iqr_rn, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "(f)", x = "Soil N (g m⁻³)",
       y = "IQR root number\namong 5%-near-optimal") + pub_theme

pStabNG <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = iqr_rl, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(g)", x = "Soil N (g m⁻³)",
       y = "IQR root length (m)\namong 5%-near-optimal") + pub_theme

pStabNG2 <- ggplot(stab_metrics, aes(x = soil_nitrogen, y = iqr_mycorrhized, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(h)", x = "Soil N (g m⁻³)",
       y = "IQR ECM colonisation (%)\namong 5%-near-optimal") + pub_theme

pStabNH <- ggplot(yy_opt, aes(x = soil_nitrogen, y = ecto_allo, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(i)", x = "Soil N (g m⁻³)",
       y = "Optimal ECM C allocation\n(year-to-year consistency)") + pub_theme

pStabNI <- ggplot(yy_opt, aes(x = soil_nitrogen, y = root_no, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "(j)", x = "Soil N (g m⁻³)",
       y = "Optimal root number\n(year-to-year consistency)") + pub_theme

pStabNJ <- ggplot(yy_opt, aes(x = soil_nitrogen, y = root_length, colour = year)) +
  pt_n + yr_col + expand_limits(y = 0) +
  labs(tag = "(k)", x = "Soil N (g m⁻³)",
       y = "Optimal root length (m)\n(year-to-year consistency)") + pub_theme

pStabNJ2 <- ggplot(yy_opt, aes(x = soil_nitrogen, y = mycorrhized, colour = year)) +
  pt_n + yr_col +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1)) +
  labs(tag = "(l)", x = "Soil N (g m⁻³)",
       y = "ECM colonisation (%)\n(year-to-year consistency)") + pub_theme

fig_stability_N <-
  make_row_title("Flatness: fraction of parameter grid within X% of maximum") /
  (pStabNA | pStabNB) /
  make_row_title("Curvature: how sharply GPP minus root costs drops away from the optimum") /
  (pStabNC | pStabND) /
  make_row_title("Parameter spread: IQR of each parameter among 5%-near-optimal combos") /
  (pStabNE | pStabNF | pStabNG | pStabNG2) /
  make_row_title("Year-to-year consistency: which parameter combo wins each year") /
  (pStabNH | pStabNI | pStabNJ | pStabNJ2) +
  plot_layout(
    heights = c(0.06, 1, 0.06, 1, 0.06, 1, 0.06, 1),
    guides  = "collect"
  ) &
  theme(
    plot.tag        = element_text(size = 18, face = "bold"),
    legend.text     = element_text(size = 14),
    legend.key.size = unit(1.2, "lines"),
    legend.position = "right"
  )
fig_stability_N

ggsave(file.path(pub_fig_dir, "16f_stability_metrics_N.png"), plot = fig_stability_N,
       width = 20, height = 28, dpi = 300)
```

## Optimal combination and model state

```{r plot17_optimal_state}
dat17 <- annual_opt_gpp %>%
  mutate(
    bga         = (C_export_to_myco + rr * 0.44 + tr * 0.44) / pmax(assim_gross, 1e-12),
    mycorrhized = 0.9
  )

yr_scale17 <- scale_colour_viridis_c(
  name  = "Year", option = "viridis",
  guide = guide_colorbar(barwidth = unit(15, "cm"), barheight = unit(0.5, "cm"))
)
pt17      <- geom_point(aes(colour = year), size = 1.8, shape = 16, alpha = 0.7, na.rm = TRUE)
viol17    <- geom_violin(fill = "grey85", colour = "grey50", linewidth = 0.4, scale = "width", na.rm = TRUE)
jit_cross <- geom_jitter(aes(colour = year), height = 0.2, size = 1.5, shape = 3,
                         stroke = 1.2, alpha = 0.8, na.rm = TRUE)

dat17_mean <- dat17 %>%
  group_by(soil_nitrogen) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop")
ln17 <- geom_line(data = dat17_mean, aes(linetype = "last", group = 1),
                  colour = "black", linewidth = 0.5, na.rm = TRUE)
lt17 <- scale_linetype_manual(values = c(last = "solid"),
                               labels = c(last = "Mean across years"), name = NULL)

# Row 1: optimised parameters (cross shape)
p17A <- ggplot(dat17, aes(x = soil_nitrogen, y = factor(ecto_allo))) +
  viol17 + jit_cross + yr_scale17 +
  scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
  labs(tag = "(a)", x = "Soil N\n(kg N m⁻³)", y = "Optimal ECM C allocation\n(% NPP)") + pub_theme

p17B <- ggplot(dat17, aes(x = soil_nitrogen, y = factor(root_no))) +
  viol17 + jit_cross + yr_scale17 +
  scale_y_discrete(labels = function(x) scales::scientific(as.numeric(x))) +
  labs(tag = "(b)", x = "Soil N\n(kg N m⁻³)", y = "Optimal root number\n(no m⁻² crown)") + pub_theme

p17C <- ggplot(dat17, aes(x = soil_nitrogen, y = factor(root_length))) +
  viol17 + jit_cross + yr_scale17 +
  labs(tag = "(c)", x = "Soil N\n(kg N m⁻³)", y = "Optimal root length\n(m)") + pub_theme

p17C2 <- ggplot(dat17, aes(x = soil_nitrogen, y = factor(mycorrhized))) +
  viol17 + jit_cross + yr_scale17 +
  scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
  labs(tag = "(d)", x = "Soil N\n(kg N m⁻³)", y = "ECM colonisation\n(% roots covered)") + pub_theme

# Row 2: ECM dynamics and N uptake
p17E <- ggplot(dat17, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  pt17 + yr_scale17 + ln17 + lt17 +
  scale_y_continuous(labels = scales::label_scientific()) + expand_limits(y = 0) +
  labs(tag = "(e)", x = "Soil N\n(kg N m⁻³)", y = "ECM biomass\n(kg)") + pub_theme

p17F <- ggplot(dat17, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  pt17 + yr_scale17 + ln17 + lt17 +
  scale_y_continuous(labels = scales::label_scientific()) + expand_limits(y = 0) +
  labs(tag = "(f)", x = "Soil N\n(kg N m⁻³)", y = "ECM biomass per crown\n(kg m⁻²)") + pub_theme

p17O <- ggplot(dat17, aes(x = soil_nitrogen, y = root_uptake_actual)) +
  pt17 + yr_scale17 + ln17 + lt17 + expand_limits(y = 0) +
  labs(tag = "(g)", x = "Soil N\n(kg N m⁻³)", y = "Root N uptake\n(kg N yr⁻¹)") + pub_theme

p17O_myco <- ggplot(dat17, aes(x = soil_nitrogen, y = myco_uptake)) +
  pt17 + yr_scale17 + ln17 + lt17 + expand_limits(y = 0) +
  labs(tag = "(h)", x = "Soil N\n(kg N m⁻³)", y = "ECM N uptake\n(kg N yr⁻¹)") + pub_theme

# Row 3: tree status
p17I <- ggplot(dat17, aes(x = soil_nitrogen, y = height_increment)) +
  pt17 + yr_scale17 + ln17 + lt17 + expand_limits(y = 0) +
  labs(tag = "(i)", x = "Soil N\n(kg N m⁻³)", y = "Height increment\n(m yr⁻¹)") + pub_theme

p17J <- ggplot(dat17, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  pt17 + yr_scale17 + ln17 + lt17 + expand_limits(y = 0) +
  labs(tag = "(j)", x = "Soil N\n(kg N m⁻³)", y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") + pub_theme

p17K <- ggplot(dat17, aes(x = leaf_nitrogen_concentration, y = optimal_leaf_nitrogen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  pt17 + yr_scale17 + expand_limits(y = 0, x = 0) +
  labs(tag = "(l)", x = "Actual leaf N\n(kg N kg⁻¹ leaf)", y = "Optimal leaf N\n(kg N kg⁻¹ leaf)") + pub_theme +
  theme(plot.background = element_rect(fill = "grey92", colour = NA))

p17BB <- ggplot(dat17, aes(x = soil_nitrogen, y = ectomycorrhiza_N_free / pmax(ectomycorrhiza_mass, 1e-12))) +
  pt17 + yr_scale17 + ln17 + lt17 + expand_limits(y = 0) +
  labs(tag = "(k)", x = "Soil N\n(kg N m⁻³)", y = "ECM free N per biomass\n(kg N kg⁻¹)") + pub_theme

fig17 <- make_row_title("Parameters to be Optimised") /
  (p17A | p17B | p17C | p17C2) /
  make_row_title("ECM Dynamics & N Uptake") /
  (p17E | p17F | p17O | p17O_myco) /
  make_row_title("Tree Status") /
  (p17I | p17J | p17BB | p17K) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") &
  theme(plot.tag         = element_text(size = 22, face = "bold"),
        legend.text      = element_text(size = 22),
        legend.key.size  = unit(0.9, "lines"),
        legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.box       = "horizontal")
ggsave(file.path(pub_fig_dir, "17_optimal_state.png"), plot = fig17, width = 22, height = 24, dpi = 300)
```

## Alternative figure 4: lifetime optimal and last 10 years

```{r plot17_alternatives}
# ── Lifetime optimal ──────────────────────────────────────────────────────────
# For each soil_N the single (ecto_allo, root_no, root_length) combo that
# maximises mean net_gpp_ca across all years (best_combo_lifetime, already
# computed).  State variables are shown year-by-year for that fixed combo.
dat17_lifetime <- annual_gs_all %>%
  semi_join(best_combo_lifetime,
            by = c("soil_nitrogen", "ecto_allo", "root_no", "root_length")) %>%
  mutate(
    bga         = (C_export_to_myco + rr * 0.44 + tr * 0.44) / pmax(assim_gross, 1e-12),
    mycorrhized = 0.9
  )

dat17_lt_mean <- dat17_lifetime %>%
  group_by(soil_nitrogen) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop")

ln17_lt <- geom_line(data = dat17_lt_mean,
                     aes(linetype = "last", group = 1),
                     colour = "black", linewidth = 0.5, na.rm = TRUE)

# Row 1: parameters are fixed per soil_N — show as single points, no violin
p17A_lt <- ggplot(best_combo_lifetime,
                  aes(x = soil_nitrogen, y = factor(ecto_allo))) +
  geom_point(size = 3, shape = 16) +
  scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
  labs(tag = "(a)", x = "Soil N\n(kg N m⁻³)", y = "Optimal ECM C allocation\n(% NPP)") + pub_theme

p17B_lt <- ggplot(best_combo_lifetime,
                  aes(x = soil_nitrogen, y = factor(root_no))) +
  geom_point(size = 3, shape = 16) +
  scale_y_discrete(labels = function(x) scales::scientific(as.numeric(x))) +
  labs(tag = "(b)", x = "Soil N\n(kg N m⁻³)", y = "Optimal root number\n(no m⁻² crown)") + pub_theme

p17C_lt <- ggplot(best_combo_lifetime,
                  aes(x = soil_nitrogen, y = factor(root_length))) +
  geom_point(size = 3, shape = 16) +
  labs(tag = "(c)", x = "Soil N\n(kg N m⁻³)", y = "Optimal root length\n(mm)") + pub_theme

p17C2_lt <- ggplot(best_combo_lifetime,
                   aes(x = soil_nitrogen, y = factor(0.9))) +
  geom_point(size = 3, shape = 16) +
  scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
  labs(tag = "(d)", x = "Soil N\n(kg N m⁻³)", y = "ECM colonisation\n(% roots covered)") + pub_theme

# Rows 2–3: state variables for the fixed combo, coloured by year
p17E_lt <- ggplot(dat17_lifetime,
                  aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  pt17 + yr_scale17 + ln17_lt + lt17 +
  scale_y_continuous(labels = scales::label_scientific()) + expand_limits(y = 0) +
  labs(tag = "(e)", x = "Soil N\n(kg N m⁻³)", y = "ECM biomass\n(kg)") + pub_theme

p17F_lt <- ggplot(dat17_lifetime,
                  aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  pt17 + yr_scale17 + ln17_lt + lt17 +
  scale_y_continuous(labels = scales::label_scientific()) + expand_limits(y = 0) +
  labs(tag = "(f)", x = "Soil N\n(kg N m⁻³)", y = "ECM biomass per crown\n(kg m⁻²)") + pub_theme

p17O_lt <- ggplot(dat17_lifetime,
                  aes(x = soil_nitrogen, y = root_uptake_actual)) +
  pt17 + yr_scale17 + ln17_lt + lt17 + expand_limits(y = 0) +
  labs(tag = "(g)", x = "Soil N\n(kg N m⁻³)", y = "Root N uptake\n(kg N yr⁻¹)") + pub_theme

p17O_myco_lt <- ggplot(dat17_lifetime,
                       aes(x = soil_nitrogen, y = myco_uptake)) +
  pt17 + yr_scale17 + ln17_lt + lt17 + expand_limits(y = 0) +
  labs(tag = "(h)", x = "Soil N\n(kg N m⁻³)", y = "ECM N uptake\n(kg N yr⁻¹)") + pub_theme

p17I_lt <- ggplot(dat17_lifetime,
                  aes(x = soil_nitrogen, y = height_increment)) +
  pt17 + yr_scale17 + ln17_lt + lt17 + expand_limits(y = 0) +
  labs(tag = "(i)", x = "Soil N\n(kg N m⁻³)", y = "Height increment\n(m yr⁻¹)") + pub_theme

p17J_lt <- ggplot(dat17_lifetime,
                  aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  pt17 + yr_scale17 + ln17_lt + lt17 + expand_limits(y = 0) +
  labs(tag = "(j)", x = "Soil N\n(kg N m⁻³)", y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") + pub_theme

p17BB_lt <- ggplot(dat17_lifetime,
                   aes(x = soil_nitrogen,
                       y = ectomycorrhiza_N_free / pmax(ectomycorrhiza_mass, 1e-12))) +
  pt17 + yr_scale17 + ln17_lt + lt17 + expand_limits(y = 0) +
  labs(tag = "(k)", x = "Soil N\n(kg N m⁻³)", y = "ECM free N per biomass\n(kg N kg⁻¹)") + pub_theme

p17K_lt <- ggplot(dat17_lifetime,
                  aes(x = leaf_nitrogen_concentration, y = optimal_leaf_nitrogen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  pt17 + yr_scale17 + expand_limits(y = 0, x = 0) +
  labs(tag = "(l)", x = "Actual leaf N\n(kg N kg⁻¹ leaf)",
       y = "Optimal leaf N\n(kg N kg⁻¹ leaf)") +
  pub_theme + theme(plot.background = element_rect(fill = "grey92", colour = NA))

fig17_lifetime <- make_row_title("Parameters to be Optimised") /
  (p17A_lt | p17B_lt | p17C_lt | p17C2_lt) /
  make_row_title("ECM Dynamics & N Uptake") /
  (p17E_lt | p17F_lt | p17O_lt | p17O_myco_lt) /
  make_row_title("Tree Status") /
  (p17I_lt | p17J_lt | p17BB_lt | p17K_lt) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") &
  theme(plot.tag         = element_text(size = 22, face = "bold"),
        legend.text      = element_text(size = 22),
        legend.key.size  = unit(0.9, "lines"),
        legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.box       = "horizontal")
ggsave(file.path(pub_fig_dir, "17_lifetime_optimals.png"),
       plot = fig17_lifetime, width = 22, height = 24, dpi = 300)

# ── Last 10 years ─────────────────────────────────────────────────────────────
# Same structure as fig17 but restricted to the final 10 years (2013--2022),
# so the violin distributions and scatter reflect mature-tree behaviour only.
dat17_last10 <- annual_opt_gpp %>%
  filter(year >= max(year, na.rm = TRUE) - 9) %>%
  mutate(
    bga         = (C_export_to_myco + rr * 0.44 + tr * 0.44) / pmax(assim_gross, 1e-12),
    mycorrhized = 0.9
  )

dat17_l10_mean <- dat17_last10 %>%
  group_by(soil_nitrogen) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop")

ln17_l10 <- geom_line(data = dat17_l10_mean,
                      aes(linetype = "last", group = 1),
                      colour = "black", linewidth = 0.5, na.rm = TRUE)

p17A_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = factor(ecto_allo))) +
  viol17 + jit_cross + yr_scale17 +
  scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
  labs(tag = "(a)", x = "Soil N\n(kg N m⁻³)", y = "Optimal ECM C allocation\n(% NPP)") + pub_theme

p17B_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = factor(root_no))) +
  viol17 + jit_cross + yr_scale17 +
  scale_y_discrete(labels = function(x) scales::scientific(as.numeric(x))) +
  labs(tag = "(b)", x = "Soil N\n(kg N m⁻³)", y = "Optimal root number\n(no m⁻² crown)") + pub_theme

p17C_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = factor(root_length))) +
  viol17 + jit_cross + yr_scale17 +
  labs(tag = "(c)", x = "Soil N\n(kg N m⁻³)", y = "Optimal root length\n(mm)") + pub_theme

p17C2_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = factor(mycorrhized))) +
  viol17 + jit_cross + yr_scale17 +
  scale_y_discrete(labels = function(x) scales::percent(as.numeric(x))) +
  labs(tag = "(d)", x = "Soil N\n(kg N m⁻³)", y = "ECM colonisation\n(% roots covered)") + pub_theme

p17E_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  pt17 + yr_scale17 + ln17_l10 + lt17 +
  scale_y_continuous(labels = scales::label_scientific()) + expand_limits(y = 0) +
  labs(tag = "(e)", x = "Soil N\n(kg N m⁻³)", y = "ECM biomass\n(kg)") + pub_theme

p17F_l10 <- ggplot(dat17_last10,
                   aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  pt17 + yr_scale17 + ln17_l10 + lt17 +
  scale_y_continuous(labels = scales::label_scientific()) + expand_limits(y = 0) +
  labs(tag = "(f)", x = "Soil N\n(kg N m⁻³)", y = "ECM biomass per crown\n(kg m⁻²)") + pub_theme

p17O_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = root_uptake_actual)) +
  pt17 + yr_scale17 + ln17_l10 + lt17 + expand_limits(y = 0) +
  labs(tag = "(g)", x = "Soil N\n(kg N m⁻³)", y = "Root N uptake\n(kg N yr⁻¹)") + pub_theme

p17O_myco_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = myco_uptake)) +
  pt17 + yr_scale17 + ln17_l10 + lt17 + expand_limits(y = 0) +
  labs(tag = "(h)", x = "Soil N\n(kg N m⁻³)", y = "ECM N uptake\n(kg N yr⁻¹)") + pub_theme

p17I_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = height_increment)) +
  pt17 + yr_scale17 + ln17_l10 + lt17 + expand_limits(y = 0) +
  labs(tag = "(i)", x = "Soil N\n(kg N m⁻³)", y = "Height increment\n(m yr⁻¹)") + pub_theme

p17J_l10 <- ggplot(dat17_last10, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  pt17 + yr_scale17 + ln17_l10 + lt17 + expand_limits(y = 0) +
  labs(tag = "(j)", x = "Soil N\n(kg N m⁻³)", y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") + pub_theme

p17BB_l10 <- ggplot(dat17_last10,
                    aes(x = soil_nitrogen,
                        y = ectomycorrhiza_N_free / pmax(ectomycorrhiza_mass, 1e-12))) +
  pt17 + yr_scale17 + ln17_l10 + lt17 + expand_limits(y = 0) +
  labs(tag = "(k)", x = "Soil N\n(kg N m⁻³)", y = "ECM free N per biomass\n(kg N kg⁻¹)") + pub_theme

p17K_l10 <- ggplot(dat17_last10,
                   aes(x = leaf_nitrogen_concentration, y = optimal_leaf_nitrogen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  pt17 + yr_scale17 + expand_limits(y = 0, x = 0) +
  labs(tag = "(l)", x = "Actual leaf N\n(kg N kg⁻¹ leaf)",
       y = "Optimal leaf N\n(kg N kg⁻¹ leaf)") +
  pub_theme + theme(plot.background = element_rect(fill = "grey92", colour = NA))

fig17_last10 <- make_row_title("Parameters to be Optimised") /
  (p17A_l10 | p17B_l10 | p17C_l10 | p17C2_l10) /
  make_row_title("ECM Dynamics & N Uptake") /
  (p17E_l10 | p17F_l10 | p17O_l10 | p17O_myco_l10) /
  make_row_title("Tree Status") /
  (p17I_l10 | p17J_l10 | p17BB_l10 | p17K_l10) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") &
  theme(plot.tag         = element_text(size = 22, face = "bold"),
        legend.text      = element_text(size = 22),
        legend.key.size  = unit(0.9, "lines"),
        legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.box       = "horizontal")
ggsave(file.path(pub_fig_dir, "17_last10_years.png"),
       plot = fig17_last10, width = 22, height = 24, dpi = 300)
```

## N pools vs soil N

```{r plot17b_N_pools}
ann17b <- annual_opt_gpp %>%
  mutate(
    root_SA_m2     = pi * (0.35 / root_length^0.2 / 1000) * (root_length / 1000) *
                       root_no * crown_area * lai,
    myco_SA_m2     = 4 * ectomycorrhiza_mass / (1200 * 2e-5),
    root_up_per_SA = root_uptake / pmax(root_SA_m2,  1e-12),
    myco_up_per_SA = myco_uptake / pmax(myco_SA_m2,  1e-12)
  )

p17b_G1 <- ggplot(ann17b, aes(x = soil_nitrogen, y = N_bar_roots)) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(a)", x = "Soil nitrogen (kg N m⁻³)", y = "N mineral soil\n(kg N m⁻³)") + pub_theme

p17b_G2 <- ggplot(ann17b, aes(x = soil_nitrogen, y = N_static)) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(b)", x = "Soil nitrogen (kg N m⁻³)",
       y = "N organic pool\n(kg N m⁻³)") + pub_theme

p17b_accN <- ggplot(ann17b, aes(x = soil_nitrogen, y = accessible_N)) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(c)", x = "Soil nitrogen (kg N m⁻³)",
       y = "Accessible N\n(kg N m⁻³)") + pub_theme

p17b_C <- ggplot(ann17b, aes(x = N_bar_roots, colour = year)) +
  geom_point(aes(y = root_up_per_SA), shape = 16, size = 1.8, na.rm = TRUE) +
  geom_point(aes(y = myco_up_per_SA), shape = 2,  size = 1.8, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(d)", x = "N mineral soil (kg N m⁻³)",
       y = "Uptake per SA\n(kg N m⁻² yr⁻¹)") + pub_theme

p17b_mine <- ggplot(ann17b, aes(x = N_static,
                               y = myco_uptake_static / pmax(ectomycorrhiza_mass, 1e-12))) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(e)", x = "Organic N pool (kg N m⁻³)",
       y = "N mineralisation by ECM\nper unit biomass\n(kg N yr⁻¹ kg C⁻¹)") + pub_theme

p17b_transfer <- ggplot(ann17b, aes(x = soil_nitrogen,
                                    y = mycorrhizal_export_to_tree / pmax(ectomycorrhiza_mass, 1e-12))) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(f)", x = "Soil N\n(kg N m⁻³)",
       y = "Transfer to tree\nper ECM biomass\n(kg N yr⁻¹ kg⁻¹)") + pub_theme

p17b_T3  <- ggplot(ann17b, aes(x = soil_nitrogen, y = tree_nitrogen)) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(g)", x = "Soil N\n(kg N m⁻³)", y = "Tree free N pool\n(kg N)") + pub_theme

p17b_BB3 <- ggplot(ann17b, aes(x = soil_nitrogen, y = ectomycorrhiza_N_free / pmax(ectomycorrhiza_mass, 1e-12))) +
  geom_point(aes(colour = year), size = 1.8, shape = 16, na.rm = TRUE) +
  scale_colour_viridis_c(name = "Year", option = "viridis") +
  expand_limits(y = 0) +
  labs(tag = "(h)", x = "Soil N\n(kg N m⁻³)", y = "ECM free N\nper biomass\n(kg N kg⁻¹)") + pub_theme

fig17b <- (p17b_G1 | p17b_G2 | p17b_accN) /
  (p17b_C | p17b_mine | p17b_transfer) /
  (p17b_T3 | p17b_BB3 | plot_spacer()) +
  plot_layout(guides = "collect") &
  theme(plot.tag         = element_text(size = 22, face = "bold"),
        legend.text      = element_text(size = 22),
        legend.key.size  = unit(1.3, "lines"),
        legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.box       = "horizontal") &
  guides(colour = guide_colorbar(barwidth = unit(18, "cm"), barheight = unit(0.7, "cm")))
ggsave(file.path(pub_fig_dir, "17b_N_pools.png"), plot = fig17b, width = 24, height = 20, dpi = 300)
```

## C:N budgets

```{r plot17c_CN_budgets}
dat17c <- annual_opt_gpp %>%
  filter(ectomycorrhiza_mass > 0.01) %>%
  mutate(
    CN          = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                    pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12),
    cn_exchange = C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12),
    transfer_eff = mycorrhizal_export_to_tree /
                    pmax(myco_uptake + ectomycorrhiza_N_free * k_mob_N, 1e-12),
    c_supply_per_mass = C_export_to_myco / pmax(ectomycorrhiza_mass, 1e-12),
    eff_rF       = 1 / pmax(CN, 1e-3)
  )

yr_scale_c <- scale_colour_viridis_c(
  name  = "Year", option = "viridis",
  guide = guide_colorbar(barwidth = unit(12, "cm"), barheight = unit(0.5, "cm"))
)
pt_c <- geom_point(aes(colour = year), size = 1.8, shape = 16, alpha = 0.7, na.rm = TRUE)

p17c_H <- ggplot(dat17c, aes(x = myco_uptake, y = CN)) +
  pt_c + yr_scale_c +
  scale_x_continuous(labels = scales::label_number(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 1)) + expand_limits(y = 0) +
  labs(tag = "(a)", x = "Fungal N uptake\n(kg N yr⁻¹)", y = "Mycorrhizal C:N\n(kg C kg⁻¹ N)") + pub_theme

p17c_L <- ggplot(dat17c, aes(x = soil_nitrogen, y = cn_exchange)) +
  pt_c + yr_scale_c +
  scale_y_log10() +
  labs(tag = "(b)", x = "Soil N\n(kg N m⁻³)",
       y = "C:N exchange cost\n(kg C kg⁻¹ N transferred)") + pub_theme

p17c_L_zoom <- ggplot(dat17c, aes(x = soil_nitrogen, y = cn_exchange)) +
  pt_c + yr_scale_c + expand_limits(y = 0) +
  coord_cartesian(ylim = c(0, quantile(dat17c$cn_exchange, 0.93, na.rm = TRUE))) +
  labs(tag = "(d)", x = "Soil N\n(kg N m⁻³)",
       y = "C:N exchange cost\n(kg C kg⁻¹ N received)\n[zoomed]") + pub_theme

p17c_phiB <- ggplot(dat17c, aes(x = eff_rF, y = mycorrhizal_export_to_tree)) +
  pt_c + yr_scale_c + expand_limits(y = 0, x = 0) +
  labs(tag = "(c)",
       x = "Effective fungal N:C\n(kg N kg C⁻¹)",
       y = "N transferred to tree\n(kg N yr⁻¹)") + pub_theme

p17c_freeN <- ggplot(dat17c, aes(x = myco_uptake, y = ectomycorrhiza_N_free)) +
  pt_c + yr_scale_c + expand_limits(y = 0) +
  labs(tag = "(d)", x = "Fungal N uptake\n(kg N yr⁻¹)", y = "ECM free N\n(kg N)") + pub_theme

p17c_freeC <- ggplot(dat17c, aes(x = myco_uptake, y = ectomycorrhiza_C_free)) +
  pt_c + yr_scale_c + expand_limits(y = 0) +
  labs(tag = "(e)", x = "Fungal N uptake\n(kg N yr⁻¹)", y = "ECM free C\n(kg C)") + pub_theme

fig17c <- (p17c_H | p17c_L | p17c_phiB) /
  (p17c_freeN | p17c_freeC | plot_spacer()) +
  plot_layout(guides = "collect") &
  theme(plot.tag         = element_text(size = 18, face = "bold"),
        legend.text      = element_text(size = 14),
        legend.key.size  = unit(0.9, "lines"),
        legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.box       = "horizontal",
        plot.margin      = margin(8, 8, 8, 8))
ggsave(file.path(pub_fig_dir, "17c_CN_budgets.png"), plot = fig17c, width = 21, height = 16, dpi = 300)
```

#################################################################################

## Life History Simulations (lh_09 / lh_10)

```{r lh_validation_data}
# Halme et al. 2022: field measurements boreal
halme_et_al_2022 <- data.frame(matrix(nrow = 9, ncol = 2))
halme_et_al_2022$height         <- c(17.1, 21.0, 20.5, 18.3, 13.2, 20.7, 16.8, 23.6, 13.6)
halme_et_al_2022$crown_diameter <- c(3.4,  4.0,  4.1,  3.3,  2.9,  3.7,  3.4,  4.2,  2.8)

# SMEAR II processed data
data_directory <- "~/Documents/CASSIA_Calibration/Processed_Data/"
loaded_data <- CASSIA::load_data(data_directory)

# Eddy covariance (Hyytiälä)
raw.directory <- "/home/josimms/Documents/CASSIA_Calibration/Raw_Data/hyytiala_weather/"
environmental.variable.list <- list()
count <- 1
for (variable in c("GPP", "NEE", "F_CO2_leaf", "F_H2O_leaf", "ET_gapf")) {
  environmental.variable.list[[count]] <- data.table::rbindlist(
    lapply(paste0(raw.directory, list.files(raw.directory, variable)), data.table::fread))
  environmental.variable.list[[count]]$Date <- paste(
    environmental.variable.list[[count]]$Year,
    environmental.variable.list[[count]]$Month,
    environmental.variable.list[[count]]$Day, sep = "-")
  environmental.variable.list[[count]]$Monthly <- paste(
    environmental.variable.list[[count]]$Year,
    environmental.variable.list[[count]]$Month, sep = "-")
  count <- count + 1
}
names(environmental.variable.list) <- c("GPP", "NEE", "F_CO2_leaf", "F_H2O_leaf", "ET")

DT  <- copy(environmental.variable.list[["GPP"]])
DT2 <- copy(environmental.variable.list[["NEE"]])
DT5 <- copy(environmental.variable.list[["ET"]])

GPP_out <- DT[,  .(GPP_mean = mean(HYY_EDDY233.GPP,    na.rm = TRUE)), by = Monthly]
NEE_out <- DT2[, .(NEE_mean = mean(HYY_EDDY233.NEE,    na.rm = TRUE)), by = Monthly]
ET_out  <- DT5[, .(ET_mean  = mean(HYY_EDDY233.ET_gapf, na.rm = TRUE)), by = Monthly]

GPP_out <- GPP_out[, Monthly := zoo::as.yearmon(Monthly, "%Y-%m")]
NEE_out <- NEE_out[, Monthly := zoo::as.yearmon(Monthly, "%Y-%m")]
ET_out  <- ET_out[,  Monthly := zoo::as.yearmon(Monthly, "%Y-%m")]
setorder(GPP_out, Monthly); setorder(NEE_out, Monthly); setorder(ET_out, Monthly)

area_per_tree        <- 1
GPP_out$GPP_mean_kg  <- GPP_out$GPP_mean * 12.011 * 1e-9 * 60 * 60 * 24 * 365.25 * area_per_tree
NEE_out$NEE_mean_kg  <- NEE_out$NEE_mean * 12.011 * 1e-9 * 60 * 60 * 24 * 365.25 * area_per_tree
GPP_out$NEE_mean     <- NEE_out$NEE_mean
GPP_out$NEE_mean_kg  <- NEE_out$NEE_mean_kg
GPP_out$ET           <- ET_out$ET_mean
Eddy_covariance      <- GPP_out
Eddy_covariance[, MonthlyDate := as.Date(paste0("01 ", Monthly), format = "%d %b %Y")]
Eddy_covariance[, Month       := as.integer(format(MonthlyDate, "%m"))]

```

```{r lh_simulations}
lh_dt         <- 1/12
lh_start_year <- 1960
lh_end_year   <- 2022
lh_years_seq  <- seq(lh_start_year, lh_end_year, lh_dt)

run_lh_sim <- function(n_val) {
  lho <- new(LifeHistoryOptimizer, param_file)
  lho$set_i_metFile(weather_file)
  lho$set_a_metFile(weather_file)
  lho$set_co2File("")
  lho$set_soil_nitrogen(n_val)
  lho$init()
  col_names <- lho$get_header()
  results <- lapply(lh_years_seq, function(t) {
    tryCatch({
      lho$grow_for_dt(t, lh_dt)
      state <- lho$get_state(t + lh_dt)
      df_row <- as.data.frame(t(state))
      names(df_row) <- col_names
      df_row
    }, error = function(e) {
      setNames(as.data.frame(as.list(rep(NA, length(col_names)))), col_names)
    })
  })
  df <- do.call(rbind, results)
  names(df) <- col_names
  df$date <- seq(as.Date(paste0(lh_start_year, "-01-01")),
                 as.Date(paste0(lh_end_year,   "-01-01")),
                 by = "month")[1:nrow(df)]
  df
}

df   <- run_lh_sim(0.2)   # N = Low
df_2 <- run_lh_sim(0.4)   # N = Medium
df_3 <- run_lh_sim(1.0)   # N = High

lh_cols     <- c("#0072B2", "#E69F00", "#D55E00")
lh_N_labels <- c("N = Low", "N = Medium", "N = High")
col_obs      <- "lightblue"
col_range    <- "lightblue"

lh_cex_axis   <- 2.0
lh_cex_lab    <- 2.2
lh_cex_main   <- 2.4
lh_cex_legend <- 1.8
lh_lwd_model  <- 3
lh_lwd_obs    <- 2
lh_pch_obs    <- 16

add_range <- function(x, y_min, y_max,
                      col = adjustcolor(col_range, alpha.f = 0.5),
                      lwd = lh_lwd_obs, cap = 0.015) {
  arrows(x, y_min, x, y_max, code = 3, angle = 90, length = cap, col = col, lwd = lwd)
}
```

```{r lh_09_main_panel}
png(file.path(pub_fig_dir, "lh_09_publication_main.png"),
    width = 16, height = 12, units = "in", res = 300)
par(mfrow = c(4, 2), mar = c(4, 6, 3, 1), oma = c(4, 0, 2, 0),
    family = "serif", las = 1, tcl = -0.4, mgp = c(4, 1, 0))

# (a) GPP
ylim_gpp <- range(df$assim_gross/df$crown_area/2.04,
                  df_3$assim_gross/df_3$crown_area/2.04,
                  Eddy_covariance$GPP_mean_kg, na.rm = TRUE)
plot(df$date, df$assim_gross/df$crown_area/2.04, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_gpp, xlab = "",
     ylab = expression("GPP (kg C m"^{-2}*" year"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, df_3$assim_gross/df_3$crown_area/2.04, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   df$assim_gross/df$crown_area/2.04,     col = lh_cols[1], lwd = lh_lwd_model)
points(Eddy_covariance$MonthlyDate, Eddy_covariance$GPP_mean_kg,
       col = col_obs, pch = lh_pch_obs, cex = 0.8)
mtext("(a)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (b) NPP
ylim_npp <- range(df$assim_net/df$crown_area/2.04,
                  df_3$assim_net/df_3$crown_area/2.04,
                  -Eddy_covariance$NEE_mean_kg, na.rm = TRUE)
plot(df$date, df$assim_net/df$crown_area/2.04, type = "n",
     ylim = ylim_npp, xlab = "",
     ylab = expression("NPP (kg C m"^{-2}*" year"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, df_3$assim_net/df_3$crown_area/2.04, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   df$assim_net/df$crown_area/2.04,     col = lh_cols[1], lwd = lh_lwd_model)
points(Eddy_covariance$MonthlyDate, -Eddy_covariance$NEE_mean_kg,
       col = col_obs, pch = lh_pch_obs, cex = 0.8)
mtext("(b)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (c) Height
smear_h_idx  <- loaded_data$smearII_data$variable == "pine height BA weighted mean"
smear_h_val  <- loaded_data$smearII_data$amount[smear_h_idx]
smear_h_date <- as.Date(paste0(loaded_data$smearII_data$date[smear_h_idx], "-01-01"))
ylim_h <- range(df$height, df_3$height, smear_h_val,
                halme_et_al_2022$height, useful_data$averageTreeHeight, na.rm = TRUE)
plot(df$date, df$height, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_h, xlab = "", ylab = "Height (m)",
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
points(smear_h_date, smear_h_val,
       col = adjustcolor(col_range, alpha.f = 0.15), pch = 4, cex = 1.5, lwd = lh_lwd_obs)
add_range(as.Date("2017-06-01"),
          min(halme_et_al_2022$height, na.rm = TRUE),
          max(halme_et_al_2022$height, na.rm = TRUE))
for (id in unique(useful_data$plotID)) {
  sub <- useful_data[useful_data$plotID == id & !is.na(useful_data$averageTreeHeight), ]
  sub <- sub[order(sub$eventYear), ]
  if (nrow(sub) < 1) next
  x <- as.Date(as.character(sub$eventYear), format = "%Y")
  y <- sub$averageTreeHeight
  x_plot <- x[1]; y_plot <- y[1]
  for (j in seq_len(nrow(sub) - 1)) {
    if (y[j + 1] < y[j]) {
      x_plot <- c(x_plot, NA, x[j + 1])
      y_plot <- c(y_plot, NA, y[j + 1])
    } else {
      x_plot <- c(x_plot, x[j + 1])
      y_plot <- c(y_plot, y[j + 1])
    }
  }
  points(x, y, col = adjustcolor(col_range, alpha.f = 0.15), pch = 16, cex = 0.4)
  lines(x_plot, y_plot, col = adjustcolor(col_range, alpha.f = 0.15), lwd = lh_lwd_obs)
}
mtext("(c)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")
lines(df_3$date, df_3$height, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   df$height,   col = lh_cols[1], lwd = lh_lwd_model)

# (d) Diameter
smear_d_idx  <- loaded_data$smearII_data$variable == "pine diameter BA weighted mean"
smear_d_val  <- 0.01 * loaded_data$smearII_data$amount[smear_d_idx]
smear_d_date <- as.Date(paste0(loaded_data$smearII_data$date[smear_d_idx], "-01-01"))
ylim_d <- range(df$diameter, df_3$diameter, smear_d_val,
                0.01 * useful_data$averageTreeDiameter, na.rm = TRUE)
plot(df_3$date, df_3$diameter, type = "l", col = lh_cols[3], lwd = lh_lwd_model,
     ylim = ylim_d, xlab = "", ylab = "Diameter (m)",
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df$date, df$diameter, col = lh_cols[1], lwd = lh_lwd_model)
points(smear_d_date, smear_d_val,
       col = adjustcolor(col_range, alpha.f = 0.15), pch = 4, cex = 1.5, lwd = lh_lwd_obs)
for (id in unique(useful_data$plotID)) {
  sub <- useful_data[useful_data$plotID == id & !is.na(useful_data$averageTreeDiameter), ]
  sub <- sub[order(sub$eventYear), ]
  if (nrow(sub) < 1) next
  x <- as.Date(as.character(sub$eventYear), format = "%Y")
  y <- 0.01 * sub$averageTreeDiameter
  x_plot <- x[1]; y_plot <- y[1]
  for (j in seq_len(nrow(sub) - 1)) {
    if (y[j + 1] < y[j]) {
      x_plot <- c(x_plot, NA, x[j + 1])
      y_plot <- c(y_plot, NA, y[j + 1])
    } else {
      x_plot <- c(x_plot, x[j + 1])
      y_plot <- c(y_plot, y[j + 1])
    }
  }
  points(x, y, col = adjustcolor(col_range, alpha.f = 0.15), pch = 16, cex = 0.4)
  lines(x_plot, y_plot, col = adjustcolor(col_range, alpha.f = 0.15), lwd = lh_lwd_obs)
}
mtext("(d)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")
lines(df_3$date, df_3$diameter, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   df$diameter,   col = lh_cols[1], lwd = lh_lwd_model)

# (e) Vcmax
ylim_vc <- range(df$vcmax/df$lai, df_3$vcmax/df_3$lai, 30, na.rm = TRUE)
plot(df$date, df$vcmax/df$lai, type = "n",
     ylim = ylim_vc, xlab = "",
     ylab = expression("V"[cmax]*" ("*mu*"mol m"^{-2}*" s"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
rect(par("usr")[1], 0, par("usr")[2], 30,
     col = adjustcolor(col_range, alpha.f = 0.15), border = NA)
lines(df_3$date, df_3$vcmax/df_3$lai, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   df$vcmax/df$lai,     col = lh_cols[1], lwd = lh_lwd_model)
mtext("(e)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (f) Optimal leaf nitrogen
korhonen_needle_n_lo <- 0.0121
korhonen_needle_n_hi <- 0.0151
ylim_oln <- range(0, df$optimal_leaf_nitrogen * 1.1, df_3$optimal_leaf_nitrogen * 1.1,
                  korhonen_needle_n_hi, na.rm = TRUE)
plot(df$date, df$optimal_leaf_nitrogen, type = "n",
     ylim = ylim_oln, xlab = "",
     ylab = expression("Optimal leaf N (g g"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
rect(par("usr")[1], korhonen_needle_n_lo, par("usr")[2], korhonen_needle_n_hi,
     col = adjustcolor(col_range, alpha.f = 0.15), border = NA)
for (i in c(3, 1)) {
  d_i <- list(df, df_2, df_3)[[i]]
  idx <- !is.na(d_i$leaf_nitrogen_concentration)
  polygon(c(d_i$date[idx], rev(d_i$date[idx])),
          c(d_i$leaf_nitrogen_concentration[idx], rep(0, sum(idx))),
          col = adjustcolor(lh_cols[i], alpha.f = 0.5), border = NA)
}
points(df_3$date, df_3$optimal_leaf_nitrogen, col = lh_cols[3], pch = 20, cex = 1)
points(df$date,   df$optimal_leaf_nitrogen,   col = lh_cols[1], pch = 20, cex = 1)
mtext("(f)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (g) N uptake
uptake_1 <- df$mycorrhizal_export_to_tree   + df$root_uptake_actual
uptake_3 <- df_3$mycorrhizal_export_to_tree + df_3$root_uptake_actual
nuptake_obs_val  <- 0.013
nbar_obs_val     <- 1/13
nbar_obs_date    <- as.Date("2008-08-01")
ylim_nu <- range(uptake_1, uptake_3, nuptake_obs_val, na.rm = TRUE)
plot(df$date, uptake_1, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_nu, xlab = "",
     ylab = expression("N uptake (kg N tree"^{-1}*" year"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, uptake_3, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   uptake_1, col = lh_cols[1], lwd = lh_lwd_model)
points(nbar_obs_date, nbar_obs_val, col = col_range, pch = lh_pch_obs, cex = 0.8)
mtext("(g)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (h) ECM / fine root biomass ratio
ratio_1 <- df$ectomycorrhiza_mass   / df$root_mass
ratio_3 <- df_3$ectomycorrhiza_mass / df_3$root_mass
ylim_nb <- range(ratio_1, ratio_3, na.rm = TRUE)
plot(df$date, ratio_1, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_nb, xlab = "",
     ylab = "ECM / fine root biomass (kg kg⁻¹)",
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, ratio_3, col = lh_cols[3], lwd = lh_lwd_model)
lines(df$date,   ratio_1, col = lh_cols[1], lwd = lh_lwd_model)
mtext("(h)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# Shared legend
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", horiz = FALSE, bty = "n", cex = lh_cex_legend, ncol = 3,
       legend = c(lh_N_labels[c(1, 3)], "Eddy covariance",
                  "Range of realistic boreal observations"),
       col  = c(lh_cols[c(1, 3)], col_obs, col_range),
       lty  = c(1, 1, NA, 1),
       pch  = c(NA, NA, lh_pch_obs, NA),
       lwd  = c(rep(lh_lwd_model, 2), NA, lh_lwd_obs))
dev.off()
```

```{r lh_10_belowground}
nbar_lower <- 0.0012
nbar_upper <- 0.003

uptake_1  <- df$mycorrhizal_export_to_tree   + df$root_uptake_actual
uptake_3  <- df_3$mycorrhizal_export_to_tree + df_3$root_uptake_actual
bg_mass_1 <- df$root_mass   + df$ectomycorrhiza_mass
bg_mass_3 <- df_3$root_mass + df_3$ectomycorrhiza_mass
eff_1     <- uptake_1 / bg_mass_1
eff_3     <- uptake_3 / bg_mass_3

k_1_par         <- 0.35
k_13_par        <- 2.5
depth_par       <- 1.1
rho_myco_par    <- 330
myco_d_par      <- 1e-5
mycorrhized_par <- 0.9
compute_sa_density <- function(d) {
  root_d_mm <- k_1_par / sqrt(d$root_length)
  SA_fr     <- pi * (root_d_mm/1000) * (d$root_length/1000) * d$root_no * d$crown_area * d$lai
  SA_m      <- 4 * d$ectomycorrhiza_mass / (rho_myco_par * myco_d_par)
  SA_active <- (1 - mycorrhized_par) * SA_fr + SA_m
  crown_r   <- sqrt(d$crown_area / pi)
  R         <- k_13_par * crown_r
  V_zone    <- (2/3) * pi * R^2 * depth_par
  list(SA_active = SA_active, V_zone = V_zone,
       density = SA_active / V_zone, SA_fr = SA_fr, SA_m = SA_m)
}
sa1 <- compute_sa_density(df)
sa3 <- compute_sa_density(df_3)

png(file.path(pub_fig_dir, "lh_10_publication_belowground.png"),
    width = 14, height = 12, units = "in", res = 300)
par(mfrow = c(3, 2), mar = c(4, 6, 3, 1), oma = c(4, 0, 2, 0),
    family = "serif", las = 1, tcl = -0.4, mgp = c(4, 1, 0))

# (a) N_bar + Korhonen reference band
ylim_nb <- range(df$N_bar, df_3$N_bar, nbar_lower, nbar_upper, na.rm = TRUE)
plot(df$date, df$N_bar, type = "n",
     ylim = ylim_nb, xlab = "",
     ylab = expression(bar(N)~"(kg N m"^{-3}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
rect(par("usr")[1], nbar_lower, par("usr")[2], nbar_upper,
     col = adjustcolor(col_range, alpha.f = 0.15), border = NA)
lines(df$date,   df$N_bar,   col = lh_cols[1], lwd = lh_lwd_model)
lines(df_3$date, df_3$N_bar, col = lh_cols[3], lwd = lh_lwd_model)
mtext("(a)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
title(sub = "Shading: Korhonen A-horizon mineral N range", col.sub = "grey40")
box(bty = "l")

# (b) Internal tree free N
ylim_tn <- range(df$tree_nitrogen, df_3$tree_nitrogen, na.rm = TRUE)
plot(df$date, df$tree_nitrogen, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_tn, xlab = "",
     ylab = expression("Internal tree N (kg N tree"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, df_3$tree_nitrogen, col = lh_cols[3], lwd = lh_lwd_model)
mtext("(b)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (c) N uptake per belowground biomass
ylim_eff <- range(eff_1, eff_3, na.rm = TRUE)
plot(df$date, eff_1, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_eff, xlab = "",
     ylab = expression("N uptake / BG mass (kg N kg C"^{-1}*" yr"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, eff_3, col = lh_cols[3], lwd = lh_lwd_model)
mtext("(c)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (d) N transferred (mycorrhizal export to tree)
ylim_tr <- range(df$mycorrhizal_export_to_tree, df_3$mycorrhizal_export_to_tree, na.rm = TRUE)
plot(df$date, df$mycorrhizal_export_to_tree, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_tr, xlab = "",
     ylab = expression("N transferred (kg N tree"^{-1}*" yr"^{-1}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, df_3$mycorrhizal_export_to_tree, col = lh_cols[3], lwd = lh_lwd_model)
mtext("(d)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# (e) SA_active / V_zone
ylim_sad <- range(sa1$density, sa3$density, na.rm = TRUE)
plot(df$date, sa1$density, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_sad, xlab = "",
     ylab = expression("SA"["active"]*" / V"["zone"]~"(m"^2~"m"^{-3}*")"),
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, sa3$density, col = lh_cols[3], lwd = lh_lwd_model)
rect(par("usr")[1], 10, par("usr")[2], 80,
     col = adjustcolor(col_range, alpha.f = 0.15), border = NA)
mtext("(e)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
title(sub = "Shading: ~10–80 m² m⁻³ typical boreal fine root + ECM density",
      col.sub = "grey40")
box(bty = "l")

# (f) Myco uptake / realised root uptake ratio
ratio_1 <- df$myco_uptake   / pmax(df$root_uptake_actual,   1e-12)
ratio_3 <- df_3$myco_uptake / pmax(df_3$root_uptake_actual, 1e-12)
ylim_rat <- range(ratio_1, ratio_3, na.rm = TRUE)
plot(df$date, ratio_1, type = "l", col = lh_cols[1], lwd = lh_lwd_model,
     ylim = ylim_rat, xlab = "",
     ylab = "Myco uptake / realised root uptake (–)",
     cex.axis = lh_cex_axis, cex.lab = lh_cex_lab)
lines(df_3$date, ratio_3, col = lh_cols[3], lwd = lh_lwd_model)
abline(h = 1, lty = 2, col = "grey50")
mtext("(f)", side = 3, adj = 0, line = 0.2, font = 2, cex = lh_cex_main)
box(bty = "l")

# Shared legend
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", horiz = TRUE, bty = "n", cex = lh_cex_legend,
       legend = c(lh_N_labels[c(1, 3)], "Range of realistic boreal observations"),
       col  = c(lh_cols[c(1, 3)], col_range),
       lty  = 1,
       lwd  = c(rep(lh_lwd_model, 2), lh_lwd_obs))
dev.off()
```