---
title: "Publication Plots"
author: "Joanna"
date: "2026-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(data.table)
library(rphydro)
library(Rprebasso)
library(patchwork)
library(PlantFATE)
library(akima)
library(future.apply)
library(future)
library(here)
```

```{r theme, include=FALSE}
pub_theme <- theme_classic(base_size = 22) +
  theme(
    axis.title       = element_text(size = 22, face = "bold"),
    axis.text        = element_text(size = 20, colour = "black"),
    axis.line        = element_line(colour = "black", linewidth = 0.6),
    axis.ticks       = element_line(colour = "black", linewidth = 0.5),
    legend.title     = element_text(size = 22, face = "bold"),
    legend.text      = element_text(size = 20),
    legend.key.size  = unit(0.9, "lines"),
    panel.grid.major = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background  = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.margin      = margin(8, 12, 8, 8),
    strip.text       = element_text(size = 22, face = "bold")
  )
```

## Data reading

```{r}
hyde_all = "~/Documents/Austria/Hyytiala_all_data/"
data <- lapply(paste0(hyde_all, list.files(hyde_all))[1:9], read.csv)
names(data) <- list.files(hyde_all)[1:9]
useful_data <- merge(data[[2]], data[[5]], c('plotID', 'eventID', 'eventYear'), all=TRUE)

```

## Simulations

```{r pressure, echo=FALSE}
# Set to TRUE for fast testing (fewer grid points, shorter run), FALSE for full publication run
testing_mode <- TRUE

plan(multicore, workers = parallel::detectCores() - 1)

# Parameter ranges
n_pts_main  <- if (testing_mode) 10 else 21
n_pts_joint <- if (testing_mode) 5  else 8   # per-axis for 3D co-optimisation sweep

nitrogen_values <- seq(0.01, 2.0, length.out = n_pts_main)

# Joint co-optimisation grid (coarser to keep run count tractable)
joint_ecto_allo_values    <- seq(0.0,  0.3, length.out = n_pts_joint)
joint_root_no_values      <- 3*10^seq(3, 6, length.out = n_pts_joint)
joint_root_length_values  <- seq(0.05, 6,   length.out = n_pts_joint)
param_file <- "/home/josimms/Documents/Austria/Plant-FATE/tests/params/p_test_boreal.ini"
weather_file <- "/home/josimms/Documents/Austria/Plant-FATE/data/ERAS_Monthly.csv"

# -------------------------------
# Core simulation function with safe error handling
# -------------------------------
run_sim_parallel <- function(rl, rn, n, eb) {
  tryCatch({
    lho <- new(LifeHistoryOptimizer, param_file)
    lho$set_i_metFile(weather_file)
    lho$set_a_metFile(weather_file)
    lho$set_co2File("")
    lho$set_soil_nitrogen(n)
    lho$traits0$investment_from_tree <- eb
    lho$init()
    lho$root_override(rn, rl)

    dt <- 1/12
    start_year <- 1960
    end_year   <- 2022
    years_seq  <- seq(start_year, end_year, dt)
    
    results <- lapply(years_seq, function(t) {
      tryCatch({
        lho$grow_for_dt(t, dt)
        state <- lho$get_state(t + dt)
        df_row <- as.data.frame(t(state))
        names(df_row) <- lho$get_header()
        df_row
      }, error = function(e) {
        setNames(as.data.frame(as.list(rep(NA, length(lho$get_header())))),
                 lho$get_header())
      })
    })
    
    df <- do.call(rbind, results)
    df$date <- seq(as.Date(paste0(start_year, "-01-01")),
                   as.Date(paste0(end_year, "-01-01")),
                   by = "month")[1:nrow(df)]
    df$root_length   <- rl
    df$root_no       <- rn
    df$soil_nitrogen <- n
    df$ecto_allo     <- eb
    df
  }, error = function(e) {
    message("Failed run_sim_parallel: rl=", rl, " rn=", rn, " n=", n, " eb=", eb)
    NULL
  })
}

# -------------------------------
# Helper: safe wrapper for run_sim_parallel
# -------------------------------
run_sim_safe <- function(rl, rn, n, eb) {
  result <- run_sim_parallel(rl, rn, n, eb)
  if (is.null(result) || nrow(result) == 0) return(NULL)
  result
}

# -------------------------------
# Joint co-optimisation: ecto_allo × root_no × root_length × nitrogen
# n_pts_joint^3 × n_pts_main runs total
# Parallelised over ecto_allo (outer axis)
# -------------------------------
all_joint_list <- future_lapply(joint_ecto_allo_values, function(eb) {
  inner_rn <- lapply(joint_root_no_values, function(rn) {
    inner_rl <- lapply(joint_root_length_values, function(rl) {
      inner_n <- lapply(nitrogen_values, function(n) {
        run_sim_safe(rl, rn, n, eb)
      })
      do.call(rbind, inner_n[!sapply(inner_n, is.null)])
    })
    do.call(rbind, inner_rl[!sapply(inner_rl, is.null)])
  })
  do.call(rbind, inner_rn[!sapply(inner_rn, is.null)])
})
all_joint_df <- do.call(rbind, all_joint_list[!sapply(all_joint_list, is.null)])
```

## Co-optimised trait optima

```{r joint_optimum}
# June–August 2021 growing-season mean per parameter combination.
# Rate variables (N uptake, GPP) are zero in winter (f_temp = 0), so January
# snapshots understate uptake; seasonal means give a fair comparison.
joint_gs <- all_joint_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  summarise(
    height                     = mean(height,                     na.rm = TRUE),
    ectomycorrhiza_mass        = mean(ectomycorrhiza_mass,        na.rm = TRUE),
    ectomycorrhiza_N_biomass   = mean(ectomycorrhiza_N_biomass,   na.rm = TRUE),
    crown_area                 = mean(crown_area,                 na.rm = TRUE),
    N_bar                      = mean(N_bar,                      na.rm = TRUE),
    C_export_to_myco           = mean(C_export_to_myco,           na.rm = TRUE),
    mycorrhizal_export_to_tree = mean(mycorrhizal_export_to_tree, na.rm = TRUE),
    root_uptake                = mean(root_uptake,                na.rm = TRUE),
    assim_gross                = mean(assim_gross,                na.rm = TRUE),
    optimal_leaf_nitrogen      = mean(optimal_leaf_nitrogen,      na.rm = TRUE),
    leaf_nitrogen_concentration = mean(leaf_nitrogen_concentration, na.rm = TRUE),
    tree_nitrogen               = mean(tree_nitrogen,              na.rm = TRUE),
    .groups = "drop"
  )

opt_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(
    max_h      = max(height, na.rm = TRUE),
    near_opt   = height >= 0.99 * max_h,
    n_total    = n(),
    n_near_opt = sum(near_opt),
    stab_frac  = n_near_opt / n_total,
    height_gain = max_h / sort(height, decreasing = TRUE)[2] - 1
  ) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup()

# --- Stability bands: all combos within 1% of maximum height per soil N ---
# This shows how flat the fitness landscape is around each optimum
stab_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_h = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_h) %>%
  filter(near_opt) %>%
  summarise(
    ecto_lo = min(ecto_allo,    na.rm = TRUE), ecto_hi = max(ecto_allo,    na.rm = TRUE),
    rn_lo   = min(root_no,      na.rm = TRUE), rn_hi   = max(root_no,      na.rm = TRUE),
    rl_lo   = min(root_length,  na.rm = TRUE), rl_hi   = max(root_length,  na.rm = TRUE),
    .groups = "drop"
  )

joint_height_range <- range(opt_joint$height, na.rm = TRUE)

shape_vals <- c("Optimal parameters" = 3, "Model output" = 1)
ltype_vals <- c("Trend line"   = "dashed")

# Produces a 2-row data frame for a straight trend-line segment
tl <- function(x1, y1, x2, y2) data.frame(x = c(x1, x2), y = c(y1, y2))

colour_vals <- c("Stability (≥99% of max height)" = "grey70")

shared_scales <- list(
  scale_shape_manual(values = shape_vals, name = NULL),
  scale_linetype_manual(values = ltype_vals, name = NULL),
  scale_colour_manual(values = colour_vals, name = NULL)
)

stab_label <- geom_text(aes(label = paste0(round(stab_frac * 100), "%")),
                        vjust = -0.8, size = 3, inherit.aes = TRUE)

# --- Top row: crosses + stability bars, with dashed trend line ---
p_joint_ecto <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ecto_allo)) +
  geom_line(data = tl(0, 0.30, 2, 0.05), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = ecto_lo, ymax = ecto_hi,
                 colour = "Stability (≥99% of max height)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(tag = "A", x = "Soil nitrogen (g N m⁻³)", y = "Optimal ectomycorrhizal\nallocation (%)") +
  pub_theme

p_joint_rn <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_no)) +
  geom_line(data = tl(0, 3e6, 2, 1e5), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rn_lo, ymax = rn_hi,
                 colour = "Stability (≥99% of max height)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(tag = "B", x = "Soil nitrogen (g N m⁻³)", y = "Optimal root number\nper crown area (no m⁻²)") +
  pub_theme

p_joint_rl <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_length)) +
  geom_line(data = tl(0, 5, 2, 1.2), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rl_lo, ymax = rl_hi,
                 colour = "Stability (≥99% of max height)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  labs(tag = "C", x = "Soil nitrogen (g N m⁻³)", y = "Optimal root length (m)") +
  pub_theme

# --- IQR ribbon across near-optimal combos (≥99% of max height) for panels D–I ---
# Ribbon shows spread of each output across the stability band;
# dots remain at the single height-maximising parameter combo.
stab_outputs <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_h    = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_h) %>%
  filter(near_opt) %>%
  summarise(
    ecm_lo        = min(ectomycorrhiza_mass, na.rm = TRUE),
    ecm_hi        = max(ectomycorrhiza_mass, na.rm = TRUE),
    ecm_area_lo   = min(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    ecm_area_hi   = max(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    height_lo     = min(height, na.rm = TRUE),
    height_hi     = max(height, na.rm = TRUE),
    nbar_lo       = min(N_bar, na.rm = TRUE),
    nbar_hi       = max(N_bar, na.rm = TRUE),
    cn_lo         = min(ectomycorrhiza_mass * 0.44 / pmax(ectomycorrhiza_N_biomass, 1e-12), na.rm = TRUE),
    cn_hi         = max(ectomycorrhiza_mass * 0.44 / pmax(ectomycorrhiza_N_biomass, 1e-12), na.rm = TRUE),
    gpp_lo        = min(assim_gross / crown_area, na.rm = TRUE),
    gpp_hi        = max(assim_gross / crown_area, na.rm = TRUE),
    n_upper_lo    = min(leaf_nitrogen_concentration, na.rm = TRUE),
    n_upper_hi    = max(leaf_nitrogen_concentration, na.rm = TRUE),
    n_realised_lo = min(optimal_leaf_nitrogen, na.rm = TRUE),
    n_realised_hi = max(optimal_leaf_nitrogen, na.rm = TRUE),
    n_free_lo     = min(tree_nitrogen, na.rm = TRUE),
    n_free_hi     = max(tree_nitrogen, na.rm = TRUE),
    .groups = "drop"
  )

# Mycorrhizal structural C:N at the optimal combo
opt_joint_cn <- opt_joint %>%
  mutate(CN = ectomycorrhiza_mass * 0.44 /
              pmax(ectomycorrhiza_N_biomass, 1e-12))

# --- Middle row: emergent biomass and height ---
p_joint_ecm <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = ecm_lo, ymax = ecm_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "D", x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_joint_ecm_area <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = ecm_area_lo, ymax = ecm_area_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "E", x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_joint_height <- ggplot(opt_joint, aes(x = soil_nitrogen, y = height)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = height_lo, ymax = height_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) +
  labs(tag = "H", x = "Soil nitrogen (g N m⁻³)", y = "Height (m)") +
  pub_theme

# --- Bottom row: N depletion and mycorrhizal economics ---
p_joint_nbar <- ggplot(opt_joint, aes(x = soil_nitrogen, y = N_bar)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = nbar_lo, ymax = nbar_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  labs(tag = "F", x = "Soil nitrogen (g N m⁻³)",
       y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  pub_theme

# Mycorrhizal structural C:N = 1/nc_myco = 10
# C (kg C) = ectomycorrhiza_mass * 0.44
# N (kg N) = ectomycorrhiza_N_biomass  (= mass * 0.44 * nc_myco, already in kg N)
p_joint_myco_cn <- ggplot(opt_joint_cn, aes(x = soil_nitrogen, y = CN)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = cn_lo, ymax = cn_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific(digits = 2)) +
  labs(tag = "G", x = "Soil nitrogen (g N m⁻³)",
       y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

p_joint_gpp <- ggplot(opt_joint, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = gpp_lo, ymax = gpp_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) +
  labs(tag = "I", x = "Soil nitrogen (g N m⁻³)",
       y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

# Nitrogen limitation diagnostic: upper bound (leaf N pool) vs realised (Phydro-optimised N)
# If the two lines converge, the tree is N-limited (all available N used for photosynthesis)
p_joint_n_limit <- ggplot(opt_joint, aes(x = soil_nitrogen)) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_upper_lo, ymax = n_upper_hi),
              fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_realised_lo, ymax = n_realised_hi),
              fill = "grey60", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
  geom_point(aes(y = leaf_nitrogen_concentration, colour = "Potential (leaf N pool)"),
             size = 4, stroke = 1.5) +
  geom_point(aes(y = optimal_leaf_nitrogen),
             colour = "grey30", shape = 1, size = 4, stroke = 1.5) +
  scale_colour_manual(values = c("Potential (leaf N pool)" = "steelblue"),
                       name = NULL) +
  expand_limits(y = 0) +
  labs(tag = "J", x = "Soil nitrogen (g N m⁻³)",
       y = "Leaf N concentration\n(kg N kg⁻¹ leaf)") +
  pub_theme

# Free nitrogen pool at maturity: if this is near nitrogen_start0 (0.039 kg) even at low soil N,
# the tree is running off its initial endowment rather than reaching a true N steady state.
p_joint_nfree <- ggplot(opt_joint, aes(x = soil_nitrogen, y = tree_nitrogen)) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_free_lo, ymax = n_free_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_hline(yintercept = 0.039, linetype = "dashed", colour = "grey40", linewidth = 0.8) +
  annotate("text", x = 0.1, y = 0.039, label = "nitrogen_start0", vjust = -0.4,
           hjust = 0, size = 3.5, colour = "grey40") +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) +
  labs(tag = "K", x = "Soil nitrogen (g N m⁻³)",
       y = "Free N pool at maturity\n(kg N tree⁻¹)") +
  pub_theme

make_row_title <- function(label) {
  wrap_elements(
    grid::textGrob(label,
                   gp = grid::gpar(fontsize = 20, fontface = "bold"),
                   hjust = 0.5, x = 0.5)
  )
}

# 4-column layout: A-C parameters, D-G mycorrhizal & N depletion, H-K growth & N status
# Stability bands (grey bars) on A-C show the range of params achieving ≥99% of max height
joint_optima <-
  guide_area() /
  make_row_title("Parameters to be Optimised") /
  (p_joint_ecto | p_joint_rn | p_joint_rl | plot_spacer()) /
  make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
  (p_joint_ecm | p_joint_ecm_area | p_joint_nbar | p_joint_myco_cn) /
  make_row_title("Model results: Tree Status") /
  (p_joint_height | p_joint_gpp | p_joint_n_limit | p_joint_nfree) +
  plot_layout(heights = c(0.15, 0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(theme = theme(plot.margin = margin(5, 30, 5, 5))) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 16),
        legend.key.size = unit(1.0, "lines"),
        legend.position    = "top",
        legend.direction   = "horizontal",
        legend.box         = "horizontal")
joint_optima

ggsave("plots/11_joint_optima.png", plot = joint_optima, width = 22, height = 18, dpi = 300)
```

## Joint optima: trajectories over time

```{r joint_optima_trajectories}
# Pull the full time series for the optimal parameter combo at each soil N level.
# opt_joint has one row per soil_nitrogen; inner_join recovers all timesteps for those combos.
opt_traj <- all_joint_df %>%
  inner_join(opt_joint %>% select(soil_nitrogen, ecto_allo, root_no, root_length),
             by = c("soil_nitrogen", "ecto_allo", "root_no", "root_length")) %>%
  mutate(date       = as.Date(date),
         CN         = ectomycorrhiza_mass * 0.44 /
                        pmax(ectomycorrhiza_N_biomass, 1e-12),
         gpp_per_ca = assim_gross / crown_area)

traj_colour <- scale_colour_viridis_c(option = "plasma",
                                      name   = "Soil N\n(g m⁻³)")

# --- Top row: optimal parameter values (constant per run, coloured by soil N) ---
p_traj_ecto <- ggplot(opt_traj,
    aes(x = date, y = ecto_allo, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = NULL, y = "Ectomycorrhizal\nallocation (%)") +
  pub_theme

p_traj_rn <- ggplot(opt_traj,
    aes(x = date, y = root_no, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Root number\nper crown area (no m⁻²)") +
  pub_theme

p_traj_rl <- ggplot(opt_traj,
    aes(x = date, y = root_length, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Root length (m)") +
  pub_theme

# --- Middle row: emergent biomass and height ---
p_traj_ecm <- ggplot(opt_traj,
    aes(x = date, y = ectomycorrhiza_mass, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_traj_ecm_area <- ggplot(opt_traj,
    aes(x = date, y = ectomycorrhiza_mass / crown_area,
        colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_traj_height <- ggplot(opt_traj,
    aes(x = date, y = height, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Height (m)") +
  pub_theme

# --- Bottom row: N depletion and mycorrhizal economics ---
p_traj_nbar <- ggplot(opt_traj,
    aes(x = date, y = N_bar, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  pub_theme

p_traj_cn <- ggplot(opt_traj,
    aes(x = date, y = CN, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

p_traj_gpp <- ggplot(opt_traj,
    aes(x = date, y = gpp_per_ca, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

p_traj_n_limit <- ggplot(opt_traj,
    aes(x = date, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(aes(y = leaf_nitrogen_concentration), linetype = "dashed", linewidth = 0.8) +
  geom_line(aes(y = optimal_leaf_nitrogen), linetype = "solid", linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Leaf N concentration\n(kg N kg⁻¹ leaf)") +
  pub_theme

p_traj_nfree <- ggplot(opt_traj,
    aes(x = date, y = tree_nitrogen, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Free N pool\n(kg N tree⁻¹)") +
  pub_theme

traj_tags <- list(c("", "A", "B", "C", "", "", "D", "E", "F", "G", "", "H", "I", "J", "K"))

joint_optima_traj <-
  make_row_title("Parameters to be Optimised") /
  (p_traj_ecto | p_traj_rn | p_traj_rl | plot_spacer()) /
  make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
  (p_traj_ecm | p_traj_ecm_area | p_traj_nbar | p_traj_cn) /
  make_row_title("Model results: Tree Status") /
  (p_traj_height | p_traj_gpp | p_traj_n_limit | p_traj_nfree) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(tag_levels = traj_tags) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 18),
        legend.key.size = unit(1.2, "lines"),
        legend.position = "right")
joint_optima_traj

ggsave("plots/11b_joint_optima_trajectories.png",
       plot = joint_optima_traj, width = 20, height = 18, dpi = 300)
```

## Data comparison

```{r data_load}
ecm_data_all <- read.csv(here("RScripts/ecm_root_data_compiled.csv"), stringsAsFactors = FALSE)

# Strip leading ~, <, > from all columns so approximate values become plain numbers
ecm_data_all <- as.data.frame(lapply(ecm_data_all, function(col) {
  if (is.character(col)) {
    cleaned <- gsub("^[~<>]+", "", trimws(col))
    num <- suppressWarnings(as.numeric(cleaned))
    if (sum(!is.na(num)) > sum(!is.na(suppressWarnings(as.numeric(col))))) num else col
  } else col
}))

# Model is calibrated for Scots pine / boreal evergreen conifer — keep only conifers
evergreen_conifers <- c("Pinus sylvestris", "Picea abies", "Picea abies (control)",
                        "Picea glauca", "Picea spp.", "Mixed (Pinus/Picea)")
ecm_data <- ecm_data_all[ecm_data_all$species %in% evergreen_conifers, ]
ecm_data <- ecm_data[is.na(ecm_data$soil_CN_ratio) | ecm_data$soil_CN_ratio < 100, ]
ecm_data$species <- factor(ecm_data$species, levels = evergreen_conifers)

species_colours <- c(
  "Pinus sylvestris"     = "#E69F00",
  "Picea abies"          = "#56B4E9",
  "Picea abies (control)"= "#009E73",
  "Picea glauca"         = "#F0E442",
  "Picea spp."           = "#0072B2",
  "Mixed (Pinus/Picea)"  = "#D55E00"
)
species_scale <- scale_colour_manual(name = "Species", values = species_colours,
                                     drop = FALSE,
                                     guide = guide_legend(ncol = 3, byrow = TRUE))

# Tip density: convert million m⁻² → m⁻² to match model root_no units
ecm_data$tip_density_per_m2 <- ecm_data$ecm_tip_density_million_per_m2_soil * 1e6
```

### Plot 07 A — Root length vs soil N
*Matches heatmap p2 (y-axis: root length, mm)*

```{r data_07a}
d_rl <- ecm_data[!is.na(ecm_data$ecm_root_tip_length_mm) &
                 !is.na(ecm_data$soil_CN_ratio), ]

p_data_rl <- ggplot(d_rl, aes(x = soil_CN_ratio, y = ecm_root_tip_length_mm,
                               colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  labs(
    x = "Soil C:N ratio",
    y = "ECM root tip\nlength (mm)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 07 B — Root / tip density vs soil N
*Matches heatmap p1 (y-axis: root number per crown area, m⁻²)*

```{r data_07b}
d_td <- ecm_data[!is.na(ecm_data$ecm_tip_density_million_per_m2_BA) &
                 !is.na(ecm_data$soil_CN_ratio), ]

p_data_td <- ggplot(d_td, aes(x = soil_CN_ratio, y = ecm_tip_density_million_per_m2_BA,
                               colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  scale_y_log10(labels = scales::label_scientific()) +
  labs(
    x = "Soil C:N ratio",
    y = "ECM tip density\n(million m⁻² basal area)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 09 A — ECM mycelium biomass vs soil N
*Matches heatmap p_ecm_mass (y-axis: ectomycorrhiza_mass)*

```{r data_09a}
d_ecm <- ecm_data[!is.na(ecm_data$ecm_mycelium_biomass_kg_ha) &
                  !is.na(ecm_data$soil_CN_ratio), ]

p_data_ecm <- ggplot(d_ecm, aes(x = soil_CN_ratio, y = ecm_mycelium_biomass_kg_ha,
                                 colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  scale_y_log10(labels = scales::label_number(accuracy = 1)) +
  labs(
    x = "Soil C:N ratio",
    y = "ECM mycelium\nbiomass (kg ha⁻¹)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 09 B — ECM root biomass vs soil N
*Proxy for ectomycorrhizal allocation (p_ecto_allo)*

```{r data_09b}
d_ecmr <- ecm_data[!is.na(ecm_data$ecm_root_biomass_kg_per_m2_basal_area) &
                   !is.na(ecm_data$soil_CN_ratio), ]

p_data_ecmr <- ggplot(d_ecmr, aes(x = soil_CN_ratio,
                                    y = ecm_root_biomass_kg_per_m2_basal_area,
                                    colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  labs(
    x = "Soil C:N ratio",
    y = "ECM root biomass\n(kg m⁻² basal area)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### ECM colonisation percentage vs soil N

```{r data_colonisation}
d_col_plot <- ecm_data[!is.na(ecm_data$ecm_colonisation_pct) &
                        !is.na(ecm_data$soil_CN_ratio), ]

p_data_col <- ggplot(d_col_plot, aes(x = soil_CN_ratio, y = ecm_colonisation_pct,
                                      colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  coord_cartesian(ylim = c(NA, 100)) +
  labs(
    x      = "Soil C:N ratio",
    y      = "ECM colonisation (%)",
    colour = "Species"
  ) +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 13)) +
  pub_theme
```

```{r data_combined}
# Standalone legend using all species so the hash is consistent
legend_plot <- ggplot(
    data.frame(species = factor(evergreen_conifers, levels = evergreen_conifers)),
    aes(x = 1, y = 1, colour = species)) +
  geom_point(alpha = 0) +
  species_scale +
  theme_void() +
  theme(legend.position   = c(0.5, 0.5),
        legend.text       = element_text(size = 22),
        legend.title      = element_text(size = 24, face = "bold"),
        legend.key.size   = unit(1.5, "lines"),
        axis.text.x       = element_blank(),
        axis.text.y       = element_blank(),
        axis.title.x      = element_blank(),
        axis.title.y      = element_blank(),
        axis.ticks        = element_blank(),
        panel.background  = element_blank()) +
  guides(colour = guide_legend(ncol = 1,
                               override.aes = list(alpha = 1, size = 5)))

data_combined <- wrap_plots(
  p_data_rl   + theme(legend.position = "none"),
  p_data_td   + theme(legend.position = "none"),
  p_data_ecm  + theme(legend.position = "none"),
  p_data_ecmr + theme(legend.position = "none"),
  p_data_col  + theme(legend.position = "none"),
  legend_plot,
  ncol = 3
) + plot_annotation(tag_levels = list(c("A","B","C","D","E",""))) &
  theme(plot.tag   = element_text(size = 28, face = "bold"),
        axis.text  = element_text(size = 24),
        axis.title = element_text(size = 26, face = "bold"))
data_combined

ggsave("plots/data_combined.png",
       plot = data_combined, width = 24, height = 14, dpi = 300)
```

