---
title: "Publication Plots"
author: "Joanna"
date: "2026-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(data.table)
library(rphydro)
library(Rprebasso)
library(patchwork)
library(PlantFATE)
library(akima)
library(here)
library(future.apply)
library(future)
```

```{r theme, include=FALSE}
pub_theme <- theme_classic(base_size = 22) +
  theme(
    axis.title       = element_text(size = 22, face = "bold"),
    axis.text        = element_text(size = 20, colour = "black"),
    axis.line        = element_line(colour = "black", linewidth = 0.6),
    axis.ticks       = element_line(colour = "black", linewidth = 0.5),
    legend.title     = element_text(size = 22, face = "bold"),
    legend.text      = element_text(size = 20),
    legend.key.size  = unit(0.9, "lines"),
    panel.grid.major = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background  = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.margin      = margin(8, 12, 8, 8),
    strip.text       = element_text(size = 22, face = "bold")
  )
```

## Data reading

```{r}
hyde_all = "~/Documents/Austria/Hyytiala_all_data/"
data <- lapply(paste0(hyde_all, list.files(hyde_all))[1:9], read.csv)
names(data) <- list.files(hyde_all)[1:9]
useful_data <- merge(data[[2]], data[[5]], c('plotID', 'eventID', 'eventYear'), all=TRUE)

```

## Simulations

```{r pressure, echo=FALSE}
# Set to TRUE for fast testing (fewer grid points, shorter run), FALSE for full publication run
testing_mode <- TRUE

plan(multicore, workers = parallel::detectCores() - 1)

# Fixed baseline values
fixed_root_length <- 1.5
fixed_root_no     <- 3e5
fixed_ecto_allo   <- 0.1

# Parameter ranges (fewer points in testing mode for speed)
n_pts_main  <- if (testing_mode) 10 else 21
n_pts_ecto  <- if (testing_mode) 10 else 11
n_pts_joint <- if (testing_mode) 5  else 8   # per-axis for 3D co-optimisation sweep

root_length_values <- seq(0.05, 6, length.out = n_pts_main)
root_no_values     <- 3*10^seq(3, 6, length.out = n_pts_main)
nitrogen_values    <- seq(0.01, 2.0, length.out = n_pts_main)
ecto_allo_values   <- seq(0.0, 0.3, length.out = n_pts_ecto)

# Joint co-optimisation grid (coarser to keep run count tractable)
joint_ecto_allo_values   <- seq(0.0, 0.3, length.out = n_pts_joint)
joint_root_no_values     <- 3*10^seq(3, 6, length.out = n_pts_joint)
joint_root_length_values <- seq(0.05, 6, length.out = n_pts_joint)

param_file <- "/home/josimms/Documents/Austria/Plant-FATE/tests/params/p_test_boreal.ini"
weather_file <- "/home/josimms/Documents/Austria/Plant-FATE/data/ERAS_Monthly.csv"

# -------------------------------
# Core simulation function with safe error handling
# -------------------------------
run_sim_parallel <- function(rl, rn, n, eb) {
  tryCatch({
    lho <- new(LifeHistoryOptimizer, param_file)
    lho$set_i_metFile(weather_file)
    lho$set_a_metFile(weather_file)
    lho$set_co2File("")
    lho$set_soil_nitrogen(n)
    lho$traits0$investment_from_tree <- eb
    #lho$uptake0$mycorrhized <- 0.7 + 0.2*(eb)/0.3
    lho$init()
    lho$root_override(rn, rl)

    dt <- 1/12
    start_year <- 1960
    end_year   <- 2022
    years_seq  <- seq(start_year, end_year, dt)
    
    results <- lapply(years_seq, function(t) {
      tryCatch({
        lho$grow_for_dt(t, dt)
        state <- lho$get_state(t + dt)
        df_row <- as.data.frame(t(state))
        names(df_row) <- lho$get_header()
        df_row
      }, error = function(e) {
        setNames(as.data.frame(as.list(rep(NA, length(lho$get_header())))),
                 lho$get_header())
      })
    })
    
    df <- do.call(rbind, results)
    df$date <- seq(as.Date(paste0(start_year, "-01-01")),
                   as.Date(paste0(end_year, "-01-01")),
                   by = "month")[1:nrow(df)]
    df$root_length   <- rl
    df$root_no       <- rn
    df$soil_nitrogen <- n
    df$ecto_allo     <- eb
    df
  }, error = function(e) {
    message("Failed run_sim_parallel: rl=", rl, " rn=", rn, " n=", n, " eb=", eb)
    NULL
  })
}

# -------------------------------
# Helper: safe wrapper for run_sim_parallel
# -------------------------------
run_sim_safe <- function(rl, rn, n, eb) {
  result <- run_sim_parallel(rl, rn, n, eb)
  if (is.null(result) || nrow(result) == 0) return(NULL)
  result
}

# -------------------------------
# Sweep Nitrogen (root traits fixed)
# -------------------------------
all_n_list <- future_lapply(nitrogen_values, function(n) {
  run_sim_safe(fixed_root_length, fixed_root_no, n, fixed_ecto_allo)
})
all_n_df <- do.call(rbind, all_n_list[!sapply(all_n_list, is.null)])

# -------------------------------
# Sweep Root Number × Nitrogen
# -------------------------------
all_rn_list <- future_lapply(root_no_values, function(rn) {
  inner <- lapply(nitrogen_values, function(n) {
    run_sim_safe(fixed_root_length, rn, n, fixed_ecto_allo)
  })
  do.call(rbind, inner[!sapply(inner, is.null)])
})
all_rn_df <- do.call(rbind, all_rn_list[!sapply(all_rn_list, is.null)])

# -------------------------------
# Sweep Root Length × Nitrogen
# -------------------------------
all_rl_list <- future_lapply(root_length_values, function(rl) {
  inner <- lapply(nitrogen_values, function(n) {
    run_sim_safe(rl, fixed_root_no, n, fixed_ecto_allo)
  })
  do.call(rbind, inner[!sapply(inner, is.null)])
})
all_rl_df <- do.call(rbind, all_rl_list[!sapply(all_rl_list, is.null)])

# -------------------------------
# Sweep Ectomycorrhizal Allocation × Nitrogen
# -------------------------------
all_eb_list <- future_lapply(ecto_allo_values, function(eb) {
  inner <- lapply(nitrogen_values, function(n) {
    run_sim_safe(fixed_root_length, fixed_root_no, n, eb)
  })
  do.call(rbind, inner[!sapply(inner, is.null)])
})
all_eb_df <- do.call(rbind, all_eb_list[!sapply(all_eb_list, is.null)])

# -------------------------------
# Joint co-optimisation: ecto_allo × root_no × root_length × nitrogen
# n_pts_joint^3 × n_pts_main runs total
# Parallelised over ecto_allo (outer axis)
# -------------------------------
all_joint_list <- future_lapply(joint_ecto_allo_values, function(eb) {
  inner_rn <- lapply(joint_root_no_values, function(rn) {
    inner_rl <- lapply(joint_root_length_values, function(rl) {
      inner_n <- lapply(nitrogen_values, function(n) {
        run_sim_safe(rl, rn, n, eb)
      })
      do.call(rbind, inner_n[!sapply(inner_n, is.null)])
    })
    do.call(rbind, inner_rl[!sapply(inner_rl, is.null)])
  })
  do.call(rbind, inner_rn[!sapply(inner_rn, is.null)])
})
all_joint_df <- do.call(rbind, all_joint_list[!sapply(all_joint_list, is.null)])

# -------------------------------
# Combine all sweeps into one dataframe
# -------------------------------
final_df <- rbind(all_rn_df, all_rl_df, all_eb_df)
final_df_no_na <- final_df[!is.na(final_df$root_length), ]
```

# Final

```{r}
# --- Convert date column to Date class if needed ---
final_df_no_na$date <- as.Date(final_df_no_na$date)

# --- Subset for plotting: each filtered to its own sweep only ---
# root_no sweep: fixed root_length=1.5 and ecto_allo=fixed_ecto_allo
subset_root_length <- all_rn_df %>%
  filter(date == as.Date("2022-01-01"))

# root_length sweep: fixed root_no=3e5 and ecto_allo=fixed_ecto_allo
subset_root_no <- all_rl_df %>%
  filter(date == as.Date("2022-01-01"))

# --- Determine height range for color scaling ---
height_range <- range(
  c(subset_root_length$height, subset_root_no$height),
  na.rm = TRUE
)

# --- Compute max height points ---
max_line_root_length <- subset_root_length %>%
  group_by(soil_nitrogen) %>%
  slice_max(height, n = 1, with_ties = FALSE, na_rm = TRUE) %>%
  ungroup()

max_line_root_no <- subset_root_no %>%
  group_by(soil_nitrogen) %>%
  slice_max(height, n = 1, with_ties = FALSE, na_rm = TRUE) %>%
  ungroup()

# --- Plot 1: Root Number heatmap ---
p1 <- ggplot(subset_root_length, aes(x = soil_nitrogen, y = root_no, fill = height)) +
  geom_tile() +
  scale_shape_manual(name = NULL, values = c("Max height" = 21)) +
  annotate("segment", x = 0, y = 3e6, xend = 2, yend = 1e5,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_line_root_length,
             aes(x = soil_nitrogen, y = root_no, shape = "Max height"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_fill_viridis_c(option = "cividis", limits = height_range,
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root number per unit\ncrown area (m⁻²)",
       fill = "Height (m)") +
  pub_theme

# --- Plot 2: Root Length heatmap ---
p2 <- ggplot(subset_root_no, aes(x = soil_nitrogen, y = root_length, fill = height)) +
  geom_tile() +
  annotate("segment", x = 0, y = 5, xend = 2, yend = 1.2,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_line_root_no,
             aes(x = soil_nitrogen, y = root_length, shape = "Max height"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max height" = 21)) +
  scale_fill_viridis_c(option = "cividis", limits = height_range,
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root length (mm)",
       fill = "Height (m)") +
  pub_theme

# --- Combine plots with panel labels and shared legend ---
length_no_N_gradient <- (p2 / p1) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
length_no_N_gradient

ggsave("plots/07_length_no_N_gradient.png",
       plot = length_no_N_gradient, width = 12, height = 10, dpi = 300)
```

## Trade offs: Root traits – Uptake and Assimilation

```{r final_uptake_assimilation}
# Rate variables (uptake, GPP) are 0 in January due to frozen soil (f_temp = 0).
# Use June-August growing-season mean instead of the single January snapshot.
gs_rn <- all_rn_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8) %>%
  group_by(soil_nitrogen, root_no) %>%
  summarise(
    root_uptake = mean(root_uptake, na.rm = TRUE),
    myco_uptake = mean(myco_uptake, na.rm = TRUE),
    assim_gross = mean(assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

gs_rl <- all_rl_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8) %>%
  group_by(soil_nitrogen, root_length) %>%
  summarise(
    root_uptake = mean(root_uptake, na.rm = TRUE),
    myco_uptake = mean(myco_uptake, na.rm = TRUE),
    assim_gross = mean(assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

root_uptake_range <- range(c(gs_rn$root_uptake, gs_rl$root_uptake), na.rm = TRUE)
myco_uptake_range <- range(c(gs_rn$myco_uptake, gs_rl$myco_uptake), na.rm = TRUE)
assim_range       <- range(c(gs_rn$assim_gross,  gs_rl$assim_gross),  na.rm = TRUE)

# Max-value markers for each variable independently
max_rn_root_uptake <- gs_rn %>% group_by(soil_nitrogen) %>%
  slice_max(root_uptake, n = 1, with_ties = FALSE) %>% ungroup()
max_rl_root_uptake <- gs_rl %>% group_by(soil_nitrogen) %>%
  slice_max(root_uptake, n = 1, with_ties = FALSE) %>% ungroup()

max_rn_myco_uptake <- gs_rn %>% group_by(soil_nitrogen) %>%
  slice_max(myco_uptake, n = 1, with_ties = FALSE) %>% ungroup()
max_rl_myco_uptake <- gs_rl %>% group_by(soil_nitrogen) %>%
  slice_max(myco_uptake, n = 1, with_ties = FALSE) %>% ungroup()

max_rn_assim <- gs_rn %>% group_by(soil_nitrogen) %>%
  slice_max(assim_gross, n = 1, with_ties = FALSE) %>% ungroup()
max_rl_assim <- gs_rl %>% group_by(soil_nitrogen) %>%
  slice_max(assim_gross, n = 1, with_ties = FALSE) %>% ungroup()

p1_root_uptake <- ggplot(gs_rn, aes(x = soil_nitrogen, y = root_no, fill = root_uptake)) +
  geom_tile() +
  annotate("segment", x = 0, y = 3e6, xend = 2, yend = 1e5,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_rn_root_uptake,
             aes(x = soil_nitrogen, y = root_no, shape = "Max root uptake"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max root uptake" = 21)) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_fill_viridis_c(option = "magma", limits = root_uptake_range,
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root number per unit\ncrown area (m⁻²)",
       fill = "Root Uptake\n(kg N yr⁻¹)") +
  pub_theme

p2_root_uptake <- ggplot(gs_rl, aes(x = soil_nitrogen, y = root_length, fill = root_uptake)) +
  geom_tile() +
  annotate("segment", x = 0, y = 5, xend = 2, yend = 1.2,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_rl_root_uptake,
             aes(x = soil_nitrogen, y = root_length, shape = "Max root uptake"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max root uptake" = 21)) +
  scale_fill_viridis_c(option = "magma", limits = root_uptake_range,
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root length (mm)",
       fill = "Root Uptake\n(kg N yr⁻¹)") +
  pub_theme

p1_myco_uptake <- ggplot(gs_rn, aes(x = soil_nitrogen, y = root_no, fill = myco_uptake)) +
  geom_tile() +
  annotate("segment", x = 0, y = 3e6, xend = 2, yend = 1e5,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_rn_myco_uptake,
             aes(x = soil_nitrogen, y = root_no, shape = "Max myco uptake"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max myco uptake" = 21)) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_fill_viridis_c(option = "magma", limits = myco_uptake_range,
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root number per unit\ncrown area (m⁻²)",
       fill = "Myco Uptake\n(kg N yr⁻¹)") +
  pub_theme

p2_myco_uptake <- ggplot(gs_rl, aes(x = soil_nitrogen, y = root_length, fill = myco_uptake)) +
  geom_tile() +
  annotate("segment", x = 0, y = 5, xend = 2, yend = 1.2,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_rl_myco_uptake,
             aes(x = soil_nitrogen, y = root_length, shape = "Max myco uptake"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max myco uptake" = 21)) +
  scale_fill_viridis_c(option = "magma", limits = myco_uptake_range,
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root length (mm)",
       fill = "Myco Uptake\n(kg N yr⁻¹)") +
  pub_theme

p1_assim <- ggplot(gs_rn, aes(x = soil_nitrogen, y = root_no, fill = assim_gross)) +
  geom_tile() +
  annotate("segment", x = 0, y = 3e6, xend = 2, yend = 1e5,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_rn_assim,
             aes(x = soil_nitrogen, y = root_no, shape = "Max GPP"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max GPP" = 21)) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_fill_viridis_c(option = "plasma", limits = assim_range,
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root number per unit\ncrown area (m⁻²)",
       fill = "GPP\n(kg C yr⁻¹)") +
  pub_theme

p2_assim <- ggplot(gs_rl, aes(x = soil_nitrogen, y = root_length, fill = assim_gross)) +
  geom_tile() +
  annotate("segment", x = 0, y = 5, xend = 2, yend = 1.2,
           colour = "white", linewidth = 1.2) +
  geom_point(data = max_rl_assim,
             aes(x = soil_nitrogen, y = root_length, shape = "Max GPP"),
             fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_shape_manual(name = NULL, values = c("Max GPP" = 21)) +
  scale_fill_viridis_c(option = "plasma", limits = assim_range,
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(x = "Soil N (g m⁻³)",
       y = "Root length (mm)",
       fill = "GPP\n(kg C yr⁻¹)") +
  pub_theme

length_no_N_uptake <- ((p2_root_uptake + p2_myco_uptake) / (p1_root_uptake + p1_myco_uptake)) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
length_no_N_uptake

ggsave("plots/07b_length_no_N_uptake.png",
       plot = length_no_N_uptake, width = 12, height = 10, dpi = 300)

length_no_N_assim <- (p2_assim / p1_assim) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
length_no_N_assim

ggsave("plots/07c_length_no_N_assimilation.png",
       plot = length_no_N_assim, width = 12, height = 10, dpi = 300)
```

## Trade offs: Root traits – N Uptake and GPP (publication format)

```{r root_trait_combined}
# Growing-season (Jun–Aug) means, matching ecto_allo section format
heatmap_rn_full <- all_rn_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8) %>%
  group_by(soil_nitrogen, root_no) %>%
  summarise(
    height          = mean(height, na.rm = TRUE),
    nitrogen_uptake = mean(root_uptake + mycorrhizal_export_to_tree, na.rm = TRUE),
    assim_gross     = mean(assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

heatmap_rl_full <- all_rl_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8) %>%
  group_by(soil_nitrogen, root_length) %>%
  summarise(
    height          = mean(height, na.rm = TRUE),
    nitrogen_uptake = mean(root_uptake + mycorrhizal_export_to_tree, na.rm = TRUE),
    assim_gross     = mean(assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

rn_uptake_range <- range(heatmap_rn_full$nitrogen_uptake, na.rm = TRUE)
rn_assim_range  <- range(heatmap_rn_full$assim_gross,     na.rm = TRUE)
rl_uptake_range <- range(heatmap_rl_full$nitrogen_uptake, na.rm = TRUE)
rl_assim_range  <- range(heatmap_rl_full$assim_gross,     na.rm = TRUE)

opt_rn_uptake <- heatmap_rn_full %>% group_by(soil_nitrogen) %>%
  slice_max(nitrogen_uptake, n = 1, with_ties = FALSE) %>% ungroup()
opt_rn_assim  <- heatmap_rn_full %>% group_by(soil_nitrogen) %>%
  slice_max(assim_gross, n = 1, with_ties = FALSE) %>% ungroup()

opt_rl_uptake <- heatmap_rl_full %>% group_by(soil_nitrogen) %>%
  slice_max(nitrogen_uptake, n = 1, with_ties = FALSE) %>% ungroup()
opt_rl_assim  <- heatmap_rl_full %>% group_by(soil_nitrogen) %>%
  slice_max(assim_gross, n = 1, with_ties = FALSE) %>% ungroup()

# --- Root number: N uptake and GPP ---
p_rn_uptake <- ggplot(heatmap_rn_full,
       aes(x = soil_nitrogen, y = root_no, fill = nitrogen_uptake)) +
  geom_tile() +
  geom_point(data = opt_rn_uptake,
             aes(x = soil_nitrogen, y = root_no),
             shape = 21, fill = "white", colour = "black", size = 3.5, stroke = 1.5,
             inherit.aes = FALSE) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_fill_viridis_c(option = "magma",
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Root number per unit\ncrown area (m⁻²)",
       fill = "N uptake\n(kg N yr⁻¹)") +
  pub_theme

p_rn_assim <- ggplot(heatmap_rn_full,
       aes(x = soil_nitrogen, y = root_no, fill = assim_gross)) +
  geom_tile() +
  geom_point(data = opt_rn_assim,
             aes(x = soil_nitrogen, y = root_no),
             shape = 21, fill = "white", colour = "black", size = 3.5, stroke = 1.5,
             inherit.aes = FALSE) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_fill_viridis_c(option = "plasma",
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Root number per unit\ncrown area (m⁻²)",
       fill = "GPP\n(kg C yr⁻¹)") +
  pub_theme

root_no_gradient <- (p_rn_uptake / p_rn_assim) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
root_no_gradient

ggsave("plots/10_root_no_gradient.png",
       plot = root_no_gradient, width = 12, height = 10, dpi = 300)

# --- Root length: N uptake and GPP ---
p_rl_uptake <- ggplot(heatmap_rl_full,
       aes(x = soil_nitrogen, y = root_length, fill = nitrogen_uptake)) +
  geom_tile() +
  geom_point(data = opt_rl_uptake,
             aes(x = soil_nitrogen, y = root_length),
             shape = 21, fill = "white", colour = "black", size = 3.5, stroke = 1.5,
             inherit.aes = FALSE) +
  scale_fill_viridis_c(option = "magma",
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Root length (mm)",
       fill = "N uptake\n(kg N yr⁻¹)") +
  pub_theme

p_rl_assim <- ggplot(heatmap_rl_full,
       aes(x = soil_nitrogen, y = root_length, fill = assim_gross)) +
  geom_tile() +
  geom_point(data = opt_rl_assim,
             aes(x = soil_nitrogen, y = root_length),
             shape = 21, fill = "white", colour = "black", size = 3.5, stroke = 1.5,
             inherit.aes = FALSE) +
  scale_fill_viridis_c(option = "plasma",
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Root length (mm)",
       fill = "GPP\n(kg C yr⁻¹)") +
  pub_theme

root_length_gradient <- (p_rl_uptake / p_rl_assim) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
root_length_gradient

ggsave("plots/10b_root_length_gradient.png",
       plot = root_length_gradient, width = 12, height = 10, dpi = 300)
```

## Trajectory

```{r}
# --- Soil N values to plot (min, median, max) ---
soilN_values <- nitrogen_values[seq(1, length(nitrogen_values), length.out = 5)]

# ---- Maximum-Fitness Root Number Trajectory (all ties) ----
trajectory_max_root_no <- all_rn_df %>%
  filter(soil_nitrogen %in% soilN_values) %>%
  group_by(date, soil_nitrogen) %>%
  slice_max(height, with_ties = TRUE, na_rm = TRUE) %>%
  ungroup() %>%
  select(date, soil_nitrogen, root_no)

# ---- Maximum-Fitness Root Length Trajectory (all ties) ----
trajectory_max_root_length <- all_rl_df %>%
  filter(soil_nitrogen %in% soilN_values) %>%
  group_by(date, soil_nitrogen) %>%
  slice_max(height, with_ties = TRUE, na_rm = TRUE) %>%
  ungroup() %>%
  select(date, soil_nitrogen, root_length)

# ---- Format soil N as a labelled factor ----
soil_n_labels <- setNames(
  paste(round(soilN_values, 2), "g m⁻³"),
  as.character(soilN_values)
)

# ---- Plot Root Number trajectory ----
p1 <- ggplot(trajectory_max_root_no,
             aes(x = date, y = root_no,
                 colour = factor(soil_nitrogen),
                 group = soil_nitrogen)) +
  geom_point(alpha = 0.7, size = 2,
             position = position_jitter(width = 0.1, height = 0)) +
  scale_y_log10(limits = range(root_no_values),
                labels = scales::label_scientific()) +
  scale_colour_viridis_d(option = "cividis",
                         labels = soil_n_labels) +
  labs(
    x = "Year",
    y = "Optimal root number\nper unit crown area (m⁻²)",
    colour = "Soil N"
  ) +
  pub_theme

# ---- Plot Root Length trajectory ----
p2 <- ggplot(trajectory_max_root_length,
             aes(x = date, y = root_length,
                 colour = factor(soil_nitrogen),
                 group = soil_nitrogen)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_y_continuous(limits = range(root_length_values)) +
  scale_colour_viridis_d(option = "cividis",
                         labels = soil_n_labels) +
  labs(
    x = "Year",
    y = "Optimal root length (mm)",
    colour = "Soil N"
  ) +
  pub_theme

# ---- Combine with panel labels and shared legend ----
trajectory_with_age <- (p2 / p1) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
trajectory_with_age

ggsave("plots/08_trajectory_with_age.png",
       plot = trajectory_with_age, width = 12, height = 10, dpi = 300)
```

## Trade offs: Ectomycorrhizal biomass

```{r}
subset_df <- final_df_no_na %>%
  filter(root_length == 1.5, root_no == 3e5)

# Use max ECM biomass across all timesteps per run (captures seasonal peaks)
sweep_max_ecm <- subset_df %>%
  group_by(soil_nitrogen, ecto_allo) %>%
  summarise(
    ectomycorrhiza_mass = max(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    height          = max(height, na.rm = TRUE),
    .groups = "drop"
  )

heatmap_ecm <- sweep_max_ecm %>%
  mutate(
    bin_soilN = cut(soil_nitrogen, breaks = 50),
    bin_ecm   = cut(log10(ectomycorrhiza_mass), breaks = 50)
  ) %>%
  group_by(bin_soilN, bin_ecm) %>%
  summarise(
    height          = mean(height, na.rm = TRUE),
    soil_nitrogen       = mean(soil_nitrogen),
    ectomycorrhiza_mass = mean(ectomycorrhiza_mass),
    .groups = "drop"
  )

ecm_height_range <- range(heatmap_ecm$height, na.rm = TRUE)

# Optimal: ecto_allo that maximises height at each soil_nitrogen
opt_ecm_mass <- sweep_max_ecm %>%
  group_by(soil_nitrogen) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup()

p_ecm_mass <- ggplot(heatmap_ecm,
       aes(x = soil_nitrogen,
           y = ectomycorrhiza_mass,
           fill = height)) +
  geom_point(shape = 21, size = 3.5, colour = NA) +
  geom_point(data = opt_ecm_mass,
             aes(x = soil_nitrogen, y = ectomycorrhiza_mass),
             shape = 21, fill = "white", colour = "black", size = 3.5,
             stroke = 1.5, inherit.aes = FALSE) +
  scale_fill_viridis_c(option = "cividis", limits = ecm_height_range,
                       labels = scales::label_number(accuracy = 0.1)) +
  scale_y_log10(labels = scales::label_number(accuracy = 0.001)) +
  labs(
    x = "Soil nitrogen (g N m⁻³)",
    y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)",
    fill = "Height (m)"
  ) +
  pub_theme

p_ecm_vs_soilN <- ggplot(opt_ecm_mass,
       aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  #geom_line(linewidth = 1.2, colour = "#404080") +
  geom_point(shape = 21, fill = "white", colour = "black", size = 3.5, stroke = 1.5) +
  scale_y_log10(labels = scales::label_number(accuracy = 0.001)) +
  labs(
    x = "Soil nitrogen (g N m⁻³)",
    y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)"
  ) +
  pub_theme

heatmap_ecto_allo <- subset_df %>%
  group_by(soil_nitrogen, ecto_allo) %>%
  summarise(
    height = mean(height, na.rm = TRUE),
    .groups = "drop"
  )

# Optimal ecto allocation: ecto_allo that maximises height at each soil_nitrogen
opt_ecto_allo <- heatmap_ecto_allo %>%
  group_by(soil_nitrogen) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup()

p_ecto_allo <- ggplot(heatmap_ecto_allo,
       aes(x = soil_nitrogen,
           y = ecto_allo,
           fill = height)) +
  geom_tile() +
  geom_point(data = opt_ecto_allo,
             aes(x = soil_nitrogen, y = ecto_allo),
             shape = 21, fill = "white", colour = "black", size = 3.5,
             stroke = 1.5, inherit.aes = FALSE) +
  annotate("segment", x = 0, y = 0.30, xend = 2, yend = 0.10,
           colour = "white", linewidth = 1.2) +
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_viridis_c(option = "cividis", limits = ecm_height_range,
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(
    x = "Soil nitrogen (g N m⁻³)",
    y = "Ectomycorrhizal\nallocation",
    fill = "Height (m)"
  ) +
  pub_theme

ectomycorrhizal_beha_gradient <- (p_ecm_mass + p_ecm_vs_soilN) / (p_ecto_allo + guide_area()) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
ectomycorrhizal_beha_gradient

ggsave("plots/09_ectomycorrhizal_beha_gradient.png",
       plot = ectomycorrhizal_beha_gradient, width = 16, height = 12, dpi = 300)
```

## Co-optimised trait optima

```{r joint_optimum}
# June–August 2021 growing-season mean per parameter combination.
# Rate variables (N uptake, GPP) are zero in winter (f_temp = 0), so January
# snapshots understate uptake; seasonal means give a fair comparison.
joint_gs <- all_joint_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  summarise(
    height                     = mean(height,                     na.rm = TRUE),
    ectomycorrhiza_mass        = mean(ectomycorrhiza_mass,        na.rm = TRUE),
    ectomycorrhiza_N_biomass   = mean(ectomycorrhiza_N_biomass,   na.rm = TRUE),
    crown_area                 = mean(crown_area,                 na.rm = TRUE),
    N_bar                      = mean(N_bar,                      na.rm = TRUE),
    C_export_to_myco           = mean(C_export_to_myco,           na.rm = TRUE),
    mycorrhizal_export_to_tree = mean(mycorrhizal_export_to_tree, na.rm = TRUE),
    root_uptake                = mean(root_uptake,                na.rm = TRUE),
    assim_gross                = mean(assim_gross,                na.rm = TRUE),
    .groups = "drop"
  )

opt_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup()

# --- Stability bands: all combos within 1% of maximum height per soil N ---
# This shows how flat the fitness landscape is around each optimum
stab_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_h = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_h) %>%
  filter(near_opt) %>%
  summarise(
    ecto_lo = min(ecto_allo,   na.rm = TRUE), ecto_hi = max(ecto_allo,   na.rm = TRUE),
    rn_lo   = min(root_no,     na.rm = TRUE), rn_hi   = max(root_no,     na.rm = TRUE),
    rl_lo   = min(root_length, na.rm = TRUE), rl_hi   = max(root_length, na.rm = TRUE),
    .groups = "drop"
  )

joint_height_range <- range(opt_joint$height, na.rm = TRUE)

shape_vals <- c("Parameters"   = 3, "Model output" = 1)
ltype_vals <- c("Trend line"   = "dashed")

# Produces a 2-row data frame for a straight trend-line segment
tl <- function(x1, y1, x2, y2) data.frame(x = c(x1, x2), y = c(y1, y2))

colour_vals <- c("Stability (≥99% of max height)" = "grey70")

shared_scales <- list(
  scale_shape_manual(values = shape_vals, name = NULL),
  scale_linetype_manual(values = ltype_vals, name = NULL),
  scale_colour_manual(values = colour_vals, name = NULL)
)

# --- Top row: crosses + stability bars, with dashed trend line ---
p_joint_ecto <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ecto_allo)) +
  geom_line(data = tl(0, 0.30, 2, 0.05), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = ecto_lo, ymax = ecto_hi,
                 colour = "Stability (≥99% of max height)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Parameters"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = "Soil nitrogen (g N m⁻³)", y = "Optimal ectomycorrhizal\nallocation (%)") +
  pub_theme

p_joint_rn <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_no)) +
  geom_line(data = tl(0, 3e6, 2, 1e5), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rn_lo, ymax = rn_hi,
                 colour = "Stability (≥99% of max height)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Parameters"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(x = "Soil nitrogen (g N m⁻³)", y = "Optimal root number\nper crown area (no m⁻²)") +
  pub_theme

p_joint_rl <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_length)) +
  geom_line(data = tl(0, 5, 2, 1.2), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rl_lo, ymax = rl_hi,
                 colour = "Stability (≥99% of max height)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Parameters"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  labs(x = "Soil nitrogen (g N m⁻³)", y = "Optimal root length (m)") +
  pub_theme

# --- Middle row: emergent biomass and height ---
p_joint_ecm <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_log10(labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_joint_ecm_area <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_log10(labels = scales::label_number(accuracy = 0.001)) +
  labs(x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_joint_height <- ggplot(opt_joint, aes(x = soil_nitrogen, y = height)) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  labs(x = "Soil nitrogen (g N m⁻³)", y = "Height (m)") +
  pub_theme

# --- Bottom row: N depletion and mycorrhizal economics ---
p_joint_nbar <- ggplot(opt_joint, aes(x = soil_nitrogen, y = N_bar)) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  pub_theme

# Mycorrhizal C:N ratio — same formula as the myco_cn_ratio chunk:
# C = ectomycorrhiza_mass * 0.44 (carbon fraction of biomass)
# N = ectomycorrhiza_N_biomass * 0.44 (as computed in gs_eb)
opt_joint_cn <- opt_joint %>%
  mutate(CN = ectomycorrhiza_mass * 0.44 /
              pmax(ectomycorrhiza_N_biomass * 0.44, 1e-12))

p_joint_myco_cn <- ggplot(opt_joint_cn, aes(x = soil_nitrogen, y = CN)) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

p_joint_gpp <- ggplot(opt_joint, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

make_row_title <- function(label) {
  wrap_elements(
    grid::textGrob(label,
                   gp = grid::gpar(fontsize = 20, fontface = "bold"),
                   hjust = 0.5, x = 0.5)
  )
}

# 3×3 layout: A-C parameters, D-F emergent biomass/height, G-I N depletion + economics
# Stability bands (grey bars) on A-C show the range of params achieving ≥99% of max height
joint_tags <- list(c("", "A", "B", "C", "", "D", "E", "F", "", "G", "H", "I"))

joint_optima <-
  make_row_title("Parameters to be Optimised") /
  (p_joint_ecto | p_joint_rn | p_joint_rl) /
  make_row_title("Emergent behaviour") /
  (p_joint_ecm | p_joint_ecm_area | p_joint_height) /
  make_row_title("N depletion & mycorrhizal economics") /
  (p_joint_nbar | p_joint_myco_cn | p_joint_gpp) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(tag_levels = joint_tags) &
  theme(plot.tag       = element_text(size = 22, face = "bold"),
        legend.text    = element_text(size = 18),
        legend.key.size = unit(1.2, "lines"),
        legend.position = "right")
joint_optima

ggsave("plots/11_joint_optima.png", plot = joint_optima, width = 20, height = 18, dpi = 300)
```

## Joint optima: trajectories over time

```{r joint_optima_trajectories}
# Pull the full time series for the optimal parameter combo at each soil N level.
# opt_joint has one row per soil_nitrogen; inner_join recovers all timesteps for those combos.
opt_traj <- all_joint_df %>%
  inner_join(opt_joint %>% select(soil_nitrogen, ecto_allo, root_no, root_length),
             by = c("soil_nitrogen", "ecto_allo", "root_no", "root_length")) %>%
  mutate(date       = as.Date(date),
         CN         = ectomycorrhiza_mass * 0.44 /
                        pmax(ectomycorrhiza_N_biomass * 0.44, 1e-12),
         gpp_per_ca = assim_gross / crown_area)

traj_colour <- scale_colour_viridis_c(option = "plasma",
                                      name   = "Soil N\n(g m⁻³)")

# --- Top row: optimal parameter values (constant per run, coloured by soil N) ---
p_traj_ecto <- ggplot(opt_traj,
    aes(x = date, y = ecto_allo, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = NULL, y = "Ectomycorrhizal\nallocation (%)") +
  pub_theme

p_traj_rn <- ggplot(opt_traj,
    aes(x = date, y = root_no, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Root number\nper crown area (no m⁻²)") +
  pub_theme

p_traj_rl <- ggplot(opt_traj,
    aes(x = date, y = root_length, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Root length (m)") +
  pub_theme

# --- Middle row: emergent biomass and height ---
p_traj_ecm <- ggplot(opt_traj,
    aes(x = date, y = ectomycorrhiza_mass, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_traj_ecm_area <- ggplot(opt_traj,
    aes(x = date, y = ectomycorrhiza_mass / crown_area,
        colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_traj_height <- ggplot(opt_traj,
    aes(x = date, y = height, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Height (m)") +
  pub_theme

# --- Bottom row: N depletion and mycorrhizal economics ---
p_traj_nbar <- ggplot(opt_traj,
    aes(x = date, y = N_bar, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = expression(bar(N)~"(kg N m"^{-3}*")")) +
  pub_theme

p_traj_cn <- ggplot(opt_traj,
    aes(x = date, y = CN, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

p_traj_gpp <- ggplot(opt_traj,
    aes(x = date, y = gpp_per_ca, colour = soil_nitrogen, group = factor(soil_nitrogen))) +
  geom_line(linewidth = 0.8) +
  traj_colour +
  labs(x = NULL, y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

traj_tags <- list(c("", "A", "B", "C", "", "D", "E", "F", "", "G", "H", "I"))

joint_optima_traj <-
  make_row_title("Parameters to be Optimised") /
  (p_traj_ecto | p_traj_rn | p_traj_rl) /
  make_row_title("Emergent behaviour") /
  (p_traj_ecm | p_traj_ecm_area | p_traj_height) /
  make_row_title("N depletion & mycorrhizal economics") /
  (p_traj_nbar | p_traj_cn | p_traj_gpp) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(tag_levels = traj_tags) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 18),
        legend.key.size = unit(1.2, "lines"),
        legend.position = "right")
joint_optima_traj

ggsave("plots/11b_joint_optima_trajectories.png",
       plot = joint_optima_traj, width = 20, height = 18, dpi = 300)
```

## Trade offs: Ectomycorrhizal – Uptake and Assimilation

```{r ecto_uptake_assimilation}
heatmap_ecto_full <- subset_df %>%
  group_by(soil_nitrogen, ecto_allo) %>%
  summarise(
    height          = mean(height, na.rm = TRUE),
    nitrogen_uptake = mean(root_uptake + mycorrhizal_export_to_tree, na.rm = TRUE),
    assim_gross     = mean(assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

opt_ecto_uptake <- heatmap_ecto_full %>%
  group_by(soil_nitrogen) %>%
  slice_max(nitrogen_uptake, n = 1, with_ties = FALSE) %>%
  ungroup()

opt_ecto_assim <- heatmap_ecto_full %>%
  group_by(soil_nitrogen) %>%
  slice_max(assim_gross, n = 1, with_ties = FALSE) %>%
  ungroup()

p_ecto_allo_uptake <- ggplot(heatmap_ecto_full,
       aes(x = soil_nitrogen, y = ecto_allo, fill = nitrogen_uptake)) +
  geom_tile() +
  geom_point(data = opt_ecto_uptake,
             aes(x = soil_nitrogen, y = ecto_allo),
             shape = 21, fill = "white", colour = "black", size = 3.5,
             stroke = 1.5, inherit.aes = FALSE) +
  annotate("segment", x = 0, y = 0.30, xend = 2, yend = 0.10,
           colour = "white", linewidth = 1.2) +
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_viridis_c(option = "magma",
                       labels = scales::label_number(accuracy = 0.001)) +
  labs(
    x    = "Soil nitrogen (g N m⁻³)",
    y    = "Ectomycorrhizal\nallocation",
    fill = "N uptake\n(kg N yr⁻¹)"
  ) +
  pub_theme

p_ecto_allo_assim <- ggplot(heatmap_ecto_full,
       aes(x = soil_nitrogen, y = ecto_allo, fill = assim_gross)) +
  geom_tile() +
  geom_point(data = opt_ecto_assim,
             aes(x = soil_nitrogen, y = ecto_allo),
             shape = 21, fill = "white", colour = "black", size = 3.5,
             stroke = 1.5, inherit.aes = FALSE) +
  annotate("segment", x = 0, y = 0.30, xend = 2, yend = 0.10,
           colour = "white", linewidth = 1.2) +
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_viridis_c(option = "plasma",
                       labels = scales::label_number(accuracy = 0.1)) +
  labs(
    x    = "Soil nitrogen (g N m⁻³)",
    y    = "Ectomycorrhizal\nallocation",
    fill = "GPP\n(kg C yr⁻¹)"
  ) +
  pub_theme

ectomycorrhizal_uptake_assim_gradient <- (p_ecto_allo_uptake / p_ecto_allo_assim) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right",
        plot.tag = element_text(size = 22, face = "bold"))
ectomycorrhizal_uptake_assim_gradient

ggsave("plots/09b_ectomycorrhizal_uptake_assim_gradient.png",
       plot = ectomycorrhizal_uptake_assim_gradient, width = 12, height = 10, dpi = 300)
```

## C:N concentration in mycorrhizal exchange

```{r myco_cn_ratio}
gs_eb <- all_eb_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8) %>%
  group_by(soil_nitrogen, ecto_allo) %>%
  summarise(
    CN    = mean(ectomycorrhiza_mass * 0.44 / pmax(ectomycorrhiza_N_biomass * 0.44, 1e-12), na.rm = TRUE),
    .groups = "drop"
  )

p_cn_scatter <- ggplot(gs_eb,
       aes(x = soil_nitrogen, y = CN, colour = ecto_allo,
           group = factor(ecto_allo))) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point(size = 2.5, alpha = 0.8) +
  scale_colour_viridis_c(option = "magma",
                         name   = "Ecto\nallocation",
                         labels = scales::label_percent()) +
  labs(
    x = "Soil nitrogen (g N m⁻³)",
    y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)"
  ) +
  pub_theme

p_cn_scatter

ggsave("plots/10_myco_cn_ratio.png",
       plot = p_cn_scatter, width = 10, height = 7, dpi = 300)
```

## Diagnostic: is max_N_transfer the bottleneck?

```{r diag_N_transfer}
# Use the ecto_allo sweep at a mid-range soil N as representative
diag_N <- nitrogen_values[ceiling(length(nitrogen_values) / 2)]

diag_df <- all_eb_df %>%
  filter(abs(soil_nitrogen - diag_N) < 1e-6) %>%
  mutate(
    root_diameter_mm  = 0.35 / root_length^0.2,
    root_SA           = pi * (root_diameter_mm / 1000) * (root_length / 1000) *
                        root_no * crown_area * lai,
    max_N_transfer    = 0.5 * 0.9 * root_SA / 12,   # u_transfer * mycorrhized / 12
    SA_m              = 4 * ectomycorrhiza_mass / (330 * 1e-5)
  )

# Growing-season (June–August) snapshot: avoids f_temp = 0 in winter
diag_snap <- diag_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(ecto_allo) %>%
  summarise(across(c(myco_uptake, mycorrhizal_export_to_tree, max_N_transfer,
                     ectomycorrhiza_mass, C_export_to_myco, N_bar, SA_m),
                   \(x) mean(x, na.rm = TRUE)),
            .groups = "drop")

# Panel A: N fluxes vs ecto_allo (growing season)
p_diag1 <- ggplot(diag_snap, aes(x = ecto_allo)) +
  geom_line(aes(y = myco_uptake,               colour = "U_myco (fungal capture)")) +
  geom_line(aes(y = mycorrhizal_export_to_tree, colour = "N_export (to tree)")) +
  geom_line(aes(y = max_N_transfer,             colour = "max_N_transfer (cap)"), linetype = "dashed") +
  scale_colour_manual(values = c(
    "U_myco (fungal capture)"  = "#0072B2",
    "N_export (to tree)"       = "#D55E00",
    "max_N_transfer (cap)"     = "black"
  )) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(x = "Ectomycorrhizal allocation", y = "kg N month⁻¹",
       colour = NULL,
       title  = paste0("N fluxes (Jun–Aug 2021),\nsoil N = ", round(diag_N, 2), " kg m⁻³")) +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel B: fungal mass and C investment vs ecto_allo
p_diag2 <- ggplot(diag_snap, aes(x = ecto_allo)) +
  geom_line(aes(y = ectomycorrhiza_mass, colour = "ECM mass (kg)")) +
  geom_line(aes(y = C_export_to_myco,   colour = "C export to myco (kg C month⁻¹)")) +
  scale_colour_manual(values = c(
    "ECM mass (kg)"                    = "#009E73",
    "C export to myco (kg C month⁻¹)" = "#CC79A7"
  )) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(x = "Ectomycorrhizal allocation", y = "kg",
       colour = NULL,
       title  = "Fungal mass and C investment\n(Jun–Aug 2021)") +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel C: N_bar (soil N at absorbing surface) — tests diffusion depletion hypothesis
p_diag3 <- ggplot(diag_snap, aes(x = ecto_allo)) +
  geom_line(aes(y = N_bar, colour = "N_bar (kg N m⁻³)")) +
  geom_hline(yintercept = diag_N, linetype = "dashed", colour = "grey50") +
  annotate("text", x = max(diag_snap$ecto_allo) * 0.05,
           y = diag_N * 1.05, label = "Soil N_s", hjust = 0, size = 5) +
  scale_colour_manual(values = c("N_bar (kg N m⁻³)" = "#E69F00")) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(x = "Ectomycorrhizal allocation", y = "kg N m⁻³",
       colour = NULL,
       title  = "N at absorbing surface\n— if << N_s, diffusion is limiting") +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel D: time series at mid ecto_allo
diag_mid_allo <- ecto_allo_values[ceiling(length(ecto_allo_values) / 2)]
diag_ts <- diag_df %>% filter(abs(ecto_allo - diag_mid_allo) < 1e-6)

p_diag4 <- ggplot(diag_ts, aes(x = as.Date(date))) +
  geom_line(aes(y = ectomycorrhiza_mass, colour = "ECM mass (kg)")) +
  geom_line(aes(y = C_export_to_myco,   colour = "C export to myco (kg C month⁻¹)")) +
  scale_colour_manual(values = c(
    "ECM mass (kg)"                    = "#009E73",
    "C export to myco (kg C month⁻¹)" = "#CC79A7"
  )) +
  labs(x = "Date", y = "kg",
       colour = NULL,
       title  = paste0("ECM mass vs C investment over time\n(ecto_allo = ",
                       round(diag_mid_allo, 2), ")")) +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

diag_combined <- (p_diag1 + p_diag2) / (p_diag3 + p_diag4) +
  plot_annotation(
    title   = "Diagnostic: what limits mycorrhizal N return?",
    caption = paste0("Cap binding: N_export ≈ max_N_transfer & U_myco >> both  |  ",
                     "Mass-limited: ECM mass tiny despite C investment  |  ",
                     "Diffusion-limited: N_bar << N_s")
  )
diag_combined

ggsave("plots/diag_N_transfer_cap.png",
       plot = diag_combined, width = 13, height = 10, dpi = 300)
```

## Diagnostic: root number per crown area

```{r diag_root_no}
diag_rn_df <- all_rn_df %>%
  filter(abs(soil_nitrogen - diag_N) < 1e-6) %>%
  mutate(
    root_diameter_mm = 0.35 / fixed_root_length^0.2,
    root_SA          = pi * (root_diameter_mm / 1000) * (fixed_root_length / 1000) *
                       root_no * crown_area * lai,
    max_N_transfer   = 0.5 * 0.9 * root_SA / 12
  )

diag_rn_snap <- diag_rn_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(root_no) %>%
  summarise(across(c(root_uptake, mycorrhizal_export_to_tree, max_N_transfer,
                     root_mass, rr, N_bar, assim_gross),
                   \(x) mean(x, na.rm = TRUE)),
            .groups = "drop")

diag_mid_rn    <- root_no_values[ceiling(length(root_no_values) / 2)]
diag_rn_ts     <- diag_rn_df %>% filter(abs(root_no - diag_mid_rn) < diag_mid_rn * 0.01)

# Panel A: N fluxes vs root_no
p_rn_diag1 <- ggplot(diag_rn_snap, aes(x = root_no)) +
  geom_line(aes(y = root_uptake,               colour = "Root uptake")) +
  geom_line(aes(y = mycorrhizal_export_to_tree, colour = "N_export (to tree)")) +
  scale_colour_manual(values = c(
    "Root uptake"        = "#0072B2",
    "N_export (to tree)" = "#D55E00"
  )) +
  scale_x_log10(labels = scales::label_scientific()) +
  labs(x = "Root number per crown area (m⁻²)", y = "kg N month⁻¹",
       colour = NULL,
       title  = paste0("N fluxes (Jun–Aug 2021),\nsoil N = ", round(diag_N, 2), " kg m⁻³")) +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel B: root mass and root respiration vs root_no
p_rn_diag2 <- ggplot(diag_rn_snap, aes(x = root_no)) +
  geom_line(aes(y = root_mass, colour = "Root mass (kg C)")) +
  geom_line(aes(y = rr,        colour = "Root respiration (kg C month⁻¹)")) +
  scale_colour_manual(values = c(
    "Root mass (kg C)"                   = "#009E73",
    "Root respiration (kg C month⁻¹)"   = "#CC79A7"
  )) +
  scale_x_log10(labels = scales::label_scientific()) +
  labs(x = "Root number per crown area (m⁻²)", y = "kg C",
       colour = NULL,
       title  = "Root mass and respiration cost\n(Jun–Aug 2021)") +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel C: N_bar vs root_no
p_rn_diag3 <- ggplot(diag_rn_snap, aes(x = root_no)) +
  geom_line(aes(y = N_bar, colour = "N_bar (kg N m⁻³)")) +
  geom_hline(yintercept = diag_N, linetype = "dashed", colour = "grey50") +
  annotate("text", x = min(diag_rn_snap$root_no) * 1.2,
           y = diag_N * 1.05, label = "Soil N_s", hjust = 0, size = 5) +
  scale_colour_manual(values = c("N_bar (kg N m⁻³)" = "#E69F00")) +
  scale_x_log10(labels = scales::label_scientific()) +
  labs(x = "Root number per crown area (m⁻²)", y = "kg N m⁻³",
       colour = NULL,
       title  = "N at absorbing surface\n— if << N_s, diffusion is limiting") +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel D: time series at mid root_no
p_rn_diag4 <- ggplot(diag_rn_ts, aes(x = as.Date(date))) +
  geom_line(aes(y = root_mass, colour = "Root mass (kg C)")) +
  geom_line(aes(y = rr,        colour = "Root respiration (kg C month⁻¹)")) +
  scale_colour_manual(values = c(
    "Root mass (kg C)"                   = "#009E73",
    "Root respiration (kg C month⁻¹)"   = "#CC79A7"
  )) +
  labs(x = "Date", y = "kg C",
       colour = NULL,
       title  = paste0("Root mass vs respiration over time\n(root_no = ",
                       format(round(diag_mid_rn), big.mark = ","), " m⁻²)")) +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

diag_rn_combined <- (p_rn_diag1 + p_rn_diag2) / (p_rn_diag3 + p_rn_diag4) +
  plot_annotation(
    title   = "Diagnostic: what limits N uptake vs root number?",
    caption = paste0("Cap binding: N_export ≈ max_N_transfer  |  ",
                     "Cost-limited: respiration >> benefit  |  ",
                     "Diffusion-limited: N_bar << N_s")
  )
diag_rn_combined

ggsave("plots/diag_root_no.png",
       plot = diag_rn_combined, width = 13, height = 10, dpi = 300)
```

## Diagnostic: root length

```{r diag_root_length}
diag_rl_df <- all_rl_df %>%
  filter(abs(soil_nitrogen - diag_N) < 1e-6) %>%
  mutate(
    root_diameter_mm = 0.35 / root_length^0.2,
    root_SA          = pi * (root_diameter_mm / 1000) * (root_length / 1000) *
                       fixed_root_no * crown_area * lai,
    max_N_transfer   = 0.5 * 0.9 * root_SA / 12
  )

diag_rl_snap <- diag_rl_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(root_length) %>%
  summarise(across(c(root_uptake, mycorrhizal_export_to_tree, max_N_transfer,
                     root_mass, rr, N_bar, assim_gross),
                   \(x) mean(x, na.rm = TRUE)),
            .groups = "drop")

diag_mid_rl <- root_length_values[ceiling(length(root_length_values) / 2)]
diag_rl_ts  <- diag_rl_df %>% filter(abs(root_length - diag_mid_rl) < 0.01)

# Panel A: N fluxes vs root_length
p_rl_diag1 <- ggplot(diag_rl_snap, aes(x = root_length)) +
  geom_line(aes(y = root_uptake,               colour = "Root uptake")) +
  geom_line(aes(y = mycorrhizal_export_to_tree, colour = "N_export (to tree)")) +
  scale_colour_manual(values = c(
    "Root uptake"        = "#0072B2",
    "N_export (to tree)" = "#D55E00"
  )) +
  labs(x = "Root length (mm)", y = "kg N month⁻¹",
       colour = NULL,
       title  = paste0("N fluxes (Jun–Aug 2021),\nsoil N = ", round(diag_N, 2), " kg m⁻³")) +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel B: root mass and root respiration vs root_length
p_rl_diag2 <- ggplot(diag_rl_snap, aes(x = root_length)) +
  geom_line(aes(y = root_mass, colour = "Root mass (kg C)")) +
  geom_line(aes(y = rr,        colour = "Root respiration (kg C month⁻¹)")) +
  scale_colour_manual(values = c(
    "Root mass (kg C)"                   = "#009E73",
    "Root respiration (kg C month⁻¹)"   = "#CC79A7"
  )) +
  labs(x = "Root length (mm)", y = "kg C",
       colour = NULL,
       title  = "Root mass and respiration cost\n(Jun–Aug 2021)") +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel C: N_bar vs root_length
p_rl_diag3 <- ggplot(diag_rl_snap, aes(x = root_length)) +
  geom_line(aes(y = N_bar, colour = "N_bar (kg N m⁻³)")) +
  geom_hline(yintercept = diag_N, linetype = "dashed", colour = "grey50") +
  annotate("text", x = min(diag_rl_snap$root_length) * 1.05,
           y = diag_N * 1.05, label = "Soil N_s", hjust = 0, size = 5) +
  scale_colour_manual(values = c("N_bar (kg N m⁻³)" = "#E69F00")) +
  labs(x = "Root length (mm)", y = "kg N m⁻³",
       colour = NULL,
       title  = "N at absorbing surface\n— if << N_s, diffusion is limiting") +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

# Panel D: time series at mid root_length
p_rl_diag4 <- ggplot(diag_rl_ts, aes(x = as.Date(date))) +
  geom_line(aes(y = root_mass, colour = "Root mass (kg C)")) +
  geom_line(aes(y = rr,        colour = "Root respiration (kg C month⁻¹)")) +
  scale_colour_manual(values = c(
    "Root mass (kg C)"                   = "#009E73",
    "Root respiration (kg C month⁻¹)"   = "#CC79A7"
  )) +
  labs(x = "Date", y = "kg C",
       colour = NULL,
       title  = paste0("Root mass vs respiration over time\n(root_length = ",
                       round(diag_mid_rl, 2), " mm)")) +
  pub_theme + theme(legend.position = "bottom", plot.title = element_text(size = 16, face = "bold")) +
  guides(colour = guide_legend(ncol = 1))

diag_rl_combined <- (p_rl_diag1 + p_rl_diag2) / (p_rl_diag3 + p_rl_diag4) +
  plot_annotation(
    title   = "Diagnostic: what limits N uptake vs root length?",
    caption = paste0("Cap binding: N_export ≈ max_N_transfer  |  ",
                     "Cost-limited: respiration >> benefit  |  ",
                     "Diffusion-limited: N_bar << N_s")
  )
diag_rl_combined

ggsave("plots/diag_root_length.png",
       plot = diag_rl_combined, width = 13, height = 10, dpi = 300)
```

## Data comparison

```{r data_load}
ecm_data_all <- read.csv(here("RScripts/ecm_root_data_compiled.csv"), stringsAsFactors = FALSE)

# Strip leading ~, <, > from all columns so approximate values become plain numbers
ecm_data_all <- as.data.frame(lapply(ecm_data_all, function(col) {
  if (is.character(col)) {
    cleaned <- gsub("^[~<>]+", "", trimws(col))
    num <- suppressWarnings(as.numeric(cleaned))
    if (sum(!is.na(num)) > sum(!is.na(suppressWarnings(as.numeric(col))))) num else col
  } else col
}))

# Model is calibrated for Scots pine / boreal evergreen conifer — keep only conifers
evergreen_conifers <- c("Pinus sylvestris", "Picea abies", "Picea abies (control)",
                        "Picea glauca", "Picea spp.", "Mixed (Pinus/Picea)")
ecm_data <- ecm_data_all[ecm_data_all$species %in% evergreen_conifers, ]
ecm_data <- ecm_data[is.na(ecm_data$soil_CN_ratio) | ecm_data$soil_CN_ratio < 100, ]
ecm_data$species <- factor(ecm_data$species, levels = evergreen_conifers)

species_colours <- c(
  "Pinus sylvestris"     = "#E69F00",
  "Picea abies"          = "#56B4E9",
  "Picea abies (control)"= "#009E73",
  "Picea glauca"         = "#F0E442",
  "Picea spp."           = "#0072B2",
  "Mixed (Pinus/Picea)"  = "#D55E00"
)
species_scale <- scale_colour_manual(name = "Species", values = species_colours,
                                     drop = FALSE,
                                     guide = guide_legend(ncol = 3, byrow = TRUE))

# Tip density: convert million m⁻² → m⁻² to match model root_no units
ecm_data$tip_density_per_m2 <- ecm_data$ecm_tip_density_million_per_m2_soil * 1e6
```

### Plot 07 A — Root length vs soil N
*Matches heatmap p2 (y-axis: root length, mm)*

```{r data_07a}
d_rl <- ecm_data[!is.na(ecm_data$ecm_root_tip_length_mm) &
                 !is.na(ecm_data$soil_CN_ratio), ]

p_data_rl <- ggplot(d_rl, aes(x = soil_CN_ratio, y = ecm_root_tip_length_mm,
                               colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  labs(
    x = "Soil C:N ratio",
    y = "ECM root tip\nlength (mm)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 07 B — Root / tip density vs soil N
*Matches heatmap p1 (y-axis: root number per crown area, m⁻²)*

```{r data_07b}
d_td <- ecm_data[!is.na(ecm_data$ecm_tip_density_million_per_m2_BA) &
                 !is.na(ecm_data$soil_CN_ratio), ]

p_data_td <- ggplot(d_td, aes(x = soil_CN_ratio, y = ecm_tip_density_million_per_m2_BA,
                               colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  scale_y_log10(labels = scales::label_scientific()) +
  labs(
    x = "Soil C:N ratio",
    y = "ECM tip density\n(million m⁻² basal area)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 09 A — ECM mycelium biomass vs soil N
*Matches heatmap p_ecm_mass (y-axis: ectomycorrhiza_mass)*

```{r data_09a}
d_ecm <- ecm_data[!is.na(ecm_data$ecm_mycelium_biomass_kg_ha) &
                  !is.na(ecm_data$soil_CN_ratio), ]

p_data_ecm <- ggplot(d_ecm, aes(x = soil_CN_ratio, y = ecm_mycelium_biomass_kg_ha,
                                 colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  scale_y_log10(labels = scales::label_number(accuracy = 1)) +
  labs(
    x = "Soil C:N ratio",
    y = "ECM mycelium\nbiomass (kg ha⁻¹)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 09 B — ECM root biomass vs soil N
*Proxy for ectomycorrhizal allocation (p_ecto_allo)*

```{r data_09b}
d_ecmr <- ecm_data[!is.na(ecm_data$ecm_root_biomass_kg_per_m2_basal_area) &
                   !is.na(ecm_data$soil_CN_ratio), ]

p_data_ecmr <- ggplot(d_ecmr, aes(x = soil_CN_ratio,
                                    y = ecm_root_biomass_kg_per_m2_basal_area,
                                    colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  labs(
    x = "Soil C:N ratio",
    y = "ECM root biomass\n(kg m⁻² basal area)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### ECM colonisation percentage vs soil N

```{r data_colonisation}
d_col_plot <- ecm_data[!is.na(ecm_data$ecm_colonisation_pct) &
                        !is.na(ecm_data$soil_CN_ratio), ]

p_data_col <- ggplot(d_col_plot, aes(x = soil_CN_ratio, y = ecm_colonisation_pct,
                                      colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  coord_cartesian(ylim = c(NA, 100)) +
  labs(
    x      = "Soil C:N ratio",
    y      = "ECM colonisation (%)",
    colour = "Species"
  ) +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 13)) +
  pub_theme
```

```{r data_combined}
# Standalone legend using all species so the hash is consistent
legend_plot <- ggplot(
    data.frame(species = factor(evergreen_conifers, levels = evergreen_conifers)),
    aes(x = 1, y = 1, colour = species)) +
  geom_point(alpha = 0) +
  species_scale +
  theme_void() +
  theme(legend.position   = c(0.5, 0.5),
        legend.text       = element_text(size = 22),
        legend.title      = element_text(size = 24, face = "bold"),
        legend.key.size   = unit(1.5, "lines"),
        axis.text.x       = element_blank(),
        axis.text.y       = element_blank(),
        axis.title.x      = element_blank(),
        axis.title.y      = element_blank(),
        axis.ticks        = element_blank(),
        panel.background  = element_blank()) +
  guides(colour = guide_legend(ncol = 1,
                               override.aes = list(alpha = 1, size = 5)))

data_combined <- wrap_plots(
  p_data_rl   + theme(legend.position = "none"),
  p_data_td   + theme(legend.position = "none"),
  p_data_ecm  + theme(legend.position = "none"),
  p_data_ecmr + theme(legend.position = "none"),
  p_data_col  + theme(legend.position = "none"),
  legend_plot,
  ncol = 3
) + plot_annotation(tag_levels = list(c("A","B","C","D","E",""))) &
  theme(plot.tag   = element_text(size = 28, face = "bold"),
        axis.text  = element_text(size = 24),
        axis.title = element_text(size = 26, face = "bold"))
data_combined

ggsave("plots/data_combined.png",
       plot = data_combined, width = 24, height = 14, dpi = 300)
```

