---
title: "Publication Plots"
author: "Joanna"
date: "2026-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(data.table)
library(rphydro)
library(Rprebasso)
library(patchwork)
library(PlantFATE)
library(akima)
library(future)
library(future.apply)
library(here)
```

```{r theme, include=FALSE}
pub_theme <- theme_classic(base_size = 22) +
  theme(
    axis.title       = element_text(size = 22, face = "bold"),
    axis.text        = element_text(size = 20, colour = "black"),
    axis.line        = element_line(colour = "black", linewidth = 0.6),
    axis.ticks       = element_line(colour = "black", linewidth = 0.5),
    legend.title     = element_text(size = 22, face = "bold"),
    legend.text      = element_text(size = 20),
    legend.key.size  = unit(0.9, "lines"),
    panel.grid.major = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background  = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.margin      = margin(8, 12, 8, 8),
    strip.text       = element_text(size = 22, face = "bold")
  )
```

## Data reading

```{r}
hyde_all = "~/Documents/Austria/Hyytiala_all_data/"
data <- lapply(paste0(hyde_all, list.files(hyde_all))[1:9], read.csv)
names(data) <- list.files(hyde_all)[1:9]
useful_data <- merge(data[[2]], data[[5]], c('plotID', 'eventID', 'eventYear'), all=TRUE)

```

## Simulations

```{r pressure, echo=FALSE}
# Set to TRUE for fast testing (fewer grid points, shorter run), FALSE for full publication run
testing_mode <- TRUE

plan(multicore, workers = parallel::detectCores() - 1)

# Parameter ranges
n_pts_main  <- if (testing_mode) 15 else 21
n_pts_joint <- if (testing_mode) 6 else 20   # per-axis for 3D co-optimisation sweep

nitrogen_values <- seq(0.1, 10, length.out = n_pts_main)

# Joint co-optimisation grid (coarser to keep run count tractable)
joint_ecto_allo_values    <- seq(0.0,  0.3, length.out = n_pts_joint)
joint_root_no_values      <- 3*10^5 # 3*10^seq(3, 6, length.out = n_pts_joint)
joint_root_length_values  <- seq(0.05, 6,   length.out = n_pts_joint)
param_file <- "/home/josimms/Documents/Austria/Plant-FATE/tests/params/p_test_boreal.ini"
weather_file <- "/home/josimms/Documents/Austria/Plant-FATE/data/ERAS_Monthly.csv"

# -------------------------------
# Core simulation function with safe error handling
# -------------------------------
run_sim_parallel <- function(rl, rn, n, eb) {
  tryCatch({
    lho <- new(LifeHistoryOptimizer, param_file)
    lho$set_i_metFile(weather_file)
    lho$set_a_metFile(weather_file)
    lho$set_co2File("")
    lho$set_soil_nitrogen(n)
    lho$traits0$investment_from_tree <- eb
    lho$init()
    lho$root_override(rn, rl)

    dt <- 1/12
    start_year <- 1960
    end_year   <- 2022
    years_seq  <- seq(start_year, end_year, dt)
    
    results <- lapply(years_seq, function(t) {
      tryCatch({
        lho$grow_for_dt(t, dt)
        state <- lho$get_state(t + dt)
        df_row <- as.data.frame(t(state))
        names(df_row) <- lho$get_header()
        df_row
      }, error = function(e) {
        setNames(as.data.frame(as.list(rep(NA, length(lho$get_header())))),
                 lho$get_header())
      })
    })
    
    df <- do.call(rbind, results)
    df$date <- seq(as.Date(paste0(start_year, "-01-01")),
                   as.Date(paste0(end_year, "-01-01")),
                   by = "month")[1:nrow(df)]
    df$root_length   <- rl
    df$root_no       <- rn
    df$soil_nitrogen <- n
    df$ecto_allo     <- eb
    df
  }, error = function(e) {
    message("Failed run_sim_parallel: rl=", rl, " rn=", rn, " n=", n, " eb=", eb)
    NULL
  })
}

# -------------------------------
# Helper: safe wrapper for run_sim_parallel
# -------------------------------
run_sim_safe <- function(rl, rn, n, eb) {
  result <- run_sim_parallel(rl, rn, n, eb)
  if (is.null(result) || nrow(result) == 0) return(NULL)
  result
}

# -------------------------------
# Joint co-optimisation: ecto_allo × root_no × root_length × nitrogen
# n_pts_joint^3 × n_pts_main runs total
# Parallelised over ecto_allo (outer axis)
# -------------------------------
all_joint_list <- future_lapply(joint_ecto_allo_values, function(eb) {
  inner_rn <- lapply(joint_root_no_values, function(rn) {
    inner_rl <- lapply(joint_root_length_values, function(rl) {
      inner_n <- lapply(nitrogen_values, function(n) {
        run_sim_safe(rl, rn, n, eb)
      })
      do.call(rbind, inner_n[!sapply(inner_n, is.null)])
    })
    do.call(rbind, inner_rl[!sapply(inner_rl, is.null)])
  })
  do.call(rbind, inner_rn[!sapply(inner_rn, is.null)])
})
all_joint_df <- do.call(rbind, all_joint_list[!sapply(all_joint_list, is.null)])
```

## Co-optimised trait optima

```{r joint_optimum}
# June–August 2021 growing-season mean per parameter combination.
# Rate variables (N uptake, GPP) are zero in winter (f_temp = 0), so January
# snapshots understate uptake; seasonal means give a fair comparison.
joint_gs <- all_joint_df %>%
  filter(as.integer(format(as.Date(date), "%m")) %in% 6:8,
         as.integer(format(as.Date(date), "%Y")) == 2021) %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  summarise(
    height                     = mean(height,                     na.rm = TRUE),
    ectomycorrhiza_mass        = mean(ectomycorrhiza_mass,        na.rm = TRUE),
    ectomycorrhiza_N_biomass   = mean(ectomycorrhiza_N_biomass,   na.rm = TRUE),
    ectomycorrhiza_C_free      = mean(ectomycorrhiza_C_free,      na.rm = TRUE),
    ectomycorrhiza_N_free      = mean(ectomycorrhiza_N_free,      na.rm = TRUE),
    crown_area                 = mean(crown_area,                 na.rm = TRUE),
    lai                        = mean(lai,                        na.rm = TRUE),
    N_bar_roots                = mean(N_bar_roots,                na.rm = TRUE),
    N_bar_myco                 = mean(N_bar_myco,                 na.rm = TRUE),
    C_export_to_myco           = mean(C_export_to_myco,           na.rm = TRUE),
    mycorrhizal_export_to_tree = mean(mycorrhizal_export_to_tree, na.rm = TRUE),
    root_uptake                = mean(root_uptake,                na.rm = TRUE),
    myco_uptake                = mean(myco_uptake,                na.rm = TRUE),
    assim_gross                = mean(assim_gross,                na.rm = TRUE),
    optimal_leaf_nitrogen      = mean(optimal_leaf_nitrogen,      na.rm = TRUE),
    leaf_nitrogen_concentration = mean(leaf_nitrogen_concentration, na.rm = TRUE),
    tree_nitrogen               = mean(tree_nitrogen,              na.rm = TRUE),
    root_mass                   = mean(root_mass,                  na.rm = TRUE),
    coarse_root_mass            = mean(coarse_root_mass,           na.rm = TRUE),
    total_mass                  = mean(total_mass,                 na.rm = TRUE),
    fineroot_lifespan           = mean(fineroot_lifespan,          na.rm = TRUE),
    total_prod                  = mean(total_prod,                 na.rm = TRUE),
    rr                          = mean(rr,                         na.rm = TRUE),
    tr                          = mean(tr,                         na.rm = TRUE),
    ecm_frac_gpp                = mean(C_export_to_myco / assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

opt_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(
    max_m        = max(height, na.rm = TRUE),
    near_opt     = height >= 0.99 * max_m,
    n_total      = n(),
    n_near_opt   = sum(near_opt),
    stab_frac    = n_near_opt / n_total,
    biomass_gain = max_m / sort(height, decreasing = TRUE)[2] - 1
  ) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup()


# --- Stability bands: all combos within 1% of maximum height per soil N ---
# This shows how flat the fitness landscape is around each optimum
stab_joint <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_m    = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_m) %>%
  filter(near_opt) %>%
  summarise(
    ecto_lo     = min(ecto_allo,     na.rm = TRUE), ecto_hi     = max(ecto_allo,     na.rm = TRUE),
    ecm_gpp_lo  = min(ecm_frac_gpp,  na.rm = TRUE), ecm_gpp_hi  = max(ecm_frac_gpp,  na.rm = TRUE),
    rn_lo       = min(root_no,       na.rm = TRUE), rn_hi       = max(root_no,       na.rm = TRUE),
    rl_lo       = min(root_length,   na.rm = TRUE), rl_hi       = max(root_length,   na.rm = TRUE),
    .groups = "drop"
  )

joint_height_range <- range(opt_joint$height, na.rm = TRUE)

shape_vals <- c("Optimal parameters" = 3, "Model output" = 1, "Calculated" = 18)
ltype_vals <- c("Trend line"   = "dashed")

# Produces a 2-row data frame for a straight trend-line segment
tl <- function(x1, y1, x2, y2) data.frame(x = c(x1, x2), y = c(y1, y2))

colour_vals <- c("Stability (≥99% of max biomass)" = "grey70")

shared_scales <- list(
  scale_shape_manual(values = shape_vals, name = NULL),
  scale_linetype_manual(values = ltype_vals, name = NULL),
  scale_colour_manual(values = colour_vals, name = NULL)
)

stab_label <- geom_text(aes(label = paste0(round(stab_frac * 100), "%")),
                        vjust = -0.8, size = 3, inherit.aes = TRUE)

# --- Top row: crosses + stability bars, with dashed trend line ---
p_joint_ecto <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ecto_allo)) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = ecto_lo, ymax = ecto_hi,
                 colour = "Stability (≥99% of max biomass)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(tag = "A", x = "Soil nitrogen (g N m⁻³)", y = "Optimal ECM C allocation\n(% NPP)") +
  pub_theme

p_joint_rn <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_no)) +
  geom_line(data = tl(0, 3e6, 2, 1e5), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rn_lo, ymax = rn_hi,
                 colour = "Stability (≥99% of max biomass)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(tag = "B", x = "Soil nitrogen (g N m⁻³)", y = "Optimal root number\nper crown area (no m⁻²)") +
  pub_theme

p_joint_rl <- ggplot(opt_joint, aes(x = soil_nitrogen, y = root_length)) +
  geom_line(data = tl(0, 5, 2, 1.2), aes(x = x, y = y, linetype = "Trend line"),
            colour = "grey40", linewidth = 1.0, inherit.aes = FALSE) +
  geom_linerange(data = stab_joint, aes(x = soil_nitrogen, ymin = rl_lo, ymax = rl_hi,
                 colour = "Stability (≥99% of max biomass)"), linewidth = 2, inherit.aes = FALSE) +
  geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 4, stroke = 1.5) +
  stab_label +
  shared_scales +
  labs(tag = "C", x = "Soil nitrogen (g N m⁻³)", y = "Optimal root length (m)") +
  pub_theme

r_coarse_root <- 0.04   # coarse root respiration rate [yr^-1], same as rs (sapwood) in p_test_boreal.ini
rr_param      <- 0.4    # fine root respiration rate [yr^-1], rr in p_test_boreal.ini

p_joint_total_bga <- ggplot(
    opt_joint,
    aes(x = soil_nitrogen,
        y = (C_export_to_myco + root_mass * 0.44 / fineroot_lifespan + rr +
               (r_coarse_root / rr_param) * rr * coarse_root_mass / root_mass) / assim_gross)
  ) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = bga_lo, ymax = bga_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Calculated"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_percent()) +
  expand_limits(y = 0) +
  labs(tag = "D", x = "Soil nitrogen (g N m⁻³)",
       y = "Total belowground C\nallocation (% GPP)") +
  pub_theme

# --- IQR ribbon across near-optimal combos (≥99% of max height) for panels D–I ---
# Ribbon shows spread of each output across the stability band;
# dots remain at the single height-maximising parameter combo.
stab_outputs <- joint_gs %>%
  group_by(soil_nitrogen) %>%
  mutate(max_m    = max(height, na.rm = TRUE),
         near_opt = height >= 0.99 * max_m) %>%
  filter(near_opt) %>%
  summarise(
    ecm_lo          = min(ectomycorrhiza_mass, na.rm = TRUE),
    ecm_hi          = max(ectomycorrhiza_mass, na.rm = TRUE),
    ecm_area_lo     = min(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    ecm_area_hi     = max(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
    total_mass_lo   = min(total_mass, na.rm = TRUE),
    total_mass_hi   = max(total_mass, na.rm = TRUE),
    height_lo       = min(height, na.rm = TRUE),
    height_hi       = max(height, na.rm = TRUE),
    nroots_lo     = min(N_bar_roots, na.rm = TRUE),
    nroots_hi     = max(N_bar_roots, na.rm = TRUE),
    nmyco_lo      = min(N_bar_myco,  na.rm = TRUE),
    nmyco_hi      = max(N_bar_myco,  na.rm = TRUE),
    cn_lo         = min((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) / pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
    cn_hi         = max((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) / pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
    gpp_lo        = min(assim_gross / crown_area, na.rm = TRUE),
    gpp_hi        = max(assim_gross / crown_area, na.rm = TRUE),
    n_upper_lo    = min(leaf_nitrogen_concentration, na.rm = TRUE),
    n_upper_hi    = max(leaf_nitrogen_concentration, na.rm = TRUE),
    n_realised_lo = min(optimal_leaf_nitrogen, na.rm = TRUE),
    n_realised_hi = max(optimal_leaf_nitrogen, na.rm = TRUE),
    n_free_lo     = min(tree_nitrogen, na.rm = TRUE),
    n_free_hi     = max(tree_nitrogen, na.rm = TRUE),
    bga_lo        = min((C_export_to_myco + root_mass * 0.44 / fineroot_lifespan + rr +
                           (r_coarse_root / rr_param) * rr * coarse_root_mass / root_mass) / assim_gross, na.rm = TRUE),
    bga_hi        = max((C_export_to_myco + root_mass * 0.44 / fineroot_lifespan + rr +
                           (r_coarse_root / rr_param) * rr * coarse_root_mass / root_mass) / assim_gross, na.rm = TRUE),
    .groups = "drop"
  )

# Mycorrhizal total C:N (structural + labile pools) at the optimal combo
opt_joint_cn <- opt_joint %>%
  mutate(CN = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
              pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12))

# --- Middle row: emergent biomass and height ---
p_joint_ecm <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = ecm_lo, ymax = ecm_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "E", x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_joint_ecm_area <- ggplot(opt_joint, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = ecm_area_lo, ymax = ecm_area_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(tag = "F", x = "Soil nitrogen (g N m⁻³)", y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_joint_height <- ggplot(opt_joint, aes(x = soil_nitrogen, y = height)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = height_lo, ymax = height_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) +
  labs(tag = "I", x = "Soil nitrogen (g N m⁻³)", y = "Height (m)") +
  pub_theme

# --- Bottom row: N depletion and mycorrhizal economics ---
# Dual y-axis: N_bar_roots (left, black) = mineral pool depleted by roots + myco;
# N_bar_myco (right, steelblue) = organic pool depleted by myco only (f_myco_pool * N_s).
nroots_range <- range(opt_joint$N_bar_roots, na.rm = TRUE)
nmyco_range  <- range(opt_joint$N_bar_myco,  na.rm = TRUE)
myco_to_roots <- function(x) nroots_range[1] + (x - nmyco_range[1]) / diff(nmyco_range) * diff(nroots_range)
roots_to_myco <- function(x) nmyco_range[1]  + (x - nroots_range[1]) / diff(nroots_range) * diff(nmyco_range)

p_joint_nbar <- ggplot(opt_joint, aes(x = soil_nitrogen, y = N_bar_roots)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = nroots_lo, ymax = nroots_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = myco_to_roots(nmyco_lo), ymax = myco_to_roots(nmyco_hi)),
              fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  geom_point(aes(y = myco_to_roots(N_bar_myco)),
             colour = "steelblue", shape = 1, size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(
    name = expression(bar(N)[roots]~"(kg N m"^{-3}*")"),
    sec.axis = sec_axis(~ roots_to_myco(.), name = expression(bar(N)[myco]~"(kg N m"^{-3}*")"))
  ) +
  labs(tag = "G", x = "Soil nitrogen (g N m⁻³)") +
  pub_theme +
  theme(
    axis.title.y.right = element_text(colour = "steelblue"),
    axis.text.y.right  = element_text(colour = "steelblue"),
    axis.ticks.y.right = element_line(colour = "steelblue"),
    axis.line.y.right  = element_line(colour = "steelblue")
  )

# Mycorrhizal total C:N (structural biomass + labile free pools)
# C (kg C) = ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free
# N (kg N) = ectomycorrhiza_N_biomass + ectomycorrhiza_N_free
p_joint_myco_cn <- ggplot(opt_joint_cn, aes(x = soil_nitrogen, y = CN)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = cn_lo, ymax = cn_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  scale_y_continuous(labels = scales::label_scientific(digits = 2)) +
  labs(tag = "H", x = "Soil nitrogen (g N m⁻³)",
       y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

p_joint_gpp <- ggplot(opt_joint, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
  geom_ribbon(data = stab_outputs, aes(x = soil_nitrogen, ymin = gpp_lo, ymax = gpp_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) +
  labs(tag = "J", x = "Soil nitrogen (g N m⁻³)",
       y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

# Nitrogen limitation diagnostic: upper bound (leaf N pool) vs realised (Phydro-optimised N)
# If the two lines converge, the tree is N-limited (all available N used for photosynthesis)
p_joint_n_limit <- ggplot(opt_joint, aes(x = soil_nitrogen)) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_upper_lo, ymax = n_upper_hi),
              fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_realised_lo, ymax = n_realised_hi),
              fill = "grey60", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
  geom_point(aes(y = leaf_nitrogen_concentration, colour = "Potential (leaf N pool)"),
             size = 4, stroke = 1.5) +
  geom_point(aes(y = optimal_leaf_nitrogen),
             colour = "grey30", shape = 1, size = 4, stroke = 1.5) +
  scale_colour_manual(values = c("Potential (leaf N pool)" = "steelblue"),
                       name = NULL) +
  expand_limits(y = 0) +
  labs(tag = "K", x = "Soil nitrogen (g N m⁻³)",
       y = "Leaf N concentration\n(kg N kg⁻¹ leaf)") +
  pub_theme

# Free nitrogen pool at maturity: if this is near nitrogen_start0 (0.039 kg) even at low soil N,
# the tree is running off its initial endowment rather than reaching a true N steady state.
p_joint_nfree <- ggplot(opt_joint, aes(x = soil_nitrogen, y = tree_nitrogen)) +
  geom_ribbon(data = stab_outputs,
              aes(x = soil_nitrogen, ymin = n_free_lo, ymax = n_free_hi),
              fill = "grey80", colour = NA, inherit.aes = FALSE) +
  geom_hline(yintercept = 0.039, linetype = "dashed", colour = "grey40", linewidth = 0.8) +
  annotate("text", x = 0.1, y = 0.039, label = "nitrogen_start0", vjust = -0.4,
           hjust = 0, size = 3.5, colour = "grey40") +
  geom_point(aes(shape = "Model output"), colour = "black", size = 4, stroke = 1.5) +
  shared_scales +
  expand_limits(y = 0) +
  labs(tag = "L", x = "Soil nitrogen (g N m⁻³)",
       y = "Free N pool at maturity\n(kg N tree⁻¹)") +
  pub_theme

make_row_title <- function(label) {
  wrap_elements(
    grid::textGrob(label,
                   gp = grid::gpar(fontsize = 20, fontface = "bold"),
                   hjust = 0.5, x = 0.5)
  )
}

# 4-column layout: A-D parameters, E-H mycorrhizal & N depletion, I-L growth & N status
# Stability bands (grey bars) on A-C show the range of params achieving ≥99% of max height
joint_optima <-
  guide_area() /
  make_row_title("Parameters to be Optimised") /
  (p_joint_ecto | p_joint_rn | p_joint_rl | p_joint_total_bga) /
  make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
  (p_joint_ecm | p_joint_ecm_area | p_joint_nbar | p_joint_myco_cn) /
  make_row_title("Model results: Tree Status") /
  (p_joint_height | p_joint_gpp | p_joint_n_limit | p_joint_nfree) +
  plot_layout(heights = c(0.15, 0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(theme = theme(plot.margin = margin(5, 30, 5, 5))) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 16),
        legend.key.size = unit(1.0, "lines"),
        legend.position    = "top",
        legend.direction   = "horizontal",
        legend.box         = "horizontal")
joint_optima

ggsave("plots/11_joint_optima.png", plot = joint_optima, width = 22, height = 18, dpi = 300)
```

## Mycorrhizal N transfer diagnostics

```{r myco_transfer_diag}
# Test whether the transfer bottleneck (max_N_transfer cap) is limiting the tree
# at low soil N.  Five panels:
#   1. Total N taken up by mycorrhiza (myco_uptake)
#   2. N actually exported to the tree (mycorrhizal_export_to_tree) vs max_N_transfer cap
#   3. N sitting in the free fungal pool (ectomycorrhiza_N_free)
#   4. Transfer efficiency: export / uptake
#   5. C:N exchange ratio: kg C paid by tree per kg N received from ECM
## Parameters matching p_test_boreal.ini
k_0_root  <- 0.2
k_1_root  <- 0.35
k_8_age   <- 2.7
u_transfer_m2 <- 0.033   # kg N m^-2 yr^-1 (matches updated ini)

diag_df <- opt_joint %>%
  mutate(
    transfer_efficiency  = mycorrhizal_export_to_tree /
                           pmax(myco_uptake, 1e-12),
    root_diameter_mm     = k_1_root / root_length^k_0_root,
    SA_tip_m2            = pi * (root_diameter_mm / 1000) * (root_length / 1000),
    root_SA_m2           = SA_tip_m2 * root_no * crown_area * lai,
    age_factor_diag      = 1 / (1 + exp(fineroot_lifespan - k_8_age)),
    max_N_transfer       = u_transfer_m2 * age_factor_diag * root_SA_m2,  # kg N yr^-1
    cn_exchange          = C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12)
  )

p_diag_uptake <- ggplot(diag_df, aes(x = soil_nitrogen, y = myco_uptake)) +
  geom_point(shape = 1, size = 3) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Mycorrhizal N uptake\n(kg N tree⁻¹ yr⁻¹)", tag = "A") +
  pub_theme

p_diag_export <- ggplot(diag_df, aes(x = soil_nitrogen)) +
  geom_point(aes(y = mycorrhizal_export_to_tree), shape = 1, size = 3) +
  geom_line(aes(y = max_N_transfer, linetype = "max_N_transfer cap"), colour = "firebrick") +
  scale_linetype_manual(values = c("max_N_transfer cap" = "dashed"), name = NULL) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "N exported to tree\n(kg N tree⁻¹ yr⁻¹)", tag = "B") +
  pub_theme

p_diag_nfree <- ggplot(diag_df, aes(x = soil_nitrogen, y = ectomycorrhiza_N_free)) +
  geom_point(shape = 1, size = 3) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Fungal free N pool\n(kg N)", tag = "C") +
  pub_theme

p_diag_eff <- ggplot(diag_df, aes(x = soil_nitrogen, y = transfer_efficiency)) +
  geom_point(shape = 1, size = 3) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Transfer efficiency\n(export / uptake)", tag = "D") +
  pub_theme

p_diag_cn <- ggplot(diag_df, aes(x = soil_nitrogen, y = cn_exchange)) +
  geom_point(shape = 1, size = 3) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "C:N exchange cost\n(kg C kg⁻¹ N received)", tag = "E") +
  pub_theme

myco_transfer_diag <- (p_diag_uptake | p_diag_export | p_diag_nfree | p_diag_eff | p_diag_cn)
myco_transfer_diag

ggsave("plots/12_myco_transfer_diag.png", plot = myco_transfer_diag,
       width = 22, height = 5, dpi = 300)
```

## Objective function comparison

```{r objective_comparison}
# Net GPP per crown area: GPP minus all N-related belowground C costs, per m2 canopy.
#   rr             = fine root respiration (kg C / time unit, direct model output)
#   tr             = fine root turnover   (kg biomass / time unit); * 0.44 = carbon
#   C_export_to_myco = C paid to mycorrhiza (kg C / time unit)
# Coarse root respiration excluded: invariant to root_no and ecto_allo.
add_net_gpp_ca <- function(df) {
  df %>% mutate(
    net_gpp_ca = (assim_gross - C_export_to_myco - rr - tr * 0.44) / crown_area
  )
}

# Last-year snapshot: 2021 growing-season mean (joint_gs already computed above)
joint_gs_lastyear <- joint_gs %>% add_net_gpp_ca()

# Multi-year mean: all growing seasons Jun-Aug across the full simulation
joint_gs_allyears <- all_joint_df %>%
  mutate(year  = as.integer(format(as.Date(date), "%Y")),
         month = as.integer(format(as.Date(date), "%m"))) %>%
  filter(month %in% 6:8) %>%
  group_by(soil_nitrogen, ecto_allo, root_no, root_length) %>%
  summarise(
    height                      = mean(height,                      na.rm = TRUE),
    ectomycorrhiza_mass         = mean(ectomycorrhiza_mass,         na.rm = TRUE),
    ectomycorrhiza_N_biomass    = mean(ectomycorrhiza_N_biomass,    na.rm = TRUE),
    ectomycorrhiza_C_free       = mean(ectomycorrhiza_C_free,       na.rm = TRUE),
    ectomycorrhiza_N_free       = mean(ectomycorrhiza_N_free,       na.rm = TRUE),
    crown_area                  = mean(crown_area,                  na.rm = TRUE),
    lai                         = mean(lai,                         na.rm = TRUE),
    N_bar_roots                 = mean(N_bar_roots,                 na.rm = TRUE),
    N_bar_myco                  = mean(N_bar_myco,                  na.rm = TRUE),
    C_export_to_myco            = mean(C_export_to_myco,            na.rm = TRUE),
    mycorrhizal_export_to_tree  = mean(mycorrhizal_export_to_tree,  na.rm = TRUE),
    root_uptake                 = mean(root_uptake,                 na.rm = TRUE),
    myco_uptake                 = mean(myco_uptake,                 na.rm = TRUE),
    assim_gross                 = mean(assim_gross,                 na.rm = TRUE),
    optimal_leaf_nitrogen       = mean(optimal_leaf_nitrogen,       na.rm = TRUE),
    leaf_nitrogen_concentration = mean(leaf_nitrogen_concentration, na.rm = TRUE),
    tree_nitrogen               = mean(tree_nitrogen,               na.rm = TRUE),
    root_mass                   = mean(root_mass,                   na.rm = TRUE),
    coarse_root_mass            = mean(coarse_root_mass,            na.rm = TRUE),
    total_mass                  = mean(total_mass,                  na.rm = TRUE),
    fineroot_lifespan           = mean(fineroot_lifespan,           na.rm = TRUE),
    total_prod                  = mean(total_prod,                  na.rm = TRUE),
    rr                          = mean(rr,                          na.rm = TRUE),
    tr                          = mean(tr,                          na.rm = TRUE),
    ecm_frac_gpp                = mean(C_export_to_myco / assim_gross, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  add_net_gpp_ca()

# Full 12-panel joint-optima figure for any objective column.
# gs_data must contain all columns from joint_gs plus tr.
make_optima_figure <- function(gs_data, var_col, fig_title) {
  var_sym <- sym(var_col)

  opt <- gs_data %>%
    group_by(soil_nitrogen) %>%
    mutate(
      max_v      = max(!!var_sym, na.rm = TRUE),
      near_opt   = !!var_sym >= 0.99 * max_v,
      n_total    = n(),
      n_near_opt = sum(near_opt),
      stab_frac  = n_near_opt / n_total
    ) %>%
    filter(near_opt) %>%
    slice_min(root_no, n = 1, with_ties = FALSE) %>%
    ungroup()

  stab <- gs_data %>%
    group_by(soil_nitrogen) %>%
    mutate(max_v = max(!!var_sym, na.rm = TRUE), near_opt = !!var_sym >= 0.99 * max_v) %>%
    filter(near_opt) %>%
    summarise(
      ecto_lo = min(ecto_allo,   na.rm = TRUE), ecto_hi = max(ecto_allo,   na.rm = TRUE),
      rn_lo   = min(root_no,     na.rm = TRUE), rn_hi   = max(root_no,     na.rm = TRUE),
      rl_lo   = min(root_length, na.rm = TRUE), rl_hi   = max(root_length, na.rm = TRUE),
      .groups = "drop"
    )

  stab_out <- gs_data %>%
    group_by(soil_nitrogen) %>%
    mutate(max_v = max(!!var_sym, na.rm = TRUE), near_opt = !!var_sym >= 0.99 * max_v) %>%
    filter(near_opt) %>%
    summarise(
      ecm_lo        = min(ectomycorrhiza_mass, na.rm = TRUE),
      ecm_hi        = max(ectomycorrhiza_mass, na.rm = TRUE),
      ecm_area_lo   = min(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
      ecm_area_hi   = max(ectomycorrhiza_mass / crown_area, na.rm = TRUE),
      height_lo     = min(height,     na.rm = TRUE),
      height_hi     = max(height,     na.rm = TRUE),
      nroots_lo     = min(N_bar_roots, na.rm = TRUE),
      nroots_hi     = max(N_bar_roots, na.rm = TRUE),
      nmyco_lo      = min(N_bar_myco,  na.rm = TRUE),
      nmyco_hi      = max(N_bar_myco,  na.rm = TRUE),
      cn_lo         = min((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                            pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
      cn_hi         = max((ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                            pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12), na.rm = TRUE),
      gpp_lo        = min(assim_gross / crown_area, na.rm = TRUE),
      gpp_hi        = max(assim_gross / crown_area, na.rm = TRUE),
      n_upper_lo    = min(leaf_nitrogen_concentration, na.rm = TRUE),
      n_upper_hi    = max(leaf_nitrogen_concentration, na.rm = TRUE),
      n_realised_lo = min(optimal_leaf_nitrogen, na.rm = TRUE),
      n_realised_hi = max(optimal_leaf_nitrogen, na.rm = TRUE),
      n_free_lo     = min(tree_nitrogen, na.rm = TRUE),
      n_free_hi     = max(tree_nitrogen, na.rm = TRUE),
      bga_lo        = min((C_export_to_myco + rr + tr * 0.44) / assim_gross, na.rm = TRUE),
      bga_hi        = max((C_export_to_myco + rr + tr * 0.44) / assim_gross, na.rm = TRUE),
      cn_exch_lo    = min(C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12), na.rm = TRUE),
      cn_exch_hi    = max(C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12), na.rm = TRUE),
      .groups = "drop"
    )

  opt_cn <- opt %>%
    mutate(CN = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                 pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12))

  fn_colour <- c("Stability band" = "grey70")
  fn_scales <- list(
    scale_shape_manual(values = shape_vals, name = NULL),
    scale_linetype_manual(values = ltype_vals, name = NULL),
    scale_colour_manual(values = fn_colour, name = NULL)
  )
  stab_lbl <- geom_text(aes(label = paste0(round(stab_frac * 100), "%")),
                        vjust = -0.8, size = 3, inherit.aes = TRUE)

  pA <- ggplot(opt, aes(x = soil_nitrogen, y = ecto_allo)) +
    geom_linerange(data = stab, aes(x = soil_nitrogen, ymin = ecto_lo, ymax = ecto_hi,
                   colour = "Stability band"), linewidth = 2, inherit.aes = FALSE) +
    geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 3, stroke = 1.2) +
    stab_lbl + fn_scales + scale_y_continuous(labels = scales::label_percent()) +
    labs(tag = "A", x = "Soil nitrogen (g N / m3)", y = "Optimal ECM C allocation
(% NPP)") +
    pub_theme

  pB <- ggplot(opt, aes(x = soil_nitrogen, y = root_no)) +
    geom_linerange(data = stab, aes(x = soil_nitrogen, ymin = rn_lo, ymax = rn_hi,
                   colour = "Stability band"), linewidth = 2, inherit.aes = FALSE) +
    geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 3, stroke = 1.2) +
    stab_lbl + fn_scales + scale_y_log10(labels = scales::label_scientific()) +
    labs(tag = "B", x = "Soil nitrogen (g N / m3)", y = "Optimal root number
(no / m2 crown)") +
    pub_theme

  pC <- ggplot(opt, aes(x = soil_nitrogen, y = root_length)) +
    geom_linerange(data = stab, aes(x = soil_nitrogen, ymin = rl_lo, ymax = rl_hi,
                   colour = "Stability band"), linewidth = 2, inherit.aes = FALSE) +
    geom_point(aes(shape = "Optimal parameters"), colour = "black", size = 3, stroke = 1.2) +
    stab_lbl + fn_scales +
    labs(tag = "C", x = "Soil nitrogen (g N / m3)", y = "Optimal root length (m)") +
    pub_theme

  pD <- ggplot(opt,
      aes(x = soil_nitrogen, y = (C_export_to_myco + rr + tr * 0.44) / assim_gross)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = bga_lo, ymax = bga_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Calculated"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_percent()) + expand_limits(y = 0) +
    labs(tag = "D", x = "Soil nitrogen (g N / m3)", y = "N-related belowground C
allocation (% GPP)") +
    pub_theme

  pE <- ggplot(opt, aes(x = soil_nitrogen, y = ectomycorrhiza_mass)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = ecm_lo, ymax = ecm_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_scientific()) +
    labs(tag = "E", x = "Soil nitrogen (g N / m3)", y = "ECM biomass (kg)") + pub_theme

  pF <- ggplot(opt, aes(x = soil_nitrogen, y = ectomycorrhiza_mass / crown_area)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = ecm_area_lo, ymax = ecm_area_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_scientific()) +
    labs(tag = "F", x = "Soil nitrogen (g N / m3)", y = "ECM biomass per crown
(kg / m2)") + pub_theme

  nr_rng <- range(opt$N_bar_roots, na.rm = TRUE)
  nm_rng <- range(opt$N_bar_myco,  na.rm = TRUE)
  m2r    <- function(x) nr_rng[1] + (x - nm_rng[1]) / diff(nm_rng) * diff(nr_rng)
  r2m    <- function(x) nm_rng[1]  + (x - nr_rng[1]) / diff(nr_rng) * diff(nm_rng)

  pG <- ggplot(opt, aes(x = soil_nitrogen, y = N_bar_roots)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = nroots_lo, ymax = nroots_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_ribbon(data = stab_out,
                aes(x = soil_nitrogen, ymin = m2r(nmyco_lo), ymax = m2r(nmyco_hi)),
                fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    geom_point(aes(y = m2r(N_bar_myco)), colour = "steelblue", shape = 1, size = 3, stroke = 1.2) +
    fn_scales +
    scale_y_continuous(
      name = "N_bar roots (kg N / m3)",
      sec.axis = sec_axis(~ r2m(.), name = "N_bar myco (kg N / m3)")
    ) +
    labs(tag = "G", x = "Soil nitrogen (g N / m3)") +
    pub_theme +
    theme(axis.title.y.right = element_text(colour = "steelblue"),
          axis.text.y.right  = element_text(colour = "steelblue"),
          axis.ticks.y.right = element_line(colour = "steelblue"),
          axis.line.y.right  = element_line(colour = "steelblue"))

  pH <- ggplot(opt_cn, aes(x = soil_nitrogen, y = CN)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = cn_lo, ymax = cn_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + scale_y_continuous(labels = scales::label_scientific(digits = 2)) +
    labs(tag = "H", x = "Soil nitrogen (g N / m3)", y = "Mycorrhizal C:N
(kg C / kg N)") + pub_theme

  pI <- ggplot(opt, aes(x = soil_nitrogen, y = height)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = height_lo, ymax = height_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = 0) +
    labs(tag = "I", x = "Soil nitrogen (g N / m3)", y = "Height (m)") + pub_theme

  pJ <- ggplot(opt, aes(x = soil_nitrogen, y = assim_gross / crown_area)) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = gpp_lo, ymax = gpp_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = 0) +
    labs(tag = "J", x = "Soil nitrogen (g N / m3)", y = "GPP per crown area
(kg C / m2 / yr)") + pub_theme

  pK <- ggplot(opt, aes(x = soil_nitrogen)) +
    geom_ribbon(data = stab_out,
                aes(x = soil_nitrogen, ymin = n_upper_lo, ymax = n_upper_hi),
                fill = "steelblue", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
    geom_ribbon(data = stab_out,
                aes(x = soil_nitrogen, ymin = n_realised_lo, ymax = n_realised_hi),
                fill = "grey60", alpha = 0.25, colour = NA, inherit.aes = FALSE) +
    geom_point(aes(y = leaf_nitrogen_concentration, colour = "Potential (leaf N pool)"),
               size = 3, stroke = 1.2) +
    geom_point(aes(y = optimal_leaf_nitrogen), colour = "grey30", shape = 1, size = 3, stroke = 1.2) +
    scale_colour_manual(values = c("Potential (leaf N pool)" = "steelblue"), name = NULL) +
    expand_limits(y = 0) +
    labs(tag = "K", x = "Soil nitrogen (g N / m3)", y = "Leaf N concentration
(kg N / kg leaf)") +
    pub_theme

  pL <- ggplot(opt, aes(x = soil_nitrogen, y = C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12))) +
    geom_ribbon(data = stab_out, aes(x = soil_nitrogen, ymin = cn_exch_lo, ymax = cn_exch_hi),
                fill = "grey80", colour = NA, inherit.aes = FALSE) +
    geom_point(aes(shape = "Model output"), colour = "black", size = 3, stroke = 1.2) +
    fn_scales + expand_limits(y = 0) +
    labs(tag = "L", x = "Soil nitrogen (g N / m3)", y = "C:N exchange cost\n(kg C / kg N received)") + pub_theme

  guide_area() /
    make_row_title("Parameters to be Optimised") /
    (pA | pB | pC | pD) /
    make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
    (pE | pF | pG | pH) /
    make_row_title("Model results: Tree Status") /
    (pI | pJ | pK | pL) +
    plot_layout(heights = c(0.15, 0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
    plot_annotation(
      title = fig_title,
      theme = theme(plot.margin = margin(5, 30, 5, 5),
                    plot.title  = element_text(size = 18, face = "bold"))
    ) &
    theme(plot.tag        = element_text(size = 18, face = "bold"),
          legend.text     = element_text(size = 14),
          legend.key.size = unit(0.9, "lines"),
          legend.position    = "top",
          legend.direction   = "horizontal",
          legend.box         = "horizontal")
}

fig_lastyear <- make_optima_figure(
  joint_gs_lastyear, "net_gpp_ca",
  "Objective: net GPP / crown area  (summer 2021)"
)
ggsave("plots/11_obj_lastyear.png", plot = fig_lastyear, width = 22, height = 18, dpi = 300)

fig_allyears <- make_optima_figure(
  joint_gs_allyears, "net_gpp_ca",
  "Objective: net GPP / crown area  (multi-year summer mean)"
)
ggsave("plots/11_obj_allyears.png", plot = fig_allyears, width = 22, height = 18, dpi = 300)
```


## Net GPP/CA landscape diagnostic

```{r netgpp_landscape}
# Plot net_gpp_ca vs root_no at representative soil N levels.
# If the landscape is flat at high N, slice_min already picks the right answer.
# If it is still rising at high N, extra roots genuinely help throughout the tree's
# life (historical N-accumulation confound) and annual re-optimisation is needed.

landscape_netgpp <- joint_gs_lastyear %>%
  add_net_gpp_ca() %>%
  group_by(soil_nitrogen, root_no) %>%
  summarise(
    net_gpp_ca_mean = mean(net_gpp_ca, na.rm = TRUE),
    net_gpp_ca_max  = max(net_gpp_ca,  na.rm = TRUE),
    .groups = "drop"
  )

n_levels_diag  <- sort(unique(landscape_netgpp$soil_nitrogen))
n_sel          <- n_levels_diag[round(seq(1, length(n_levels_diag),
                                          length.out = min(6, length(n_levels_diag))))]

p_netgpp_landscape <- ggplot(
    landscape_netgpp %>% filter(soil_nitrogen %in% n_sel),
    aes(x = root_no, y = net_gpp_ca_mean,
        colour = factor(round(soil_nitrogen, 4)),
        group  = factor(round(soil_nitrogen, 4)))
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_colour_viridis_d(name = "Soil N\n(g N m⁻³)") +
  labs(x = "Root number per crown area (no m⁻²)",
       y = "Net GPP / crown area\n(mean across ecto_allo)") +
  pub_theme

# Also show per ecto_allo to check whether EcM allocation changes the shape
landscape_netgpp_full <- joint_gs_lastyear %>%
  add_net_gpp_ca() %>%
  filter(soil_nitrogen %in% n_sel)

p_netgpp_landscape_ecto <- ggplot(
    landscape_netgpp_full,
    aes(x = root_no, y = net_gpp_ca,
        colour = factor(round(soil_nitrogen, 4)),
        group  = interaction(factor(round(soil_nitrogen, 4)), ecto_allo))
  ) +
  geom_line(linewidth = 0.5, alpha = 0.6) +
  geom_point(size = 2, aes(shape = factor(round(ecto_allo, 2)))) +
  scale_colour_viridis_d(name = "Soil N\n(g N m⁻³)") +
  scale_shape_discrete(name = "EcM alloc") +
  labs(x = "Root number per crown area (no m⁻²)",
       y = "Net GPP / crown area") +
  pub_theme

p_netgpp_landscape / p_netgpp_landscape_ecto
ggsave("plots/15_landscape.png", plot = p_netgpp_landscape / p_netgpp_landscape_ecto, width = 22, height = 18, dpi = 300)

# ECM C vs N limitation diagnostic.
# ECM growth: C_for_growth = min(C_available, N_available / nc_myco).
# nc_myco = 0.1, so 1 kg C invested needs 0.1 kg N to build biomass.
# cn_free = free_C / (free_N / nc_myco): >1 means C is in excess → ECM is N-limited.
# p_nbar: compare N_bar_myco (N at fungal surface after depletion) against k_15 (0.5).
# If N_bar << k_15 everywhere, the MM term never saturates and k_15 is irrelevant.
ecm_lim <- joint_gs_lastyear %>%
  filter(soil_nitrogen %in% n_sel) %>%
  group_by(soil_nitrogen, root_no) %>%
  summarise(
    C_free  = mean(ectomycorrhiza_C_free,  na.rm = TRUE),
    N_free  = mean(ectomycorrhiza_N_free,  na.rm = TRUE),
    N_bar   = mean(N_bar_myco,             na.rm = TRUE),
    cn_exch = mean(C_export_to_myco / pmax(mycorrhizal_export_to_tree, 1e-12), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(cn_free = C_free / pmax(N_free / 0.1, 1e-12))

p_ecm_lim <- ggplot(ecm_lim,
    aes(x = soil_nitrogen, y = cn_free, colour = factor(round(root_no, -4)))) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey50") +
  geom_point(size = 2.5) + geom_line(aes(group = factor(round(root_no, -4)))) +
  scale_colour_viridis_d(name = "Root no") +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "ECM free C / (free N / nc_myco)\n(> 1 = N-limited, < 1 = C-limited)") +
  pub_theme

p_nbar <- ggplot(ecm_lim,
    aes(x = soil_nitrogen, y = N_bar, colour = factor(round(root_no, -4)))) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "red") +
  annotate("text", x = min(ecm_lim$soil_nitrogen, na.rm = TRUE), y = 0.52,
           label = "k_15 = 0.5", hjust = 0, size = 5, colour = "red") +
  geom_point(size = 2.5) + geom_line(aes(group = factor(round(root_no, -4)))) +
  scale_colour_viridis_d(name = "Root no") +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "N_bar at fungal surface\n(must be in same units as k_15 to be meaningful)") +
  pub_theme

(p_ecm_lim | p_nbar)
```


## Root tip number landscape

```{r root_tip_landscape}
# Show the fitness landscape across root_no values at key soil N levels.
# This reveals whether the jump in optimal root_no (Panel B of the main figure)
# is a genuine sharp optimum or an artefact of the coarse grid.

# For each soil_nitrogen × root_no combination, take the best (max height)
# across all ecto_allo and root_length values. Also pull rr at that combo so we
# can show the cost side of the root_no trade-off.
root_landscape <- joint_gs %>%
  group_by(soil_nitrogen, root_no) %>%
  summarise(
    best_mass = max(height,    na.rm = TRUE),
    best_ecto = ecto_allo[which.max(height)],
    best_rl   = root_length[which.max(height)],
    best_rr   = rr[which.max(height)],
    .groups = "drop"
  ) %>%
  group_by(soil_nitrogen) %>%
  mutate(rel_mass = best_mass / max(best_mass, na.rm = TRUE)) %>%
  ungroup()

# Pick a handful of soil N levels that span the range including the jump
n_levels    <- sort(unique(root_landscape$soil_nitrogen))
n_highlight <- n_levels[round(seq(1, length(n_levels), length.out = min(6, length(n_levels))))]

p_root_landscape <- ggplot(
    root_landscape %>% filter(soil_nitrogen %in% n_highlight),
    aes(x = root_no, y = best_mass,
        colour = factor(round(soil_nitrogen, 2)),
        group  = factor(round(soil_nitrogen, 2)))
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_log10(labels = scales::label_scientific()) +
  scale_colour_viridis_d(name = "Soil N\n(g N m⁻³)") +
  labs(x = "Root number per crown area (no m⁻²)",
       y = "Height (m)") +
  pub_theme

# Normalised benefit vs cost: height gain and rr cost both scaled to [0,1]
# within each N level. Where rr (cost, dashed) rises faster than height
# (benefit, solid), adding more roots is no longer worthwhile.
# At high N the benefit still dominates across the whole grid, which is why
# the optimum sits at the maximum root_no value rather than an interior point.
root_tradeoff <- root_landscape %>%
  filter(soil_nitrogen %in% n_highlight) %>%
  group_by(soil_nitrogen) %>%
  mutate(
    height_gain = (best_mass - min(best_mass)) /
                    (max(best_mass) - min(best_mass) + 1e-12),
    rr_gain     = (best_rr   - min(best_rr))   /
                    (max(best_rr)   - min(best_rr)   + 1e-12)
  ) %>%
  ungroup() %>%
  pivot_longer(cols      = c(height_gain, rr_gain),
               names_to  = "metric",
               values_to = "value") %>%
  mutate(metric = recode(metric,
                         height_gain = "Height benefit",
                         rr_gain     = "Respiration cost"))

p_root_tradeoff <- ggplot(
    root_tradeoff,
    aes(x        = root_no,
        y        = value,
        colour   = factor(round(soil_nitrogen, 2)),
        linetype = metric,
        group    = interaction(factor(round(soil_nitrogen, 2)), metric))
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_log10(labels = scales::label_scientific()) +
  scale_colour_viridis_d(name = "Soil N\n(g N m⁻³)") +
  scale_linetype_manual(values = c("Height benefit"   = "solid",
                                   "Respiration cost" = "dashed"),
                        name = NULL) +
  labs(x = "Root number per crown area (no m⁻²)",
       y = "Normalised gain\n(fraction of range)") +
  pub_theme

# Show how the optimal root_no and its nearest competitor compare across N
root_rank <- joint_gs %>%
  group_by(soil_nitrogen, root_no) %>%
  summarise(best_mass = max(height, na.rm = TRUE), .groups = "drop") %>%
  group_by(soil_nitrogen) %>%
  arrange(desc(best_mass)) %>%
  mutate(rank     = row_number(),
         max_m    = max(best_mass, na.rm = TRUE),
         rel_mass = best_mass / max_m) %>%
  ungroup()

p_root_rank <- ggplot(root_rank %>% filter(rank <= 3),
    aes(x = soil_nitrogen, y = root_no, colour = factor(rank), shape = factor(rank))
  ) +
  geom_point(size = 4) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_colour_manual(values = c("1" = "black", "2" = "steelblue", "3" = "grey60"),
                      labels = c("1st", "2nd", "3rd"), name = "Rank") +
  scale_shape_manual(values  = c("1" = 16, "2" = 1, "3" = 2),
                     labels  = c("1st", "2nd", "3rd"), name = "Rank") +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Root number per crown area (no m⁻²)") +
  pub_theme

p_root_margin <- ggplot(
    root_rank %>% filter(rank == 2) %>%
      inner_join(root_rank %>% filter(rank == 1) %>%
                   select(soil_nitrogen, best_mass) %>%
                   rename(top_mass = best_mass),
                 by = "soil_nitrogen") %>%
      mutate(margin = (top_mass - best_mass) / top_mass),
    aes(x = soil_nitrogen, y = margin)
  ) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "Fitness gap to 2nd-best\nroot_no (% height)") +
  pub_theme

root_diag <- (p_root_landscape | p_root_tradeoff) /
             (p_root_rank      | p_root_margin)
root_diag

ggsave("plots/14_root_tip_landscape.png", plot = root_diag, width = 18, height = 12, dpi = 300)
```

## Belowground C allocation components

```{r bga_components}
# tr (model root turnover) vs root_mass * 0.44 / fineroot_lifespan (R recomputation).
# tr is what the model uses in NPP = GPP - R - T, so it is the authoritative value.
# root_mass * 0.44 / fineroot_lifespan is annual C; tr is per time-unit — they should
# agree after accounting for years_per_tunit, but are shown side by side for verification.
bga_diag <- opt_joint %>%
  select(soil_nitrogen, C_export_to_myco, root_mass, coarse_root_mass,
         fineroot_lifespan, rr, tr, assim_gross) %>%
  mutate(
    myco_term        = C_export_to_myco / assim_gross,
    fine_root_resp   = rr / assim_gross,
    fine_root_turn_R = (root_mass * 0.44 / fineroot_lifespan) / assim_gross,  # R recomputation
    fine_root_turn_m = tr / assim_gross,                                        # model tr
    coarse_root_term = (r_coarse_root / rr_param) * rr * coarse_root_mass / root_mass / assim_gross,
    total_bga        = myco_term + fine_root_resp + fine_root_turn_R + coarse_root_term
  ) %>%
  tidyr::pivot_longer(cols = c(myco_term, fine_root_resp, fine_root_turn_R,
                               fine_root_turn_m, coarse_root_term, total_bga),
                      names_to = "component", values_to = "fraction") %>%
  mutate(component = factor(component,
                            levels = c("myco_term", "fine_root_resp", "fine_root_turn_R",
                                       "fine_root_turn_m", "coarse_root_term", "total_bga"),
                            labels = c("ECM C export / GPP",
                                       "Fine root respiration / GPP",
                                       "Fine root turnover (R calc) / GPP",
                                       "Fine root turnover (model tr) / GPP",
                                       "Coarse root respiration / GPP",
                                       "Total belowground")))

p_bga_diag <- ggplot(bga_diag, aes(x = soil_nitrogen, y = fraction, colour = component)) +
  geom_point(size = 3) +
  geom_line() +
  scale_colour_manual(values = c("ECM C export / GPP"                    = "darkorange",
                                  "Fine root respiration / GPP"          = "steelblue",
                                  "Fine root turnover (R calc) / GPP"    = "steelblue4",
                                  "Fine root turnover (model tr) / GPP"  = "dodgerblue",
                                  "Coarse root respiration / GPP"        = "forestgreen",
                                  "Total belowground"                    = "black"),
                      name = NULL) +
  expand_limits(y = 0) +
  labs(x = "Soil nitrogen (g N m⁻³)",
       y = "C allocation (fraction of GPP)") +
  pub_theme

p_bga_diag

ggsave("plots/13_bga_components.png", plot = p_bga_diag, width = 8, height = 5, dpi = 300)
```

## Joint optima: trajectories over time

```{r joint_optima_trajectories}
# Low and high N endpoints of the sweep
n_low  <- nitrogen_values[1]
n_high <- nitrogen_values[length(nitrogen_values)]

# For each year × N level, find the parameter combo that maximises height at
# that year (growing-season mean), then pull all outputs from that combo.
# This shows how optimal root_no, root_length, and ecto_allo shift as the tree ages.
annual_opt <- all_joint_df %>%
  filter(soil_nitrogen %in% c(n_low, n_high)) %>%
  mutate(year  = as.integer(format(as.Date(date), "%Y")),
         month = as.integer(format(as.Date(date), "%m"))) %>%
  filter(month %in% 6:8) %>%
  group_by(soil_nitrogen, year, ecto_allo, root_no, root_length) %>%
  summarise(
    height                      = mean(height,                      na.rm = TRUE),
    ectomycorrhiza_mass         = mean(ectomycorrhiza_mass,         na.rm = TRUE),
    ectomycorrhiza_N_biomass    = mean(ectomycorrhiza_N_biomass,    na.rm = TRUE),
    ectomycorrhiza_N_free       = mean(ectomycorrhiza_N_free,       na.rm = TRUE),
    ectomycorrhiza_C_free       = mean(ectomycorrhiza_C_free,       na.rm = TRUE),
    crown_area                  = mean(crown_area,                  na.rm = TRUE),
    assim_gross                 = mean(assim_gross,                 na.rm = TRUE),
    leaf_nitrogen_concentration = mean(leaf_nitrogen_concentration, na.rm = TRUE),
    optimal_leaf_nitrogen       = mean(optimal_leaf_nitrogen,       na.rm = TRUE),
    tree_nitrogen               = mean(tree_nitrogen,               na.rm = TRUE),
    N_bar_roots                 = mean(N_bar_roots,                 na.rm = TRUE),
    N_bar_myco                  = mean(N_bar_myco,                  na.rm = TRUE),
    C_export_to_myco            = mean(C_export_to_myco,           na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(soil_nitrogen, year) %>%
  slice_max(height, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    CN           = (ectomycorrhiza_mass * 0.44 + ectomycorrhiza_C_free) /
                     pmax(ectomycorrhiza_N_biomass + ectomycorrhiza_N_free, 1e-12),
    gpp_per_ca   = assim_gross / crown_area,
    ecm_frac_gpp = C_export_to_myco / assim_gross,
    N_label      = factor(
      ifelse(soil_nitrogen == n_low, "Low N", "High N"),
      levels = c("Low N", "High N"),
      labels = c(paste0("Low N (",  round(n_low,  1), " g m⁻³)"),
                 paste0("High N (", round(n_high, 1), " g m⁻³)"))
    )
  )

traj_colour2 <- scale_colour_manual(
  name   = "Soil N",
  values = c("#2166ac", "#d6604d")
)

# --- Top row: how optimal parameter values shift as the tree grows ---
p_traj_ecto <- ggplot(annual_opt,
    aes(x = year, y = ecto_allo, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) + geom_point(size = 1.5) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = NULL, y = "Optimal ECM C allocation\n(% NPP)") +
  pub_theme

p_traj_rn <- ggplot(annual_opt,
    aes(x = year, y = root_no, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) + geom_point(size = 1.5) +
  traj_colour2 +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Optimal root number\nper crown area (no m⁻²)") +
  pub_theme

p_traj_rl <- ggplot(annual_opt,
    aes(x = year, y = root_length, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) + geom_point(size = 1.5) +
  traj_colour2 +
  labs(x = NULL, y = "Optimal root length (m)") +
  pub_theme

# --- Middle row: emergent mycorrhizal outputs ---
p_traj_ecm <- ggplot(annual_opt,
    aes(x = year, y = ectomycorrhiza_mass, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Ectomycorrhizal\nbiomass (kg)") +
  pub_theme

p_traj_ecm_area <- ggplot(annual_opt,
    aes(x = year, y = ectomycorrhiza_mass / crown_area,
        colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(x = NULL, y = "Ectomycorrhizal biomass\nper crown area (kg m⁻²)") +
  pub_theme

p_traj_nbar <- ggplot(annual_opt,
    aes(x = year, y = N_bar_roots, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  labs(x = NULL, y = expression(bar(N)[roots]~"(kg N m"^{-3}*")")) +
  pub_theme

p_traj_cn <- ggplot(annual_opt,
    aes(x = year, y = CN, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  labs(x = NULL, y = "Mycorrhizal C:N ratio\n(kg C kg⁻¹ N)") +
  pub_theme

# --- Bottom row: tree status ---
p_traj_height <- ggplot(annual_opt,
    aes(x = year, y = height, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  labs(x = NULL, y = "Height (m)") +
  pub_theme

p_traj_gpp <- ggplot(annual_opt,
    aes(x = year, y = gpp_per_ca, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  labs(x = NULL, y = "GPP per crown area\n(kg C m⁻² yr⁻¹)") +
  pub_theme

p_traj_n_limit <- ggplot(annual_opt,
    aes(x = year, colour = N_label, group = N_label)) +
  geom_line(aes(y = leaf_nitrogen_concentration), linetype = "dashed", linewidth = 0.8) +
  geom_line(aes(y = optimal_leaf_nitrogen),       linetype = "solid",  linewidth = 0.8) +
  traj_colour2 +
  labs(x = NULL, y = "Leaf N concentration\n(kg N kg⁻¹ leaf)") +
  pub_theme

p_traj_nfree <- ggplot(annual_opt,
    aes(x = year, y = tree_nitrogen, colour = N_label, group = N_label)) +
  geom_line(linewidth = 0.8) +
  traj_colour2 +
  labs(x = NULL, y = "Free N pool\n(kg N tree⁻¹)") +
  pub_theme

traj_tags <- list(c("", "A", "B", "C", "", "", "D", "E", "F", "G", "", "H", "I", "J", "K"))

joint_optima_traj <-
  make_row_title("Parameters to be Optimised") /
  (p_traj_ecto | p_traj_rn | p_traj_rl | plot_spacer()) /
  make_row_title("Model results: Mycorrhizal dynamics & N depletion") /
  (p_traj_ecm | p_traj_ecm_area | p_traj_nbar | p_traj_cn) /
  make_row_title("Model results: Tree Status") /
  (p_traj_height | p_traj_gpp | p_traj_n_limit | p_traj_nfree) +
  plot_layout(heights = c(0.07, 1, 0.07, 1, 0.07, 1), guides = "collect") +
  plot_annotation(tag_levels = traj_tags) &
  theme(plot.tag        = element_text(size = 22, face = "bold"),
        legend.text     = element_text(size = 18),
        legend.key.size = unit(1.2, "lines"),
        legend.position = "right")
joint_optima_traj

ggsave("plots/11b_joint_optima_trajectories.png",
       plot = joint_optima_traj, width = 20, height = 18, dpi = 300)
```

## Data comparison

```{r data_load}
ecm_data_all <- read.csv(here("RScripts/ecm_root_data_compiled.csv"), stringsAsFactors = FALSE)

# Strip leading ~, <, > from all columns so approximate values become plain numbers
ecm_data_all <- as.data.frame(lapply(ecm_data_all, function(col) {
  if (is.character(col)) {
    cleaned <- gsub("^[~<>]+", "", trimws(col))
    num <- suppressWarnings(as.numeric(cleaned))
    if (sum(!is.na(num)) > sum(!is.na(suppressWarnings(as.numeric(col))))) num else col
  } else col
}))

# Model is calibrated for Scots pine / boreal evergreen conifer — keep only conifers
evergreen_conifers <- c("Pinus sylvestris", "Picea abies", "Picea abies (control)",
                        "Picea glauca", "Picea spp.", "Mixed (Pinus/Picea)")
ecm_data <- ecm_data_all[ecm_data_all$species %in% evergreen_conifers, ]
ecm_data <- ecm_data[is.na(ecm_data$soil_CN_ratio) | ecm_data$soil_CN_ratio < 100, ]
ecm_data$species <- factor(ecm_data$species, levels = evergreen_conifers)

species_colours <- c(
  "Pinus sylvestris"     = "#E69F00",
  "Picea abies"          = "#56B4E9",
  "Picea abies (control)"= "#009E73",
  "Picea glauca"         = "#F0E442",
  "Picea spp."           = "#0072B2",
  "Mixed (Pinus/Picea)"  = "#D55E00"
)
species_scale <- scale_colour_manual(name = "Species", values = species_colours,
                                     drop = FALSE,
                                     guide = guide_legend(ncol = 3, byrow = TRUE))

# Tip density: convert million m⁻² → m⁻² to match model root_no units
ecm_data$tip_density_per_m2 <- ecm_data$ecm_tip_density_million_per_m2_soil * 1e6
```

### Plot 07 A — Root length vs soil N
*Matches heatmap p2 (y-axis: root length, mm)*

```{r data_07a}
d_rl <- ecm_data[!is.na(ecm_data$ecm_root_tip_length_mm) &
                 !is.na(ecm_data$soil_CN_ratio), ]

p_data_rl <- ggplot(d_rl, aes(x = soil_CN_ratio, y = ecm_root_tip_length_mm,
                               colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  labs(
    x = "Soil C:N ratio",
    y = "ECM root tip\nlength (mm)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 07 B — Root / tip density vs soil N
*Matches heatmap p1 (y-axis: root number per crown area, m⁻²)*

```{r data_07b}
d_td <- ecm_data[!is.na(ecm_data$ecm_tip_density_million_per_m2_BA) &
                 !is.na(ecm_data$soil_CN_ratio), ]

p_data_td <- ggplot(d_td, aes(x = soil_CN_ratio, y = ecm_tip_density_million_per_m2_BA,
                               colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  scale_y_log10(labels = scales::label_scientific()) +
  labs(
    x = "Soil C:N ratio",
    y = "ECM tip density\n(million m⁻² basal area)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 09 A — ECM mycelium biomass vs soil N
*Matches heatmap p_ecm_mass (y-axis: ectomycorrhiza_mass)*

```{r data_09a}
d_ecm <- ecm_data[!is.na(ecm_data$ecm_mycelium_biomass_kg_ha) &
                  !is.na(ecm_data$soil_CN_ratio), ]

p_data_ecm <- ggplot(d_ecm, aes(x = soil_CN_ratio, y = ecm_mycelium_biomass_kg_ha,
                                 colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  scale_y_log10(labels = scales::label_number(accuracy = 1)) +
  labs(
    x = "Soil C:N ratio",
    y = "ECM mycelium\nbiomass (kg ha⁻¹)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### Plot 09 B — ECM root biomass vs soil N
*Proxy for ectomycorrhizal allocation (p_ecto_allo)*

```{r data_09b}
d_ecmr <- ecm_data[!is.na(ecm_data$ecm_root_biomass_kg_per_m2_basal_area) &
                   !is.na(ecm_data$soil_CN_ratio), ]

p_data_ecmr <- ggplot(d_ecmr, aes(x = soil_CN_ratio,
                                    y = ecm_root_biomass_kg_per_m2_basal_area,
                                    colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  labs(
    x = "Soil C:N ratio",
    y = "ECM root biomass\n(kg m⁻² basal area)",
    colour = "Species"
  ) +
  species_scale +
  pub_theme
```

### ECM colonisation percentage vs soil N

```{r data_colonisation}
d_col_plot <- ecm_data[!is.na(ecm_data$ecm_colonisation_pct) &
                        !is.na(ecm_data$soil_CN_ratio), ]

p_data_col <- ggplot(d_col_plot, aes(x = soil_CN_ratio, y = ecm_colonisation_pct,
                                      colour = species)) +
  geom_point(size = 4.5, alpha = 0.8) +

  coord_cartesian(ylim = c(NA, 100)) +
  labs(
    x      = "Soil C:N ratio",
    y      = "ECM colonisation (%)",
    colour = "Species"
  ) +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 13)) +
  pub_theme
```

```{r data_combined}
# Standalone legend using all species so the hash is consistent
legend_plot <- ggplot(
    data.frame(species = factor(evergreen_conifers, levels = evergreen_conifers)),
    aes(x = 1, y = 1, colour = species)) +
  geom_point(alpha = 0) +
  species_scale +
  theme_void() +
  theme(legend.position   = c(0.5, 0.5),
        legend.text       = element_text(size = 22),
        legend.title      = element_text(size = 24, face = "bold"),
        legend.key.size   = unit(1.5, "lines"),
        axis.text.x       = element_blank(),
        axis.text.y       = element_blank(),
        axis.title.x      = element_blank(),
        axis.title.y      = element_blank(),
        axis.ticks        = element_blank(),
        panel.background  = element_blank()) +
  guides(colour = guide_legend(ncol = 1,
                               override.aes = list(alpha = 1, size = 5)))

data_combined <- wrap_plots(
  p_data_rl   + theme(legend.position = "none"),
  p_data_td   + theme(legend.position = "none"),
  p_data_ecm  + theme(legend.position = "none"),
  p_data_ecmr + theme(legend.position = "none"),
  p_data_col  + theme(legend.position = "none"),
  legend_plot,
  ncol = 3
) + plot_annotation(tag_levels = list(c("A","B","C","D","E",""))) &
  theme(plot.tag   = element_text(size = 28, face = "bold"),
        axis.text  = element_text(size = 24),
        axis.title = element_text(size = 26, face = "bold"))
data_combined

ggsave("plots/data_combined.png",
       plot = data_combined, width = 24, height = 14, dpi = 300)
```

