---
title: "Plant-FATE: Stability of the Optimisation -- Before vs After Growth"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 14
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Motivation

Manuscript Figures S10/S11 (page 41) test whether the yearly local-pattern-search
re-optimisation converges to the same trait choice regardless of the tree's *starting*
`root_no0` (3 values, constant vs real climate) -- a "before growth" perturbation: the
parameter is changed before any growth happens, then the whole run proceeds from seedling.

This script asks the same stability question for all **four parameters the search itself
optimises** (`root_no0`, `root_length0`, `investment_from_tree` = ECM allocation,
`mycorrhized` = ECM colonisation), and -- the new part -- compares that against perturbing the
*same* parameter **after** the tree has already grown for 50 years under baseline conditions
(checkpoint, then replace the parameter, then continue). If the optimisation is genuinely
stable, "before" and "after" perturbations of the same size should converge to the same place;
if they diverge, the optimum depends on history, not just on the parameter value. This is
**not** a redo of Figure S8 (the OAT sensitivity index on raw physiological outputs under a
fixed parameter) -- no raw-parameter OAT table, no sensitivity index, no physiology-only
`grow_for_dt()` loop. Every run here uses the real re-optimisation search throughout.

```{r libraries}
library(PlantFATE)
library(tidyverse)
library(patchwork)
```

## Settings

```{r settings}
param_file   <- "/home/josimms/Documents/Austria/Plant-FATE/tests/params/p_test_boreal.ini"
weather_file <- "/home/josimms/Documents/Austria/Plant-FATE/data/ERAS_Monthly_constant1974.csv"

dt                  <- 1/12
start_year          <- 1960
spinup_years        <- 50
response_years      <- 30   # widened per Joanna's request, more room to observe reconvergence
perturb_t0          <- start_year + spinup_years
end_year            <- perturb_t0 + response_years   # 2030; total horizon for both conditions
trial_horizon_years <- 3

soil_n_default <- 0.32

STEP_ARGS <- list(
  root_no_step_factor = 3.0,  root_no_min = 1e3,  root_no_max = 5e7,
  root_length_step_factor = 1.5, root_length_min = 0.05, root_length_max = 6.0,
  ecto_step = 0.05, ecto_min = 0.0, ecto_max = 1.0,
  mycorrhized_step = 0.05, mycorrhized_min = 0.0, mycorrhized_max = 1.0
)

out_dir <- here::here("vignettes/plots/delayed_perturbation_stability")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

perturb_fracs <- c(0.10, 0.25, 0.50)   # widened per Joanna's request

ini_lines <- readLines(param_file)
read_ini_val <- function(key) {
  line <- grep(paste0("^", key, "\\s*="), ini_lines, value = TRUE)[1]
  if (is.na(line)) stop("ini key not found: ", key)
  as.numeric(sub(paste0("^", key, "\\s*=\\s*([0-9.eE+\\-]+).*"), "\\1", line))
}

decision_params <- list(
  list(struct = "par0",    field = "root_no0",             label = "Root tip density (n, root_no0)"),
  list(struct = "par0",    field = "root_length0",         label = "Root tip length (l, root_length0)"),
  list(struct = "traits0", field = "investment_from_tree", label = "ECM C allocation fraction (φₘ)"),
  list(struct = "uptake0", field = "mycorrhized",          label = "ECM colonisation fraction (m)")
)
decision_params <- map(decision_params, function(p) { p$baseline <- read_ini_val(p$field); p })

root_no0_baseline     <- read_ini_val("root_no0")
root_length0_baseline <- read_ini_val("root_length0")
lai0_val              <- read_ini_val("lai0")   # root_no0 is per m^2 LEAF area; root_no0*lai0 = per m^2 CROWN area (ini's own convention)
```

## Simulation helpers

```{r sim_helpers}
run_search <- function(lho, t0, t1) {
  log <- do.call(lho$run_with_relaxed_local_reopt_trajectory,
                 c(list(t0, t1, dt, trial_horizon_years), STEP_ARGS))
  df <- as.data.frame(do.call(rbind, log))
  names(df) <- c(lho$get_header(), "ecto_allo", "mycorrhized_committed", "root_no_target", "root_length_target")
  df$gpp_ca         <- df$assim_gross / pmax(df$crown_area, 1e-12)
  df$total_n_uptake <- df$root_uptake + df$myco_uptake
  df$date <- as.Date(df$i - 2440588, origin = "1970-01-01")
  df
}

fresh_lho <- function() {
  lho <- new(LifeHistoryOptimizer, param_file)
  lho$set_i_metFile(weather_file); lho$set_a_metFile(weather_file); lho$set_co2File("")
  lho$set_soil_nitrogen(soil_n_default)
  lho
}

# "BEFORE growth": the parameter is set to its perturbed value at construction
# (before lho$init() for investment_from_tree/mycorrhized -- live traits0/
# uptake0 fields copied in by Plant::init(); via root_override() right after
# init() for root_no0/root_length0, which otherwise have no live effect -- see
# notes "Step 1 status"), then the search runs continuously start_year..end_year.
# Generalises Fig S10/S11's existing method (root_no0 only) to all 4 params.
run_before <- function(field = NULL, value = NULL) {
  lho <- fresh_lho()
  if (!is.null(field) && field == "investment_from_tree") lho$traits0$investment_from_tree <- value
  if (!is.null(field) && field == "mycorrhized")          lho$uptake0$mycorrhized          <- value
  lho$init()
  rn <- if (!is.null(field) && field == "root_no0")     value else root_no0_baseline
  rl <- if (!is.null(field) && field == "root_length0") value else root_length0_baseline
  lho$root_override(rn, rl)
  run_search(lho, start_year, end_year)
}

# "AFTER growth": baseline search for 50y, checkpoint, THEN replace one
# decision variable's CURRENT converged value, continue the search 20y more.
spin_up_shared <- function() {
  lho <- fresh_lho()
  lho$init()
  lho$root_override(root_no0_baseline, root_length0_baseline)
  spinup_df <- run_search(lho, start_year, perturb_t0)
  lho$save_checkpoint()
  list(lho = lho, checkpoint = as.list(tail(spinup_df, 1)))
}

run_after <- function(lho, checkpoint, field = NULL, value = NULL) {
  lho$restore_checkpoint()
  # Reset all 4 decision vars to their checkpoint (converged) values first --
  # restore_checkpoint() already does this for the live P, this just makes the
  # staging fields consistent -- then override exactly one if requested.
  lho$traits0$investment_from_tree <- checkpoint$ecto_allo
  lho$uptake0$mycorrhized          <- checkpoint$mycorrhized_committed
  lho$reinit_params()
  lho$root_override(checkpoint$root_no, checkpoint$root_length)

  if (!is.null(field)) {
    if (field == "root_no0") {
      lho$root_override(value, checkpoint$root_length)
    } else if (field == "root_length0") {
      lho$root_override(checkpoint$root_no, value)
    } else if (field == "investment_from_tree") {
      lho$traits0$investment_from_tree <- value; lho$reinit_params()
    } else if (field == "mycorrhized") {
      lho$uptake0$mycorrhized <- value; lho$reinit_params()
    }
  }
  run_search(lho, perturb_t0, end_year)
}

reference_value <- function(p, checkpoint) {
  switch(p$field,
    root_no0             = checkpoint$root_no,
    root_length0         = checkpoint$root_length,
    investment_from_tree = checkpoint$ecto_allo,
    mycorrhized           = checkpoint$mycorrhized_committed
  )
}
```

## Baselines (no perturbation, either timing)

```{r baselines}
# NOT cache=TRUE: lho is an Rcpp Module object (external pointer). knitr's
# cache serializes chunk results to disk and reloads them in later renders'
# fresh R sessions -- an external pointer deserialized that way is dead
# ("NULL value passed as symbol address" on any method call), which silently
# broke every "after" run in an earlier version of this script (caught by
# the per-run tryCatch below, but invisible because the global `message =
# FALSE` chunk option suppressed the "FAILED: ..." diagnostic -- see
# notes_delayed_perturbation_sensitivity.md). The spin-up is cheap (~1-2 min)
# so just always recompute it fresh; the 48 perturbation runs still use their
# own manual per-run .rds cache (file.exists() check below), which is safe
# since those are plain data frames, not Rcpp objects.
cat("BEFORE-condition baseline: continuous", end_year - start_year, "year search...\n")
t0 <- Sys.time()
baseline_before <- run_before()
cat("  done in", round(as.numeric(Sys.time() - t0, units = "secs"), 1), "s\n")

cat("AFTER-condition spin-up + no-shock continuation...\n")
t0 <- Sys.time()
spin <- spin_up_shared()
lho <- spin$lho; checkpoint <- spin$checkpoint
baseline_after <- run_after(lho, checkpoint)
cat("  done in", round(as.numeric(Sys.time() - t0, units = "secs"), 1), "s\n")
cat("Converged-at-checkpoint (year", perturb_t0, "): root_no=", checkpoint$root_no,
    " root_length=", checkpoint$root_length, " ecto_allo=", checkpoint$ecto_allo,
    " mycorrhized=", checkpoint$mycorrhized_committed, "\n")
```

## Perturbation runs: 4 parameters x {before, after} x 3 magnitudes x {high, low}

```{r perturbation_runs}
# Physical bounds per decision variable (same as the search's own candidate
# bounds in STEP_ARGS). A kick must be clamped to these -- unlike the search's
# own candidate generation (make_levels_mult/make_levels_add, which always
# clamps), a value assigned directly via root_override()/reinit_params() has
# no such protection, and mycorrhized's checkpoint value (~0.994) sits close
# enough to its upper bound (1.0) that an unclamped "+50%" kick (1.49) feeds
# a physically impossible colonisation fraction straight into the uptake
# model -- confirmed by direct test to produce pathological trajectories
# (collapsing to 0 within the response window). root_length0's checkpoint
# (~5.66) is similarly close to its bound (6.0) and has the same problem.
field_bounds <- list(
  root_no0             = c(STEP_ARGS$root_no_min,     STEP_ARGS$root_no_max),
  root_length0         = c(STEP_ARGS$root_length_min, STEP_ARGS$root_length_max),
  investment_from_tree = c(STEP_ARGS$ecto_min,        STEP_ARGS$ecto_max),
  mycorrhized           = c(STEP_ARGS$mycorrhized_min, STEP_ARGS$mycorrhized_max)
)

run_table <- map_dfr(decision_params, function(p) {
  before_ref <- p$baseline
  after_ref  <- reference_value(p, checkpoint)
  bounds <- field_bounds[[p$field]]
  map_dfr(c("before", "after"), function(timing) {
    ref <- if (timing == "before") before_ref else after_ref
    map_dfr(perturb_fracs, function(frac) {
      delta <- frac * abs(ref)
      tibble(struct = p$struct, field = p$field, label = p$label, timing = timing,
             magnitude = frac, direction = c("high", "low"),
             value = pmin(pmax(c(ref + delta, ref - delta), bounds[1]), bounds[2]))
    })
  })
})
cat("Running", nrow(run_table), "perturbed searches (~",
    round(sum(if_else(run_table$timing == "before", end_year - start_year, response_years)) * 0.74 / 60, 0),
    "min estimated)...\n")

t0 <- Sys.time()
run_dfs <- map(seq_len(nrow(run_table)), function(i) {
  row <- run_table[i, ]
  cache_path <- file.path(out_dir, sprintf("run_%s_%s_%.2f_%s.rds", row$field, row$timing, row$magnitude, row$direction))
  if (file.exists(cache_path)) return(readRDS(cache_path))

  df <- tryCatch({
    if (row$timing == "before") run_before(row$field, row$value)
    else                        run_after(lho, checkpoint, row$field, row$value)
  }, error = function(e) {
    # cat(), not message() -- the global `message = FALSE` chunk option
    # silently swallows message() output, which is exactly how the broken-
    # cached-lho bug above went undetected the first time.
    cat("FAILED: ", row$label, " ", row$timing, " (", row$magnitude, ", ", row$direction, "): ", conditionMessage(e), "\n")
    NULL
  })
  if (!is.null(df)) {
    df$field <- row$field; df$timing <- row$timing
    df$magnitude <- row$magnitude; df$direction <- row$direction
    saveRDS(df, cache_path)
  }
  df
})
cat("Done in", round(as.numeric(Sys.time() - t0, units = "mins"), 1), "min\n")

baseline_before$field <- NA_character_; baseline_before$timing <- "before"
baseline_before$magnitude <- NA_real_;  baseline_before$direction <- "baseline"
baseline_after$field  <- NA_character_; baseline_after$timing  <- "after"
baseline_after$magnitude  <- NA_real_;  baseline_after$direction  <- "baseline"

all_df <- bind_rows(baseline_before, baseline_after, Filter(Negate(is.null), run_dfs)) |>
  mutate(
    root_no_crown = root_no * lai,       # per-leaf-area -> per-crown-area, using each row's own LAI
    ecto_allo_pct      = ecto_allo * 100,
    mycorrhized_pct    = mycorrhized_committed * 100
  )
```

## Stability plots: before vs after, per parameter

Left column: perturbed at t=0 (before growth), full `r end_year - start_year`y trajectory.
Right column: baseline grown 50y then perturbed (after growth), last `r response_years`y shown.
Both columns share the y-axis per row so the converged level is directly comparable.

```{r stability_plots, fig.height=7, fig.width=13}
# Publication set: 6 panels. ECM biomass and total N uptake dropped (not
# needed for this figure). Root tip density converted from per-leaf-area to
# per-crown-area (root_no * lai -- the ini's own convention, see
# p_test_boreal.ini's root_no0 comment). ECM allocation/colonisation given
# explicit units, matching the manuscript's own Fig S10/S11 convention
# ("% NPP" / "% roots covered") rather than the dimensionless phi_m/m symbols.
trace_vars <- c("root_no_crown", "root_length", "ecto_allo_pct", "mycorrhized_pct",
                 "height", "gpp_ca")
trace_labels <- c(
  root_no_crown = "Root tip density (m⁻² crown area)", root_length = "Root tip length (mm)",
  ecto_allo_pct = "ECM allocation (% NPP)", mycorrhized_pct = "ECM colonisation (% roots covered)",
  height = "Height (m)", gpp_ca = "GPP/crown area (kg C m⁻² yr⁻¹)"
)

# Search bounds (STEP_ARGS) for the 4 variables the optimisation itself
# controls -- injected as invisible points so each of those panels' free_y
# axis is forced to span the full range the search was allowed to explore
# (height/GPP have no such "optimisation bounds", so they're left on their
# own data-driven range). root_no's bound is per leaf area; scaled by lai0
# (not each row's own instantaneous lai) to get a fixed, single per-crown-area
# bound for the axis -- an approximation, since actual lai varies over the run.
bound_df <- tibble(
  var = factor(c("root_no_crown", "root_no_crown", "root_length", "root_length",
                 "ecto_allo_pct", "ecto_allo_pct", "mycorrhized_pct", "mycorrhized_pct"),
               levels = trace_vars, labels = trace_labels[trace_vars]),
  y = c(STEP_ARGS$root_no_min * lai0_val, STEP_ARGS$root_no_max * lai0_val,
        STEP_ARGS$root_length_min, STEP_ARGS$root_length_max,
        STEP_ARGS$ecto_min * 100, STEP_ARGS$ecto_max * 100,
        STEP_ARGS$mycorrhized_min * 100, STEP_ARGS$mycorrhized_max * 100)
)

mag_colours <- c(setNames(RColorBrewer::brewer.pal(3, "Dark2"), c("10%", "25%", "50%")), Baseline = "black")

bound_wide <- bound_df |>
  group_by(var) |>
  summarise(bound_min = min(y), bound_max = max(y), .groups = "drop")

# Shared by plot_timing() (for the in-panel annotation) and the summary
# export below (for the manuscript caption text) -- computed once so the two
# can never drift out of sync with each other.
get_plot_data <- function(pfield, tm) {
  dat <- all_df |>
    filter((field == pfield & timing == tm) | (is.na(field) & timing == tm)) |>
    mutate(run_label = if_else(is.na(field), "Baseline", sprintf("%s%%", magnitude * 100))) |>
    pivot_longer(cols = all_of(trace_vars), names_to = "var", values_to = "y") |>
    mutate(var = factor(var, levels = trace_vars, labels = trace_labels[trace_vars]))

  # Excludes the very first logged point: that value is exactly the forced
  # kicked starting value by construction (not something the search chose),
  # so including it would measure the size of the artificial kick rather
  # than how far the search itself subsequently roamed. All runs within one
  # pfield/tm call share the same first date, so a single global filter works.
  ann_df <- dat |>
    filter(var %in% bound_wide$var, date > min(date)) |>
    group_by(var) |>
    summarise(data_min = min(y, na.rm = TRUE), data_max = max(y, na.rm = TRUE), .groups = "drop") |>
    left_join(bound_wide, by = "var") |>
    mutate(
      pct_used = 100 * (data_max - data_min) / (bound_max - bound_min),
      label = sprintf("range used: %.2g%% of [%.3g, %.3g]", pct_used, bound_min, bound_max),
      top = (tm == "after" & var == trace_labels[["root_no_crown"]]) |
            (tm == "before" & var == trace_labels[["ecto_allo_pct"]]) |
            (tm == "before" & var == trace_labels[["root_no_crown"]]),
      ann_y = if_else(top, Inf, -Inf),
      ann_vjust = case_when(
        top ~ 1,
        # Nudged up from the bottom edge, between the 25% and 50% lines --
        # sits a little too close to the data right at the edge otherwise.
        pfield == "investment_from_tree" & tm == "before" & var == trace_labels[["root_length"]] ~ -2,
        TRUE ~ 0
      )
    )
  list(dat = dat, ann_df = ann_df)
}

plot_timing <- function(pfield, tm) {
  pd <- get_plot_data(pfield, tm)
  dat <- pd$dat; ann_df <- pd$ann_df

  # Auto-scaled (free_y) axes everywhere, zoomed to the actual data -- but for
  # the 4 variables the search itself bounds, annotate each panel with how
  # much of that allowed range the trajectory actually used. Keeps every
  # panel equally readable while still making the "stays within a sensible
  # range" point, without compressing detail the way forcing the axis to the
  # full bound did (root_no_crown's bound alone spans ~5 orders of magnitude).
  # Default bottom-right placement, except where the trajectory itself sits
  # low for most of the window (so bottom-right would sit on top of the
  # data): "after growth" root tip density (any parameter), and the ECM
  # allocation panel specifically in the investment_from_tree figure.
  ggplot(dat, aes(date, y, colour = run_label, linetype = direction)) +
    geom_line(data = filter(dat, run_label != "Baseline"), linewidth = 0.6) +
    geom_line(data = filter(dat, run_label == "Baseline"), linewidth = 0.9) +
    # geom_label (not geom_text): the white fill sits on top of the data lines
    # instead of competing with them, so the annotation stays legible
    # regardless of where the trajectory happens to sit in the panel.
    geom_label(data = ann_df, aes(x = Inf, y = ann_y, label = label, vjust = ann_vjust), inherit.aes = FALSE,
               hjust = 1, size = 7.5, colour = "grey20", fill = "white",
               alpha = 0.85, label.size = 0, fontface = "plain") +
    facet_wrap(~var, scales = "free_y", ncol = 3) +
    scale_colour_manual(values = mag_colours, name = "Magnitude") +
    scale_linetype_manual(values = c(high = 1, low = 2, baseline = 1),
                           breaks = c("high", "low"), labels = c("High", "Low"), name = "Direction") +
    guides(
      colour   = guide_legend(override.aes = list(linewidth = 3)),
      linetype = guide_legend(override.aes = list(linewidth = 3))
    ) +
    labs(x = NULL, y = NULL, title = toupper(paste0(tm, " growth"))) +
    pub_theme
}

# Matches the house style used throughout sensitivity_analysis.Rmd's own
# published tornado/heatmap figures (theme_classic(base_size = 22), bold
# titles, bold legend title) -- scaled up further on top of that since this
# figure's panels are denser (6 vars x 2 timing rows).
pub_theme <- theme_classic(base_size = 26) +
  theme(
    axis.text       = element_text(size = 24),
    strip.text      = element_text(size = 24, face = "bold"),
    legend.text     = element_text(size = 24),
    legend.title    = element_text(face = "bold", size = 26),
    plot.title      = element_text(face = "bold", size = 32),
    legend.position = "bottom",
    legend.key.width = unit(2.2, "cm")
  )

stability_plots <- map(decision_params, function(p) {
  plt <- (plot_timing(p$field, "before") / plot_timing(p$field, "after")) +
    plot_layout(guides = "collect") & theme(legend.position = "bottom")
  ggsave(file.path(out_dir, sprintf("stability_%s.png", p$field)), plt, width = 28, height = 19, dpi = 300)
  plt
})
names(stability_plots) <- map_chr(decision_params, "field")

stability_plots[["root_no0"]]
stability_plots[["root_length0"]]
stability_plots[["investment_from_tree"]]
stability_plots[["mycorrhized"]]
```

## Range-used summary (for the manuscript caption text)

Same numbers shown as text on each panel, pulled out as a table so the
MAIN.tex captions can report them without re-deriving by eye.

```{r range_summary}
range_summary <- map_dfr(decision_params, function(p) {
  map_dfr(c("before", "after"), function(tm) {
    get_plot_data(p$field, tm)$ann_df |>
      transmute(field = p$field, timing = tm, var = as.character(var), pct_used,
                bound_min, bound_max)
  })
})
write_csv(range_summary, file.path(out_dir, "range_used_summary.csv"))
print(range_summary, n = Inf)
```
