---
title: "Plant-FATE: One-at-a-time Sensitivity Analysis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 14
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Libraries

```{r libraries}
library(PlantFATE)
library(tidyverse)
library(furrr)
library(patchwork)

plan(multicore, workers = max(1, parallel::detectCores() - 1))
```

## Settings

```{r settings}
param_file    <- "/home/josimms/Documents/Austria/Plant-FATE/tests/params/p_test_boreal.ini"
weather_file  <- "/home/josimms/Documents/Austria/Plant-FATE/data/ERAS_Monthly.csv"

testing_mode  <- FALSE   # set TRUE for a quick sanity check (shorter run)
start_year    <- 1960
end_year      <- if (testing_mode) 1985 else 2022
dt            <- 1/12

soil_n_default <- 0.405   # kg N m⁻³ — N_s default from p_test_boreal.ini
perturb_frac   <- 0.10    # ±10% of |baseline|

# Outputs to track: summer (Jun–Aug) mean over the last 10 years of the run
output_vars <- c(
  "assim_gross",
  "height",
  "leaf_nitrogen_concentration",
  "ectomycorrhiza_mass",
  "total_n_uptake",
  "gpp_ca"
)

output_labels <- c(
  assim_gross                 = "GPP (kg C yr⁻¹)",
  height                      = "Height (m)",
  leaf_nitrogen_concentration = "Leaf N conc. (kg N kg⁻¹)",
  ectomycorrhiza_mass         = "ECM biomass (kg)",
  total_n_uptake              = "Total N uptake (kg N yr⁻¹)",
  gpp_ca                      = "GPP / crown area (kg C m⁻² yr⁻¹)"
)

dir.create("plots", showWarnings = FALSE)
```

## Parameter definitions

Parameters are grouped by process; baselines match `p_test_boreal.ini`.

```{r param_definitions}
params <- list(
  # --- Photosynthesis ---
  list(struct = "par0",    field = "kphio",               label = "Quantum yield (kphio)",          group = "Photosynthesis", baseline = 0.08),
  list(struct = "par0",    field = "alpha",               label = "Photo. capacity cost (α)",       group = "Photosynthesis", baseline = 0.1008),
  list(struct = "par0",    field = "gamma",               label = "Hydraulic path cost (γ)",        group = "Photosynthesis", baseline = 0.1805),
  list(struct = "par0",    field = "a_jmax",              label = "Jmax:Vcmax scaling (a_jmax)",    group = "Photosynthesis", baseline = 3600),
  # --- Plant traits ---
  list(struct = "traits0", field = "lma",                 label = "Leaf mass per area (LMA)",       group = "Plant traits",   baseline = 0.086),
  list(struct = "traits0", field = "hmat",                label = "Max height (hmat, m)",           group = "Plant traits",   baseline = 35),
  list(struct = "traits0", field = "wood_density",        label = "Wood density (kg m⁻³)",          group = "Plant traits",   baseline = 412),
  list(struct = "traits0", field = "p50_xylem",           label = "Xylem P50 (MPa)",                group = "Plant traits",   baseline = -7),
  # --- ECM / mycorrhiza ---
  list(struct = "traits0", field = "investment_from_tree",          label = "ECM C allocation fraction",         group = "ECM", baseline = 0.15),
  list(struct = "traits0", field = "mycorrhizal_turnover",          label = "ECM turnover rate (yr⁻¹)",         group = "ECM", baseline = 2.0),
  list(struct = "traits0", field = "nc_myco",                       label = "ECM N:C ratio (nc_myco)",           group = "ECM", baseline = 0.1),
  list(struct = "traits0", field = "mycorrhizal_biomass_conversion",label = "ECM biomass conversion (kg kg⁻¹)", group = "ECM", baseline = 2.27),
  list(struct = "uptake0", field = "rho_myco",                      label = "ECM tissue density (kg m⁻³)",      group = "ECM", baseline = 1200),
  list(struct = "uptake0", field = "myco_diameter",                  label = "Hyphal diameter (m)",              group = "ECM", baseline = 2e-5),
  list(struct = "uptake0", field = "investment_from_mycorrhiza",     label = "ECM N return fraction",            group = "ECM", baseline = 0.1),
  # --- Respiration ---
  list(struct = "par0",    field = "rr",                  label = "Fine root resp. rate (rr)",      group = "Respiration",    baseline = 0.4),
  list(struct = "par0",    field = "r_myco",              label = "ECM maint. resp. rate (r_myco)", group = "Respiration",    baseline = 0.3),
  list(struct = "par0",    field = "rs",                  label = "Sapwood resp. rate (rs)",        group = "Respiration",    baseline = 0.04),
  # --- N uptake ---
  list(struct = "uptake0", field = "u_max",               label = "Max area-based uptake (u_max)",  group = "N uptake",       baseline = 0.0088),
  list(struct = "uptake0", field = "k_15",                label = "MM half-saturation (k_15)",      group = "N uptake",       baseline = 0.7),
  list(struct = "uptake0", field = "D",                   label = "Soil N diffusion (D)",           group = "N uptake",       baseline = 0.025),
  list(struct = "uptake0", field = "mycorrhized",         label = "Colonisation fraction",          group = "N uptake",       baseline = 1.0),
  list(struct = "uptake0", field = "u_transfer",          label = "N transfer rate (u_transfer)",   group = "N uptake",       baseline = 3.8e-3),
  list(struct = "par0",    field = "nitrogen_start0",     label = "Initial tree N (nitrogen_start0)", group = "N uptake",     baseline = 0.0039)
)

# Ordered factor levels for plots (group order preserved)
param_levels <- map_chr(params, "label")

group_levels <- c("Photosynthesis", "Plant traits", "ECM", "Respiration", "N uptake")
group_pal    <- c(
  "Photosynthesis" = "#4dac26",
  "Plant traits"   = "#7b3294",
  "ECM"            = "#d01c8b",
  "Respiration"    = "#f1a340",
  "N uptake"       = "#2166ac"
)
```

## Simulation helpers

```{r sim_helpers}
# Run the LifeHistoryOptimizer with an optional single-parameter perturbation.
# The parameter is set on the appropriate struct (par0 / traits0 / uptake0)
# before lho$init() is called, matching the pattern in Publication Plots.Rmd.
run_lho <- function(struct = NULL, field = NULL, value = NULL) {
  lho <- new(LifeHistoryOptimizer, param_file)
  lho$set_i_metFile(weather_file)
  lho$set_a_metFile(weather_file)
  lho$set_co2File("")
  lho$set_soil_nitrogen(soil_n_default)

  if (!is.null(struct)) {
    switch(struct,
      par0    = { lho$par0[[field]]    <- value },
      traits0 = { lho$traits0[[field]] <- value },
      uptake0 = { lho$uptake0[[field]] <- value }
    )
  }

  lho$init()

  header    <- lho$get_header()
  years_seq <- seq(start_year, end_year, dt)

  results <- lapply(years_seq, function(t) {
    tryCatch({
      lho$grow_for_dt(t, dt)
      setNames(as.data.frame(t(lho$get_state(t + dt))), header)
    }, error = function(e) NULL)
  })

  df <- do.call(rbind, Filter(Negate(is.null), results))
  df$date <- seq(as.Date(paste0(start_year, "-01-01")),
                 by = "month", length.out = nrow(df))
  df
}

# Summer (Jun–Aug) mean over the last 10 years.
summarise_run <- function(df) {
  df |>
    mutate(
      year           = as.integer(format(date, "%Y")),
      month          = as.integer(format(date, "%m")),
      total_n_uptake = root_uptake + myco_uptake,
      gpp_ca         = assim_gross / pmax(crown_area, 1e-12)
    ) |>
    filter(year >= max(year) - 9, month %in% 6:8) |>
    summarise(across(all_of(output_vars), ~ mean(.x, na.rm = TRUE)))
}
```

## Baseline run

```{r baseline_run, cache=TRUE}
cat("Running baseline...\n")
baseline_sim     <- run_lho()
baseline_summary <- summarise_run(baseline_sim)
print(round(baseline_summary, 5))
```

## OAT perturbation runs

Results are cached to CSV so you can re-run the notebook without repeating simulations.

```{r perturbation_runs, cache=TRUE}
sa_cache <- "plots/sa_perturb_results.csv"

if (file.exists(sa_cache)) {
  perturb_df <- read_csv(sa_cache, show_col_types = FALSE)
  cat("Loaded", nrow(perturb_df), "rows from cache:", sa_cache, "\n")

} else {
  # Build flat table of all (param, direction, value) combinations
  perturbation_table <- map_dfr(params, function(p) {
    delta <- perturb_frac * abs(p$baseline)
    tibble(
      param           = p$field,
      label           = p$label,
      group           = p$group,
      struct          = p$struct,
      direction       = c("high", "low"),
      perturbed_value = c(p$baseline + delta, p$baseline - delta)
    )
  })

  cat("Running", nrow(perturbation_table), "perturbed simulations in parallel...\n")

  perturb_results <- future_map(
    seq_len(nrow(perturbation_table)),
    function(i) {
      row <- perturbation_table[i, ]
      tryCatch({
        df   <- run_lho(row$struct, row$param, row$perturbed_value)
        smry <- summarise_run(df)
        bind_cols(
          tibble(param = row$param, label = row$label, group = row$group,
                 direction = row$direction),
          smry
        )
      }, error = function(e) {
        message("FAILED: ", row$label, " (", row$direction, "): ", conditionMessage(e))
        NULL
      })
    },
    .options = furrr_options(seed = NULL)
  )

  perturb_df <- bind_rows(perturb_results)
  write_csv(perturb_df, sa_cache)
  cat("Saved results to", sa_cache, "\n")
}
```

## Sensitivity indices

The normalized sensitivity index S gives the percentage change in each output per 1% change in the parameter:

$$S = \frac{y_{+} - y_{-}}{y_{\text{baseline}}} \cdot \frac{1}{2 \delta}$$

where $\delta = 0.10$ (10% perturbation).

```{r sensitivity_indices}
# Pivot so each row has both high and low values per (param, output_var)
sens_long <- perturb_df |>
  pivot_longer(cols = all_of(output_vars), names_to = "output_var", values_to = "y") |>
  pivot_wider(names_from = direction, values_from = y)

baseline_long <- baseline_summary |>
  pivot_longer(cols = everything(), names_to = "output_var", values_to = "y_baseline")

sens_df <- sens_long |>
  left_join(baseline_long, by = "output_var") |>
  mutate(
    S = (high - low) / pmax(abs(y_baseline), 1e-12) / (2 * perturb_frac),
    label      = factor(label, levels = rev(param_levels)),
    group      = factor(group, levels = group_levels),
    output_var = factor(output_var, levels = output_vars)
  )

# Top 20 by |S|
sens_df |>
  arrange(desc(abs(S))) |>
  transmute(group, Parameter = label, Output = output_var,
            S = round(S, 3),
            Baseline = round(y_baseline, 5),
            High = round(high, 5),
            Low  = round(low,  5)) |>
  head(20) |>
  print()
```

## Tornado charts

Parameters are ordered by S within each panel so the most influential appear at the top. Colours indicate parameter group.

```{r tornado_charts, fig.height=22, fig.width=18}
tornado_plots <- map(output_vars, function(ov) {
  dat <- sens_df |>
    filter(output_var == ov) |>
    mutate(label = fct_reorder(label, S))

  ggplot(dat, aes(x = S, y = label, fill = group)) +
    geom_col(width = 0.7) +
    geom_vline(xintercept = 0, linewidth = 0.5, colour = "grey30") +
    scale_fill_manual(values = group_pal, name = "Group",
                      drop = FALSE) +
    labs(title = output_labels[[ov]],
         x = "Sensitivity index",
         y = NULL) +
    theme_classic(base_size = 22) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold", size = 22))
})

wrap_plots(tornado_plots, ncol = 3, guides = "collect") +
  plot_annotation(
    #subtitle = "S = % change in output per 1% change in parameter  |  summer mean, last 10 simulation years",
    theme    = theme(plot.subtitle = element_text(size = 18, colour = "grey40"))
  ) &
  theme(legend.position = "bottom",
        legend.title    = element_text(face = "bold"))

ggsave("plots/sensitivity_tornado.png", width = 26, height = 20, dpi = 300)
```

## Sensitivity heatmap

All parameters × outputs in one view. Red = positive S (higher parameter → higher output), blue = negative.

```{r sensitivity_heatmap, fig.height=10, fig.width=13}
heat_df <- sens_df |>
  mutate(
    out_lab = factor(unname(output_labels[as.character(output_var)]),
                     levels = unname(output_labels))
  )

s_lim <- heat_df |>
  filter(as.character(label) != "Quantum yield (kphio)") |>
  pull(S) |> abs() |> max(na.rm = TRUE)

ggplot(heat_df |> filter(as.character(label) != "Quantum yield (kphio)"),
       aes(x = out_lab, y = label, fill = S)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.2f", S)),
            colour = "grey20", size = 5.5, fontface = "bold") +
  scale_fill_distiller(palette  = "PuOr", direction = -1,
                       limits   = c(-s_lim, s_lim),
                       oob      = scales::squish,
                       name     = "Sensitivity\nindex (S)") +
  scale_x_discrete(position = "bottom") +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 22) +
  theme(
    axis.text.x   = element_text(angle = -25, hjust = 0, size = 20),
    axis.text.y   = element_text(size  = 20),
    strip.text.y  = element_text(angle = 0, face = "bold", size = 22),
    panel.grid    = element_blank(),
    legend.position = "right",
    plot.title    = element_text(face = "bold")
  )

ggsave("plots/sensitivity_heatmap.png", width = 18, height = 14, dpi = 300)
```

## Parameter trajectory plots

For the 3 parameters with highest |S| on `gpp_ca`, show the time series of all outputs
under baseline, +10%, and −10% conditions.

```{r trajectory_plots, fig.height=14, fig.width=16}
# Identify the 3 most influential parameters for gpp_ca
top3 <- sens_df |>
  filter(output_var == "gpp_ca") |>
  arrange(desc(abs(S))) |>
  slice_head(n = 3) |>
  pull(label) |> as.character()

cat("Top 3 parameters for gpp_ca:", paste(top3, collapse = ", "), "\n")

# Re-run (or use cached data) for these parameters + baseline
top3_params <- keep(params, ~ .x$label %in% top3)

traj_list <- c(
  list(list(label = "Baseline", struct = NULL, field = NULL, value = NULL)),
  map(top3_params, function(p) {
    delta <- perturb_frac * abs(p$baseline)
    list(
      list(label = paste0(p$label, " +10%"), struct = p$struct, field = p$field, value = p$baseline + delta),
      list(label = paste0(p$label, " −10%"), struct = p$struct, field = p$field, value = p$baseline - delta)
    )
  }) |> list_flatten()
)

traj_runs <- map(traj_list, function(r) {
  df <- run_lho(r$struct, r$field, r$value)
  df |>
    mutate(
      run_label      = r$label,
      year           = as.integer(format(date, "%Y")),
      month          = as.integer(format(date, "%m")),
      total_n_uptake = root_uptake + myco_uptake,
      gpp_ca         = assim_gross / pmax(crown_area, 1e-12)
    ) |>
    filter(month %in% 6:8) |>
    group_by(run_label, year) |>
    summarise(across(all_of(output_vars), ~ mean(.x, na.rm = TRUE)), .groups = "drop")
})

traj_df <- bind_rows(traj_runs) |>
  mutate(
    param_group = case_when(
      run_label == "Baseline"                     ~ "Baseline",
      str_detect(run_label, top3[1])              ~ top3[1],
      length(top3) >= 2 & str_detect(run_label, top3[2]) ~ top3[2],
      length(top3) >= 3 & str_detect(run_label, top3[3]) ~ top3[3],
      TRUE                                        ~ run_label
    ),
    linetype = ifelse(str_detect(run_label, "\\+10%"), "dashed",
               ifelse(str_detect(run_label, "−10%"),   "dotted", "solid"))
  )

traj_color <- c("Baseline" = "black",
                setNames(c("#e41a1c", "#377eb8", "#4dac26"), top3[seq_along(top3)]))

traj_ltype <- c("solid", "dashed", "dotted")

traj_panel <- function(var) {
  ggplot(traj_df, aes(x = year, y = .data[[var]],
                       colour = param_group, linetype = run_label)) +
    geom_line(linewidth = 0.8) +
    scale_colour_manual(values = traj_color, name = "Parameter") +
    scale_linetype_manual(
      values = setNames(
        c("solid", rep(c("dashed", "dotted"), length(top3_params))),
        c("Baseline", map(top3_params, function(p)
            c(paste0(p$label, " +10%"), paste0(p$label, " −10%"))) |> unlist())
      ),
      guide = "none"
    ) +
    labs(x = "Year", y = output_labels[[var]]) +
    expand_limits(y = 0) +
    theme_classic(base_size = 22)
}

traj_plots <- map(output_vars, traj_panel)

wrap_plots(traj_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title    = "Trajectories for 3 most sensitive parameters (net GPP / crown area)",
    subtitle = "Solid = baseline; dashed = +10%; dotted = −10%",
    theme    = theme(plot.title    = element_text(face = "bold"),
                     plot.subtitle = element_text(colour = "grey40"))
  ) &
  theme(legend.position = "right")

ggsave("plots/sensitivity_trajectories.png", width = 16, height = 18, dpi = 300)
```

## kphio diagnostic: total N uptake over time

Does the large negative sensitivity of total N uptake to kphio reflect a life-history-stage shift rather than a direct parameter effect? If so, the high-kphio plant will grow faster (greater height) but show lower N uptake in later years as it passes through a different allocation phase.

```{r kphio_diagnostic, fig.height=8, fig.width=14}
kphio_baseline <- params[[which(map_chr(params, "field") == "kphio")]]$baseline
kphio_delta    <- perturb_frac * kphio_baseline

kphio_runs <- list(
  list(label = "Baseline",   value = NULL),
  list(label = "kphio +10%", value = kphio_baseline + kphio_delta),
  list(label = "kphio −10%", value = kphio_baseline - kphio_delta)
)

kphio_traj <- map(kphio_runs, function(r) {
  df <- if (is.null(r$value)) baseline_sim else run_lho("par0", "kphio", r$value)
  df |>
    mutate(
      run_label      = r$label,
      year           = as.integer(format(date, "%Y")),
      month          = as.integer(format(date, "%m")),
      total_n_uptake = root_uptake + myco_uptake
    ) |>
    filter(month %in% 6:8) |>
    group_by(run_label, year) |>
    summarise(total_n_uptake = mean(total_n_uptake, na.rm = TRUE),
              height         = mean(height,         na.rm = TRUE),
              .groups = "drop")
}) |> bind_rows() |>
  mutate(run_label = factor(run_label, levels = c("Baseline", "kphio +10%", "kphio −10%")))

kphio_pal      <- c("Baseline" = "black", "kphio +10%" = "#e41a1c", "kphio −10%" = "#377eb8")
kphio_linetypes <- c("Baseline" = "solid", "kphio +10%" = "dashed", "kphio −10%" = "dotted")

p_n <- ggplot(kphio_traj, aes(x = year, y = total_n_uptake,
                               colour = run_label, linetype = run_label)) +
  geom_line(linewidth = 1) +
  scale_colour_manual(values = kphio_pal,      name = NULL) +
  scale_linetype_manual(values = kphio_linetypes, name = NULL) +
  labs(x = "Year", y = "Total N uptake (kg N yr⁻¹)") +
  expand_limits(y = 0) +
  theme_classic(base_size = 22)

p_h <- ggplot(kphio_traj, aes(x = year, y = height,
                               colour = run_label, linetype = run_label)) +
  geom_line(linewidth = 1) +
  scale_colour_manual(values = kphio_pal,      name = NULL) +
  scale_linetype_manual(values = kphio_linetypes, name = NULL) +
  labs(x = "Year", y = "Height (m)") +
  expand_limits(y = 0) +
  theme_classic(base_size = 22)

(p_n | p_h) +
  plot_annotation(
    subtitle = "Summer (Jun–Aug) means  |  solid = baseline, dashed = +10%, dotted = −10%",
    theme    = theme(plot.subtitle = element_text(size = 16, colour = "grey40"))
  ) &
  theme(legend.position = "bottom")

ggsave("plots/kphio_n_uptake_diagnostic.png", width = 14, height = 8, dpi = 300)
```
